### 2012-07-26 artigo.tex
Adicionado dois arquivos

\documentclass[a4paper,11pt]{article}

\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[brazil]{babel}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{graphicx}
\pagestyle{empty}
\usepackage{amsmath}%
\usepackage{amsfonts}%
\usepackage{mathrsfs}%
\usepackage{amssymb}%
\usepackage{xcolor}
\usepackage{indentfirst}

\headheight 20mm      %
\oddsidemargin 2.0mm  %
\evensidemargin 2.0mm %
\topmargin -40mm      %
\textheight 250mm     %
\textwidth 160mm      %

\hyphenation{es-pa-lha-men-to po-pu-la-\c c\~ao}

\begin{document}

\section{Compactação de matrizes}

O processo de compactação de matrizes consiste em reduzir o número de bytes
necessários para alocar uma matriz na memória. Esse processo permite que uma
maior quantidade de informações seja armazenada, utilizando o mesmo hardware 
disponível para o método de alocação de memória tradicional. Uma das vantagens 
em aplicar esse método é a possibilidade de trabalhar com
grandes quantidades de informações sobre dados diretamente da memória sem a
necessidade de acesso a disco, leitura e escrita, o que onera o processamento.

As etapas e as implementações necessárias para utilizar o método proposto são
descritas a seguir. O processo de
compactação de uma matriz é realizado por linha, ou seja, 
o número de colunas é reduzido. Considere uma matriz $M_{l \times c}$, na qual
$l$ é o número de linhas e $c$ o número de colunas. Cada elemento dessa
matriz, denominado
$m_{ij}$, armazena apenas um número inteiro. Para um número inteiro ser
armazenado na memória de um processador de 64 bits, utilizando o maior número de
bits disponíveis,
demandam-se 64 bits. Logo, o número de bits ($nb_T$) necessários para
alocar todos os elementos de uma linha da matriz M da forma tradicional pode
ser definido por:

\begin{equation}
  nb_T = c \times 64
\end{equation}

O $nb_T$ calculado representa um número maior de bits do que realmente é utilizado. 
Considere que em uma linha da matriz, por exemplo, existam apenas elementos com valores iguais a 0 ou 1. Tais
valores são representados em binário apenas por 1 bit. Portanto, quando
alocados em uma variável que demande 64 bits para ser armazenada em memória,
ocorre uma perda de 63 bits,
que poderiam ser utilizados para armazenar outros números ou informações. Esse
exemplo trata do caso mais expressivo com relação ao desperdício de memória.
Calculando o número de bits $nb_C$ realmente necessário para implementação desse
exemplo, tem-se

\begin{equation}
  nb_C = c \times 1
\end{equation}

A economia $E_c$ de bits pode ser representada pela diferença entre $nb_T$ e $nb_C$,

\begin{equation}
 E_c = nb_T - nb_C
\end{equation}

\begin{equation}
 E_c =  c \times 64 - c \times 1
\end{equation}

\begin{equation}
 E_c =  c \times (64 -  1)
\end{equation}

\begin{equation}
 E_c =  c \times 63
\end{equation}

Com base nas equações descritas anteriormente, pode-se constatar que a
economia $E_c$ é elevada, pois $c$ colunas demandam apenas $c \times 1$ bits ao
invés de $c \times 64$ bits. O mesmo ocorre com outros valores que não
necessitem de 
64 bits para serem representados em memória, pois nem sempre todos os bits são
utilizados.

Existem duas formas de realizar as manipulações desses bits para minimizar
a perda de espa\c co. Cada
uma dessas formas segue uma metodologia específica que deve ser aplicada
à determinados casos.

A primeira metodologia consiste em determinar o valor do maior elemento que deve 
ser armazenado na matriz e como consequência determina o maior número de bits ($B$) realmente necessário para 
armazenar esse valor. Cada elemento $m_{ij}$ da matriz M é representado por uma tira composta por 64 bits. No 
exemplo mencionado, para os valores 0 e 1, o valor de $B$ é igual a 1, ou seja,
em uma tira de 64 bits, podem-se alocar 64 valores. Cada bit
dessa tira irá representar um dos valores a serem alocados. Esse fato é
responsável pela economia de bits, pois 64 valores, 
que eram armazenados tradicionalmente em 64 colunas, são 
armazenados em apenas uma.

Se o maior valor a ser armazenado é 32, então $B = 5$, pois o número
32 demanda
5 bits para ser representado na memória. Assim, a tira é dividida em 12
regiões, totalizando 60 bits. Os primeiros 12 
valores são armazenados nessas regiões, sendo que  13$^o$ valor 
deve ser dividido, para que seja armazenada parte nos quatro bits que
sobram da tira e o bit restante em outra tira. A divisão de um valor em dois
segmentos faz com que o processador execute um número maior de 
operações para armazenar os valores, o que em determinados casos pode ser
inviável. 
Essa metodologia deve ser aplicada em matrizes que possuem valores próximos,
para que não ocorra perda de bits.

A segunda metodologia consiste em tentar minimizar as perdas quando os números
possuem valores distantes. Por exemplo, quando deseja-se
armazenar valores de magnitudes diferentes, como 0, 1 e 1023. Nessa
faixa de valores, pode-se verificar que os valores 0 e 1 necessitam apenas de 1
bit, enquanto o valor 1023 demanda 10 bits. Seguindo a primeira
metodologia, teríamos que dividir a tira em regiões de 10 bits, o que para
armazenar os valores 0 e 1, faz com que ocorra a perda de 9 bits, não
permitindo uma compactação mais eficaz. Para contornar esse problema é sugerida
uma proposta de compactação dos dados.
O primeiro passo é determinar o maior número de bits necessários $B$. Logo em
seguida, determinar o número de bits necessários para representar $B$, sendo
esse número definido por $S$. A tira não
será mais representada em regiões de mesmo tamanho, mas em regiões com tamanhos
distintos. Dessa maneira, um número para ser representado terá que ter o número
de bits necessários para representá-lo armazenado em uma região $S$, e
na região logo em seguida, o seu respectivo valor armazenado.
Para representar um número, por exemplo 1, nessa nova metodologia, com o número de bits do maior
elemento igual a 10, demanda-se o número 5, em que 4 bits é utilizando para
representar o tamanho da região a ser usada e mais um bit para armazenar o valor
nesse caso.

Para ilustrar essa metodologia, os valores 0, 1023 e 1 serão
alocados em uma tira. Supondo que esses valores fazem parte de uma linha de
uma matriz e que são armazenados em três colunas pelo método tradicional. Na
tabela \ref{tab:1} esses valores estão representados em binário.

\begin{table}[htpb]
  \centering
  \caption{Representação em binário.}
  \begin{tabular}{rrr}
    \hline
    Número & Binário & Quantidade de bits \\
    \hline
    1		&           1	& 1 \\
    1023	& 11111111111   & 10 \\
    0		&           0	& 1 \\
    \hline
  \end{tabular}
  \label{tab:1}
\end{table}

Seguindo a metodologia proposta tem-se a tira

\begin{equation}
 000011111111111101010001
\end{equation}

\noindent em que os valores 0, 1023 e 1  já foram
adicionados. Uma representação, na qual é possível verificar as regiões em que
ocorrem o armazenamento está descrita a seguir

\begin{equation}
 \underbrace{0}_{0}\underbrace{0001}_{1}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1}_{1}\underbrace{0001}_{1}
\end{equation}

\noindent em que as regiões para armazenar os número de bits e os valores em
binários estão armazenados na tira.

O processo de compactação de uma Matriz M, com $c$ colunas é realizado linha a linha. Dado uma linha $j$ da matriz M e $nb(x)$ uma função que
retorna o número de bits necessário para representar o valor $x$, o número total de bits para
compactar essa linha é defindo por:

\begin{equation}
  nb_T = \sum_{i=1}^{c} nb(m_{ij})+ nb(S)
\end{equation}

\noindent em que $B$  é o número de bits necessários para representar o maior valor alocado na matriz M  e $S = nb(B)$, o número de bits $nb_T$ pode ser definido como:

\begin{equation}
  nb_T = c \times nb(S) + \sum_{i=1}^{c} nb(m_{ij})
\end{equation}

Como a variável $m_{ij}$ deve assumir um valor, ela será
considerada como uma variável aleatória, tal que $m_{ij}
\sim U_d(0,N)$, sendo $N$ o maior valor possível a ser armazenado.  Assim, $S =
nb(N)$. Logo, pode-se calcular o valor esperado para $nb_T$, que pode ser 
definido por

\begin{equation}
  E(nb_T) = E(c \times nb(S) + \sum_{i=1}^{c} nb(m_{ij}))
\end{equation}

\begin{equation}
  E(nb_T) = E(c \times nb(S)) + E(\sum_{i=1}^{c} nb(m_{ij}))
\end{equation}

\begin{equation}
  E(nb_T) = E(\sum_{i=1}^{c} nb(m_{ij}))
\end{equation}

\begin{equation}
  E(nb_T) = \sum_{i=1}^{c} E(nb(m_{ij}))
\end{equation}

\end{document}

### 2012-12-27 manuscript.tex
versão revisada

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

\usepackage{psfrag}
\usepackage{subfigure}

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Computable Compressed Matrices}
}
% Insert Author names, affiliations and corresponding author email.
\\
Crysttian Arantes Paix\~{a}o$^{1\ast}$, 
Fl\'{a}vio Code\c{c}o Coelho$^{2}$, 
\\
\bf{1} Crysttian Arantes Paix\~{a}o Applied Mathematics School, Getulio Vargas Foundation, Rio de Janeiro, RJ, Brazil
\\
\bf{2} Fl\'{a}vio Code\c{c}o Coelho Applied Mathematics School, Getulio Vargas Foundation, Rio de Janeiro, RJ, Brazil
\\
$\ast$ E-mail: Corresponding author@institute.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}

\section*{Introduction}

Data compression is traditionally used to reduce storage resouces usage and/or transmission costs\cite{salomon}. Compression techniques can be classified into lossy and lossless. Goos examples of lossy data compression are sound (mp3), image (JPEG) and movie (mpeg) compression. In this paper we are concerned exclusively with lossless compression. In lossless compression, no information is lost, during the compression process, and they are the method of choice to compress numbers and text.

Any kind of compression incurs in some computational cost. such costs often have to be paid twice since the data needs to be decompressed to be used for its original purpose. Sometimes computational costs are irrelevant, but the need to decompress for usage, can signify that savings in storage costs are applicable to storage situation since we still require the uncompressed space allocation for usage.

Lossless compression methods usually try to exploit redundancies present in the data in order to find a shorter form for of describe the same information content, for example a dictionary based compression, can only store the positions in which a given word occurs in a document, thus saving the space required to store all its repeatitions\cite{salomon2}.

Most if not all existing lossless compression methods were develop under the following paradigm: produce $\rightarrow$ compress $\rightarrow$ store $\rightarrow$ uncompress $\rightarrow$ use. However in the brave new world of big data, the need to analyze data immediately after its coming into existence became the norm. And this analysis must take place, efficiently, within the confines of (RAM) memory. Given a sufficiently big flow of data, compression an decompression costs may become prohibitive as it will slow down data transmission. Moreover, the need to decompress before analyzing removes any potential gains in memory usage.

With the growth of data and analytical demands, creative solutions are need to not only efficiently store large data as consume it on demand. This phenomenon is present in many areas of aplication, ranging from business to science\cite{lynch}, and is being called the big data phenomenon.

One data type which is very common across areas of application is numerical data. Most common statistical and machine learning algorithms operate over large matrices, frequently making use of linear algebra methods. So it would be very desirable to have a compression method in which the compressed object retained the mathematical properties as the original, i.e. could be subject to the same mathematical operations. The idea of operating with compressed array is not new\cite{yemliha2007compiler}, but has yet to applied consistently to the field of numerical computations. One application which employs a form compression is the sparse matrix linear algebra algorithms\cite{dodson1991sparse}, in this case there is no encoding of the data, but only the non-zero elements of the matrices are stored and operated upon.

Currently, the technique most commonly used when dealing with large matrices for numerical computations, is memory mapping\cite{van2011numpy,big}. In memory mapping the matrix is allocated in a virtual contiguous address space which overflows from memory into disk, this way larger than memory datastructures can be manipulated as if they were in memory.  This technique has a big performance penalty due to much lower access speeds of disk when compared to RAM.

In this paper we present two methods that allow the lossless compression of (numerical) arrays, which do not require decompression to be manipulated mathematically. The method involves the encoding of the numbers as strings of bits of variable length. The methods resemble the arithmetic coding\cite{bodden2007arithmetic} algorithm, but is cheaper to compute and does not require decompression. We describe the process of compression and decompression, and how to operate with the arrays in compressed format. We discuss the efficiency of the compression as a function of the distribution of the elements of the matrix.

\section*{Methods}

\subsection*{Matrix compression}

The main goal of our algorithm in the context of this paper was to achieved the best compression ratio possible while maintaining mathematical equivalence with the original data regarding all arithmetic operations. In order to achieve this, we decided we needed to maintain the structure of the matrix, i.e., the ability to acess any element given its row $i$ and column $j$. In order to achieve compression we decided to exploit inefficiencies in the conventional way matrices are allocated in memory.

The compression method is as follows. Consider a matrix $M_{r \times c}$ in which $r$ is the rows number and $c$ the columns number. Each element of this matrix, called $m_{ij}$, is an integer. Since all information is stored as binary code in digital computers, the conventional way to store arrays of integers in most programming languages is to chose a maximum size for the elements (up to 64bits in a typical CPU), and then allocating a block of memory sufficient to store all the elements of $M$:

\begin{equation}\label{eq:01}
  \mathcal{B} = r \times c \times 64
\end{equation}

The number of bits allocated $\mathcal{B}$, is larger than the absolute minimum number of bits required to represent all the elements of $M$. To help illustrate this fact let's consider as an extreme example: a matrix composed exclusively of 0s and 1s. If we set the element to be 64-bit integers we willbe wasting 63 bits per element of the matrix, since the number of bits really needed to store the this matrix, $\mathfrak{b}$ is given by:

\begin{equation}\label{eq:02}
  \mathfrak{b} = l \times c \times 1
\end{equation}

The potential economy of bits $\xi$ can be represented by the difference between $\mathcal{B}$ and $\mathfrak{b}$ (Equations \ref{eq:03} to \ref{eq:06}).

\begin{equation}\label{eq:03}
 \xi = \mathcal{B} - \mathfrak{b}
\end{equation}

\begin{equation} \label{eq:06}
 \xi =  r \times c \times 63
\end{equation}

From equations \ref{eq:01} to \ref{eq:06}, it can be seen that for any matrix whose greatest element requires less than 64 bits (or the fixed type of the matrix) to be represented, memory savings $\xi$ grow linearly with the size of the matrix.

Now we will introduce two simple methods to exploit these inefficiencies in conventional memory allocation, and save a lot of memory in the process. In this section we will use capital roman letters to denote the original uncompressed matrices and the corresponding lower case letter for the compressed version.

\subsubsection*{Method 1: The Supreme Minimum}

The first methodology consists in determining the value of the greatest element of matrix $M$, which coincides with its supremum, $max M == \sup M$  and determine the minimum number of bits required to store it ($b$) .

\begin{equation} \label{eq:07}
 b(\sup M) \approx \begin{cases}
	1, &  \text{ if } \sup M \in \{0,1\} \\ 
	\lfloor \log_2(\sup M)  \rfloor + 1,  & \text{ if } \sup M > 0  
	\end{cases}
\end{equation}

The allocation of memory still happens in the usual way, only that now, in the space required for a single 64 bit integer, we can store an entire $8\times 8$ matrix of $0$s and $1$s.

To make it clearer, lets look at a concrete example: suppose that the greatest value to be stored in a matrix $M$ is $\max M=1023$. Therefore, the number of bits required to represent it is $10$ $(1111111111)$. Let the first 8 elements of $M$ be:

\begin{equation} \label{eq:08}
  M = \begin{bmatrix}
  900 & 1023 & 721 & 256 & 1 & 10 & 700 & 20 & \hdots\\ 
  \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \ddots 
  \end{bmatrix}
\end{equation}

These elements of $M$, in binary, are shown in Table \ref{tab:01}. It is evident that the number of bits required to represent any other element must be $\leq 10$.

\begin{table}[h]
 \centering
 \caption{Some elements of $M$ represented in binary base.}
 \begin{tabular}{cccc} 
  \hline 
  Element & Value  & Binary & Bit length\\
  \hline
  $M_{1,1}$ & 900  & 1110000100 & 10\\
  $M_{1,2}$ & 1023 & 1111111111 & 10\\
  $M_{1,3}$ & 721  & 1011010001 & 10\\
  $M_{1,4}$ & 256  & 100000000  & 9\\
  $M_{1,5}$ & 1    & 1          & 1\\
  $M_{1,6}$ & 10   & 1010       & 4\\
  $M_{1,7}$ & 700  & 1010111100 & 10\\
  $M_{1,8}$ & 20   & 10100      & 5\\
  \hline
 \end{tabular}
 \label{tab:01}
\end{table}

Now consider a matrix $m$, which corresponds to $M$ converted to base $2$ ($M_2$). How do we store such matrix in memory? The conventional way to allocate matrices in memory is to unravel its structure by column (column major, e.g. in Fortran) or by row (row major, e.g. in C) and line them up in a contiguous block of memory. Each element of the matrix is allocated in turn according to the type associated with the matrix.

For our example, since we have determined that each element will require at most 10 bits, we can divide the memory block corresponding to a single 64 bit integer into 6 10-bit chunks which can each hold a single element of $M$. Each of these blocks we will call a bitstring. The remaining 4 bits will be used later. The number of bit strings needed will be $\lfloor \frac{\dim(M) *b} {64} \rfloor +1$.

The final layout of the first 6 elements of $m$ in the first bitstring can be seen in equation \ref{eq:09}.

\begin{equation}\label{eq:09}
 bitstring 1 = 0000\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\underbrace{0100000000}_{256}\underbrace{1011010001}_{721}\underbrace{1111111111}_{1023}\underbrace{1110000100}_{900} 
\end{equation}

Here is a step-by-step description of the allocation process:
\begin{enumerate}
 \item Element $M_{1,1}=900 = m_{1,1} = 1110000100$ is stored in the first 10-bit chunk of the element strip $bitstring[1]$ , which corresponds to bits 0 to 9.
 \item Element $M_{1,2}=1023$ is allocated in the second chunk, from bit 10 to bit 19.
 \item And so forth: Elements $M_{1,i}$ with $i=1,\ldots,6$ re stored on the remaining chunks (eq. \ref{eq:10}).
 \item Element 7 does not fit on the remaining 4 bits of the first bitstring. So it will straddle two bitstrings, i.e., we will start writing it on the first bitstring and finish on the second (equation \ref{eq:11}).
\end{enumerate}

 Please note that bitstrings are written from right to left.

\begin{equation}\label{eq:10}
  \underbrace{\underbrace{0000010100}_{20}\overbrace{101011}^{a}}_{\phi_{1,2}}|\underbrace{\overbrace{1100}^{b}\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\hdots}_{\phi_{1,1}}
\end{equation}

\begin{equation}\label{eq:11}
  \underbrace{0000010100}_{20}\underbrace{\overbrace{101011}^a|\overbrace{1100}^{b}}_{700}\underbrace{0000001010}_{10}\ldots
\end{equation}

In conclusion, the final space required to store matrix $m = M_2$ is much smaller than would be required for the conventional storage of $M$ as a 64-bit integer array. In the results section, we will examine in detail the efficiency of this method on real use cases.

\subsubsection*{Method: Variable Bit Length}

In method 1, there is still waste of space since for elements smaller than the supremum, a number of bits may remain unused.

In this second method, we strive to use absolute minimal number of bits necessary to store each value. However, if we are going to divide the biststrings into variable length chunks, we will also need to reserve some extra bits to represent the size of the chunks, otherwise we cannot recover the elements once they are stored.

To illustrate this method, lets consider again the matrix described in equation \ref{eq:08}. Suppose the largest value stored in it is number 1023. Now instead of assigning one chunk of the bitstring to each element of $m$, we will assign two chunks: the first will store the number of bits required to store the value and the second  will store the respective value. The first region will have a fixed size, in the example cited, 4 bits. These 4 bits are the required space to store the bit-length of $\sup M $, in this case, 10.

To illustrate this second method, lets go through the compression process step-by-step. The largest element of $M$ is 1023. The number of bits required to represent it is 10 which in turn requires 4 bits to be stored. Thus the fixed size chunk is 4 bits long for every element.

\begin{enumerate}
 \item The first element $M_{1,1}=900$ requires 10 bits to store, so we write $10_2$ in the first chunk and $900_2$ in the second (Eq. \ref{eq:13}).
 \item We apply the same logic with the following element, $M_{1,2}=1023$.
 \item Element $M_{1,2}=721$ is also added taking the bitstring to the state illustrated in \ref{eq:14}.
\end{enumerate}

\begin{equation} \label{eq:13}
 bitstring_1 = 00000000000000000000000000000000000000000000000000\underbrace{\underbrace{1110000100}_{\text{value} = 900}\underbrace{1010}_{\text{bitlength=10}}}_{900}
\end{equation}

\begin{equation} \label{eq:14}
 bitstring_1 = 
 0000000000000000000000\underbrace{1011010001}_{721}\underbrace{1010}_{10}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1110000100}_{900}\underbrace{1010}_{10}
\end{equation}

So far the second method is proving more wasteful than the first, but when we add $M_{1,4} =256$  we start to save some space.

\begin{enumerate}
 \item[4.] Element $M_{1,4} =256$ is added.  
 \item [5.] Elements $M_{1,5} =1$ and $M_{1,6} =1$ are added requiring a total of 13 bits instead of 20 with the first method (eq. \ref{eq:16} and \ref{eq:17} ).
 \item [6.] The remaining two elements are added $M_{1,7} =1$ and $M_{1,8} =1$ (eq. \ref{eq:18}).
\end{enumerate}

\begin{equation} \label{eq:16}
 bitstring_1 =
 \underbrace{0100}_{4}\underbrace{1}_{1}\underbrace{0001}_{1}\underbrace{100000000}_{256}\underbrace{1001}_{9}\underbrace{1011010001}_{721}\underbrace{1010}_{10}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1110000100}_{900}\underbrace{1010}_{10}
\end{equation}

\begin{equation} \label{eq:17}
 bitstring_2 = 000000000000000000000000000000000000000000000000000000000000\underbrace{1010}_{10}
\end{equation}

\begin{equation} \label{eq:18}
 bitstring_2 = 00000000000000000000000000000000000\underbrace{10100}_{20}\underbrace{0101}_{5}\underbrace{1010111100}_{700}\underbrace{1010}_{10}\underbrace{1010}_{10}
\end{equation}

 We used a total of 87 bits to store matrix $m$ with the second method instead of 80 bits using method 1. However, method 2 will be the most efficient for most matrices as show below.

 \subsection*{Efficiency of Compression}
 To compare the two compression methods proposed, we can calculate their efficiencies. However since they depend of the data being compressed, we will describe below how to  calculate the efficienty given the data for both methods. Figure

 \subsubsection*{Method 1}
 Let $M_{r \times c}$ be the matrix we wish to compress. we can define the efficiency of the compression as:
 \begin{equation}\label{eq:eff1}
  \eta_1=\frac{bits\, alocated-bits\, used}{bits\,alocated}
 \end{equation}

Equation \ref{eq:eff1} represent the compression ratio achieved by the method. In comparison with a conventional allocation (64-bit integers) we can rewrite equation \ref{eq:eff1} as:

\begin{align}\label{eq:22}
 \eta_1 &= \frac{64 \times rc - b(max M) \times rc}{64 \times rc}\nonumber\\
 &= \frac{64  - b }{64}
\end{align}

As we can see in (\ref{eq:22}), $\eta_1$ does not depend on size of the matrix, only on the bit-length of $\max M$ (equation \ref{eq:19}). According to equation \ref{eq:22} if $b(\max M)=64$, $\eta_1$ will be 0, i.e. no compression is possible. This leads to a very interesting optimization oportunity: divide the entire matrix by the greatest common divisor  of its elements. On the other extreme, if the matrix is composed exclusively of 0s and 1s, maximal compression is achievable, $\eta_1=1$.

\begin{equation} \label{eq:19}
 0000000000000000000000000000000000\underbrace{1111111111}_{1023}\underbrace{\overbrace{000000000}^{\text{Wasted}}1}_{1}\underbrace{\overbrace{000000000}^{\text{Wasted}}0}_{0}
\end{equation}

\begin{figure}[h]
 \centering
 \psfrag{e}{$\xi$}
 \includegraphics[scale=0.5]{fig01}
 \caption{Compression efficiency of method 1 behavior with respect to bit number.}
 \label{fig:01}
\end{figure}

 \subsubsection*{Method 2}

 For the second method, compression depends on the value of each element of the matrix. Thus to compute its efficiency, we shal modify equation \ref{eq:eff1}. We'll start by grouping the elements by bit length, since in this method, bit-lentgh variability greatly affects the compression ratio.

Again consider the matrix $M_{r \times c}$, only the $rc$ elements are now divided into $g$ groups, with $f_i$ numbers of bit-length $b_i = b(m_i)$ each. Let $k = b(b(\max M))$, i.e., the bit length of the bit-length of $\max M$. The efficiency $\eta_2$ is then defined as shown in \ref{eq:24}. Equation \ref{eq:20} can help us visualize it.

\begin{equation} \label{eq:20}
 0000000000000000000000000000000000000000\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1}_{1}\underbrace{0001}_{1}\underbrace{0}_{0}\underbrace{0001}_{1}
\end{equation}

\begin{equation}\label{eq:24}
 \eta_2 = \frac{64 \times rc - \sum_{i=1}^{g} ( b_i + k ) \times f_i }{64 \times rc} 
\end{equation}

We can further simplify equation \ref{eq:24} to get at shorter expression for the compression ratio.

\begin{align} \label{eq:25}
 \eta_2 &= \frac{64 \times rc - \sum_{i=1}^{g} ( b_i \times f_i + k \times f_i )}{64 \times rc} \nonumber \\
  &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  -\sum_{i=1}^{g}  k \times f_i }{64 \times rc}\nonumber \\
  &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  - k \times\sum_{i=1}^{g}  f_i }{64 \times rc}\nonumber \\
 &\sum_{i=1}^{g} f_i = rc \nonumber \\
 \eta_2 &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  - k \times rc }{64 \times rc}\nonumber \\
  &= 1 - \frac{\sum_{i=1}^{g}  b_i \times f_i }{64 \times rc} - \frac{k}{64}
\end{align}

% Results and Discussion can be combined.
\section*{Results}

Consider a matrix $M$ such that $m_{i,j} \in \mathbb{Z}$. In this section we will examine and compare the efficiency of the two compression methods described.

Since the efficiency of both methods depend on the distribution of the bit lengths $b(m_{i,j})$, we will explore a variety of distributions and try to cover the most common cases.

For simplicity we will model the distribution of  as a mixture $X$ of two Beta distributions ,$ B_1 \sim Beta(\alpha_1,\beta_1)$ and $B_2 \sim Beta(\alpha_2,\beta_2)$, whose probability function is shown in equation (eq.\ref{eq:mixture}). Since the Beta distribution is defined only in the interval $[0,1] \subset \mathbb{R}$ , we applied a simple transformation (\ref{eq:transf}) to the mixture in order to map it to the interval of $[1,64] \subset \mathbb{Z}$

\begin{equation}\label{eq:mixture}
 F(x) =  w \, Beta(\alpha_1,\beta_1) + (1-w)\,Beta(\alpha_2,\beta_2)
\end{equation}

The intention of using this mixture was to find a simple way to represent a large variety of bit length distributions. The first two central moments of this mixture are given in \ref{eq:53} and will be used later to summarize our numerical results.

\begin{align}\label{eq:53}
 E(X) &= w E(B_1) + (1-w) E(B_2) \nonumber \\
 &= w \frac{\alpha_1}{\alpha_1+\beta_1} + (1-w)\frac{\alpha_2}{\alpha_2+\beta_2}\nonumber \\
 Var(X) &= w Var(B_1) + (1-w) Var(B_2) + w (1-w) (E(B_1)^2 - E(B_2)^2)
\end{align}

\begin{equation}\label{eq:transf}
   \lfloor 64 \times x \rfloor + 1
\end{equation}

In order to explore the compression efficiency of both methods, we generated samples from the mixture defined above, varying its parameters. From now on, when we mention Beta distribution we will mean the transformed version defined above.

Given any matrix generated by our mixture distribution, we can use equation \ref{eq:22} and \ref{eq:25}, to determine the compression efficiency of methods 1 and 2, respectively.

With $w=0$, ie, our mixture is reduce to a single Beta distribution. In Figure \ref{fig:02},
we show some distributions of bit-lengths for some combinations of $\alpha_1$ and $\beta_1$. From the figure it can be seen that we can generate just about any kind of unimodal distributions in the interval $[1,64]$.

\begin{figure}[ht]
  \centering
  \subfigure[$\alpha=1,\beta_1=1$]{
  \includegraphics[scale=0.25,angle=-90,clip]{fig02}
  }
  \subfigure[$\alpha=1,\beta=32$]{
  \includegraphics[scale=0.25,angle=-90,clip]{fig03}
  }
  \subfigure[$\alpha=32,\beta=1$]{
  \includegraphics[scale=0.25,angle=-90,clip]{fig04}
  }
  \subfigure[$\alpha=64,\beta=64$]{
  \includegraphics[scale=0.25,angle=-90]{fig05}
  }
  \caption{Histograms constructed from samples with 10,000 elements, generated from a Beta distribution. Below each histogram is possible to verify the parameters used.}
  \label{fig:02}
\end{figure}

As we are sampling the from set of all possible distributions of bit-length, represented byt the mixture of betas presented above, in order to make our results more general, we will base our analysis on the expected bit-length of a sample, since the efficiency of both methods  depend on it. So, from equations \ref{eq:22} and \ref{eq:25}, the expected efficiencies become:

\begin{equation}\label{eq:56}
 E(\eta_1) = 1 - \frac{k}{64}
\end{equation}

\begin{equation}\label{eq:57}
 E(\eta_2) = 1 - \frac{E(b)}{64} - \frac{k}{64}
\end{equation}

where k in the expression of $\eta_2$ is set to 7 (the bit-length required to represent the largest possible bit-length: 64). In \ref{eq:56}, $k$ is the bit-length of the greatest element, or in the worst case, 64.

We will use the difference $D=E(\eta_1)-E(\eta_2)$ to compare the efficiency of the two methods. Thus a positive $D$ will favor method $1$ while a negative $D$ favors method $2$.

For the numeric experiments, we generate $3$ samples of size $10000$, from which we calculate the expected compression efficiency, using equations \ref{eq:56} and \ref{eq:57}. the values presented in tables and figures below are averages of the efficiencies of the three samples.

In Figure \ref{fig:03}, we can see the distribution of efficiencies and their difference for a sample generated from a single Beta distribution of bit-lengths. We can see that both methods can achieve efficiencies greater than 80\% for matrices with very small numbers. We also see in figure \ref{fig:03} that method 2 is more efficient in the majority of cases.

\begin{figure}[h]
  \centering
%   \subfigure[$\alpha=1,\beta_1=1$]{
%   \includegraphics[scale=0.28,angle=-90,clip]{fig02}
%   }
%   \subfigure[$\alpha=1,\beta=32$]{
%   \includegraphics[scale=0.28,angle=-90,clip]{fig03}
%   }
%   \subfigure[$\alpha=32,\beta=1$]{
%   \includegraphics[scale=0.28,angle=-90,clip]{fig04}
%   }
%   \subfigure[$\alpha=64,\beta=64$]{
%   \includegraphics[scale=0.28,angle=-90]{fig05}
%   }
  \includegraphics[scale=0.7]{simples.eps}
  \caption{Comparing compression efficiency of methods 1 and 2. Color scale in (a), (b) and (c) represent average bit-length. In (a) we can see the difference $D = \eta_1 - \eta_2$. It can be seen that for most combination of $\alpha$ and $\beta$, $D<0$, meaning the second method is more efficent to compress a sample of numbers with bit-lengths coming from a $Beta(\alpha,\beta)$ distribution. However, there is a small region in parameter space, which is shown in white on (d), where the first method is more efficient. This region corresponds to the dots in red in (a), where the average bit-length is higher. In panels (b) and (c), we can see the efficiencies of methods 1 and 2, respectively.}
  \label{fig:03}
\end{figure}

Now let's analyze matrices whose elements have bit lengths sampled from a mixture of beta distributions, $B_1$ and $B_2$. For this mixture $B$ we set $w = 0.5$, i.e., the mixture will have half the numbers generated from the distribution $B_1 \sim Beta(\alpha_1,\beta_1)$ and the other part distribution $B_2 \sim Beta(\alpha_2,\beta_2)$. The expected value for this mixture is shown in equation \ref{eq:58}.

\begin{equation}\label{eq:58}
 E(B) = 0.5 E(B_1) + 0.5 E(B_2)
\end{equation}

We must also redefine the calculations of efficiency for the mixture samples. Equation \ref{eq:59} shows the expected efficiency for method 1. Now, instead of having the efficiency being a function of greatest bit-length in the sample (denoted as  $k$ in \ref{eq:56} and \ref{eq:57}), it will be a function of $max\{E(B_1),E(B_2)\}$. We do the same for method​ 2 (equation \ref{eq:60})​.

\begin{equation}\label{eq:59}
 E(\eta_1) = 1 - \frac{max\{E(B_1),E(B_2)\}}{64}
\end{equation}

\begin{equation}\label{eq:60}
 E(\eta_2) = 1 - 0.5\frac{E(B_1)}{64} - 0.5\frac{E(B_2)}{64} - \frac{max\{E(B_1),E(B_2)\}}{64}
\end{equation}

As before, we generate 3 samples of size $10000$ for each parameterization, calculate the average efficiencies (equations \ref{eq:59} and \ref{eq:60}) and their diference $D$.

Before we move on to the efficiency results and analyses, let's first inspect our samples from the mixture of transformed Beta distributions. Figures \ref{fig:02A} and \ref{fig:02B}, show a few parameterizations and their resulting sample distributions. It is important to note that from the mixture we can now generate bimodal distributions as well as the unimodal types tested before. Since we are making statements about efficiency as a function of the expected bit-length, it is important to verify if these statements hold for bimodal  distributions as well.

After sampling uniformly ($[1,5,9,\ldots,64]$, $n=65536$) the parameter space and comparing efficiencies, we summarized the results on table \ref{tab:06}. In it we see how many parameterizations (from our sample) favor each method. We can also look at the distribution of efficiencies on our samples for each method (figure \ref{fig:1415}), which clearly demonstrate the greater expected efficiency of method 2 (figure \ref{fig:15}).

\begin{figure}[h]
  \centering
  \subfigure[$\alpha_1=1,\beta_1=1,\alpha_2=1,\beta_2=1$]{
  \includegraphics[scale=0.28,angle=-90,clip]{fig06}
  }
  \subfigure[$\alpha_1=1,\beta_1=32,\alpha_2=32,\beta_2=1$]{
  \includegraphics[scale=0.28,angle=-90,clip]{fig07}
  }
  \subfigure[$\alpha_1=32,\beta_1=32,\alpha_2=32,\beta_2=32$]{
  \includegraphics[scale=0.28,angle=-90,clip]{fig08}
  }
  \subfigure[$\alpha_1=64,\beta_1=32,\alpha_2=32,\beta_2=64$]{
  \includegraphics[scale=0.28,angle=-90]{fig09}
  }
  \caption{Histograms constructed from samples with 10,000 elements, generated from the combination of two distributions Betas. Below each histogram is possible to verify the parameters used..}
  \label{fig:02A}
\end{figure}

\begin{figure}[h]
  \centering
  \subfigure[$\alpha_1=64,\beta_1=48,\alpha_2=1,\beta_2=48$]{
  \includegraphics[scale=0.28,angle=-90]{fig10}
  }
  \subfigure[$\alpha_1=16,\beta_1=46,\alpha_2=49,\beta_2=64$]{
  \includegraphics[scale=0.28,angle=-90]{fig11}
  }
  \subfigure[$\alpha_1=16,\beta_1=16,\alpha_2=16,\beta_2=49$]{
  \includegraphics[scale=0.28,angle=-90]{fig12}
  }
  \subfigure[$\alpha_1=1,\beta_1=16,\alpha_2=1,\beta_2=49$]{
  \includegraphics[scale=0.28,angle=-90]{fig13}
  }
  \caption{Histograms constructed from samples with 10,000 elements, generated from the combination of two distributions Betas. Below each histogram is possible to verify the parameters used.}
  \label{fig:02B}
\end{figure}

\begin{table}[h]
 \centering
 \caption{Efficiency comparison of methods I and II for parameters covering uniformly the support of $B$. Column $n$ shows the number of parameter combinations with which each method has superior compression.}
 \begin{tabular}{ccc}
  \hline 
  Methods  & n   & Percentage \\
  \hline
  I	   & 592	& 0.9034\% \\
  II	   & 64944	& 99.0966\% \\
  \hline
  Total    & 65536	& 100\% \\
  \hline
 \end{tabular}
 \label{tab:06}
\end{table}

\begin{figure}[h]
  \centering
  \subfigure[Efficiency histogram of the method I]{
  \includegraphics[scale=0.4,angle=-90]{fig14}
  \label{fig:14}
  }
  \subfigure[Efficiency histogram of the method II]{
  \includegraphics[scale=0.4,angle=-90]{fig15}
  \label{fig:15}
  }
  \caption{Efficiency histograms of the methods I and II. Note that the method II has a greater average efficiency than method I.}
  \label{fig:1415}
\end{figure}

As we have shown, method is more effective compressing most integer datasets up to 64bits in size. This is due to its ability to exploit the variance in the data set and reduce the waste of bits in the representation of some numbers. In specific cases where the variance in the data null or too small, method I will be more efficient. As a matter of fact, for matrices where all elements have the same bit-length, method I will always be better, regardless of bit-length (figures \ref{fig18} and \ref{fig19}), The only exception if for bit-length 64 where neither method is able to compress the data.

\begin{figure}[h]
  \centering
  \subfigure[Efficiencies of the methods I e II, for constant bit-length matrices.]{\includegraphics[scale=0.3]{fig18}\label{fig18}}
  \subfigure[$\eta_1-\eta_2$ for matrices of constant bit-length.]{\includegraphics[scale=0.3]{fig19}\label{fig19}}
  \caption{Compression efficiency of the methods I and II  for matrices of constant bit-length.}
  \label{fig:1819}
\end{figure}

The two methods presented allow the storage and information manipulation in the compressed form, which provides an data managing large amounts optimization. The next step in the work development is the implementation of the proposed methods, as well as tests to be performed on data large volumes.

\section*{Discussion}

\subsection*{Which method to use?}

To determine the best compression method to apply, it's necessary to inspect the distribution of bit-lengths of matrix elements. When matrix elements are small or have nearly-constant bit-length, Method I is best, otherwise, the second method should be chosen.

As an example, let $M_{r\times c}$ be a integer matrix such that the half of its elements have bit-length $1$ and the other half $64$. Recalling equation \ref{eq:25}, now we have two groups of elements (by bit-length), $b_1=1 $, $b_2=64$ and $f_i = \frac{rc}{2}$ for $i = 1 \text{ and } 2$. As the greatest bit-length is $64$, then $k=7$. Compression efficiency $\eta_2$ can be calculated using Equation \ref{eq:25}. After plugging in our numbers, we obtain a compression of $38.29\%$.

\begin{equation*} %\label{eq:32}
 \eta_2 = 1 - \frac{\sum_{i=1}^{2}  b_i \times f_i }{64 \times rc} - \frac{7}{64} 
\end{equation*}

\begin{equation*} %\label{eq:33}
 \eta_2 = 1 - \frac{  1 \times \frac{rc}{2} + 64 \times \frac{rc}{2} }{64 \times rc} - \frac{7}{64} 
\end{equation*}

\begin{equation*} %\label{eq:34}
 \eta_2 = 1 - \frac{  32.5  \times rc }{64 \times rc} - \frac{7}{64} 
\end{equation*}

\begin{equation*} %\label{eq:35}
 \eta_2 \approx 1 - 0.5078 - 0.1093
\end{equation*}

\begin{equation*} %\label{eq:36}
 \eta_2 \approx 38.29\%
\end{equation*}

The efficiency of method II is influenced by the relative size of the bit-length groups. In this first example we considered only two groups, each comprised of half the matrix elements. Let's now vary the relative frequency of the groups, $\frac{f_i}{rc}$, while sticking to two groups. Let's also assume that $\frac{f_i}{rc}$ is a good approximation to the probability of a given bit-length in a matrix, which we will denote by $p_i$.

With this definition we can rewrite the equation \ref{eq:25}, which becomes \ref{eq:37}. In Equation \ref{eq:37}, the $\frac{f_i}{rc}$ is replaced by $p_i$, representing the probability of elements from group $i$ in matrix M.

\begin{equation}\label{eq:37}
 \eta_2 = 1 - \frac{\sum_{i=1}^{g}  b_i \times p_i }{64} - \frac{k}{64} 
\end{equation}

\noindent with

\begin{equation}\label{eq:38}
 p_i = \frac{f_i}{rc}
\end{equation}

With the equation \ref{eq:37} can analyze the influence of bit-length probability in compression efficiency. In this example, $p_1$ and $p_2$ represent the probability of elements of bit-lengths 1 and 64, respectively. Thus, efficiency is defined in Equation \ref{eq:40}.

\begin{equation}\label{eq:40}
 \eta_2 = 1 - \frac{1 \times p_1  + 64 \times p_2}{64} - \frac{7}{64} 
\end{equation}

Now, we can determine which probabilities give us the best and worst compression levels. When $\eta_2=1$, then the efficiency is maximal and if $\eta_2=0$, a efficiency is minimal. To calculate the values ​​of $p_1$ and $p_2$ for both extreme values of $\eta_2$,  we must solve the linear systems shown in equations \ref{eq:41} and \ref{eq:42}. The first equation on both systems come from the law of total probability. The second comes from \ref{eq:40} after setting $\eta_2$ to $1$ and $0$, respectively.

\begin{equation}\label{eq:41}
  \eta_2 = 1 :\left
  \{\begin{matrix}
    p_1 + p_2 = 1\\ 
    p_1+64p_2 = 7
  \end{matrix}
  \right.
\end{equation}

Solving the system above, we find that  when $p_1=0.9047$ and $p_2=0.0953$, efficiency is maximal, and in this particular case is equal to 87.5\%.

\begin{equation}\label{eq:42}
  \eta_2 = 0 :\left
  \{\begin{matrix}
    p_1 + p_2 = 1\\ 
    p_1+64p_2 = 57
  \end{matrix}
  \right.
\end{equation}

Thus, when $p_1=0.1111$ and $p_2=0.8889$ the efficiency is minimal for the second method. For other combinations see table \ref{tab:02}. Looking at this table, one can see two negative efficiencies, when ($p_1$,$p_2$) assume the values ​​(0,1) and (0.1,0.9). This correspond th cases when the method increases the memory requirements instead of decreasing it.

\begin{table}[h]
 \centering
 \caption{Combinations $p_1$ and $p_2$ to calculate the efficiency.}
 \begin{tabular}{ccc}
  \hline 
  $p_1$  & $p_2$ & Efficiency \\
  \hline
  0.0	&1.0    &-0.109 \\
  0.1	&0.9	&-0.010 \\
  0.2	&0.8	&0.087 \\
  0.3	&0.7	&0.185 \\
  0.4	&0.6	&0.284 \\
  0.5	&0.5	&0.382 \\
  0.6	&0.4	&0.481 \\
  0.7	&0.3	&0.579 \\ 
  0.8	&0.2	&0.678 \\
  0.9	&0.1	&0.776 \\ 
  1.0	&0.0	&0.875 \\
  \hline
 \end{tabular}
 \label{tab:02}
\end{table}

So far, we have examined only two groups (hence two probabilities) of bit-length for the sake of simplicity. Before we generalize to probability distributions let's take a quick look at the efficiencies for more groups, with uniform probability:

\begin{itemize}
  \item 3 number groups that require numbers 1, 32 and 64 bits, efficiency $\zeta=0.3854$,
  \item 5 number groups that require numbers 1, 16, 32, 48 and 64 bits, efficiency $\zeta=0.3875$,
  \item 8 number groups that require numbers 1, 8, 16, 24, 32, 40, 48, 56 and 64 bits, efficiency $\zeta=0.3888$
\end{itemize}

When the distribution of the group probabilities is uniform, i.e. the groups have approximately the same size, efficiency is basically the same, regardless of the number of groups.

Now we can leverage the notion of bit-length probabilities, and study efficiency when bit-lengths follow some  commonly used discrete probability distributions: Discrete Uniform, Binomial and Poisson. For all the experiments, we assume $k=7$, that is, the maximum possible bit-length is 64 bits. Thus, efficiency obtained will not be the best possible, since for that we would need assume small values of $k$ (equation \ref{eq:37}).

\subsection*{Discrete Uniform}
The discrete uniform distribution has two parameters, $a$ and $b$ representing the minimum and maximum value of the distribution. For our case we will use $U(a=1,b=64)$, which means bit-lengths may take values in the set $\{1, 2, 3, $\ldots$, 64\}$ with equal probability, i.e., $\frac{1}{64}$.

For the Poisson distribution, we use Poisson($\lambda$) with $\lambda$ being the expected bit-length in the sample.

To calculate the efficiency, we generated a sample with 100 ($M_{10 \times 10}$), 10,000 ($M_{100 \times 100}$) and 1,000,000 ($M_{1,000 \times 1.000}$) numbers . The average efficiency (table \ref{tab:03}) is calculated from a 1000 replicates of each sample size.

\begin{table}[h]
  \centering
  \caption{Compression efficiency of method II of samples with bit-lengths comming from a Discrete Uniform distribution $U(a=1,b=64)$. Average efficiency were calculated over a 1000 replicates.}
 \begin{tabular}{cc}
    \hline
    Sample size & Avg. efficiency \\
    \hline
     100	& 0.3760938 \\
     10000 	& 0.3795953 \\
     1000000 	& 0.3826869 \\
    \hline
 \end{tabular}
 \label{tab:03}
\end{table}

\subsection*{Binomial Distribution}

For the binomial distribution, we will use $Bin(n, p=0.5)$, with the number of trials $n$ representing the greatest possible bit-length in the matrix, and $p=0.5$ giving us an expected bit-length of 32.

For these experiments, the parameter n represents the maximum bit-length of matrix elements and takes values in $\{1, 8, 16, 32, 64\}$. In this case, we evaluate the efficiency as a function of the parameter $n$, and sample size. Even though efficiency does not depend on matrix size, we tried different sizes to test the stability of the compression algorithm. Results are shown in Table \ref{tab:04}. As expected smaller expected bit lengths lead to higher compression efficiencies. However, for the case in which the elements require 64 bits, the efficiency is negative. [REMOVE AFTER FIXING TABLE]This result indicates that using the second method, when elements require 64 bit this becomes impractical, because it would be require more memory to allocate the compressed matrix.[REMOVE]

\begin{table}[h]
  \centering
  \caption{Compression efficiency with bit-lengths distributed according to a binomial distribution $B(n,0.5)$. Parameter $n \in \{1, 8, 16, 32, 64\}$ represents the maximum bit-length. Since $p=0.5$ the expected bit-length is $n/2$ (first column).[REMOVE LAST LINE SINCE WE CANNOT HAVE AN EXPECTED VALUE OF 64 and check for correct values]} 
 \begin{tabular}{cccc}
    \hline
			& 		&Efficiency         & \\
    \hline
			&  		&Sample size & \\
    Expected bit-length $(n/2)$		& 100		& 1,000		    & 1,000,000 \\
    \hline
     1 			&  0.9760937	& 0.9764797 	    & 0.9765566 \\
     8			&  0.8109375	& 0.8128453 	    & 0.8125079 \\
     16			&  0.626875	& 0.6251859 	    & 0.6249333 \\
     32			&  0.2476562	& 0.2504953 	    & 0.2499712 \\
     64			&  -0.5004688	& -0.4999828 	    & -0.5000366 \\
    \hline
 \end{tabular}
 \label{tab:04}
\end{table}
\subsection*{Poisson Distribution}
With bit-length derived from a Poisson($\lambda$), the parameter $\lambda$ corresponds to the expected bit-length. Values ​​greater than 64 generated in this simulation were considered as being equal to 64. The results for this simulation can be seen in Table \ref{tab:05}. Note that increasing the bit number to represent numbers increases, there is a loss of efficiency in the compression process. Another interesting fact is the emergence of a negative efficiency. In this case, for values ​​generated with a $\lambda=64$, the second method requires more memory to allocate the matrix formed by elements generated from this distribution.

\begin{table}[h]
  \centering
  \caption{Compression efficiency with bit-lengths distributed according to a Poisson distribution ($\lambda$), where $\lambda$ represents the expected bit-length. [REMOVE LAST LINE]}
 \begin{tabular}{cccc}
    \hline
			& 	&Efficiency         & \\
    \hline
			&  	&Size sample& \\
    Expected bit-length ($\lambda$)		& 100	& 1,000		    & 1,000,000 \\
    \hline
     1	& 0.8746875 & 0.8749859   & 0.8749834 \\
     8 	& 0.7657812 & 0.7656016   & 0.7656138 \\
     16	& 0.6300000 & 0.6400594   & 0.6405952 \\
     32	& 0.3851562 & 0.3903531   & 0.3903477 \\
     64	& -0.0528125& -0.05984375 & -0.05953145 \\
    \hline
 \end{tabular}
 \label{tab:05}
\end{table}

With bit-length as a random variable, we can generalize the results about the compression efficiency as a function of expected bit-length

Let the random variable $B \sim U(a=1,b=64)$ represent the  bit-length of the elements of matrix $M$. Then $E(b_i) = \sum_{i} b_i \times p(b_i) = \frac{a+b}{2}$.  Applying this result to the expected compression efficiency of method II (equation \ref{eq:37}), we have

\begin{equation}\label{eq:44}
 E(\eta_2) = 1 - \frac{E(B)}{64} - \frac{k}{64} 
\end{equation}

assuming all bit-lengths are possible, i.e., $a=1$ and $b=64$, and hence $k=7$, we can calculate $\eta_2$:

\begin{equation}\label{eq:45}
 E(\eta_2) = 1 - \frac{\frac{1+64}{2}}{64} - \frac{7}{64} \approx 38.28\% 
\end{equation}

This result agrees with the numerical estimates presented in table \ref{tab:03}. We can do the same for the case where bit-length $(B)$ is a random variable with Binomial and Poisson distributions. In the case of Binomial, $B \sim B(n=64,p=0.5)$, $E(b_i) = \sum_{i} b_i \times p(b_i) = n \times p$ and the eficiency becomes (with $k=7$):

\begin{align}\label{eq:47}
 E(\eta_2) &= 1 - \frac{64 \times p }{64} - \frac{k}{64}  \\
  &= 1 - \frac{64 \times 0.5 }{64} - \frac{7}{64} \approx 39.05\% \nonumber
\end{align}

Which again agrees with estimates in Table \ref{tab:04}.

Lastly, for $B$ distributed as Poisson$(\lambda=32)$ case,$E(b_i) = \lambda$, $k=7$, and the efficiency becomes:

\begin{align}\label{eq:49}
 E(\eta_2) &= 1 - \frac{\lambda}{64} - \frac{k}{64} \\
 &= 1 - \frac{32}{64} - \frac{7}{64} \approx 39.06\%
\end{align}

This result is again in accordance to Table \ref{tab:05}. These results show that a good compression is guaranteed when bit-lengths are distributed according to the tested distributions regardless of sample size.

[devemos incluir um estudo sobre a variância?]

In this paper we have focused in the compression of matrix data, since this is one of the most important application the authors foresee. However, the compression methodology presented can be applied to any numerical data structure, with gains to performance and memory footprint[citar tese Crysttian e possíveis artigos derivados]. Further discussions about doing computation with such compressed data-structures will be the subject of another manuscript (in preparation) in which we will present details about the implementation of the compression algorithm, and benchmarks on classical linear algebra tasks such as those in Linpack

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis"

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}
We would like to thank Paulo Cezar Carvalho and Moacyr silva for fruitful discussions and key ideas which helped to improve this paper.

%\section*{References}
% The bibtex filename
\bibliography{plos.bib}

\section*{Figure Legends}
%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

\section*{Tables}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

\end{document}

### 2012-12-27 manuscript.tex
adicionei um parágrafo à discussão.

@@
 However, the compression methodology presented can be applied to any numerical data structure, with gains to performance and memory footprint[citar tese Crysttian e possíveis artigos derivados].
-Further discussions about doing computation with such compressed data-structures will be the subject of another manuscript (in preparation) in which we will present details about the implementation of the compression algorithm, and benchmarks on classical linear algebra tasks such as those in Linpack
+
+Further discussions about doing computation with such compressed data-structures will be the subject of another manuscript (in preparation) in which we will present details about the implementation of the compression algorithm, and benchmarks on classical linear algebra tasks such as those in Linpack[ref].
+
+Representation of floating point numbers is also possible within the proposed compression framework, but at the expense of precision in their representation.
+Although this may sound like a limitation, when we take into consideration that most experimental data have fewer ``significant'' digits than the maximal precision available in modern computers, fairly good compression may still be achievable for floats.
 

### 2012-12-29 manuscript.tex
quase terminando a segunda revisão

@@
 
-Data compression is traditionally used to reduce storage resouces usage and/or transmission costs\cite{salomon}.
+Data compression is traditionally used to reduce storage resources usage and/or transmission costs\cite{salomon}.
 Compression techniques can be classified into lossy and lossless.
-Goos examples of lossy data compression are sound (mp3), image (JPEG) and movie (mpeg) compression.
-In this paper we are concerned exclusively with lossless compression.
-In lossless compression, no information is lost, during the compression process, and they are the method of choice to compress numbers and text.
+Examples of lossy data compression are MP3 (audio), JPEG (image) and MPEG (video).
+In this paper we discuss the use of lossless compression for numerical data structures such as numerical arrays to achieve compression without losing the mathematical properties of the original data.
 
-Any kind of compression incurs in some computational cost.
-such costs often have to be paid twice since the data needs to be decompressed to be used for its original purpose.
-Sometimes computational costs are irrelevant, but the need to decompress for usage, can signify that savings in storage costs are applicable to storage situation since we still require the uncompressed space allocation for usage.
+Lossless compression methods usually exploit redundancies present in the data in order to find a shorter form of describing the same information content.
+For example, a dictionary-based compression, only stores the positions in which a given word occurs in a document, thus saving the space required to store all its repetitions\cite{salomon2}.
 
-Lossless compression methods usually try to exploit redundancies present in the data in order to find a shorter form for of describe the same information content, for example a dictionary based compression, can only store the positions in which a given word occurs in a document, thus saving the space required to store all its repeatitions\cite{salomon2}.
+Any kind of compression incurs some computational cost.
+Such costs often have to be paid twice since the data needs to be decompressed to be used for its original purpose.
+Sometimes computational costs are irrelevant, but the need to decompress for usage, can signify that the space saved with compression must be available when data is decompressed for usage, thus partially negating the advantages of compression.
 
-Most if not all existing lossless compression methods were develop under the following paradigm: produce $\rightarrow$ compress $\rightarrow$ store $\rightarrow$ uncompress $\rightarrow$ use.
-However in the brave new world of big data, the need to analyze data immediately after its coming into existence became the norm.
+Most if not all existing lossless compression methods were developed under the following usage paradigm: \textit{produce} $\rightarrow$ \textit{compress} $\rightarrow$ \textit{store} $\rightarrow$ \textit{uncompress} $\rightarrow$ \textit{use}.
+The focus of the present work is to allow a slightly different usage:  \textit{produce} $\rightarrow$ \textit{compress} $\rightarrow$ \textit{perform mathematical manipulations} $\rightarrow$ \textit{decompress} (only for human reading).
+
+With the growth of data volumes and analytical demands, creative solutions are needed to efficiently store as well as consume it on demand.
+This issue is present in many areas of application, ranging from business to science\cite{lynch}, and is being called the \textit{Big Data} phenomenon.
+In the world of Big Data, the need to analyze data immediately after its coming into existence became the norm.
 And this analysis must take place, efficiently, within the confines of (RAM) memory.
-Given a sufficiently big flow of data, compression an decompression costs may become prohibitive as it will slow down data transmission.
-Moreover, the need to decompress before analyzing removes any potential gains in memory usage.
+This kind of analyses are what is now known as streaming data analysis\cite{gaber2005mining}.
+Given a sufficiently dense stream of data, compression an decompression costs may become prohibitive.
+So having a way to compress data and keeping it compressed for the entire course of the analytical pipeline, is very desirable.
 
-With the growth of data and analytical demands, creative solutions are need to not only efficiently store large data as consume it on demand.
-This phenomenon is present in many areas of aplication, ranging from business to science\cite{lynch}, and is being called the big data phenomenon.
+This paper will focus solely on numerical data which for the purpose of the applications is organized as   matrices.
+This is a most common data structure found in computational data analysis environments.
+The matrices compressed according to the methodologies proposed here should be able to undergo the same mathematical operations as the original uncompressed matrices.
+e.g.
+linear algebra manipulations.
+This way, the cost of compression is reduced to a single event of compression and no need of decompression except when displaying the results for human reading.
+The idea of operating with compressed arrays is relatively new\cite{yemliha2007compiler}, and has yet to find mainstream applications to the field of numerical computations.
+One application which employs a form of compression is the sparse matrix linear algebra algorithms\cite{dodson1991sparse}, in this case there is no alteration in the standard encoding of the data, but only the non-zero elements of the matrices are stored and operated upon.
 
-One data type which is very common across areas of application is numerical data.
-Most common statistical and machine learning algorithms operate over large matrices, frequently making use of linear algebra methods.
-So it would be very desirable to have a compression method in which the compressed object retained the mathematical properties as the original, i.e.
-could be subject to the same mathematical operations.
-The idea of operating with compressed array is not new\cite{yemliha2007compiler}, but has yet to applied consistently to the field of numerical computations.
-One application which employs a form compression is the sparse matrix linear algebra algorithms\cite{dodson1991sparse}, in this case there is no encoding of the data, but only the non-zero elements of the matrices are stored and operated upon.
+Larger than RAM data structures can render traditional analytical algorithms impracticable.
+Currently, the technique most commonly used when dealing with large matrices for numerical computations, is memory mapping\cite{van2011numpy,big}.
+In memory mapping the matrix is allocated in a virtual contiguous address space which extends from memory into disk.
+Thus, larger than memory data structures can be manipulated as if they were in memory.
+This technique has a big performance penalty due to lower access speeds of disk when compared to RAM.
 
-Currently, the technique most commonly used when dealing with large matrices for numerical computations, is memory mapping\cite{van2011numpy,big}.
-In memory mapping the matrix is allocated in a virtual contiguous address space which overflows from memory into disk, this way larger than memory datastructures can be manipulated as if they were in memory.
-This technique has a big performance penalty due to much lower access speeds of disk when compared to RAM.
-
-In this paper we present two methods that allow the lossless compression of (numerical) arrays, which do not require decompression to be manipulated mathematically.
-The method involves the encoding of the numbers as strings of bits of variable length.
-The methods resemble the arithmetic coding\cite{bodden2007arithmetic} algorithm, but is cheaper to compute and does not require decompression.
-We describe the process of compression and decompression, and how to operate with the arrays in compressed format.
-We discuss the efficiency of the compression as a function of the distribution of the elements of the matrix.
+In this paper we present two methods for the lossless compression of (numerical) arrays.
+The methods involve the encoding of the numbers as strings of bits of variable length.
+The methods resemble the arithmetic coding\cite{bodden2007arithmetic} algorithm, but is cheaper to compute.
+We describe the process of compression and decompression, and study their efficiency under different applications.
+We also discuss the efficiency of the compression as a function of the distribution of the elements of the matrix.
 
@@
 
-The main goal of our algorithm in the context of this paper was to achieved the best compression ratio possible while maintaining mathematical equivalence with the original data regarding all arithmetic operations.
-In order to achieve this, we decided we needed to maintain the structure of the matrix, i.e., the ability to acess any element given its row $i$ and column $j$.
+To maintain mathematical equivalence with the original data for any arithmetic operations, we need to  maintain the structure of the matrix, i.e., the ability to acess any element given its row $i$ and column $j$ and also the numeric nature of its elements.
 In order to achieve compression we decided to exploit inefficiencies in the conventional way matrices are allocated in memory.
+The exemples in this paper will be restricted to matrices with positive integer elements.
 
 The compression method is as follows.
-Consider a matrix $M_{r \times c}$ in which $r$ is the rows number and $c$ the columns number.
-Each element of this matrix, called $m_{ij}$, is an integer.
-Since all information is stored as binary code in digital computers, the conventional way to store arrays of integers in most programming languages is to chose a maximum size for the elements (up to 64bits in a typical CPU), and then allocating a block of memory sufficient to store all the elements of $M$:
+Let $M_{r \times c}$ be a matrix, in which $r$ is the number of rows and $c$ the number of columns.
+Each element of this matrix, called $m_{ij}$, is a positive integer.
+In digital computers, all information is stored as binary code (base 2 numbers).
+However the conventional way to store arrays of integers is on a sequence of memory blocks of fixed size (power of 2 number of bits), one for each element.
+The maximum size of a block is equal to the word size of the processor, which for most current CPUs is 64 bits.
+Some special number such as complex number may be encoded as two blocks instead of one.
+The size of the chunk of memory allocated to each number will determine their maximum  size (for integers) or their precision (for floating-point numbes).
+So for matrix $M$, the total memory allocated, assuming chunks of 64 bits, is given by $\mathcal{B} = r \times c \times 64$.
 
-\begin{equation}\label{eq:01}
-  \mathcal{B} = r \times c \times 64
-\end{equation}
+The number of bits allocated $\mathcal{B}$, is larger than the absolute minimum number of bits required to represent all the elements of $M$, since smaller integers, when converted to base 2, require less digits.
+From now on, when the numerical base will be explicitly notated when necessary to avoid confusion between binary and decimal integers.
 
-The number of bits allocated $\mathcal{B}$, is larger than the absolute minimum number of bits required to represent all the elements of $M$.
-To help illustrate this fact let's consider as an extreme example: a matrix composed exclusively of 0s and 1s.
-If we set the element to be 64-bit integers we willbe wasting 63 bits per element of the matrix, since the number of bits really needed to store the this matrix, $\mathfrak{b}$ is given by:
+Let's consider an extreme example: a matrix composed exclusively of 0s and 1s (base 10).
+If the matrix type is set to 64-bit integers, 63 bits will be wasted per element of the matrix, since the minimum number of bits needed to store such a matrix is $\mathfrak{b} = r \times c \times 1$.
+The potential economy of bits $\xi$ can be represented by $\xi = \mathcal{B} - \mathfrak{b} =  r \times c \times 63$.
 
-\begin{equation}\label{eq:02}
-  \mathfrak{b} = l \times c \times 1
-\end{equation}
+So it is evident that for any matrix whose greatest element requires less than 64 bits (or the fixed type of the matrix) to be represented, potential memory savings will grow linearly with the size of the matrix.
 
-The potential economy of bits $\xi$ can be represented by the difference between $\mathcal{B}$ and $\mathfrak{b}$ (Equations \ref{eq:03} to \ref{eq:06}).
+\subsubsection*{Method 1: The Supreme Minimum (SM)}
 
-\begin{equation}\label{eq:03}
- \xi = \mathcal{B} - \mathfrak{b}
-\end{equation}
-
-\begin{equation} \label{eq:06}
- \xi =  r \times c \times 63
-\end{equation}
-
-From equations \ref{eq:01} to \ref{eq:06}, it can be seen that for any matrix whose greatest element requires less than 64 bits (or the fixed type of the matrix) to be represented, memory savings $\xi$ grow linearly with the size of the matrix.
-
-Now we will introduce two simple methods to exploit these inefficiencies in conventional memory allocation, and save a lot of memory in the process.
-In this section we will use capital roman letters to denote the original uncompressed matrices and the corresponding lower case letter for the compressed version.
-
-\subsubsection*{Method 1: The Supreme Minimum}
-
-The first methodology consists in determining the value of the greatest element of matrix $M$, which coincides with its supremum, $max M == \sup M$  and determine the minimum number of bits required to store it ($b$) .
+The SM method consists in determining the value of the greatest element of matrix $M$, which coincides with its supremum, $max M == \sup M$  and determine the minimum number of bits, $b(\sup M$), required to store it.
+We will use capital roman letters to denote uncompressed matrices and the corresponding lower case letter for the compressed version.
 
@@
 	1, &  \text{ if } \sup M \in \{0,1\} \\ 
-	\lfloor \log_2(\sup M)  \rfloor + 1,  & \text{ if } \sup M > 0  
+	\lfloor \log_2(\sup M)  \rfloor + 1,  & \text{ if } \sup M > 1  
 	\end{cases}
@@
 
-The allocation of memory still happens in the usual way, only that now, in the space required for a single 64 bit integer, we can store an entire $8\times 8$ matrix of $0$s and $1$s.
+The allocation of memory still happens in the usual way, i.e.
+in fixed size 64-bit chunks, only that now, in the space required for a single 64 bit integer, we can store for example, an entire $8\times 8$ matrix of $0_{10}$ and $1_{10}$.
 
-To make it clearer, lets look at a concrete example: suppose that the greatest value to be stored in a matrix $M$ is $\max M=1023$.
+Let's look at a concrete example: suppose that the greatest value to be stored in a matrix $M$ is $\max M=1023$.
 Therefore, the number of bits required to represent it is $10$ $(1111111111)$.
@@
 
-These elements of $M$, in binary, are shown in Table \ref{tab:01}. It is evident that the number of bits required to represent any other element must be $\leq 10$.
+These elements of $M$, in binary, are shown in Table \ref{tab:01}.
+It is evident that the number of bits required to represent any other element must be $\leq 10$.
+From now on the minimum number of bits required to represent a base 10 integer will be refered to as its bit-length.
 
@@
 
-Now consider a matrix $m$, which corresponds to $M$ converted to base $2$ ($M_2$).
-How do we store such matrix in memory?
-The conventional way to allocate matrices in memory is to unravel its structure by column (column major, e.g.
+to store matrix $M$ it first has to be converted to base $2$ ($M_2$).
+Then it will be unraveled by column (column major, e.g.
 in Fortran) or by row (row major, e.g.
-in C) and line them up in a contiguous block of memory.
-Each element of the matrix is allocated in turn according to the type associated with the matrix.
+in C) and its elements will be written as  fixed size adjacent chunks of memory.
+The size of each chunk is determined by the type associated with the matrix (typically 64 bits, but always a power of 2).
 
-For our example, since we have determined that each element will require at most 10 bits, we can divide the memory block corresponding to a single 64 bit integer into 6 10-bit chunks which can each hold a single element of $M$.
-Each of these blocks we will call a bitstring.
+According to the SM method, having determined that each element will require at most 10 bits, we can divide the memory block corresponding to a single 64 bit integer into 6 10-bit chunks which can each hold a single element of $M$.
+These 64-bit blocks will be called a bitstring.
 The remaining 4 bits will be used later.
-The number of bit strings needed will be $\lfloor \frac{\dim(M) *b} {64} \rfloor +1$.
+The number of bitstrings needed will be $\lfloor \frac{\dim(M) *b(\sup M)} {64} \rfloor +1$.
 
-The final layout of the first 6 elements of $m$ in the first bitstring can be seen in equation \ref{eq:09}.
+The final layout of the first 6 elements of $m$ in the first bitstring can be seen in \ref{eq:09}.
 
 \begin{equation}\label{eq:09}
- bitstring 1 = 0000\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\underbrace{0100000000}_{256}\underbrace{1011010001}_{721}\underbrace{1111111111}_{1023}\underbrace{1110000100}_{900} 
+ bitstring_1 = 0000\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\underbrace{0100000000}_{256}\underbrace{1011010001}_{721}\underbrace{1111111111}_{1023}\underbrace{1110000100}_{900} 
 \end{equation}
 
-Here is a step-by-step description of the allocation process:
+Here is a step-by-step description of the aplicationof the SM method to matrix $M$:
 \begin{enumerate}
@@
  \item Element $M_{1,2}=1023$ is allocated in the second chunk, from bit 10 to bit 19.
- \item And so forth: Elements $M_{1,i}$ with $i=1,\ldots,6$ re stored on the remaining chunks (eq. \ref{eq:10}).
+ \item Repeat for elements $M_{1,i}$ with $i=1,\ldots,6$ which are stored on the remaining chunks.
+ \begin{equation*}\label{eq:10}
+  \underbrace{\underbrace{0000010100}_{20}\overbrace{101011}^{b}}_{bitstring_2}|\underbrace{\overbrace{1100}^{a}\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\hdots}_{bitstring_1}
+\end{equation*}
  \item Element 7 does not fit on the remaining 4 bits of the first bitstring.
-So it will straddle two bitstrings, i.e., we will start writing it on the first bitstring and finish on the second (equation \ref{eq:11}).
+So it will straddle two bitstrings, i.e., it is divided in two segments $a$ and $b$, $a$ is written on the first bitstring $b$ on the second.
+ \begin{equation*}\label{eq:10}
+  \underbrace{\underbrace{0000010100}_{20}\overbrace{101011}^{b}}_{bitstring_2}|\underbrace{\overbrace{1100}^{a}\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\hdots}_{bitstring_1}
+\end{equation*}
 \end{enumerate}
@@
 
-\begin{equation}\label{eq:10}
-  \underbrace{\underbrace{0000010100}_{20}\overbrace{101011}^{a}}_{\phi_{1,2}}|\underbrace{\overbrace{1100}^{b}\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\hdots}_{\phi_{1,1}}
-\end{equation}
+\begin{equation*}\label{eq:11}
+  \underbrace{0000010100}_{20}\underbrace{\overbrace{101011}^a|\overbrace{1100}^{b}}_{700}\underbrace{0000001010}_{10}\ldots
+\end{equation*}
 
-\begin{equation}\label{eq:11}
-  \underbrace{0000010100}_{20}\underbrace{\overbrace{101011}^a|\overbrace{1100}^{b}}_{700}\underbrace{0000001010}_{10}\ldots
-\end{equation}
+Thus the compressed matrix $m = M_2$ requires less memory than the conventional storage of $M$ as a 64-bit integer array.
 
-In conclusion, the final space required to store matrix $m = M_2$ is much smaller than would be required for the conventional storage of $M$ as a 64-bit integer array.
-In the results section, we will examine in detail the efficiency of this method on real use cases.
+\subsubsection*{Method 2: Variable Length Blocks (VLB)}
 
-\subsubsection*{Method: Variable Bit Length}
+In the SM method, there is still waste of space since for elements smaller than the supremum, a number of bits remain unused.
 
-In method 1, there is still waste of space since for elements smaller than the supremum, a number of bits may remain unused.
+In the VBL method, the absolute minimal number of bits are used to store each value.
+However, if we are going to divide the biststrings into variable length chunks, we also need to reserve some extra bits to represent the size of each chunk, otherwise the elements cannot be recovered once they are stored.
 
-In this second method, we strive to use absolute minimal number of bits necessary to store each value.
-However, if we are going to divide the biststrings into variable length chunks, we will also need to reserve some extra bits to represent the size of the chunks, otherwise we cannot recover the elements once they are stored.
-
-To illustrate this method, lets consider again the matrix described in equation \ref{eq:08}.
-Suppose the largest value stored in it is number 1023.
-Now instead of assigning one chunk of the bitstring to each element of $m$, we will assign two chunks: the first will store the number of bits required to store the value and the second  will store the respective value.
-The first region will have a fixed size, in the example cited, 4 bits.
+Lets use again the matrix described in equation \ref{eq:08}, where  the largest element is number 1023.
+Now instead of assigning one chunk of the bitstring to each element of $m$, we will assign two chunks: the first will store the number of bits required to store the element and the second  will store the actual element.
+The first chunk will have a fixed size, in this case, 4 bits.
 These 4 bits are the required space to store the bit-length of $\sup M $, in this case, 10.
 
-To illustrate this second method, lets go through the compression process step-by-step.
+Lets go through VLB compression step-by-step.
 The largest element of $M$ is 1023.
-The number of bits required to represent it is 10 which in turn requires 4 bits to be stored.
+Its bit-length is 10 which in turn is 4 bits long in base 2 (1010).
 Thus the fixed size chunk is 4 bits long for every element.
@@
 \begin{enumerate}
- \item The first element $M_{1,1}=900$ requires 10 bits to store, so we write $10_2$ in the first chunk and $900_2$ in the second (Eq. \ref{eq:13}).
- \item We apply the same logic with the following element, $M_{1,2}=1023$.
- \item Element $M_{1,2}=721$ is also added taking the bitstring to the state illustrated in \ref{eq:14}.
+ \item The first element $M_{1,1}=900$ requires 10 bits to store, so we write $10$ in the first chunk and $900$ in the second.
+ \begin{equation*} \label{eq:13}
+ bitstring_1 = 00000000000000000000000000000000000000000000000000\underbrace{\underbrace{1110000100}_{\text{element} = 900}\underbrace{1010}_{\text{bit-length=10}}}_{M_{1,1}}
+\end{equation*}
+ \item Do the same for the next element, $M_{1,2}=1023$.
+ \item Element $M_{1,2}=721$ is also added taking the bitstring to the state.
+ \begin{equation*} \label{eq:14}
+ bitstring_1 = 
+ 0000000000000000000000\underbrace{1011010001}_{721}\underbrace{1010}_{10}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1110000100}_{900}\underbrace{1010}_{10}
+\end{equation*}
 \end{enumerate}
 
-\begin{equation} \label{eq:13}
- bitstring_1 = 00000000000000000000000000000000000000000000000000\underbrace{\underbrace{1110000100}_{\text{value} = 900}\underbrace{1010}_{\text{bitlength=10}}}_{900}
-\end{equation}
-
-\begin{equation} \label{eq:14}
- bitstring_1 = 
- 0000000000000000000000\underbrace{1011010001}_{721}\underbrace{1010}_{10}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1110000100}_{900}\underbrace{1010}_{10}
-\end{equation}
-
-So far the second method is proving more wasteful than the first, but when we add $M_{1,4} =256$  we start to save some space.
+So far the VLB method is more wasteful than the SM, but when we add $M_{1,4} =256$  we start to save some space.
 
@@
  \item[4.] Element $M_{1,4} =256$ is added.  
- \item [5.] Elements $M_{1,5} =1$ and $M_{1,6} =1$ are added requiring a total of 13 bits instead of 20 with the first method (eq. \ref{eq:16} and \ref{eq:17} ).
- \item [6.] The remaining two elements are added $M_{1,7} =1$ and $M_{1,8} =1$ (eq. \ref{eq:18}).
+ \item [5.] Elements $M_{1,5} =1$ and $M_{1,6} =10$ are added requiring a total of 13 bits instead of 20 with the SM method.
+ \begin{align*} \label{eq:16}
+ bitstring_1 &=
+ \underbrace{0100}_{4}\underbrace{1}_{1}\underbrace{0001}_{1}\underbrace{100000000}_{256}\underbrace{1001}_{9}\underbrace{1011010001}_{721}\underbrace{1010}_{10}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1110000100}_{900}\underbrace{1010}_{10} \\
+ bitstring_2 &= 000000000000000000000000000000000000000000000000000000000000\underbrace{1010}_{10}
+\end{align*}
+ \item [6.] The remaining two elements are added $M_{1,7} =700$ and $M_{1,8} =20$.
+ \begin{equation*} \label{eq:18}
+ bitstring_2 = 00000000000000000000000000000000000\underbrace{10100}_{20}\underbrace{0101}_{5}\underbrace{1010111100}_{700}\underbrace{1010}_{10}\underbrace{1010}_{10}
+\end{equation*}
 \end{enumerate}
 
-\begin{equation} \label{eq:16}
- bitstring_1 =
- \underbrace{0100}_{4}\underbrace{1}_{1}\underbrace{0001}_{1}\underbrace{100000000}_{256}\underbrace{1001}_{9}\underbrace{1011010001}_{721}\underbrace{1010}_{10}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1110000100}_{900}\underbrace{1010}_{10}
+ We used a total of 87 bits to store matrix $m$ with the VLB method instead of 80 bits using the SM method.
+However, as shall be seen later, the VLB method will be the most efficient for most matrices.
+
+ \subsection*{Compression Efficiency}
+ Compression efficiency depends of the data being compressed. Below, a formula for calculating compression efficiency is derived for both methods. They will be based on the following ratio:
+
+ \begin{equation}\label{eq:eff1}
+  \eta=\frac{bits\, alocated-bits\, used}{bits\,alocated}
  \end{equation}
-
-\begin{equation} \label{eq:17}
- bitstring_2 = 000000000000000000000000000000000000000000000000000000000000\underbrace{1010}_{10}
-\end{equation}
-
-\begin{equation} \label{eq:18}
- bitstring_2 = 00000000000000000000000000000000000\underbrace{10100}_{20}\underbrace{0101}_{5}\underbrace{1010111100}_{700}\underbrace{1010}_{10}\underbrace{1010}_{10}
-\end{equation}
-
- We used a total of 87 bits to store matrix $m$ with the second method instead of 80 bits using method 1. However, method 2 will be the most efficient for most matrices as show below.
-
- \subsection*{Efficiency of Compression}
- To compare the two compression methods proposed, we can calculate their efficiencies.
-However since they depend of the data being compressed, we will describe below how to  calculate the efficienty given the data for both methods.
-Figure
-
  \subsubsection*{Method 1}
  Let $M_{r \times c}$ be the matrix we wish to compress. we can define the efficiency of the compression as:
- \begin{equation}\label{eq:eff1}
-  \eta_1=\frac{bits\, alocated-bits\, used}{bits\,alocated}
- \end{equation}
 
@@
 \section*{Acknowledgments}
-We would like to thank Paulo Cezar Carvalho and Moacyr silva for fruitful discussions and key ideas which helped to improve this paper.
+We would like to thank Claudia Torres Codeço, Paulo Cezar Carvalho and Moacyr silva for fruitful discussions and key ideas which helped to improve this paper.
 

### 2012-12-29 manuscript.tex
a meio caminha dos resultados

@@
  \end{equation}
- \subsubsection*{Method 1}
- Let $M_{r \times c}$ be the matrix we wish to compress. we can define the efficiency of the compression as:
 
-Equation \ref{eq:eff1} represent the compression ratio achieved by the method. In comparison with a conventional allocation (64-bit integers) we can rewrite equation \ref{eq:eff1} as:
+ Where \textit{bits alocated} above mean total bits required for standard storage of the matrix, without compression, while bits used mean total bits requires to store the matrix after compression.
+ \subsubsection*{SM Method}
+ Let $M_{r \times c}$ be the matrix we wish to compress.
+In comparison with a conventional allocation (64-bit integers) we can apply equation \ref{eq:eff1} to caluculate the efficiency of the SM method:
 
@@
  \eta_1 &= \frac{64 \times rc - b(max M) \times rc}{64 \times rc}\nonumber\\
- &= \frac{64  - b }{64}
+ &= \frac{64  - b(max M) }{64}
 \end{align}
 
-As we can see in (\ref{eq:22}), $\eta_1$ does not depend on size of the matrix, only on the bit-length of $\max M$ (equation \ref{eq:19}).
-According to equation \ref{eq:22} if $b(\max M)=64$, $\eta_1$ will be 0, i.e.
+As we see in \ref{eq:22}, $\eta_1$ does not depend on size of the matrix, only on the bit-length of $\max M$.
+If $b(\max M)=64$, $\eta_1$ is 0, i.
+e.
 no compression is possible.
-This leads to a very interesting optimization oportunity: divide the entire matrix by the greatest common divisor  of its elements.
 On the other extreme, if the matrix is composed exclusively of 0s and 1s, maximal compression is achievable, $\eta_1=1$.
 
-\begin{equation} \label{eq:19}
- 0000000000000000000000000000000000\underbrace{1111111111}_{1023}\underbrace{\overbrace{000000000}^{\text{Wasted}}1}_{1}\underbrace{\overbrace{000000000}^{\text{Wasted}}0}_{0}
-\end{equation}
+ \subsubsection*{VLB Method}
 
-\begin{figure}[h]
- \centering
- \psfrag{e}{$\xi$}
- \includegraphics[scale=0.5]{fig01}
- \caption{Compression efficiency of method 1 behavior with respect to bit number.}
- \label{fig:01}
-\end{figure}
+ For the VLB method, compression depends on the value of each element of the matrix.
+In this method bit-lentgh variability affects the compression ratio, so the formula will have to include this information.
 
- \subsubsection*{Method 2}
-
- For the second method, compression depends on the value of each element of the matrix.
-Thus to compute its efficiency, we shal modify equation \ref{eq:eff1}.
-We'll start by grouping the elements by bit length, since in this method, bit-lentgh variability greatly affects the compression ratio.
-
-Again consider the matrix $M_{r \times c}$, only the $rc$ elements are now divided into $g$ groups, with $f_i$ numbers of bit-length $b_i = b(m_i)$ each.
+Let the $rc$ elements of the matrix $M_{r \times c}$ be divided into $g$ groups, each with $f_i$ numbers of bit-length $b_i = b(m_i)$.
+Thus $f_i$ is the frequency of each bit-length present in $M$.
 Let $k = b(b(\max M))$, i.e., the bit length of the bit-length of $\max M$.
-The efficiency $\eta_2$ is then defined as shown in \ref{eq:24}.
-Equation \ref{eq:20} can help us visualize it.
-
-\begin{equation} \label{eq:20}
- 0000000000000000000000000000000000000000\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1}_{1}\underbrace{0001}_{1}\underbrace{0}_{0}\underbrace{0001}_{1}
-\end{equation}
+The efficiency $\eta_2$ is shown below.
 
@@
 
-Consider a matrix $M$ such that $m_{i,j} \in \mathbb{Z}$. In this section we will examine and compare the efficiency of the two compression methods described.
-
-Since the efficiency of both methods depend on the distribution of the bit lengths $b(m_{i,j})$, we will explore a variety of distributions and try to cover the most common cases.
+\paragraph{Random Matrix Generation.}
+In both methods, compression eficiency depends on the distribution of the bit-lengths $b(m_{i,j})$.
+Thus, in this section, a method to generate a variety of random bit-length distributions is proposed.
 
 For simplicity we will model the distribution of  as a mixture $X$ of two Beta distributions ,$ B_1 \sim Beta(\alpha_1,\beta_1)$ and $B_2 \sim Beta(\alpha_2,\beta_2)$, whose probability function is shown in equation (eq.\ref{eq:mixture}).
-Since the Beta distribution is defined only in the interval $[0,1] \subset \mathbb{R}$ , we applied a simple transformation (\ref{eq:transf}) to the mixture in order to map it to the interval of $[1,64] \subset \mathbb{Z}$
+Since the Beta distribution is defined only in the interval $[0,1] \subset \mathbb{R}$ , we applied a simple transformation ($\lfloor 64 \times x \rfloor + 1$) to the mixture in order to map it to the interval of $[1,64] \subset \mathbb{Z}$
 
@@
 
-\begin{equation}\label{eq:transf}
-   \lfloor 64 \times x \rfloor + 1
-\end{equation}
-
 In order to explore the compression efficiency of both methods, we generated samples from the mixture defined above, varying its parameters.
@@
 
-Given any matrix generated by our mixture distribution, we can use equation \ref{eq:22} and \ref{eq:25}, to determine the compression efficiency of methods 1 and 2, respectively.
+From now on we will apply equations \ref{eq:22} and \ref{eq:25}, to determine the compression efficiency of methods SM and VLB for random matrices generated as describe above.
 
-With $w=0$, ie, our mixture is reduce to a single Beta distribution. In Figure \ref{fig:02},
+With $w=0$, a single Beta distribution is used. In Figure \ref{fig:02},
 we show some distributions of bit-lengths for some combinations of $\alpha_1$ and $\beta_1$.
-From the figure it can be seen that we can generate just about any kind of unimodal distributions in the interval $[1,64]$.
+From the figure it can be seen that a large variety of unimodal distributions can be generated in the interval $[1,64]$.
 
@@
 
-As we are sampling the from set of all possible distributions of bit-length, represented byt the mixture of betas presented above, in order to make our results more general, we will base our analysis on the expected bit-length of a sample, since the efficiency of both methods  depend on it.
+As we are sampling the from set of all possible distributions of bit-length, represented by the mixture of betas presented above, in order to make our results more general, we will base our analysis on the expected bit-length of a sample, since the efficiency of both methods  depend on it.
 So, from equations \ref{eq:22} and \ref{eq:25}, the expected efficiencies become:

### 2012-12-30 manuscript.tex
terminei segunda revisão

@@
  Where \textit{bits alocated} above mean total bits required for standard storage of the matrix, without compression, while bits used mean total bits requires to store the matrix after compression.
+From nowon the efficiencies are denoted by $\eta_1$ for the SM method and by $\eta_2$ for the VLB method.
  \subsubsection*{SM Method}
@@
 
-As we are sampling the from set of all possible distributions of bit-length, represented by the mixture of betas presented above, in order to make our results more general, we will base our analysis on the expected bit-length of a sample, since the efficiency of both methods  depend on it.
+As we are sampling from a large set distributions of bit-length, represented by the mixture of betas presented above, in order to make our results more general, we will base our analysis on the expected bit-length of a sample, since the efficiency of both methods depends on it.
 So, from equations \ref{eq:22} and \ref{eq:25}, the expected efficiencies become:
@@
 
-where k in the expression of $\eta_2$ is set to 7 (the bit-length required to represent the largest possible bit-length: 64).
-In \ref{eq:56}, $k$ is the bit-length of the greatest element, or in the worst case, 64.
+where k in \ref{eq:57}, is set to 7 (the bit-length required to represent the largest possible bit-length: 64). In \ref{eq:56}, $k$ is the bit-length of the greatest element, or in the worst case, 64.
 
@@
 
-For the numeric experiments, we generate $3$ samples of size $10000$, from which we calculate the expected compression efficiency, using equations \ref{eq:56} and \ref{eq:57}.
-the values presented in tables and figures below are averages of the efficiencies of the three samples.
+The expected compression efficiency in the following numeric experiments, will be calculated from $3$ matrices of dimension $10000$, generated as described, and presented in tables and figures below.
 
 In Figure \ref{fig:03}, we can see the distribution of efficiencies and their difference for a sample generated from a single Beta distribution of bit-lengths.
-We can see that both methods can achieve efficiencies greater than 80\% for matrices with very small numbers.
-We also see in figure \ref{fig:03} that method 2 is more efficient in the majority of cases.
+Note that both methods can achieve efficiencies greater than 80\% for matrices with very small numbers.
+Also note that the VLB method is more efficient in the majority of cases.
 
@@
 
-Now let's analyze matrices whose elements have bit lengths sampled from a mixture of beta distributions, $B_1$ and $B_2$.
-For this mixture $B$ we set $w = 0.5$, i.e., the mixture will have half the numbers generated from the distribution $B_1 \sim Beta(\alpha_1,\beta_1)$ and the other part distribution $B_2 \sim Beta(\alpha_2,\beta_2)$.
+Now let $w = 0.5$, i.
+e., matrices will have elements with bit-lengths comming from a mixture of beta distributions, $B_1\sim Beta(\alpha_1,\beta_1)$ and $B_2\sim Beta(\alpha_2,\beta_2)$.
 The expected value for this mixture is shown in equation \ref{eq:58}.
@@
 
-We must also redefine the calculations of efficiency for the mixture samples.
-Equation \ref{eq:59} shows the expected efficiency for method 1.
-Now, instead of having the efficiency being a function of greatest bit-length in the sample (denoted as  $k$ in \ref{eq:56} and \ref{eq:57}), it will be a function of $max\{E(B_1),E(B_2)\}$.
-We do the same for method​ 2 (equation \ref{eq:60})​.
+For bit-lengths coming from a mixture ($w>0$), let the expected efficiencies for the SM and VLB methods be as given by equations \ref{eq:59} and \ref{eq:60}.
+So now, instead of having the efficiency be a function of greatest bit-length in the sample (denoted as  $k$ in \ref{eq:56} and \ref{eq:57}), it will be a function of $max\{E(B_1),E(B_2)\}$.
 
@@
 
-As before, we generate 3 samples of size $10000$ for each parameterization, calculate the average efficiencies (equations \ref{eq:59} and \ref{eq:60}) and their diference $D$.
+As before, we generate 3 matrices of dimension $10000$ for each parameterization, calculate the average efficiencies (equations \ref{eq:59} and \ref{eq:60}) and their diference $D$.
 
-Before we move on to the efficiency results and analyses, let's first inspect our samples from the mixture of transformed Beta distributions.
+Before moving on to efficiency results and analyses, let's first inspect samples from the mixture of transformed Beta distributions.
 Figures \ref{fig:02A} and \ref{fig:02B}, show a few parameterizations and their resulting sample distributions.
@@
 
-After sampling uniformly ($[1,5,9,\ldots,64]$, $n=65536$) the parameter space and comparing efficiencies, we summarized the results on table \ref{tab:06}.
+After sampling uniformly ($[1,5,9,\ldots,64]$, $n=65536$) the bit-length space and comparing efficiencies, we summarized the results on table \ref{tab:06}.
 In it we see how many parameterizations (from our sample) favor each method.
@@
   \hline 
-  Methods  & n   & Percentage \\
+  Methods  & n   & Method wins \\
   \hline
@@
 
-As we have shown, method is more effective compressing most integer datasets up to 64bits in size.
+As we have shown, the VLB method is more effective compressing most integer datasets up to 64bits in size.
 This is due to its ability to exploit the variance in the data set and reduce the waste of bits in the representation of some numbers.
@@
 
-The two methods presented allow the storage and information manipulation in the compressed form, which provides an data managing large amounts optimization.
-The next step in the work development is the implementation of the proposed methods, as well as tests to be performed on data large volumes.
-
 \section*{Discussion}
 
-\subsection*{Which method to use?}
+\subsection*{Calculating Efficiencies}
 
 To determine the best compression method to apply, it's necessary to inspect the distribution of bit-lengths of matrix elements.
-When matrix elements are small or have nearly-constant bit-length, Method I is best, otherwise, the second method should be chosen.
+When matrix elements are small or have nearly-constant bit-length, the SM Method is best, otherwise, the VLB method should be chosen.
 
@@
 
-The efficiency of method II is influenced by the relative size of the bit-length groups.
+The efficiency of the VLB method is influenced by the relative size of the bit-length groups.
 In this first example we considered only two groups, each comprised of half the matrix elements.
@@
   \hline 
-  $p_1$  & $p_2$ & Efficiency \\
+  $p_1$  & $p_2$ & $\eta_2$ \\
   \hline
@@
 \begin{itemize}
-  \item 3 number groups that require numbers 1, 32 and 64 bits, efficiency $\zeta=0.3854$,
-  \item 5 number groups that require numbers 1, 16, 32, 48 and 64 bits, efficiency $\zeta=0.3875$,
-  \item 8 number groups that require numbers 1, 8, 16, 24, 32, 40, 48, 56 and 64 bits, efficiency $\zeta=0.3888$
+  \item 3 groups with bit-lengths 1, 32 and 64 bits, efficiency $\eta_2=0.3854$,
+  \item 5 groups with bit-lengths 1, 16, 32, 48 and 64 bits, efficiency $\eta_2=0.3875$,
+  \item 8 groups with bit-lengths 1, 8, 16, 24, 32, 40, 48, 56 and 64 bits, efficiency $\eta_2=0.3888$
 \end{itemize}
@@
 \subsection*{Discrete Uniform}
-The discrete uniform distribution has two parameters, $a$ and $b$ representing the minimum and maximum value of the distribution.
-For our case we will use $U(a=1,b=64)$, which means bit-lengths may take values in the set $\{1, 2, 3, $\ldots$, 64\}$ with equal probability, i.e., $\frac{1}{64}$.
+Let the bit-lengths of the matrices be distribute according to the Uniform distribution $U(a=1,b=64)$, which means bit-lengths may take values in the set $\{1, 2, 3, $\ldots$, 64\}$ with equal probability, i.e., $\frac{1}{64}$.
 
-For the Poisson distribution, we use Poisson($\lambda$) with $\lambda$ being the expected bit-length in the sample.
+\paragraph{Theoretical Efficiency:}
+Let the random variable $B \sim U(a=1,b=64)$ represent the  bit-length of the elements of matrix $M$.
+Then $E(b_i) = \sum_{i} b_i \times p(b_i) = \frac{a+b}{2}$.
+Applying this result to the expected compression efficiency of method II (equation \ref{eq:37}), we have
 
-To calculate the efficiency, we generated a sample with 100 ($M_{10 \times 10}$), 10,000 ($M_{100 \times 100}$) and 1,000,000 ($M_{1,000 \times 1.000}$) numbers .
-The average efficiency (table \ref{tab:03}) is calculated from a 1000 replicates of each sample size.
+\begin{equation}\label{eq:44}
+ E(\eta_2) = 1 - \frac{E(B)}{64} - \frac{k}{64} 
+\end{equation}
+
+assuming all bit-lengths are possible, i.e., $a=1$ and $b=64$, and hence $k=7$, we can calculate $\eta_2$:
+
+\begin{equation}\label{eq:45}
+ E(\eta_2) = 1 - \frac{\frac{1+64}{2}}{64} - \frac{7}{64} \approx 38.28\% 
+\end{equation}
+
+This result agrees with the numerical estimates presented in table \ref{tab:03}.
+
+\paragraph{Numerical Estimates:}
+To calculate the efficiency, we generated a matrices with 100 ($M_{10 \times 10}$), 10,000 ($M_{100 \times 100}$) and 1,000,000 ($M_{1,000 \times 1.000}$) elements .
+The average efficiency (table \ref{tab:03}) is calculated from a 1000 replicates of each matrix size.
 
@@
     \hline
-    Sample size & Avg. efficiency \\
+    Matrix size & Avg. efficiency \\
     \hline
@@
 
-For the binomial distribution, we will use $Bin(n, p=0.5)$, with the number of trials $n$ representing the greatest possible bit-length in the matrix, and $p=0.5$ giving us an expected bit-length of 32.
+For the binomial distribution, we will use $Bin(n, p)$, with the number of trials $n$ representing the greatest possible bit-length in the matrix, with $np$ giving us an expected bit-length.
 
+\paragraph{Theoretical Efficiency:}
+Let bit-length $(B)$ be a random variable with Binomial distribution, $B \sim B(n=64,p=0.5)$, $E(b_i) = \sum_{i} b_i \times p(b_i) = n \times p$ and the eficiency becomes (with $k=7$):
+
+\begin{align}\label{eq:47}
+ E(\eta_2) &= 1 - \frac{64 \times p }{64} - \frac{k}{64}  \\
+  &= 1 - \frac{64 \times 0.5 }{64} - \frac{7}{64} \approx 39.05\% \nonumber
+\end{align}
+
+Which again agrees with estimates in Table \ref{tab:04}.
+\paragraph{Numerical Estimates:}
 For these experiments, the parameter n represents the maximum bit-length of matrix elements and takes values in $\{1, 8, 16, 32, 64\}$.
-In this case, we evaluate the efficiency as a function of the parameter $n$, and sample size.
+In this case, we evaluate the efficiency as a function of the parameter $n$, and matrix size.
 Even though efficiency does not depend on matrix size, we tried different sizes to test the stability of the compression algorithm.
 Results are shown in Table \ref{tab:04}.
-As expected smaller expected bit lengths lead to higher compression efficiencies.
+As expected, smaller bit-lengths lead to higher compression efficiencies.
 However, for the case in which the elements require 64 bits, the efficiency is negative.
@@
 \subsection*{Poisson Distribution}
-With bit-length derived from a Poisson($\lambda$), the parameter $\lambda$ corresponds to the expected bit-length.
-Values ​​greater than 64 generated in this simulation were considered as being equal to 64.
+With bit-length derived from a Poisson($\lambda$), the parameter $\lambda$ corresponds to the expected bit-length. For the purpose of this analysis this Poisson distribution is truncated at 64. 
+\paragraph{Theoretical Efficiency:}
+Let bit-length $B \sim (\lambda=32)$. In this case,$E(b_i) = \lambda$, with $k=7$,  the efficiency becomes:
+
+\begin{align}\label{eq:49}
+ E(\eta_2) &= 1 - \frac{\lambda}{64} - \frac{k}{64} \\
+ &= 1 - \frac{32}{64} - \frac{7}{64} \approx 39.06\%
+\end{align}
+
+This result is in accordance to Table \ref{tab:05}.
+\paragraph{Numerical Estimates:}
 The results for this simulation can be seen in Table \ref{tab:05}.
@@
     \hline
-			& 	&Efficiency         & \\
+			& 	&Efficiency $\eta_2$      & \\
     \hline
-			&  	&Size sample& \\
+			&  	&Matrix Size& \\
     Expected bit-length ($\lambda$)		& 100	& 1,000		    & 1,000,000 \\
@@
 
-With bit-length as a random variable, we can generalize the results about the compression efficiency as a function of expected bit-length
-
-Let the random variable $B \sim U(a=1,b=64)$ represent the  bit-length of the elements of matrix $M$.
-Then $E(b_i) = \sum_{i} b_i \times p(b_i) = \frac{a+b}{2}$.
-Applying this result to the expected compression efficiency of method II (equation \ref{eq:37}), we have
-
-\begin{equation}\label{eq:44}
- E(\eta_2) = 1 - \frac{E(B)}{64} - \frac{k}{64} 
-\end{equation}
-
-assuming all bit-lengths are possible, i.e., $a=1$ and $b=64$, and hence $k=7$, we can calculate $\eta_2$:
-
-\begin{equation}\label{eq:45}
- E(\eta_2) = 1 - \frac{\frac{1+64}{2}}{64} - \frac{7}{64} \approx 38.28\% 
-\end{equation}
-
-This result agrees with the numerical estimates presented in table \ref{tab:03}.
-We can do the same for the case where bit-length $(B)$ is a random variable with Binomial and Poisson distributions.
-In the case of Binomial, $B \sim B(n=64,p=0.5)$, $E(b_i) = \sum_{i} b_i \times p(b_i) = n \times p$ and the eficiency becomes (with $k=7$):
-
-\begin{align}\label{eq:47}
- E(\eta_2) &= 1 - \frac{64 \times p }{64} - \frac{k}{64}  \\
-  &= 1 - \frac{64 \times 0.5 }{64} - \frac{7}{64} \approx 39.05\% \nonumber
-\end{align}
-
-Which again agrees with estimates in Table \ref{tab:04}.
-
-Lastly, for $B$ distributed as Poisson$(\lambda=32)$ case,$E(b_i) = \lambda$, $k=7$, and the efficiency becomes:
-
-\begin{align}\label{eq:49}
- E(\eta_2) &= 1 - \frac{\lambda}{64} - \frac{k}{64} \\
- &= 1 - \frac{32}{64} - \frac{7}{64} \approx 39.06\%
-\end{align}
-
-This result is again in accordance to Table \ref{tab:05}.
 These results show that a good compression is guaranteed when bit-lengths are distributed according to the tested distributions regardless of sample size.
@@
 [devemos incluir um estudo sobre a variância?]
-
+\section{Conclusion}
 In this paper we have focused in the compression of matrix data, since this is one of the most important application the authors foresee.
@@
 \section*{Acknowledgments}
-We would like to thank Claudia Torres Codeço, Paulo Cezar Carvalho and Moacyr silva for fruitful discussions and key ideas which helped to improve this paper.
+We would like to thank Claudia Torres Codeço, Paulo Cezar Carvalho and Moacyr silva for fruitful discussions and key ideas which helped improve the manuscript.
 

### 2012-12-30 manuscript.tex
corrigi pequenos erros de ortigrafia

@@
 \section*{Acknowledgments}
-We would like to thank Claudia Torres Codeço, Paulo Cezar Carvalho and Moacyr silva for fruitful discussions and key ideas which helped improve the manuscript.
+We would like to thank Claudia Torres Code\c{c}o, Paulo Cezar Carvalho and Moacyr Silva for fruitful discussions and key ideas which helped improve the manuscript.
 

### 2013-01-11 manuscript.tex
adicionei as modificações

@@
 Most if not all existing lossless compression methods were developed under the following usage paradigm: \textit{produce} $\rightarrow$ \textit{compress} $\rightarrow$ \textit{store} $\rightarrow$ \textit{uncompress} $\rightarrow$ \textit{use}.
-The focus of the present work is to allow a slightly different usage:  \textit{produce} $\rightarrow$ \textit{compress} $\rightarrow$ \textit{perform mathematical manipulations} $\rightarrow$ \textit{decompress} (only for human reading).
+The focus of the present work is to allow a slightly different usage:  \textit{produce} $\rightarrow$ \textit{compress} $\rightarrow$ \textit{perform mathematical manipulations to use} and \textit{decompress} only for human reading.
 
@@
 This way, the cost of compression is reduced to a single event of compression and no need of decompression except when displaying the results for human reading.
-The idea of operating with compressed arrays is relatively new\cite{yemliha2007compiler}, and has yet to find mainstream applications to the field of numerical computations.
+The idea of operating with compressed arrays is relatively new\cite{yemliha2007compiler}, and it has yet to find mainstream applications to the field of numerical computations.
 One application which employs a form of compression is the sparse matrix linear algebra algorithms\cite{dodson1991sparse}, in this case there is no alteration in the standard encoding of the data, but only the non-zero elements of the matrices are stored and operated upon.
@@
 In order to achieve compression we decided to exploit inefficiencies in the conventional way matrices are allocated in memory.
-The exemples in this paper will be restricted to matrices with positive integer elements.
+The examples in this paper will be restricted to matrices with positive integer elements.
 
@@
 In digital computers, all information is stored as binary code (base 2 numbers).
-However the conventional way to store arrays of integers is on a sequence of memory blocks of fixed size (power of 2 number of bits), one for each element.
+However the conventional way to store arrays of integers is on a memory block sequence of fixed size (power of 2 bit numbers), one for each element.
 The maximum size of a block is equal to the word size of the processor, which for most current CPUs is 64 bits.
@@
 
-The SM method consists in determining the value of the greatest element of matrix $M$, which coincides with its supremum, $max M == \sup M$  and determine the minimum number of bits, $b(\sup M$), required to store it.
+The SM method consists in determining the value of the greatest element of matrix $M$, which coincides with its supremum, $max M == \sup M$  and determine the minimum bit number, $b(\sup M$)(Equation \ref{eq:01}), required to store it.
 We will use capital roman letters to denote uncompressed matrices and the corresponding lower case letter for the compressed version.
 
-\begin{equation} \label{eq:07}
+\begin{equation} \label{eq:01}
  b(\sup M) \approx \begin{cases}
@@
 
-The allocation of memory still happens in the usual way, i.e.
-in fixed size 64-bit chunks, only that now, in the space required for a single 64 bit integer, we can store for example, an entire $8\times 8$ matrix of $0_{10}$ and $1_{10}$.
+The allocation of memory still happens in the usual way, i.e., in fixed size 64-bit chunks, only that now, in the space required for a single 64 bit integer, we can store for example, an entire $8\times 8$ matrix of $0_{10}$ and $1_{10}$.
 
 Let's look at a concrete example: suppose that the greatest value to be stored in a matrix $M$ is $\max M=1023$.
-Therefore, the number of bits required to represent it is $10$ $(1111111111)$.
+Therefore, the bit number required to represent it is $10$ $(1111111111)$.
 Let the first 8 elements of $M$ be:
 
-\begin{equation} \label{eq:08}
+\begin{equation} \label{eq:02}
   M = \begin{bmatrix}
@@
 These elements of $M$, in binary, are shown in Table \ref{tab:01}.
-It is evident that the number of bits required to represent any other element must be $\leq 10$.
+It is evident that the bit number required to represent any other element must be lower or equal to $10$.
 From now on the minimum number of bits required to represent a base 10 integer will be refered to as its bit-length.
@@
 
-to store matrix $M$ it first has to be converted to base $2$ ($M_2$).
+To store matrix $M$ it first has to be converted to base $2$ ($M_2$).
 Then it will be unraveled by column (column major, e.g.
@@
 
-According to the SM method, having determined that each element will require at most 10 bits, we can divide the memory block corresponding to a single 64 bit integer into 6 10-bit chunks which can each hold a single element of $M$.
+According to the SM method, having determined that each element will require at most 10 bits, we can divide the memory block corresponding to a single 64 bit integer into six 10-bit chunks which can each hold a single element of $M$.
 These 64-bit blocks will be called a bitstring.
@@
 
-The final layout of the first 6 elements of $m$ in the first bitstring can be seen in \ref{eq:09}.
+The final layout of the first 6 elements of $m$ in the first bitstring can be seen in \ref{eq:03}.
 
-\begin{equation}\label{eq:09}
+\begin{equation}\label{eq:03}
  bitstring_1 = 0000\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\underbrace{0100000000}_{256}\underbrace{1011010001}_{721}\underbrace{1111111111}_{1023}\underbrace{1110000100}_{900} 
@@
 
-Here is a step-by-step description of the aplicationof the SM method to matrix $M$:
+Here is a step-by-step description of the application of the SM method to matrix $M$:
+
 \begin{enumerate}
  \item Element $M_{1,1}=900 = m_{1,1} = 1110000100$ is stored in the first 10-bit chunk of the element strip $bitstring[1]$ , which corresponds to bits 0 to 9.
+\begin{equation*}\label{eq:04}
+bitstring_1 = 000000000000000000000000000000000000000000000000000000\underbrace{1110000100}_{900} 
+\end{equation*}
  \item Element $M_{1,2}=1023$ is allocated in the second chunk, from bit 10 to bit 19.
+\begin{equation*}\label{eq:05}
+ bitstring_1 = 00000000000000000000000000000000000000000000\underbrace{1111111111}_{1023}\underbrace{1110000100}_{900} 
+\end{equation*}
  \item Repeat for elements $M_{1,i}$ with $i=1,\ldots,6$ which are stored on the remaining chunks.
- \begin{equation*}\label{eq:10}
-  \underbrace{\underbrace{0000010100}_{20}\overbrace{101011}^{b}}_{bitstring_2}|\underbrace{\overbrace{1100}^{a}\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\hdots}_{bitstring_1}
+ \begin{equation*}\label{eq:06}
+bitstring_1 = 0000\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\underbrace{0100000000}_{256}\underbrace{1011010001}_{721}\underbrace{1111111111}_{1023}\underbrace{1110000100}_{900} 
  \end{equation*}
  \item Element 7 does not fit on the remaining 4 bits of the first bitstring.
-So it will straddle two bitstrings, i.e., it is divided in two segments $a$ and $b$, $a$ is written on the first bitstring $b$ on the second.
- \begin{equation*}\label{eq:10}
+So it will straddle two bitstrings, i.e., it is divided in two segments $a$ and $b$, $a$ is written on the first bitstring and $b$ on the second.
+ \begin{equation*}\label{eq:07}
   \underbrace{\underbrace{0000010100}_{20}\overbrace{101011}^{b}}_{bitstring_2}|\underbrace{\overbrace{1100}^{a}\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\hdots}_{bitstring_1}
@@
 
-\begin{equation*}\label{eq:11}
+\begin{equation*}\label{eq:8}
   \underbrace{0000010100}_{20}\underbrace{\overbrace{101011}^a|\overbrace{1100}^{b}}_{700}\underbrace{0000001010}_{10}\ldots
@@
 
-Lets use again the matrix described in equation \ref{eq:08}, where  the largest element is number 1023.
+Lets use again the matrix described in Equation \ref{eq:02}, where  the largest element is number 1023.
 Now instead of assigning one chunk of the bitstring to each element of $m$, we will assign two chunks: the first will store the number of bits required to store the element and the second  will store the actual element.
@@
  \item The first element $M_{1,1}=900$ requires 10 bits to store, so we write $10$ in the first chunk and $900$ in the second.
- \begin{equation*} \label{eq:13}
+ \begin{equation*} \label{eq:09}
  bitstring_1 = 00000000000000000000000000000000000000000000000000\underbrace{\underbrace{1110000100}_{\text{element} = 900}\underbrace{1010}_{\text{bit-length=10}}}_{M_{1,1}}
@@
  \item Do the same for the next element, $M_{1,2}=1023$.
- \item Element $M_{1,2}=721$ is also added taking the bitstring to the state.
- \begin{equation*} \label{eq:14}
+  \begin{equation*} \label{eq:10}
+ bitstring_1 = \ldots00000000000000000000000000\underbrace{\underbrace{1111111111}_{\text{element} = 1023}\underbrace{1010}_{\text{bit-length=10}}}_{M_{1,2}}\underbrace{\underbrace{1110000100}_{\text{element} = 900}\underbrace{1010}_{\text{bit-length=10}}}_{M_{1,1}}
+\end{equation*}
+ \item Element $M_{1,3}=721$ is also added taking the bitstring to the state.
+ \begin{equation*} \label{eq:11}
  bitstring_1 = 
@@
  \item[4.] Element $M_{1,4} =256$ is added.  
- \item [5.] Elements $M_{1,5} =1$ and $M_{1,6} =10$ are added requiring a total of 13 bits instead of 20 with the SM method.
- \begin{align*} \label{eq:16}
+ \item [5.] Elements $M_{1,5} =1$ and $M_{1,6} =10$ are added requiring a total of 13 bits instead of 20 with the SM method. Now, it is used 2 bit strips.
+ \begin{align*} \label{eq:12}
  bitstring_1 &=
@@
 \end{align*}
- \item [6.] The remaining two elements are added $M_{1,7} =700$ and $M_{1,8} =20$.
- \begin{equation*} \label{eq:18}
+ \item [6.] The remaining two elements are added $M_{1,7} =700$ and $M_{1,8} =20$ in the second bit strip.
+ \begin{equation*} \label{eq:13}
  bitstring_2 = 00000000000000000000000000000000000\underbrace{10100}_{20}\underbrace{0101}_{5}\underbrace{1010111100}_{700}\underbrace{1010}_{10}\underbrace{1010}_{10}
@@
 
- \begin{equation}\label{eq:eff1}
+ \begin{equation}\label{eq:14}
   \eta=\frac{bits\, alocated-bits\, used}{bits\,alocated}
@@
  Where \textit{bits alocated} above mean total bits required for standard storage of the matrix, without compression, while bits used mean total bits requires to store the matrix after compression.
-From nowon the efficiencies are denoted by $\eta_1$ for the SM method and by $\eta_2$ for the VLB method.
+From now on the efficiencies are denoted by $\eta_1$ for the SM method and by $\eta_2$ for the VLB method.
  \subsubsection*{SM Method}
- Let $M_{r \times c}$ be the matrix we wish to compress.
-In comparison with a conventional allocation (64-bit integers) we can apply equation \ref{eq:eff1} to caluculate the efficiency of the SM method:
+ Let $M_{r \times c}$ be the matrix we wish to compress. In comparison with a conventional allocation (64-bit integers) we can apply Equation \ref{eq:14} to caluculate the efficiency of the SM method:
 
-\begin{align}\label{eq:22}
+\begin{align}\label{eq:15}
  \eta_1 &= \frac{64 \times rc - b(max M) \times rc}{64 \times rc}\nonumber\\
@@
 
-As we see in \ref{eq:22}, $\eta_1$ does not depend on size of the matrix, only on the bit-length of $\max M$.
+As we see in \ref{eq:15}, $\eta_1$ does not depend on size of the matrix, only on the bit-length of $\max M$.
 If $b(\max M)=64$, $\eta_1$ is 0, i.
-e.
-no compression is possible.
+e., no compression is possible.
 On the other extreme, if the matrix is composed exclusively of 0s and 1s, maximal compression is achievable, $\eta_1=1$.
@@
 
-\begin{equation}\label{eq:24}
+\begin{equation}\label{eq:16}
  \eta_2 = \frac{64 \times rc - \sum_{i=1}^{g} ( b_i + k ) \times f_i }{64 \times rc} 
@@
 
-We can further simplify equation \ref{eq:24} to get at shorter expression for the compression ratio.
+We can further simplify Equation \ref{eq:16} to get at shorter expression for the compression ratio.
 
-\begin{align} \label{eq:25}
+\begin{align} \label{eq:17}
  \eta_2 &= \frac{64 \times rc - \sum_{i=1}^{g} ( b_i \times f_i + k \times f_i )}{64 \times rc} \nonumber \\
@@
   &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  - k \times\sum_{i=1}^{g}  f_i }{64 \times rc}\nonumber \\
- &\sum_{i=1}^{g} f_i = rc \nonumber \\
+ &\textrm{Please note that }\sum_{i=1}^{g} f_i = rc, \nonumber \\
  \eta_2 &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  - k \times rc }{64 \times rc}\nonumber \\
@@
 \paragraph{Random Matrix Generation.}
-In both methods, compression eficiency depends on the distribution of the bit-lengths $b(m_{i,j})$.
+In both methods, compression efficiency depends on the distribution of the bit-lengths $b(m_{i,j})$.
 Thus, in this section, a method to generate a variety of random bit-length distributions is proposed.
 
-For simplicity we will model the distribution of  as a mixture $X$ of two Beta distributions ,$ B_1 \sim Beta(\alpha_1,\beta_1)$ and $B_2 \sim Beta(\alpha_2,\beta_2)$, whose probability function is shown in equation (eq.\ref{eq:mixture}).
-Since the Beta distribution is defined only in the interval $[0,1] \subset \mathbb{R}$ , we applied a simple transformation ($\lfloor 64 \times x \rfloor + 1$) to the mixture in order to map it to the interval of $[1,64] \subset \mathbb{Z}$
+For simplicity we will model the distribution of  as a mixture $X$ of two Beta distributions, $B_1 \sim Beta(\alpha_1,\beta_1)$ and $B_2 \sim Beta(\alpha_2,\beta_2)$, whose probability function is shown in Equation \ref{eq:18}.
+Since the Beta distribution is defined only in the interval $[0,1] \subset \mathbb{R}$ , we applied a simple transformation ($\lfloor 64 \times x \rfloor + 1$) to the mixture in order to map it to the interval of $[1,64] \subset \mathbb{Z}$.
 
-\begin{equation}\label{eq:mixture}
+\begin{equation}\label{eq:18}
  F(x) =  w \, Beta(\alpha_1,\beta_1) + (1-w)\,Beta(\alpha_2,\beta_2)
@@
 The intention of using this mixture was to find a simple way to represent a large variety of bit length distributions.
-The first two central moments of this mixture are given in \ref{eq:53} and will be used later to summarize our numerical results.
+The first two central moments of this mixture are given in \ref{eq:19} and will be used later to summarize our numerical results.
 
-\begin{align}\label{eq:53}
+\begin{align}\label{eq:19}
  E(X) &= w E(B_1) + (1-w) E(B_2) \nonumber \\
@@
 
-From now on we will apply equations \ref{eq:22} and \ref{eq:25}, to determine the compression efficiency of methods SM and VLB for random matrices generated as describe above.
+From now on we will apply Equations \ref{eq:15} and \ref{eq:17}, to determine the compression efficiency of SM and VLB methods for random matrices generated as describe above.
 
-With $w=0$, a single Beta distribution is used. In Figure \ref{fig:02},
-we show some distributions of bit-lengths for some combinations of $\alpha_1$ and $\beta_1$.
+With $w=0$, a single Beta distribution is used.
+In Figure \ref{fig:01020304}, we show some distributions of bit-lengths for some combinations of $\alpha_1$ and $\beta_1$.
 From the figure it can be seen that a large variety of unimodal distributions can be generated in the interval $[1,64]$.
@@
   \subfigure[$\alpha=1,\beta_1=1$]{
+  \includegraphics[scale=0.25,angle=-90,clip]{fig01}
+  }
+  \subfigure[$\alpha=1,\beta=32$]{
   \includegraphics[scale=0.25,angle=-90,clip]{fig02}
   }
-  \subfigure[$\alpha=1,\beta=32$]{
+  \subfigure[$\alpha=32,\beta=1$]{
   \includegraphics[scale=0.25,angle=-90,clip]{fig03}
   }
-  \subfigure[$\alpha=32,\beta=1$]{
-  \includegraphics[scale=0.25,angle=-90,clip]{fig04}
-  }
   \subfigure[$\alpha=64,\beta=64$]{
-  \includegraphics[scale=0.25,angle=-90]{fig05}
+  \includegraphics[scale=0.25,angle=-90]{fig04}
   }
   \caption{Histograms constructed from samples with 10,000 elements, generated from a Beta distribution. Below each histogram is possible to verify the parameters used.}
-  \label{fig:02}
+  \label{fig:01020304}
 \end{figure}
@@
 As we are sampling from a large set distributions of bit-length, represented by the mixture of betas presented above, in order to make our results more general, we will base our analysis on the expected bit-length of a sample, since the efficiency of both methods depends on it.
-So, from equations \ref{eq:22} and \ref{eq:25}, the expected efficiencies become:
+So, from Equations \ref{eq:15} and \ref{eq:17}, the expected efficiencies become:
 
-\begin{equation}\label{eq:56}
+\begin{equation}\label{eq:20}
  E(\eta_1) = 1 - \frac{k}{64}
@@
 
-\begin{equation}\label{eq:57}
+\begin{equation}\label{eq:21}
  E(\eta_2) = 1 - \frac{E(b)}{64} - \frac{k}{64}
@@
 
-where k in \ref{eq:57}, is set to 7 (the bit-length required to represent the largest possible bit-length: 64). In \ref{eq:56}, $k$ is the bit-length of the greatest element, or in the worst case, 64.
+\noindent where k in \ref{eq:21}, is set to 7 (the bit-length required to represent the largest possible bit-length: 64).
+In \ref{eq:20}, $k$ is the bit-length of the greatest element, or in the worst case, 64.
 
-We will use the difference $D=E(\eta_1)-E(\eta_2)$ to compare the efficiency of the two methods. Thus a positive $D$ will favor method $1$ while a negative $D$ favors method $2$.
+We will use the difference $D=E(\eta_1)-E(\eta_2)$ to compare the efficiency of the two methods. Thus a positive $D$ will favor SM method while a negative $D$ favors VLB method.
 
@@
 
-In Figure \ref{fig:03}, we can see the distribution of efficiencies and their difference for a sample generated from a single Beta distribution of bit-lengths.
+In Figure \ref{fig:06070809}, we can see the distribution of efficiencies and their difference for a sample generated from a single Beta distribution of bit-lengths.
 Note that both methods can achieve efficiencies greater than 80\% for matrices with very small numbers.
@@
   \centering
-%   \subfigure[$\alpha=1,\beta_1=1$]{
-%   \includegraphics[scale=0.28,angle=-90,clip]{fig02}
-%   }
-%   \subfigure[$\alpha=1,\beta=32$]{
-%   \includegraphics[scale=0.28,angle=-90,clip]{fig03}
-%   }
-%   \subfigure[$\alpha=32,\beta=1$]{
-%   \includegraphics[scale=0.28,angle=-90,clip]{fig04}
-%   }
-%   \subfigure[$\alpha=64,\beta=64$]{
-%   \includegraphics[scale=0.28,angle=-90]{fig05}
-%   }
-  \includegraphics[scale=0.7]{simples.eps}
+   \subfigure[SM-VLB  Difference ($D = \eta_1 - \eta_2$)]{ 
+   \includegraphics[scale=0.18,clip]{fig07}
+   }
+   \subfigure[SM Efficiency  ($\eta_1$)]{
+   \includegraphics[scale=0.18,clip]{fig06}
+   }
+   \subfigure[VBL Efficiency ($\eta_2$)]{
+   \includegraphics[scale=0.18,clip]{fig05}
+   }
+   \subfigure[SM$\geq$VLB Efficiency ($\eta_1 \geq \eta_2$)]{
+   \includegraphics[scale=0.18,clip]{fig08}
+   }
   \caption{Comparing compression efficiency of methods 1 and 2.
@@
 This region corresponds to the dots in red in (a), where the average bit-length is higher.
-In panels (b) and (c), we can see the efficiencies of methods 1 and 2, respectively.}
-  \label{fig:03}
+In panels (b) and (c), we can see the efficiencies of SM and VLB methods, respectively.}
+  \label{fig:06070809}
 \end{figure}
 
-Now let $w = 0.5$, i.
-e., matrices will have elements with bit-lengths comming from a mixture of beta distributions, $B_1\sim Beta(\alpha_1,\beta_1)$ and $B_2\sim Beta(\alpha_2,\beta_2)$.
-The expected value for this mixture is shown in equation \ref{eq:58}.
+Now let $w = 0.5$, i.e., matrices will have elements with bit-lengths comming from a mixture of beta distributions, $B_1\sim Beta(\alpha_1,\beta_1)$ and $B_2\sim Beta(\alpha_2,\beta_2)$.
+The expected value for this mixture is shown in Equation \ref{eq:22}.
 
-\begin{equation}\label{eq:58}
+\begin{equation}\label{eq:22}
  E(B) = 0.5 E(B_1) + 0.5 E(B_2)
@@
 
-For bit-lengths coming from a mixture ($w>0$), let the expected efficiencies for the SM and VLB methods be as given by equations \ref{eq:59} and \ref{eq:60}.
-So now, instead of having the efficiency be a function of greatest bit-length in the sample (denoted as  $k$ in \ref{eq:56} and \ref{eq:57}), it will be a function of $max\{E(B_1),E(B_2)\}$.
+For bit-lengths coming from a mixture ($w>0$), let the expected efficiencies for the SM and VLB methods be as given by Equations \ref{eq:23} and \ref{eq:24}.
+So now, instead of having the efficiency be a function of greatest bit-length in the sample (denoted as  $k$ in \ref{eq:15} and \ref{eq:17}), it will be a function of $max\{E(B_1),E(B_2)\}$.
 
-\begin{equation}\label{eq:59}
+\begin{equation}\label{eq:23}
  E(\eta_1) = 1 - \frac{max\{E(B_1),E(B_2)\}}{64}
@@
 
-\begin{equation}\label{eq:60}
+\begin{equation}\label{eq:24}
  E(\eta_2) = 1 - 0.5\frac{E(B_1)}{64} - 0.5\frac{E(B_2)}{64} - \frac{max\{E(B_1),E(B_2)\}}{64}
@@
 
-As before, we generate 3 matrices of dimension $10000$ for each parameterization, calculate the average efficiencies (equations \ref{eq:59} and \ref{eq:60}) and their diference $D$.
+As before, we generate 3 matrices of dimension $10,000$ for each parameterization, calculate the average efficiencies (Equations \ref{eq:23} and \ref{eq:24}) and their diference $D$.
 
 Before moving on to efficiency results and analyses, let's first inspect samples from the mixture of transformed Beta distributions.
-Figures \ref{fig:02A} and \ref{fig:02B}, show a few parameterizations and their resulting sample distributions.
+Figures \ref{fig:09101112} and \ref{fig:13141516}, show a few parameterizations and their resulting sample distributions.
 It is important to note that from the mixture we can now generate bimodal distributions as well as the unimodal types tested before.
@@
 
-After sampling uniformly ($[1,5,9,\ldots,64]$, $n=65536$) the bit-length space and comparing efficiencies, we summarized the results on table \ref{tab:06}.
+After sampling uniformly ($[1,5,9,\ldots,64]$, $n=65536$) the bit-length space and comparing efficiencies, we summarized the results on Table \ref{tab:02}.
 In it we see how many parameterizations (from our sample) favor each method.
-We can also look at the distribution of efficiencies on our samples for each method (figure \ref{fig:1415}), which clearly demonstrate the greater expected efficiency of method 2 (figure \ref{fig:15}).
+We can also look at the distribution of efficiencies on our samples for each method (figure \ref{fig:1718}), which clearly demonstrate the greater expected efficiency of method VLB (figure \ref{fig:18}).
 
@@
   \subfigure[$\alpha_1=1,\beta_1=1,\alpha_2=1,\beta_2=1$]{
-  \includegraphics[scale=0.28,angle=-90,clip]{fig06}
+  \includegraphics[scale=0.28,angle=-90,clip]{fig09}
   }
   \subfigure[$\alpha_1=1,\beta_1=32,\alpha_2=32,\beta_2=1$]{
-  \includegraphics[scale=0.28,angle=-90,clip]{fig07}
+  \includegraphics[scale=0.28,angle=-90,clip]{fig10}
   }
   \subfigure[$\alpha_1=32,\beta_1=32,\alpha_2=32,\beta_2=32$]{
-  \includegraphics[scale=0.28,angle=-90,clip]{fig08}
+  \includegraphics[scale=0.28,angle=-90,clip]{fig11}
   }
   \subfigure[$\alpha_1=64,\beta_1=32,\alpha_2=32,\beta_2=64$]{
-  \includegraphics[scale=0.28,angle=-90]{fig09}
+  \includegraphics[scale=0.28,angle=-90]{fig12}
   }
-  \caption{Histograms constructed from samples with 10,000 elements, generated from the combination of two distributions Betas. Below each histogram is possible to verify the parameters used..}
-  \label{fig:02A}
+  \caption{Histograms constructed from samples with 10,000 elements, generated from the combination of two distributions Betas with $\lambda=0.5$.
+Below each histogram is possible to verify the parameters used..}
+  \label{fig:09101112}
 \end{figure}
@@
   \subfigure[$\alpha_1=64,\beta_1=48,\alpha_2=1,\beta_2=48$]{
-  \includegraphics[scale=0.28,angle=-90]{fig10}
+  \includegraphics[scale=0.28,angle=-90]{fig13}
   }
   \subfigure[$\alpha_1=16,\beta_1=46,\alpha_2=49,\beta_2=64$]{
-  \includegraphics[scale=0.28,angle=-90]{fig11}
+  \includegraphics[scale=0.28,angle=-90]{fig14}
   }
   \subfigure[$\alpha_1=16,\beta_1=16,\alpha_2=16,\beta_2=49$]{
-  \includegraphics[scale=0.28,angle=-90]{fig12}
+  \includegraphics[scale=0.28,angle=-90]{fig15}
   }
   \subfigure[$\alpha_1=1,\beta_1=16,\alpha_2=1,\beta_2=49$]{
-  \includegraphics[scale=0.28,angle=-90]{fig13}
+  \includegraphics[scale=0.28,angle=-90]{fig16}
   }
-  \caption{Histograms constructed from samples with 10,000 elements, generated from the combination of two distributions Betas. Below each histogram is possible to verify the parameters used.}
-  \label{fig:02B}
+  \caption{Histograms constructed from samples with 10,000 elements, generated from the combination of two distributions Betas  with $\lambda=0.5$.
+Below each histogram is possible to verify the parameters used.}
+  \label{fig:13141516}
 \end{figure}
@@
  \centering
- \caption{Efficiency comparison of methods I and II for parameters covering uniformly the support of $B$.
+ \caption{Efficiency comparison of SM and VLB methods for parameters covering uniformly the support of $B$.
 Column $n$ shows the number of parameter combinations with which each method has superior compression.}
@@
   \hline
-  I	   & 592	& 0.9034\% \\
-  II	   & 64944	& 99.0966\% \\
+  SM	   & 592	& 0.9034\% \\
+  VLB	   & 64944	& 99.0966\% \\
   \hline
@@
  \end{tabular}
- \label{tab:06}
+ \label{tab:02}
 \end{table}
@@
   \centering
-  \subfigure[Efficiency histogram of the method I]{
-  \includegraphics[scale=0.4,angle=-90]{fig14}
-  \label{fig:14}
+  \subfigure[Efficiency histogram of the SM method ]{
+  \includegraphics[scale=0.25,angle=-90]{fig17}
+  \label{fig:17}
   }
-  \subfigure[Efficiency histogram of the method II]{
-  \includegraphics[scale=0.4,angle=-90]{fig15}
-  \label{fig:15}
+  \subfigure[Efficiency histogram of the VLB method ]{
+  \includegraphics[scale=0.25,angle=-90]{fig18}
+  \label{fig:18}
   }
-  \caption{Efficiency histograms of the methods I and II. Note that the method II has a greater average efficiency than method I.}
-  \label{fig:1415}
+  \caption{Efficiency histograms of the SM and VLB methods. Note that the VLB method has a greater average efficiency than SM method .}
+  \label{fig:1718}
 \end{figure}
 
-As we have shown, the VLB method is more effective compressing most integer datasets up to 64bits in size.
+As we have shown, the VLB method is more effective compressing most integer datasets up to 64 bits in size.
 This is due to its ability to exploit the variance in the data set and reduce the waste of bits in the representation of some numbers.
 In specific cases where the variance in the data null or too small, method I will be more efficient.
-As a matter of fact, for matrices where all elements have the same bit-length, method I will always be better, regardless of bit-length (figures \ref{fig18} and \ref{fig19}), The only exception if for bit-length 64 where neither method is able to compress the data.
+As a matter of fact, for matrices where all elements have the same bit-length, SM method  will always be better, regardless of bit-length (figures \ref{fig19} and \ref{fig20}), The only exception if for bit-length 64 where neither method is able to compress the data.
 
@@
   \centering
-  \subfigure[Efficiencies of the methods I e II, for constant bit-length matrices.]{\includegraphics[scale=0.3]{fig18}\label{fig18}}
-  \subfigure[$\eta_1-\eta_2$ for matrices of constant bit-length.]{\includegraphics[scale=0.3]{fig19}\label{fig19}}
-  \caption{Compression efficiency of the methods I and II  for matrices of constant bit-length.}
-  \label{fig:1819}
+  \subfigure[Efficiencies of the SM and VLB methods, for constant bit-length matrices.]{\includegraphics[scale=0.3,clip]{fig19}\label{fig19}}
+  \subfigure[$\eta_1-\eta_2$ for matrices of constant bit-length.]{\includegraphics[scale=0.3,clip]{fig20}\label{fig20}}
+  \caption{Compression efficiency of the SM and VLB methods for matrices of constant bit-length.}
+  \label{fig:1920}
 \end{figure}
@@
 As an example, let $M_{r\times c}$ be a integer matrix such that the half of its elements have bit-length $1$ and the other half $64$.
-Recalling equation \ref{eq:25}, now we have two groups of elements (by bit-length), $b_1=1 $, $b_2=64$ and $f_i = \frac{rc}{2}$ for $i = 1 \text{ and } 2$.
+Recalling Equation \ref{eq:17}, now we have two groups of elements (by bit-length), $b_1=1 $, $b_2=64$ and $f_i = \frac{rc}{2}$ for $i = 1 \text{ and } 2$.
 As the greatest bit-length is $64$, then $k=7$.
-Compression efficiency $\eta_2$ can be calculated using Equation \ref{eq:25}.
+Compression efficiency $\eta_2$ can be calculated using Equation \ref{eq:17}.
 After plugging in our numbers, we obtain a compression of $38.29\%$.
 
-\begin{equation*} %\label{eq:32}
+\begin{equation*}\label{eq:25}
  \eta_2 = 1 - \frac{\sum_{i=1}^{2}  b_i \times f_i }{64 \times rc} - \frac{7}{64} 
@@
 
-\begin{equation*} %\label{eq:33}
+\begin{equation*}\label{eq:26}
  \eta_2 = 1 - \frac{  1 \times \frac{rc}{2} + 64 \times \frac{rc}{2} }{64 \times rc} - \frac{7}{64} 
@@
 
-\begin{equation*} %\label{eq:34}
+\begin{equation*}\label{eq:26}
  \eta_2 = 1 - \frac{  32.5  \times rc }{64 \times rc} - \frac{7}{64} 
@@
 
-\begin{equation*} %\label{eq:35}
+\begin{equation*}\label{eq:27}
  \eta_2 \approx 1 - 0.5078 - 0.1093
@@
 
-\begin{equation*} %\label{eq:36}
+\begin{equation*}\label{eq:28}
  \eta_2 \approx 38.29\%
@@
 
-With this definition we can rewrite the equation \ref{eq:25}, which becomes \ref{eq:37}.
-In Equation \ref{eq:37}, the $\frac{f_i}{rc}$ is replaced by $p_i$, representing the probability of elements from group $i$ in matrix M.
+With this definition we can rewrite the Equation \ref{eq:17}, which becomes \ref{eq:29}.
+In Equation \ref{eq:29}, the $\frac{f_i}{rc}$ is replaced by $p_i$, representing the probability of elements from group $i$ in matrix M.
 
-\begin{equation}\label{eq:37}
+\begin{equation}\label{eq:29}
  \eta_2 = 1 - \frac{\sum_{i=1}^{g}  b_i \times p_i }{64} - \frac{k}{64} 
@@
 
-\begin{equation}\label{eq:38}
+\begin{equation}\label{eq:30}
  p_i = \frac{f_i}{rc}
@@
 
-With the equation \ref{eq:37} can analyze the influence of bit-length probability in compression efficiency.
+With the Equation \ref{eq:29} can analyze the influence of bit-length probability in compression efficiency.
 In this example, $p_1$ and $p_2$ represent the probability of elements of bit-lengths 1 and 64, respectively.
-Thus, efficiency is defined in Equation \ref{eq:40}.
+Thus, efficiency is defined in Equation \ref{eq:31}.
 
-\begin{equation}\label{eq:40}
+\begin{equation}\label{eq:31}
  \eta_2 = 1 - \frac{1 \times p_1  + 64 \times p_2}{64} - \frac{7}{64} 
@@
 When $\eta_2=1$, then the efficiency is maximal and if $\eta_2=0$, a efficiency is minimal.
-To calculate the values ​​of $p_1$ and $p_2$ for both extreme values of $\eta_2$,  we must solve the linear systems shown in equations \ref{eq:41} and \ref{eq:42}.
+To calculate the values ​​of $p_1$ and $p_2$ for both extreme values of $\eta_2$,  we must solve the linear systems shown in Equations \ref{eq:32} and \ref{eq:33}.
 The first equation on both systems come from the law of total probability.
-The second comes from \ref{eq:40} after setting $\eta_2$ to $1$ and $0$, respectively.
+The second comes from \ref{eq:31} after setting $\eta_2$ to $1$ and $0$, respectively.
 
-\begin{equation}\label{eq:41}
+\begin{equation}\label{eq:32}
   \eta_2 = 1 :\left
@@
 
-\begin{equation}\label{eq:42}
+\begin{equation}\label{eq:33}
   \eta_2 = 0 :\left
@@
 
-Thus, when $p_1=0.1111$ and $p_2=0.8889$ the efficiency is minimal for the second method.
-For other combinations see table \ref{tab:02}.
+Thus, when $p_1=0.1111$ and $p_2=0.8889$ the efficiency is minimal for the VLB method.
+For other combinations see Table \ref{tab:02}.
 Looking at this table, one can see two negative efficiencies, when ($p_1$,$p_2$) assume the values ​​(0,1) and (0.1,0.9).
@@
 
-When the distribution of the group probabilities is uniform, i.e. the groups have approximately the same size, efficiency is basically the same, regardless of the number of groups.
+When the distribution of the group probabilities is uniform, i.e., the groups have approximately the same size, efficiency is basically the same, regardless of the number of groups.
 
@@
 For all the experiments, we assume $k=7$, that is, the maximum possible bit-length is 64 bits.
-Thus, efficiency obtained will not be the best possible, since for that we would need assume small values of $k$ (equation \ref{eq:37}).
+Thus, efficiency obtained will not be the best possible, since for that we would need assume small values of $k$ (Equation \ref{eq:29}).
 
@@
 Then $E(b_i) = \sum_{i} b_i \times p(b_i) = \frac{a+b}{2}$.
-Applying this result to the expected compression efficiency of method II (equation \ref{eq:37}), we have
+Applying this result to the expected compression efficiency of VLB method (Equation \ref{eq:29}), we have
 
@@
 
-This result agrees with the numerical estimates presented in table \ref{tab:03}.
+This result agrees with the numerical estimates presented in Table \ref{tab:03}.
 
 \paragraph{Numerical Estimates:}
-To calculate the efficiency, we generated a matrices with 100 ($M_{10 \times 10}$), 10,000 ($M_{100 \times 100}$) and 1,000,000 ($M_{1,000 \times 1.000}$) elements .
-The average efficiency (table \ref{tab:03}) is calculated from a 1000 replicates of each matrix size.
+To calculate the VLB efficiency, we generated a matrices with 100 ($M_{10 \times 10}$), 10,000 ($M_{100 \times 100}$) and 1,000,000 ($M_{1,000 \times 1,000}$) elements with 1, 8, 16, 32 and 64 bit numbers.
+The average efficiency (Table \ref{tab:03}) is calculated from a 1,000 replicates of each matrix size.
+É possível verificar que quanto menor for o número de bits necessários para representar os elementos da matriz, maior será o nivel de compactação.
 
@@
   \centering
-  \caption{Compression efficiency of method II of samples with bit-lengths comming from a Discrete Uniform distribution $U(a=1,b=64)$. Average efficiency were calculated over a 1000 replicates.}
- \begin{tabular}{cc}
+  \caption{Compression efficiency of VLB method of samples with bit-lengths coming from a Discrete Uniform distribution $U(a=1,b=64)$.
+Average efficiency ($\overline{\eta_2}\pm\textrm{standard deviation}$) were calculated over a 1,000 replicates.}
+ \begin{tabular}{cccc}
     \hline
-    Matrix size & Avg. efficiency \\
+    & &Matrix sizes& \\
     \hline
-     100	& 0.3760938 \\
-     10000 	& 0.3795953 \\
-     1000000 	& 0.3826869 \\
+    & &Sample size & \\
+    Expected bit-length & 100 & 10,000 & 1,000,000 \\
+    \hline
+     1&0.8750$\pm$0.0000& 0.8750$\pm$0.0000&0.8750$\pm$0.0000\\ 
+     8&0.8202$\pm$0.0031& 0.8203$\pm$0.0003&0.8203$\pm$0.0000\\ 
+     16&0.7580$\pm$0.0066& 0.7578$\pm$0.0007&0.7578$\pm$0.0001\\ 
+     32&0.6330$\pm$0.0142& 0.6329$\pm$0.0014&0.6328$\pm$0.0001\\ 
+     64&0.3826$\pm$0.0288& 0.3828$\pm$0.0028&0.3828$\pm$0.0003\\ 
     \hline
@@
 Which again agrees with estimates in Table \ref{tab:04}.
+
 \paragraph{Numerical Estimates:}
@@
 As expected, smaller bit-lengths lead to higher compression efficiencies.
-However, for the case in which the elements require 64 bits, the efficiency is negative.
-[REMOVE AFTER FIXING TABLE]This result indicates that using the second method, when elements require 64 bit this becomes impractical, because it would be require more memory to allocate the compressed matrix.[REMOVE]
 
@@
   \caption{Compression efficiency with bit-lengths distributed according to a binomial distribution $B(n,0.5)$.
-Parameter $n \in \{1, 8, 16, 32, 64\}$ represents the maximum bit-length.
-Since $p=0.5$ the expected bit-length is $n/2$ (first column).[REMOVE LAST LINE SINCE WE CANNOT HAVE AN EXPECTED VALUE OF 64 and check for correct values]}
+Parameter $n \in \{1, 8, 16, 32, 64 \} $ represents the maximum bit-length.
+Since $p=0.5$ the expected bit-length is $n/2$ (first column)}
    \begin{tabular}{cccc}
      \hline
-			& 		&Efficiency         & \\
+     & &Efficiency& \\
      \hline
-			&  		&Sample size & \\
-    Expected bit-length $(n/2)$		& 100		& 1,000		    & 1,000,000 \\
+     & &Sample size& \\
+     Expected bit-length $(n/2)$ &100 &1,000 &1,000,000 \\
      \hline
-     1 			&  0.9760937	& 0.9764797 	    & 0.9765566 \\
-     8			&  0.8109375	& 0.8128453 	    & 0.8125079 \\
-     16			&  0.626875	& 0.6251859 	    & 0.6249333 \\
-     32			&  0.2476562	& 0.2504953 	    & 0.2499712 \\
-     64			&  -0.5004688	& -0.4999828 	    & -0.5000366 \\
+     1 &0.8828$\pm$0.0004&0.8828$\pm$0.0004&0.8828$\pm$0.0000\\ 
+     8 &0.8283$\pm$0.0022&0.8281$\pm$0.0002&0.8281$\pm$0.0000\\ 
+     16&0.7656$\pm$0.0032&0.7656$\pm$0.0003&0.7656$\pm$0.0000\\ 
+     32&0.6406$\pm$0.0045&0.6406$\pm$0.0004&0.6406$\pm$0.0000\\ 
+     64&0.3910$\pm$0.0065&0.3906$\pm$0.0006&0.3906$\pm$0.0001\\
      \hline
@@
 \end{table}
+
 \subsection*{Poisson Distribution}
@@
 \paragraph{Theoretical Efficiency:}
-Let bit-length $B \sim (\lambda=32)$. In this case,$E(b_i) = \lambda$, with $k=7$,  the efficiency becomes:
+Let bit-length $B \sim Poisson(\lambda=32)$. In this case,$E(b_i) = \lambda$, with $k=7$,  the efficiency becomes:
 
@@
 \paragraph{Numerical Estimates:}
-The results for this simulation can be seen in Table \ref{tab:05}.
+The results for this simulation can be seen in \ref{tab:05}.
 Note that increasing the bit number to represent numbers increases, there is a loss of efficiency in the compression process.
@@
    \centering
-  \caption{Compression efficiency with bit-lengths distributed according to a Poisson distribution ($\lambda$), where $\lambda$ represents the expected bit-length. [REMOVE LAST LINE]}
+   \caption{Compression efficiency with bit-lengths distributed according to a Poisson distribution ($\lambda$), where $\lambda$ represents the expected bit-length.}
   \begin{tabular}{cccc}
       \hline
-			& 	&Efficiency $\eta_2$      & \\
+      &&Efficiency $\eta_2$      & \\
       \hline
-			&  	&Matrix Size& \\
+      &&Matrix Size& \\
       Expected bit-length ($\lambda$)	& 100	& 1,000		    & 1,000,000 \\
       \hline
-     1	& 0.8746875 & 0.8749859   & 0.8749834 \\
-     8 	& 0.7657812 & 0.7656016   & 0.7656138 \\
-     16	& 0.6300000 & 0.6400594   & 0.6405952 \\
-     32	& 0.3851562 & 0.3903531   & 0.3903477 \\
-     64	& -0.0528125& -0.05984375 & -0.05953145 \\
+      1 	& 0.8751$\pm$0.0015 	& 0.8750$\pm$0.0004 & 0.8750$\pm$0.0000 \\ 
+      8 	& 0.7654$\pm$0.0046 	& 0.7656$\pm$0.0005 & 0.7656$\pm$0.0000 \\ 
+      16 	& 0.6405$\pm$0.0064 	& 0.6406$\pm$0.0006 & 0.6406$\pm$0.0001 \\ 
+      32 	& 0.3908$\pm$0.009 	& 0.3906$\pm$0.0009 & 0.3906$\pm$0.0001 \\ 
+      64 	& 0.4086$\pm$0.0456 	& 0.4073$\pm$0.0046 & 0.4072$\pm$0.0005 \\ 
       \hline
@@
 
-[devemos incluir um estudo sobre a variância?]
 \section{Conclusion}
-In this paper we have focused in the compression of matrix data, since this is one of the most important application the authors foresee.
+In this paper, we have focused in the compression of matrix data, since this is one of the most important application the authors foresee.
 However, the compression methodology presented can be applied to any numerical data structure, with gains to performance and memory footprint[citar tese Crysttian e possíveis artigos derivados].

### 2013-01-11 compilar.sh

@@ -1,3 +1,0 @@
-#/bin/bash
-
-ifort matriz.f90 -o matriz.x -xhost -i8 -O3 -parallel -fast

### 2013-01-11 manuscript.tex
correções com crysttian

@@
 Most if not all existing lossless compression methods were developed under the following usage paradigm: \textit{produce} $\rightarrow$ \textit{compress} $\rightarrow$ \textit{store} $\rightarrow$ \textit{uncompress} $\rightarrow$ \textit{use}.
-The focus of the present work is to allow a slightly different usage:  \textit{produce} $\rightarrow$ \textit{compress} $\rightarrow$ \textit{perform mathematical manipulations to use} and \textit{decompress} only for human reading.
+The focus of the present work is to allow a slightly different usage:  \textit{produce} $\rightarrow$ \textit{compress} $\rightarrow$ \textit{perform mathematical manipulations} and \textit{decompress} (only for human reading).
 
@@
 The remaining 4 bits will be used later.
-The number of bitstrings needed will be $\lfloor \frac{\dim(M) *b(\sup M)} {64} \rfloor +1$.
+The number of bitstrings needed will be $\lfloor \frac{\dim(M) *b(\sup M)} {64} \rfloor +1$, where $dim(M)$ is the dimension of the matrix or its number of elements.
 
@@
 \begin{enumerate}
- \item Element $M_{1,1}=900 = m_{1,1} = 1110000100$ is stored in the first 10-bit chunk of the element strip $bitstring[1]$ , which corresponds to bits 0 to 9.
+ \item Element $M_{1,1}=900 = m_{1,1} = 1110000100$ is stored in the first 10-bit chunk of the element strip $bitstring[1]$ , which corresponds to bits 0 to 9 (read from right to left).
 \begin{equation*}\label{eq:04}
@@
 
- Please note that bitstrings are written from right to left.
+ Please notice that bitstrings are written from right to left.
 
@@
  \item[4.] Element $M_{1,4} =256$ is added.  
- \item [5.] Elements $M_{1,5} =1$ and $M_{1,6} =10$ are added requiring a total of 13 bits instead of 20 with the SM method. Now, it is used 2 bit strips.
+ \item [5.] Elements $M_{1,5} =1$ and $M_{1,6} =10$ are added requiring a total of 13 bits instead of 20 with the SM method. With the addition of these elements we require a second bitstring.
  \begin{align*} \label{eq:12}
@@
 
-\begin{align} \label{eq:17}
+\begin{align}
  \eta_2 &= \frac{64 \times rc - \sum_{i=1}^{g} ( b_i \times f_i + k \times f_i )}{64 \times rc} \nonumber \\
   &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  -\sum_{i=1}^{g}  k \times f_i }{64 \times rc}\nonumber \\
-  &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  - k \times\sum_{i=1}^{g}  f_i }{64 \times rc}\nonumber \\
- &\textrm{Please note that }\sum_{i=1}^{g} f_i = rc, \nonumber \\
+  &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  - k \times\sum_{i=1}^{g}  f_i }{64 \times rc}\nonumber 
+\end{align}
+
+ Knowing that $$\sum_{i=1}^{g} f_i = rc, $$ we can simplify the equation above, obtaining (\ref{eq:17}).
+
+\begin{align}\label{eq:17}
  \eta_2 &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  - k \times rc }{64 \times rc}\nonumber \\
@@
 
-After sampling uniformly ($[1,5,9,\ldots,64]$, $n=65536$) the bit-length space and comparing efficiencies, we summarized the results on Table \ref{tab:02}.
+After sampling uniformly ($[1,5,9,\ldots,64]$, $n=65,536$) the bit-length space and comparing efficiencies, we summarized the results on Table \ref{tab:02}.
 In it we see how many parameterizations (from our sample) favor each method.
@@
   }
-  \caption{Histograms constructed from samples with 10,000 elements, generated from the combination of two distributions Betas with $\lambda=0.5$.
-Below each histogram is possible to verify the parameters used..}
+  \caption{Histograms constructed from samples with 10,000 elements, generated from the mixture of two Beta distributions with $w=0.5$. Below each histogram are the parameters of the mixture.}
   \label{fig:09101112}
@@
   }
-  \caption{Histograms constructed from samples with 10,000 elements, generated from the combination of two distributions Betas  with $\lambda=0.5$.
-Below each histogram is possible to verify the parameters used.}
+  \caption{Histograms constructed from samples with 10,000 elements, generated from the mixture of two Beta distributions with $w=0.5$. Below each histogram are the parameters of the mixture.}
   \label{fig:13141516}
@@
   \hline 
-  Methods  & n   & Method wins \\
+  Methods  & n   & Percentage \\
   \hline
@@
 The average efficiency (Table \ref{tab:03}) is calculated from a 1,000 replicates of each matrix size.
-É possível verificar que quanto menor for o número de bits necessários para representar os elementos da matriz, maior será o nivel de compactação.
+As expected the compression effiency gets better with lower expected bit-length.
 
@@
   \caption{Compression efficiency of VLB method of samples with bit-lengths coming from a Discrete Uniform distribution $U(a=1,b=64)$.
-Average efficiency ($\overline{\eta_2}\pm\textrm{standard deviation}$) were calculated over a 1,000 replicates.}
+Average efficiency ($\overline{\eta_2}\pm\textrm{SD}$) were calculated over a 1,000 replicates.}
  \begin{tabular}{cccc}
@@
 
-For the binomial distribution, we will use $Bin(n, p)$, with the number of trials $n$ representing the greatest possible bit-length in the matrix, with $np$ giving us an expected bit-length.
+For the binomial distribution, we will use $Bin(n, p)$, with the number of trials $n$ representing the greatest possible bit-length in the matrix, and $np$ giving us the expected bit-length.
 

### 2013-01-11 manuscript.tex
Veja as alteracoes para a simuacao da Poisson.

@@
 \\
-$\ast$ E-mail: Corresponding author@institute.edu
+$\ast$ E-mail: Corresponding crysttian.paixao@fgv.br
 \end{flushleft}
@@
 In digital computers, all information is stored as binary code (base 2 numbers).
-However the conventional way to store arrays of integers is on a memory block sequence of fixed size (power of 2 bit numbers), one for each element.
+However the conventional way to store arrays of integers is on a memory block sequence of fixed size (power of 2 numbers of bit), one for each element.
 The maximum size of a block is equal to the word size of the processor, which for most current CPUs is 64 bits.
@@
 
-The SM method consists in determining the value of the greatest element of matrix $M$, which coincides with its supremum, $max M == \sup M$  and determine the minimum bit number, $b(\sup M$)(Equation \ref{eq:01}), required to store it.
+The SM method consists in determining the value of the greatest element of matrix $M$, which coincides with its supremum, $max M == \sup M$  and determine the minimum number of bits, $b(\sup M$)(Equation \ref{eq:01}), required to store it.
 We will use capital roman letters to denote uncompressed matrices and the corresponding lower case letter for the compressed version.
@@
 Let's look at a concrete example: suppose that the greatest value to be stored in a matrix $M$ is $\max M=1023$.
-Therefore, the bit number required to represent it is $10$ $(1111111111)$.
+Therefore, the number of bits required to represent it is $10$ $(1111111111)$.
 Let the first 8 elements of $M$ be:
@@
 These elements of $M$, in binary, are shown in Table \ref{tab:01}.
-It is evident that the bit number required to represent any other element must be lower or equal to $10$.
+It is evident that the number of bits required to represent any other element must be lower or equal to $10$.
 From now on the minimum number of bits required to represent a base 10 integer will be refered to as its bit-length.
@@
 \paragraph{Numerical Estimates:}
-To calculate the VLB efficiency, we generated a matrices with 100 ($M_{10 \times 10}$), 10,000 ($M_{100 \times 100}$) and 1,000,000 ($M_{1,000 \times 1,000}$) elements with 1, 8, 16, 32 and 64 bit numbers.
+To calculate the VLB efficiency, we generated a matrices with 100 ($M_{10 \times 10}$), 10,000 ($M_{100 \times 100}$) and 1,000,000 ($M_{1,000 \times 1,000}$) elements with 1, 8, 16, 32 and 64 number of bitss.
 The average efficiency (Table \ref{tab:03}) is calculated from a 1,000 replicates of each matrix size.
@@
 \paragraph{Theoretical Efficiency:}
-Let bit-length $B \sim Poisson(\lambda=32)$. In this case,$E(b_i) = \lambda$, with $k=7$,  the efficiency becomes:
+Let bit-length $B \sim Poisson(\lambda=32)$. In this case, $E(b_i) = \lambda$, with $k=7$,  the efficiency becomes:
 
-\begin{align}\label{eq:49}
- E(\eta_2) &= 1 - \frac{\lambda}{64} - \frac{k}{64} \\
- &= 1 - \frac{32}{64} - \frac{7}{64} \approx 39.06\%
+\begin{align}
+ E(\eta_2) &= 1 - \frac{\lambda}{64} - \frac{k}{64} \label{eq:49} \\
+ &= 1 - \frac{32}{64} - \frac{7}{64} \approx 39.06\% \label{eq:50}
 \end{align}
@@
 The results for this simulation can be seen in \ref{tab:05}.
-Note that increasing the bit number to represent numbers increases, there is a loss of efficiency in the compression process.
+Note that increasing the number of bits to represent numbers increases, there is a loss of efficiency in the compression process.
 Another interesting fact is the emergence of a negative efficiency.
 In this case, for values ​​generated with a $\lambda=64$, the second method requires more memory to allocate the matrix formed by elements generated from this distribution.
+Para realizar essa simulação para $\lambda=64$, inúmeros valores seriam gerados acima de 64.
+O maior valor de $\lambda$, o qual faz com que a eficiênica seja mínima é 57.
+Esse valor pode ser calculado igualando a Equação \ref{eq:50} a 0.
+Logo, o maior valor esperado para o número de bits utilizado na Tabela \ref{tab:05} é o 57.
+(Eu trunquei a série que era gerada, para valores maiores que 64, eles eram convertidos para 64, e o resultado é o mesmo do 57)
 
@@
       16 	& 0.6405$\pm$0.0064 	& 0.6406$\pm$0.0006 & 0.6406$\pm$0.0001 \\ 
-      32 	& 0.3908$\pm$0.009 	& 0.3906$\pm$0.0009 & 0.3906$\pm$0.0001 \\ 
-      64 	& 0.4086$\pm$0.0456 	& 0.4073$\pm$0.0046 & 0.4072$\pm$0.0005 \\ 
+      32 	& 0.3908$\pm$0.0087 	& 0.3906$\pm$0.0009 & 0.3906$\pm$0.0001 \\ 
+      57 	& 0.0121$\pm$0.0099 	& 0.0118$\pm$0.0010 & 0.0118$\pm$0.0001 \\
       \hline

### 2013-01-14 manuscript.tex
adicionei pequenas correções do Crysttian

@@
 This is a most common data structure found in computational data analysis environments.
-The matrices compressed according to the methodologies proposed here should be able to undergo the same mathematical operations as the original uncompressed matrices.
-e.g.
+The matrices compressed according to the methodologies proposed here should be able to undergo the same mathematical operations as the original uncompressed matrices, e.g.
 linear algebra manipulations.
@@
  \end{equation*}
- \item Element 7 does not fit on the remaining 4 bits of the first bitstring.
+ \item Element $M_{1,7}=700=1010111100$ does not fit on the remaining 4 bits of the first bitstring.
 So it will straddle two bitstrings, i.e., it is divided in two segments $a$ and $b$, $a$ is written on the first bitstring and $b$ on the second.
@@
  \subsubsection*{SM Method}
- Let $M_{r \times c}$ be the matrix we wish to compress. In comparison with a conventional allocation (64-bit integers) we can apply Equation \ref{eq:14} to caluculate the efficiency of the SM method:
+ Let $M_{r \times c}$ be the matrix we wish to compress.
+In comparison with a conventional allocation (64-bit integers), we can apply Equation \ref{eq:14} to caluculate the efficiency of the SM method:
 
@@
 As we see in \ref{eq:15}, $\eta_1$ does not depend on size of the matrix, only on the bit-length of $\max M$.
-If $b(\max M)=64$, $\eta_1$ is 0, i.
-e., no compression is possible.
+If $b(\max M)=64$, $\eta_1$ is 0, i.e., no compression is possible.
 On the other extreme, if the matrix is composed exclusively of 0s and 1s, maximal compression is achievable, $\eta_1=1$.
@@
 
-\noindent where k in \ref{eq:21}, is set to 7 (the bit-length required to represent the largest possible bit-length: 64).
-In \ref{eq:20}, $k$ is the bit-length of the greatest element, or in the worst case, 64.
+\noindent where k, in (\ref{eq:21}), is set to 7 (the bit-length required to represent the largest possible bit-length: 64).
+In (\ref{eq:20}), $k$ is the bit-length of the greatest element, or in the worst case, 64.
 
@@
 It can be seen that for most combination of $\alpha$ and $\beta$, $D<0$, meaning the second method is more efficent to compress a sample of numbers with bit-lengths coming from a $Beta(\alpha,\beta)$ distribution.
-However, there is a small region in parameter space, which is shown in white on (d), where the first method is more efficient.
+However, there is a small region in parameter space, which is shown in white on (d), where the SM method is more efficient.
 This region corresponds to the dots in red in (a), where the average bit-length is higher.
@@
 In it we see how many parameterizations (from our sample) favor each method.
-We can also look at the distribution of efficiencies on our samples for each method (figure \ref{fig:1718}), which clearly demonstrate the greater expected efficiency of method VLB (figure \ref{fig:18}).
+We can also look at the distribution of efficiencies on our samples for each method (Figure \ref{fig:1718}), which clearly demonstrate the greater expected efficiency of method VLB (Figure \ref{fig:18}).
 
@@
 In specific cases where the variance in the data null or too small, method I will be more efficient.
-As a matter of fact, for matrices where all elements have the same bit-length, SM method  will always be better, regardless of bit-length (figures \ref{fig19} and \ref{fig20}), The only exception if for bit-length 64 where neither method is able to compress the data.
+As a matter of fact, for matrices where all elements have the same bit-length, SM method  will always be better, regardless of bit-length (Figures \ref{fig19} and \ref{fig20}), The only exception if for bit-length 64 where neither method is able to compress the data.
 
@@
      & &Sample size& \\
-     Expected bit-length $(n/2)$ &100 &1,000 &1,000,000 \\
+     Expected bit-length $(np)$ &100 &1,000 &1,000,000 \\
      \hline
@@
 Note that increasing the number of bits to represent numbers increases, there is a loss of efficiency in the compression process.
-Another interesting fact is the emergence of a negative efficiency.
-In this case, for values ​​generated with a $\lambda=64$, the second method requires more memory to allocate the matrix formed by elements generated from this distribution.
-Para realizar essa simulação para $\lambda=64$, inúmeros valores seriam gerados acima de 64.
-O maior valor de $\lambda$, o qual faz com que a eficiênica seja mínima é 57.
-Esse valor pode ser calculado igualando a Equação \ref{eq:50} a 0.
-Logo, o maior valor esperado para o número de bits utilizado na Tabela \ref{tab:05} é o 57.
-(Eu trunquei a série que era gerada, para valores maiores que 64, eles eram convertidos para 64, e o resultado é o mesmo do 57)
+In this case we did not simulate for $\lambda=64$ since a large portion of the samples would fall above the maximum bit-length we are considering for this analysis.
 
@@
       32 	& 0.3908$\pm$0.0087 	& 0.3906$\pm$0.0009 & 0.3906$\pm$0.0001 \\ 
-      57 	& 0.0121$\pm$0.0099 	& 0.0118$\pm$0.0010 & 0.0118$\pm$0.0001 \\
       \hline
@@
 Further discussions about doing computation with such compressed data-structures will be the subject of another manuscript (in preparation) in which we will present details about the implementation of the compression algorithm, and benchmarks on classical linear algebra tasks such as those in Linpack[ref].
+
+For the compression calculations presented in this paper we limited bit-lenght of integers to 64 bits.
+However the compression would work in the same way as discussed for computer architectures with larger word sizes.
 

### 2013-02-20 suppl.tex
adicionei o material suplementar em português

%
% Hello! Here's how this works:
%
% You edit the source code here on the left, and the preview on the
% right shows you the result within a few seconds.
%
% Bookmark this page and share the URL with your co-authors. They can
% edit at the same time!
%
% You can upload figures, bibliographies, custom classes and
% styles using the files menu.
%
% If you're new to LaTeX, the wikibook at
% http://en.wikibooks.org/wiki/LaTeX
% is a great place to start, and there are some examples in this
% document, too.
%
% We're still in beta. Please leave some feedback using the link at
% the top left of this page. Enjoy!
%
\documentclass[12pt]{article}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}

\usepackage[normalem]{ulem}

\title{Your Paper}
\author{You}

\begin{document}
\maketitle

\begin{abstract}
Your abstract.
\end{abstract}

\section{Introduction}

Inicialmente vamos considerar as operações elementares (soma, subtração, divisão
e multiplicação) a serem realizadas nos elementos armazenados na matriz de
bitstring. Sendo essa matriz gerada pelos métodos SM ou VLB. Ressalta-se que os
números e resultados das operações são descritos no formato inteiro.

Para realizar essas operações torna-se necessário avaliar o comportamento dos
bits dos números envolvidos. Quando uma das operações mencionadas é utilizada,
os bits dos números envolvidos são manipulados para chegar ao resultados final.
Em um computador essas operações são realizadas na base binária, diferente do
caso tradicional, no qual para fazer uma conta, a realizamos na base decimal. A
principal característica que deve-se considerar é a quantidade de bits
nacessários para armazenar o valor resultante da operação. Para isso, vamos
utilizar exemplos, de tal forma que as operações gerem resultados, os quais
demandem a maior quantidade de bits.

A primeira operação a ser avaliada é a soma. Na base binária o comportamento do
carry é o mesmo da decimal, porém o que muda é a base. A soma 1+1 resulta em 0 e
gera um carry de 1. Em decimal, seria a mesma coisa de realizar a operação 5 +
5, que gera um carry de 1 para ser somado a dezena.  As demais contas são $1 + 0
= 1$, $0 + 1 = 1$ e $0 + 0 = 0$. Considere $n_1=7$ e $n_2=8$, a soma em decimal
e binário é realizada na Tabela \ref{tab:01}. Observe que quando ocorre a soma
$1 + 1$, gera-se um carry para ser adicionado a soma dos bits à esquerda. Ao
todo foram gerados 3 bits de carry. Já na Tabela \ref{tab:02}, podemos verificar
a soma de $7 + 7$, que gera um número com 4 bits. Conclui-se que o resultado
soma, demanda 1 bit a mais do que o número de bits para representar o maior
valor envolvido na operação. Para representar o resultado, nesse caso o número
7, demanda 3 bits,logo o resultado demanda 4. Isso só é válido para o pior caso,
note que 7 demanda 3 bits, o qual é adicionado a 7, que também demanda 3 
bits, logo resultado irá demandar 4 bits. Pode ocorrer que na soma, a mesma
quantidade de bits do maior elemento, por exemplo a soma 1 + 2, seja mentida,
sem o acréscimo de 1 bits ao resultado. Por exemplo, o número 1 demanda 1 bit, o
número 2, 2 bits, o resultado 3, demanda 2 bits.

\begin{table}[ht]
	\centering
    \caption{Comparacao da soma em base decimal e binaria.}
    \begin{tabular}{crrrc}
    \hline
    	  & Decimal & & Binario & Numero de Bits \\
    \hline      
    Carry & 	& & \textit{11100} & \\
    $n_1$ & 7   & & 111 & 3 \\
    $n_2$ & \underline{10}	& & \underline{1010} & 4 \\
    Total & 17	& & 10001 & 5 \\
    \hline
	\end{tabular}
    \label{tab:01}
\end{table}

\begin{table}[ht]
	\centering
    \caption{Comparacao da soma em base decimal e binaria.}
    \begin{tabular}{crrrc}
    \hline
    	  & Decimal & & Binario & Numero de Bits \\
    \hline      
    Carry & 	& & \textit{1110} & \\
    $n_1$ & 7   & & 111 & 3 \\
    $n_1$ & \underline{~7}	& & \underline{111} & 3 \\
    Total & 14	& & 1110 & 4 \\
    \hline
	\end{tabular}
    \label{tab:02}
\end{table}

A próxima operação a ser representada é a subtração. Dependendo dos números
envolvidos na operação, o resultado pode ser negativo. Com isso, gera-se a
demanda de armazenar os números em um formato que contemple o sinal. A forma
mais utilizada para realizar isso é o complemento de 2 - C2 (ref). O número
representado nessa forma, demanda 1 bit a mais para representar o seu sinal.
Vamos realizar a operação $n_2 - n_1$, que pode ser representada por $n_2 +
(-n_1)$. Isso permite que as mesmas instruções para a soma, possam ser
utilizadas para a subtração. Para que isso possa ser feito, torna-se necessário
representar os números envolvidos no formato C2. Para passar um número para o
formato C2, deve-se seguir seguintes passos:

\begin{itemize}
	\item Representa-se o número com um bit a mais que o necessário, por
exemplo o número 7, demanda 3 bits, no formato C2, demandará 4 bits. Como o -7
será adicionado ao 10, os números devem ser representados por 5 bits ( 4 bits
para representar o 10 mais um bit para o sinal). Logo, 7 será representados por
00111. Aqui o número está no formato C1. Esse é outro formato utilizado, o
problema é que o número 0 possui duas representações nesse formato. Essas duas
representações gera uma demanda para verificar qual dos formatos de zeros está
presente, já o C2 possui apenas uma.
    \item Inverte-se os valores dos bits do número 7, nesse caso de 00111 para
11000
	\item Ao resultado da inversão soma-se 1. Nesse caso 11000 + 1 irá
resultar em 11001. Logo 11001 é a representação para -7 no formato C2.
\end{itemize}

Uma vez realizada a conversão, basta executar o processo de soma como já
mencionado. O resultado está representado na Tabela \ref{tab:03}. Observe que é
executado uma soma com os números envolvidos. Ambos demandam 5 bits para serem
representados por causa do C2. O mais importante é o bit, que é gerado mais à
esquerda pelo carry no resultado, é descartado(overflow, número 1 cortado). Isso
ocorre pois os números envolvidos na operação demandam 5 bits e não 6 bits. Para
usar o formato C2, deve-se converter os números para o mesmo tamanho.

Essa operação seria possível, sem que fosse utilizado o C2, pois 10 é maior que
7, logo não demandaria sinal. A operação está representada na Tabela
\ref{tab:04}. Observe que ocorre um carry para compensar os valores. Na operação
0 - 1, demanda um carry do número a esquerda, esse carry tem o valor de 2, ao
ser removido o valor 1, resultado em 1.

\begin{table}[ht]
	\centering
    \caption{Comparacao da soma em base decimal e binaria.}
    \begin{tabular}{crrrc}
    \hline
    	  & Decimal & & Binario & Numero de Bits \\
    \hline      
    Carry & 	& & \textit{110000} & \\
    $n_2$ & 10	& & 01010 & 5 \\
    $+(-n_1)$ & \underline{+(-7)} & & \underline{11001} & 5 \\
    Total & 3	& & \sout{1}00011 & 5 \\
    \hline
	\end{tabular}
    \label{tab:03}
\end{table}

\begin{table}[ht]
	\centering
    \caption{Comparacao da soma em base decimal e binaria.}
    \begin{tabular}{crrrc}
    \hline
    	  & Decimal & & Binario & Numero de Bits \\
    \hline      
    Carry & 	& & \textit{0111} & \\
    $n_2$ & 10	& & 1010 & 4 \\
    $-n_1$ & \underline{-7} & & \underline{111} & 3 \\
    Total & 3	& & 11 & 2\\
    \hline
	\end{tabular}
    \label{tab:04}
\end{table}

Para a subtração, o resultado no pior caso, irá demandar o mesmo número de bits
do maior valor envolvido na operação, considerando o formato C2 ou não. Como
exemplo, considere $n_3=1$, logo o resutlado da operação $n_2 - n_1$ irá
demandar 3 bits, como pode ser verificado na Tabela \ref{tab:05}.

\begin{table}[ht]
	\centering
    \caption{Comparacao da soma em base decimal e binaria.}
    \begin{tabular}{crrrc}
    \hline
    	  & Decimal & & Binario & Numero de Bits \\
    \hline      
    Carry & 	& & \textit{000} & \\
    $n_1$ & 7	& & 111 & 3 \\
    $-n_3$ & \underline{-1} & & \underline{1} & 1 \\
    Total & 6	& & 110 & 3\\
    \hline
	\end{tabular}
    \label{tab:05}
\end{table}

A multiplicação pode ser realizada da mesma forma que a soma e a subtração. Ela
nada mais é que o resultado de múltiplas somas. Por exemplo, a operação $2
\times 3 = 3 + 3 = 2 + 2 + 2$, ou seja 3 somado 2 vezes ou 2 somados 3 vezes.
Vamos calcular a multiplicação $n_1 \times n_2$, a qual a operação está
representada na Tabela \ref{tab:06}. Observe que são realizadas duas a duas, as
somas, da mesma forma que o computador realiza a operação.

\begin{table}[ht]
	\centering
    \caption{Comparacao da soma em base decimal e binaria.}
    \begin{tabular}{crrrc}
    \hline
    	  & Decimal & & Binario & Numero de Bits \\
    \hline      
    $n_2$ & 10	& & 1010 & 3 \\
    $n_1$ & \underline{$\times 7$} & & \underline{111} & 1 \\
    & & & 1010  & \\
    & & & \underline{1010\ } & \\
    & & & $11110$ & \\
    & & & \underline{1010\ \ } & \\
    Total& 70 & & 1000110 & 7\\
    \hline
	\end{tabular}
    \label{tab:06}
\end{table}

Na multiplicação o comportamento do número de bits é diferente da soma e da
subtração. No pior caso, o resultado é representado com o número de bits
resultante da soma do número de bits de cada número envolvido na operação. Na
operação $n_2 \times n_2$, ou seja $7 \times 7$, o resultado da operação, 49,
demanda 6 bits para ser representado, pois 49 em binário é 110001.

Caso na multiplicação esteja envolvido algum número com sinal, torna-se
necessário utilizar o complemento C2. Para ilustrar essa operação, vamos
realizar a operação $n_2 \times -n_1$. Lembrando que 10 e -7, no formato C2, são
 iguais à 01010 e 11001, respectivamente. Na Tabela \ref{tab:07} é possível
verificar a operação. A diferença básica é que os bits de sinal (em negrito),
posicionados mais à esquerda, não são utilizados para executar as somas
sucessivas. Apenas no final, esses bits são utilizados. Como 0 representa o
positivo e 1, o negativo, o resultado terá sinal negativo. O resto das operações
são iguais as operações já comentadas anteriormente. O número de bits
necessários para representar o resultado, no pior caso, é igual a soma do número
de bits envolvidos na multiplicação, sem contar o sinal, mais um bit, em
decorrência do sinal do resultado.

\begin{table}[ht]
	\centering
    \caption{Comparacao da soma em base decimal e binaria.}
    \begin{tabular}{crrrc}
    \hline
    	  & Decimal & & Binario & Numero de Bits \\
    \hline      
    $n_2$ & 10	& & \textbf{0}1010 & 5 \\
    $\times -1n_1$ & \underline{$\times -7$} & & \underline{\textbf{1}1001} & 5
\\
    & & & 1010  & \\
    & & & \underline{0000\ } & \\
    & & & $01010$ & \\
    & & & \underline{0000\ \ } & \\
    & & & 001010 & \\
    & & & \underline{1010\ \ \ } & \\
    Total& -70 & & \textbf{1}1011010 & 8\\
    \hline
	\end{tabular}
    \label{tab:07}
\end{table}

A divisão, ao contrário da multiplicação, demanda que seja executada sucessivas
subtrações até encontrar um determinado resto. Quando exato, esse resto vale
zero. Para ilustar o comportamento dos bits, considere $n_4=50$ na operação $n_4
/ n_2$. O resultado está descrito na Tabela \ref{tab:08}. Observe que ocorre
sucessivas subtrações em binário. O número de bits necessário para representar o
resultado da divisão, no pior caso, é igual o número de bits do dividento,
quando o dividendo é maior que o divisor. A cada subtração executada, soma-se 1
ao quociente, que inicialmente vale 0. O processo de subtração acaba quando o
valor do resto é zero.

\begin{table}[ht]
	\centering
    \caption{Comparacao da soma em base decimal e binaria.}
    \begin{tabular}{crrrrr}
    \hline
    	  & Decimal & & Binario & Quociente Decimal & Quociente Binario \\
    \hline  
    $n_4$ & 50 					& & 110010
&0& 000 \\
    $n_2$ & \underline{-10} 	& & \underline{001010}  &\underline{+1}&
\underline{+001} \\
    Resto & 40      			& & 101000  			&1& 001
\\
    $n_2$ & \underline{-10} 	& & \underline{001010}  &\underline{+1}&
\underline{+001} \\
    Resto & 30      			& & 011110  			&2& 010
\\
    $n_2$ & \underline{-10} 	& & \underline{001010}  &\underline{+1}&
\underline{+001} \\
    Resto & 20      			& & 010100  			&3& 011
\\
    $n_2$ & \underline{-10} 	& & \underline{001010}  &\underline{+1}&
\underline{+001} \\
    Resto & 10      			& & 001010  			&4& 100
\\
    $n_2$ & \underline{-10} 	& & \underline{001010}  &\underline{+1}&
\underline{+001} \\
    Resto & 0      				& & 000000
&5& 101 \\
    \hline
	\end{tabular}
    \label{tab:07}
\end{table}

O mesmo ocorre quando o dividendo ou dividor possuem sinal. Para efetuar a
divisão com números que possuam sinal, torna-se necessário utilizar o C2. Vamos
ilustrar o processo, considerando a divisão $-n_4 / n_2$. Quando o dividendo vai
ser subtraído o valor do divisor, quando no formato C2, o divisor deve ser
convertido para C2. E ao invés de usar a subtração, realiza-se a soma. E como
feito no caso da divisão sem sinal, a operação de soma termina, quando o resto é
igual a zero. O processo de divisão de número com sinais pode ser visualizado no
Tabela \ref{tab:08}. O quociente é 5, mas devido a diferença de sinais, esse
deve ser convertido para -5.

\begin{table}[ht]
	\centering
    \caption{Comparacao da soma em base decimal e binaria.}
    \begin{tabular}{crrrrr}
    \hline
    	  & Decimal & & Binario & Quociente Decimal & Quociente Binario \\
    \hline  
    $n_4$ & -50 					& & 1001110
&0& 000 \\
    $n_2$ & \underline{+10} 	& & \underline{001010}  &\underline{+1}&
\underline{+001} \\
    Resto & -40      			& & 1011000  			&1& 001
\\
    $n_2$ & \underline{+10} 	& & \underline{001010}  &\underline{+1}&
\underline{+001} \\
    Resto & -30      			& & 1100010 			&2& 010
\\
    $n_2$ & \underline{+10} 	& & \underline{001010}  &\underline{+1}&
\underline{+001} \\
    Resto & -20      			& & 1101100  			&3& 011
\\
    $n_2$ & \underline{+10} 	& & \underline{001010}  &\underline{+1}&
\underline{+001} \\
    Resto & -10      			& & 1110110  			&4& 100
\\
    $n_2$ & \underline{+10} 	& & \underline{001010}  &\underline{+1}&
\underline{+001} \\
    Resto & 0      				& & 0000000
&5& 101 \\
    \hline
	\end{tabular}
    \label{tab:08}
\end{table}

Apresentamos até agora as quatro operações elementares, as quais verificamos o
comportamento dos bits. Resumidamente podemos inferir sobre o comportamento do
número de bits, sem sinal, no pior caso que reflete a utilização de bits, da
seguinte forma:

\begin{itemize}
\item Soma: demanda 1 bits a mais que o número de bits para representar o maior
valor presente na operação;
\item Subtração: demanda a mesma quantidade de bits para representar o do maior
valor presente na operação;
\item Multiplicação: demanda é igual a soma dos bits dos números envolvidos na
operação;
\item Divisão: demanda a mesma quantidade de bits do dividendo;
\end{itemize}

Quando os números apresentam sinal, eles demandam 1 bits a mais para o sinal.
Nesse caso, na multiplicação, para o pior caso, o número de bits para
representar o resultando será a soma do número de bits dos números envolvidos,
menos 1.

Até o momento, as operações elementares foram executas apenas com números. Agora
será demonstrado a sua aplicação sobre as tiras de bits. Essas tiras de bits são
geradas pelos métodos SM ou VLB, os quais promovem a compactação dos dados e
armazenam diversos números com economia de espaço em tiras de bits.

Antes de demonstrarmos o procedimento, algumas regras devem ser contempladas
para que as operações possam ser realizadas. A primeira é que quando existir um
número com sinal, todos os números deverão ser representados na forma de
complemento de 2, como consequencia, existirá a demanda de 1 bit para cada
número armazenado na tira. As operações só podem resultar números inteiros, pois
a forma de armazenamento discutida atende apenas essa configuração.

Para demonstrar a aplicação envolvendo tiras de bits, vamos considerar duas
matrizes A e B, ambos as dimensões $3 \times 4$. Sendo definidas por:

\begin{equation}
	A = \begin{bmatrix}
			1 & 3 & 5 & 8\\ 
			12 &14  & 6 & 9\\ 
			3 & 7 & 10 & 11
		\end{bmatrix}
\end{equation}

e

\begin{equation}
	B = \begin{bmatrix}
			1  & 3  & 7 & 9\\ 
			1  &3   & 14 & 15\\ 
			20 & 30 & 2 & 1
		\end{bmatrix}
\end{equation}

Após o processo de compressão, as matrizes A e B, antes com dimensão $3 \times
4$, passam a ter dimensão $3 \times 1$. A matriz A e B após a compressão pelo
método SM, podem ser visualizada em decinal e binário nas Equações \ref{eq:01} e
\ref{eq:02}. As tiras possuem tamanhos de 4 bits e 5 bits, para as regiões de
armazenamento, para a matriz A e B, respectivamente. Os "$\ldots$" representam
os zeros omitidos em cada tira, pois os elementos são armazenados apenas em uma
tira com 64 bits, na qual apenas alguns bits são utilizados, restando apenas
zeros em desuso.

\begin{equation}\label{eq:01}
	AA = \begin{bmatrix}
			4952\\ 
			52841\\ 
			14251
		\end{bmatrix} 
        =
        \begin{bmatrix}
			\ldots0001001101011000\\ 
			\ldots1100111001101001\\ 
			\ldots0011011110101011
		\end{bmatrix}
        =
        \begin{bmatrix}
\ldots\underbrace{0001}_{1}\underbrace{0011}_{3}\underbrace{0101}_{5}\underbrace
{1000}_{8}\\
\ldots\underbrace{1100}_{12}\underbrace{1110}_{14}\underbrace{0110}_{6}
\underbrace{1001}_{9}\\ 
\ldots\underbrace{0011}_{3}\underbrace{0111}_{7}\underbrace{1010}_{10}
\underbrace{1011}_{11}
		\end{bmatrix}
\end{equation}

e

\begin{equation}\label{eq:02}
	BB = \begin{bmatrix}
			36073\\ 
			36303\\ 
			686145
		\end{bmatrix}
        =
         \begin{bmatrix}
			\ldots00001000110011101001\\ 
			\ldots00001000110111001111\\ 
			\ldots10100111100001000001
		\end{bmatrix}
        =
        \begin{bmatrix}

\ldots\underbrace{00001}_{1}\underbrace{00011}_{3}\underbrace{00111}_{7}
\underbrace{01001}_{59}\\
\ldots\underbrace{00001}_{1}\underbrace{00011}_{3}\underbrace{01110}_{14}
\underbrace{01111}_{15}\\
\ldots\underbrace{10100}_{20}\underbrace{11110}_{30}\underbrace{00010}_{2}
\underbrace{00001}_{1}
		\end{bmatrix}
\end{equation}

Considere apenas a tira armazenada na primeira linha de cada matriz (AA e BB).
Observe que para obter os valores de ambas as tiras, torna-se necessário
conhecer o tamanho da região que os valores foram armazenados. Isso ocorre, pois
elas foram geradas pelo método SM, o qual determina essa região. Para obter
esses valores, de forma otimizada, torna-se necessário utilizar o que
denominamos de máscara. A máscará irá permitir obter os valores armazenados
dentro das tirar, de forma otimizada. Para esse exemplo, para otimizar as
operações, a máscara irá coletar os valores da primeira e terceira região ao
mesmo tempo, o que otimiza o processo. A máscara é armazenada em uma tira com o
mesmo tamanho da tira que está armazenada nas matrizes. A máscara mencionada,
para ser aplicada na matriz AA, está representada na Equação \ref{eq:06}. Para
coletar as informações da segunda e quarta região da matriz AA, deve-se usar a
máscara descrita na equação \ref{eq:07}. Já para a matriz BB deve-se usar as
máscaras das equações \
ref{eq:08} e \ref{eq:09}, para coletar as informações das primeira e terceira
regiões e da segunda e quarta regiões, respectivamente.

\begin{equation}\label{eq:06}
\text{máscara da matriz AA para posições 1 e 3} = \ldots0000111100001111
\end{equation}

\begin{equation}\label{eq:07}
\text{máscara da matriz AA para posições 2 e 4} = \ldots1111000011110000
\end{equation}

\begin{equation}\label{eq:08}
\text{máscara da matriz BB para posições 1 e 3} = \ldots00000111110000011111
\end{equation}

\begin{equation}\label{eq:09}
\text{máscara da matriz BB para posições 2 e 4} = \ldots11111000001111100000
\end{equation}

Uma vez definida a máscara, vamos aplicá-la na matriz A e B, para obter os
valores armazenados na primeira e terceira regiões. Para usar a máscará, deve-se
utilizar a função booleana AND. Vamos chamar de tempA e tempB, os valores
coletados. O processo de coleta está ilustrado nas equações \ref{eq:10} e
\ref{eq:11}. As região está destacadas. Verifique que apenas as regiões, na qual
a máscara possui apenas 1 são preservadas, isso faz com que apenas esses valores
sejam obtidos.

%\begin{equation}\label{eq:10}
  \begin{align*}
   A[1,1]			&=	\ldots0001|0011|0101|1000\\
   \text{máscara A}	&=	\ldots0000|1111|0000|1111\\
   tempA 			&=	\ldots0000|0011|0000|1000
  \end{align*}
%\end{equation}

%\begin{equation}\label{eq:11}
  \begin{align*}
   B[1,1]			&= \ldots00001|00011|00111|01001 \\
   \text{máscara B}	&= \ldots00000|11111|00000|11111 \\
   tempB 			&= \ldots00000|00011|00000|01001
  \end{align*}
%\end{equation}

Suponha que desejamos calcular a soma das duas matrizes e armazená-las em uma
matriz S. Para isso as regiões das tiras tempA e tempB devem ter o mesmo
tamanho. Como para a soma, na pior das hipóteses, demanda-se 1 bit a mais que o
maior valor armazenado, aqui esse valor é representado pela região na qual os
valores são armazenados. Nesse caso, a região da tira resultante irá ter tamanho
de 6 bits. As tiras, com as regiões corrigidas, são representadas nas equações
\ref{eq:12} e \ref{eq:13}. Os bits inseridos para compensar o tamanho das
regiões estão em negrito.

\begin{equation}\label{eq:12}
tempA =  \ldots\textbf{00}0000\textbf{00}0011\textbf{00}0000\textbf{00}1000
\end{equation}

\begin{equation}\label{eq:13}
tempB =  \ldots\textbf{0}00000\textbf{0}00011\textbf{0}00000\textbf{0}01001
\end{equation}

Com as regiões igualadas, basta executar a operacão $S[1,1] = tempA + tempB$.
Realizando o mesmo processo com tempA e tempB, mas na segunda e quarta região,
podemos efetuar a operação soma de forma paralela, pois torna-se possível
realizar 2 operações em paralelo, e o melhor, o resultado, nesse caso aa tira
S[1,1], já está na forma compactada.

\begin{align*}
 tempA&= \ldots000001|000011|000101|001000\\
 tempB&= \ldots000001|000011|000111|001001\\
 S[1,1]&=\ldots000010|000110|001100|010001
\end{align*}

Na equação X, pode-se observar o resultado da soma para S[2,1] e S[3,1].

\begin{align*}
 tempA&= \ldots001100|001110|000110|001001\\
 tempB&= \ldots000001|000011|001110|001111\\
 S[2,1]&=\ldots001101|010001|010100|011000
\end{align*}

\begin{align*}
 tempA&= \ldots000011|000111|001010|001011 \\
 tempB&= \ldots010100|011110|000010|000001 \\
 S[3,1]&=\ldots010111|100101|001100|001100
\end{align*}

Logo, como resultado da operação S= A + B, temos:

\begin{equation}
	S = \begin{bmatrix}
			549649\\ 
			274889\\ 
			6181644
		\end{bmatrix}
        =
         \begin{bmatrix}
			\ldots000010000110001100010001\\
			\ldots000001000011000111001001\\
			\ldots010111100101001100001100
		\end{bmatrix}
        =
        \begin{bmatrix}

\ldots\underbrace{000010}_{2}\underbrace{000110}_{6}\underbrace{001100}_{12}
\underbrace{010001}_{17}\\	
 \ldots\underbrace{001101}_{13}\underbrace{010001}_{17}\underbrace{010100}_{20}
\underbrace{011000}_{24}\\
 \ldots\underbrace{010111}_{23}\underbrace{100101}_{37}\underbrace{001100}_{12}
\underbrace{001100}_{12}
		\end{bmatrix}
\end{equation}

A operação de subtração pode ser realizada, mas para isso será utilizado a
notação de complemento 2 - C2. Suponha que se deseja calcular $D = B - A$. Essa
operação pode ser realizada da mesma forma, utilizando uma operação de soma $D =
B + (-A)$, em que $-A$ será a matriz na forma C2. A matriz A na forma de
complemento de 2, pode ser visualizada na Equação X. Ressalta-se que a os
elementos serão calculados com 6 dígitos para serem armazenados nas tiras.

\begin{equation}
	-A = \begin{bmatrix}
			4952\\ 
			52841\\ 
			14251
		\end{bmatrix} 
        =
        \begin{bmatrix}
   			\ldots111111111101111011111000\\ 
			\ldots110100110010111010110111\\ 
			\ldots111101111001110110110101
		\end{bmatrix}
        =
        \begin{bmatrix}
 \ldots\underbrace{111111}_{-1}\underbrace{111101}_{-3}\underbrace{111011}_{-5}
\underbrace{111000}_{-8}\\
\ldots\underbrace{110100}_{-12}\underbrace{110010}_{-14}\underbrace{111010}_{-6}
\underbrace{110111}_{-9}\\
\ldots\underbrace{111101}_{-3}\underbrace{111001}_{-7}\underbrace{110110}_{-10}
\underbrace{110101}_{-11}
		\end{bmatrix}
\end{equation}

Para realizar a operação  $D = B + (-A)$, basta usar uma máscara para adequar as
regiões que armazenam número na matriz B e realizar a operação como soma de C2.

Como exemplo, vamos ilustrar a soma $B[1,1] + (-A[1,1])$, $B[2,1] + (-A[2,1])$ e
$B[3,1] + (-A[3,1])$ . Utilizaremos o tempB, que correnponde a primeira linha da
matriz B, ajustada para C2. Os números 1, destacados, são overflows e são
removidos para não influenciarem no processamento. O resultado da operação $D =
B + (-A)$ está representado abaixo:

\begin{align*}
	tempB	&=\ldots000001|000011|000111|001001\\
+(-A[1,1])	&=\ldots111111|111101|111011|111000\\
	D[1,1]
&=\ldots\sout{1}000000|\sout{1}000000|\sout{1}000010|\sout{1}000001
\end{align*}

\begin{align*}
	tempB &= \ldots000001|000011|001110|001111\\
+(-A[2,1])&= \ldots110100|110010|111010|110111\\ 
D[2,1]	  &= \ldots110101|110101|\sout{1}001000|\sout{1}000110
\end{align*}

\begin{align*}
tempB	  &= \ldots010100|011110|000010|000001 \\
+(-A[3,1])&= \ldots111101|111001|110110|110101\\ 
D[3,1]	  &= \ldots\sout{1}010001|\sout{1}110111|\sout{1}111000|110110
\end{align*}

Removendo os overflows a matriz D, essa fica definida como:

\begin{equation}
	D = \begin{bmatrix}
			129\\ 
			14111238\\ 
			4685366
		\end{bmatrix} 
        =
        \begin{bmatrix}
   			\ldots000000000000000010000001 \\
			\ldots110101110101001000000110 \\
			\ldots010001110111111000110110
		\end{bmatrix}
        =
        \begin{bmatrix}
 \ldots\underbrace{000000}_{0}\underbrace{000000}_{0}\underbrace{000010}_{2}
\underbrace{000001}_{1} \\ 			 
 \ldots\underbrace{110101}_{-11}\underbrace{110101}_{-11}\underbrace{001000}_{8}
\underbrace{000110}_{6} \\
 \ldots\underbrace{010001}_{17}\underbrace{110111}_{23}\underbrace{111000}_{-8}
\underbrace{110110}_{-10}
		\end{bmatrix}
\end{equation}

A multiplicação de matrizes é outra operação que é possível realizar com
matrizes de bitstring. Agora vamos calcuar o produto matricial $P = A \times
B^{t}$, sendo $B^t$ a matriz transposta da matriz B. Logo, temos que solucionar:

\begin{equation}
	P = A \times B = \begin{bmatrix}
			1 & 3 & 5 & 8\\ 
			12 &14  & 6 & 9\\ 
			3 & 7 & 10 & 11
		\end{bmatrix}
        \times
		\begin{bmatrix}
			1 & 1  & 20\\ 
            3 & 3  & 30\\
            7 & 14 & 2\\
            9 & 15 & 1
		\end{bmatrix}
\end{equation}

Utilizando as matrizes AA e BB, que correspondem as formas comprimidas das
matrizes A e B, respectivamente, vamos cacular o produto. A multiplicação de
matrizes bitstring ($\times\times$) fica definida como:

\begin{equation}
	P = AA \times BB = \begin{bmatrix}
			4952\\ 
			52841\\ 
			14251
		\end{bmatrix} 
        \times\times
        \begin{bmatrix}
			36073 & 36303 & 686145
		\end{bmatrix}
\end{equation}

Na equação X, as matrizes AA e BB estão no formato de bitstring.

\begin{equation}
	P = AA \times BB = 
    	\begin{bmatrix}

\ldots\underbrace{0001}_{1}\underbrace{0011}_{3}\underbrace{0101}_{5}\underbrace
{1000}_{8}\\

\ldots\underbrace{1100}_{12}\underbrace{1110}_{14}\underbrace{0110}_{6}
\underbrace{1001}_{9}\\

\ldots\underbrace{0011}_{3}\underbrace{0111}_{7}\underbrace{1010}_{10}
\underbrace{1011}_{11}
		\end{bmatrix}
\end{equation}
\begin{equation}
        \times\times \\
        \begin{bmatrix}

\ldots\underbrace{00001}_{1}\underbrace{00011}_{3}\underbrace{00111}_{7}
\underbrace{01001}_{9}  &

\ldots\underbrace{00001}_{1}\underbrace{00011}_{3}\underbrace{01110}_{14}
\underbrace{01111}_{15} &

\ldots\underbrace{10100}_{20}\underbrace{11110}_{30}\underbrace{00010}_{2}
\underbrace{00001}_{1}
		\end{bmatrix}
\end{equation}

Agora vamos demonstrar o cálculo do primeiro elemento da matriz P. A operação a
ser realizada envolvendo as matrizes A e B, para calcular o primeiro elemento, é
$1 \times 1 + 3 \times 3 + 5 \times 7 + 8 \times 9$. O mesmo ocorre em uma
matriz de bitstring. Mas nesse caso os elementos devem ser obtidos das tiras,
multiplicados e somandos para gerar o novo elemento. Para isso vamos ilustrar o
processo. Novamente podemos utilizar a máscara A e máscara B para realizar a
obtenção dos valores, como realizado em X e X. Uma vez que os valores foram
obtidos, pode-se executar a multiplicação deles e posterior armazenagem. Suponha
que tempA e tempB, armazenam os valores referentes as regiões 1 e 3, dos
elementos A[1,1] e B[1,1]. Para que seja possível armazenar os valores, torna-se
necessário calcular o tamanho da região que irá conter os valores. Como já
demosntrado, para a multiplicação, no pior caso, a região equivale a soma das
regiões das tiras de bitstring que estão sendo multiplicadas, nesse caso 4 e 5.
Logo, a região teria que ter tamanho 9. Porém são realizadas 3 somas, e para
cada uma delas, a região aumenta em 1 bit, logo a região para armazenar os
produtos demanda 
12 bits. Definido o tamano da região, vamos apenas realizar os cálculos e
armazenálos na tira do produto. A multiplicação dos valores armazenados em tempA
e tempB estão ilustrados nas Equações de  X até XX. Esses valores são
armazenados nas variáveis temporárias tempPi, i = 1 até 4, que e que i
corresponde a região dos números envolvidos no processo. Uma vez calculado os
produtos, basta somá-los para obter o resultado.

\begin{align*}
 tempA&= \ldots0001|0011|0101|1000\\
 tempB&= \ldots00001|00011|00111|01001\\
\end{align*}

\begin{align*}
 tempA&= \ldots\textbf{0001}|0011|0101|1000\\
 tempB&= \ldots\textbf{00001}|00011|00111|01001\\
 tempP1&= \ldots000000001
\end{align*}

\begin{align*}
 tempA&= \ldots0001|\textbf{0011}|0101|1000\\
 tempB&= \ldots00001|\textbf{00011}|00111|01001\\
 tempP2&= \ldots000001001
\end{align*}

\begin{align*}
 tempA&= \ldots0001|0011|\textbf{0101}|1000\\
 tempB&= \ldots00001|00011|\textbf{00111}|01001\\
 tempP3&= \ldots000100011
\end{align*}

\begin{align*}
 tempA&= \ldots0001|0011|0101|\textbf{1000}\\
 tempB&= \ldots00001|00011|00111|\textbf{01001}\\
 tempP4&= \ldots001001000
\end{align*}

\begin{align*}
 tempP1&= \ldots0000000001 \\
 tempP2&= \ldots0000001001 \\
   soma&= \ldots0000001010 \text{ (10 bits)}
\end{align*}

\begin{align*}
 tempP3&= \ldots00000100011 \\
   soma&= \ldots00000001010 \\
   soma&= \ldots00000101101 \text{ (11 bits)}
\end{align*}

\begin{align*}
 tempP4&= \ldots000001001000 \\
 soma&=   \ldots000000101101 \\
 soma&=   \ldots000001110101 \text{ (12 bits)}
\end{align*}

Portanto, o primeiro valor a ser armazenado na matriz P[1,1], na quarta região,
é $000001010101$. A matriz P, após realizar todas as operações, fica definida
como

\begin{equation}
	P = \begin{bmatrix}
			1426882688\\ 
			2970686121\\ 
			3239350573
		\end{bmatrix} 
        =
        \begin{bmatrix}
   			\ldots000001010101000011001000000010000000 \\
			\ldots000010110001000100010001001010101001\\
			\ldots000011000001000101001001000100101101
		\end{bmatrix}
\end{equation}
\begin{equation}
        =
        \begin{bmatrix}

\ldots\underbrace{000001010101}_{117}\underbrace{000011001000}_{200}\underbrace{
000010000000}_{128}\\

\ldots\underbrace{000010110001}_{177}\underbrace{000100010001}_{273}\underbrace{
001010101001}_{681}\\

\ldots\underbrace{000011000001}_{193}\underbrace{000101001001}_{329}\underbrace{
000100101101}_{301} 			 
		\end{bmatrix}
\end{equation}

Com os elementos armazenados na matriz P, pode-se verificar que  já estão no
formato comprimido, o que torna a utilização da memória otimizada. A operação de
divisão para matrizes é realizada usando uma inversa em um produto. Como até
agora tratamos apenas de números inteiros, essa operação será estudada em outro
trabalho.

Nos exemplo demonstrados, foi utilizado apenas as matrizes de bitstring criadas
a partir do método SM, mas o mesmo procedimento pode ser aplicado ao método VLB.
A diferença é que o resultado será armazenado no formato VLB.

Assim, foi demonstrado que para as operações elementares e para as operações de
soma, subtração e multiplicação de matrizes o formato de compressão de bits
string opde ser realizado.

\end{document}

### 2013-03-25 manuscript.tex
terminei de traduzir o material supplementar

@@
 \section*{Abstract}
-
+The biggest cost of
+computing with large matrices in any modern computer is related to memory
+latency and bandwidth. The average latency of modern RAM reads is 150 times
+greater than a clock step of the processor\cite{alted2010modern}. Throughput is
+a little better but still 25 times slower than the CPU can consume. The
+application of bitstring compression allows for larger matrices to be moved
+entirely to the cache memory of the computer, which has much better latency and
+bandwidth (average latency of L1 cache is 3 to 4 clock steps). This allows for
+massive performance gains as well as the ability to simulate much larger models
+efficiently. In this work, we propose a methodology to compress matrices in such
+a way that they retain their mathematical properties. Considerable compression
+of the data is also achieved in the process Thus allowing for the computation of
+much larger linear problems within the same memory constraints when compared
+with the traditional representation of matrices.
 % Please keep the Author Summary between 150 and 200 words
@@
 
-\section*{Figure Legends}
+%\section*{Figure Legends}
 %\begin{figure}[!ht]
@@
 
-\section*{Tables}
+%\section*{Tables}
 %\begin{table}[!ht]

### 2013-03-25 suppl.tex
terminei de traduzir o material supplementar

@@
 \usepackage{graphicx}
+\usepackage{cite}
 
@@
 
-\title{Your Paper}
-\author{You}
+\title{Computable Compressed Matrices\\Suplementary Material}
+\author{Crysttian A. Paixão}
 
+\bibliographystyle{plos2009}
 \begin{document}
@@
 
-Inicialmente vamos considerar as operações elementares (soma, subtração, divisão
-e multiplicação) a serem realizadas nos elementos armazenados na matriz de
-bitstring. Sendo essa matriz gerada pelos métodos SM ou VLB. Ressalta-se que os
-números e resultados das operações são descritos no formato inteiro.
+In this supplementary material we will explore basic arithmetic operations
+(addition, subtraction, division and multiplication) over the elements of
+bitstring compressed arrays. We will consider integer matrices compressed via
+both the SM and the VLB methods.
 
-Para realizar essas operações torna-se necessário avaliar o comportamento dos
-bits dos números envolvidos. Quando uma das operações mencionadas é utilizada,
-os bits dos números envolvidos são manipulados para chegar ao resultados final.
-Em um computador essas operações são realizadas na base binária, diferente do
-caso tradicional, no qual para fazer uma conta, a realizamos na base decimal. A
-principal característica que deve-se considerar é a quantidade de bits
-nacessários para armazenar o valor resultante da operação. Para isso, vamos
-utilizar exemplos, de tal forma que as operações gerem resultados, os quais
-demandem a maior quantidade de bits.
+In these operations, the most important aspect is the bit-length of the result
+in relation to those of the operands. As the operations are all done in
+binary, when the result bit-length increases, the resulting matrix will require
+more space to store and the operation in itself gets more complex, as in place
+operations are not possible. In the following examples we will explore these
+kinds of operations.
 
-A primeira operação a ser avaliada é a soma. Na base binária o comportamento do
-carry é o mesmo da decimal, porém o que muda é a base. A soma 1+1 resulta em 0 e
-gera um carry de 1. Em decimal, seria a mesma coisa de realizar a operação 5 +
-5, que gera um carry de 1 para ser somado a dezena.  As demais contas são $1 + 0
-= 1$, $0 + 1 = 1$ e $0 + 0 = 0$. Considere $n_1=7$ e $n_2=8$, a soma em decimal
-e binário é realizada na Tabela \ref{tab:01}. Observe que quando ocorre a soma
-$1 + 1$, gera-se um carry para ser adicionado a soma dos bits à esquerda. Ao
-todo foram gerados 3 bits de carry. Já na Tabela \ref{tab:02}, podemos verificar
-a soma de $7 + 7$, que gera um número com 4 bits. Conclui-se que o resultado
-soma, demanda 1 bit a mais do que o número de bits para representar o maior
-valor envolvido na operação. Para representar o resultado, nesse caso o número
-7, demanda 3 bits,logo o resultado demanda 4. Isso só é válido para o pior caso,
-note que 7 demanda 3 bits, o qual é adicionado a 7, que também demanda 3 
-bits, logo resultado irá demandar 4 bits. Pode ocorrer que na soma, a mesma
-quantidade de bits do maior elemento, por exemplo a soma 1 + 2, seja mentida,
-sem o acréscimo de 1 bits ao resultado. Por exemplo, o número 1 demanda 1 bit, o
-número 2, 2 bits, o resultado 3, demanda 2 bits.
+\subsection{Operations with Scalars}
+
+\paragraph{Addition}
+
+In base 2 the carry over  behavior is the same as we observer when operating
+with decimals, only the base is different. Adding $1+1$ results in $0$ and a
+carry over of 1, this is similar to the decimal sum $5+5$ where we also have a
+carry over of 1. Other single digit additions in base 2 are: $1 + 0
+= 1$, $0 + 1 = 1$, and $0 + 0 = 0$. Consider the following addition $n_1+n_2$
+where $n_1=7$ and $n_2=10$. This addition is illustrated in table \ref{tab:01}
+in both decimal and binary. Note that when we add $1+1$, a carry over is
+generated to be added to bit on the left. In total we generate three bits in
+carry over.
+
+In Table \ref{tab:02} we show the addition of $7 + 7$ which generates a 4 bits
+long number. In integer additions, the results may, in the very least, be as
+long (in bits) as the greatest operand (for example $1+2=3$), but is frequently
+longer.
 
@@
 	\centering
-    \caption{Comparacao da soma em base decimal e binaria.}
+    \caption{Comparing binary and decimal sum ($7+10$).}
     \begin{tabular}{crrrc}
     \hline
-    	  & Decimal & & Binario & Numero de Bits \\
+    	  & Decimal & & Binary & bit-length \\
     \hline      
-    Carry & 	& & \textit{11100} & \\
+    Carry Over& 	& & \textit{11100} & \\
     $n_1$ & 7   & & 111 & 3 \\
@@
     \hline
-    	  & Decimal & & Binario & Numero de Bits \\
+    	  & Decimal & & Binario & Bit-length \\
     \hline      
@@
 
-A próxima operação a ser representada é a subtração. Dependendo dos números
-envolvidos na operação, o resultado pode ser negativo. Com isso, gera-se a
-demanda de armazenar os números em um formato que contemple o sinal. A forma
-mais utilizada para realizar isso é o complemento de 2 - C2 (ref). O número
-representado nessa forma, demanda 1 bit a mais para representar o seu sinal.
-Vamos realizar a operação $n_2 - n_1$, que pode ser representada por $n_2 +
-(-n_1)$. Isso permite que as mesmas instruções para a soma, possam ser
-utilizadas para a subtração. Para que isso possa ser feito, torna-se necessário
-representar os números envolvidos no formato C2. Para passar um número para o
-formato C2, deve-se seguir seguintes passos:
+\paragraph{Subtraction}
+Depending of the number involved in this operation, the result may be negative.
+This fact alone generates the need for an extra bit to represent the sign($+$ or
+$-$). The usual way to achieve this is to use the  two's complement
+\cite{flores1963logic} representation which has the advantage of making the
+addition subtraction and multiplication operation the same as those for unsigned
+binary numbers. To exemplify let's subtract $7$ from $10$. To make use of the
+same algorithm of the sum we represent the operation as $10+(-7)$. First we
+need to convert the operands to two's complement representation (C2):
 
 \begin{itemize}
-	\item Representa-se o número com um bit a mais que o necessário, por
-exemplo o número 7, demanda 3 bits, no formato C2, demandará 4 bits. Como o -7
-será adicionado ao 10, os números devem ser representados por 5 bits ( 4 bits
-para representar o 10 mais um bit para o sinal). Logo, 7 será representados por
-00111. Aqui o número está no formato C1. Esse é outro formato utilizado, o
-problema é que o número 0 possui duas representações nesse formato. Essas duas
-representações gera uma demanda para verificar qual dos formatos de zeros está
-presente, já o C2 possui apenas uma.
-    \item Inverte-se os valores dos bits do número 7, nesse caso de 00111 para
-11000
-	\item Ao resultado da inversão soma-se 1. Nesse caso 11000 + 1 irá
-resultar em 11001. Logo 11001 é a representação para -7 no formato C2.
+    \item The numbers will require an extra bit in C2. Thus 7 which is 3
+bits long in binary, will require 4 bits in C2. As $7$ will be added to $10$,
+the operands must be represented by 5 bits (4 bits to represent 10 plus one for
+the signal). Therefore $7$ becomes $00111$ (in one's complement).
+
+    \item Now the bits in 7 are flipped, going from $00111$ to $11000$.
+
+    \item to the flipped number we add 1: $11000+1=11001$, which is the C2
+representation of $-7$
 \end{itemize}
 
-Uma vez realizada a conversão, basta executar o processo de soma como já
-mencionado. O resultado está representado na Tabela \ref{tab:03}. Observe que é
-executado uma soma com os números envolvidos. Ambos demandam 5 bits para serem
-representados por causa do C2. O mais importante é o bit, que é gerado mais à
-esquerda pelo carry no resultado, é descartado(overflow, número 1 cortado). Isso
-ocorre pois os números envolvidos na operação demandam 5 bits e não 6 bits. Para
-usar o formato C2, deve-se converter os números para o mesmo tamanho.
+Once completed the conversion, we add the numbers (see table  \ref{tab:03}).
+Both operands require 5 bits because of C2. Most importantly, the left most
+bit generated by the carry over, is discarded. That happens because to operate
+in C2, the operands must have the same bit-length so any overflow in the result
+must be discarded.
 
-Essa operação seria possível, sem que fosse utilizado o C2, pois 10 é maior que
-7, logo não demandaria sinal. A operação está representada na Tabela
-\ref{tab:04}. Observe que ocorre um carry para compensar os valores. Na operação
-0 - 1, demanda um carry do número a esquerda, esse carry tem o valor de 2, ao
-ser removido o valor 1, resultado em 1.
+This operation would be possible without the use of C2 representation as 10 is
+greater than 7 (see table \ref{tab:04} for this).
 
@@
 	\centering
-    \caption{Comparacao da soma em base decimal e binaria.}
+    \caption{Two's complement addition(subtraction) in decimal and binary.}
     \begin{tabular}{crrrc}
     \hline
-    	  & Decimal & & Binario & Numero de Bits \\
+    	  & Decimal & & C2 Binary & Bit-length \\
     \hline      
-    Carry & 	& & \textit{110000} & \\
+    Carry over & 	& & \textit{110000} & \\
     $n_2$ & 10	& & 01010 & 5 \\
@@
 	\centering
-    \caption{Comparacao da soma em base decimal e binaria.}
+    \caption{Standard binary subtraction.}
     \begin{tabular}{crrrc}
     \hline
-    	  & Decimal & & Binario & Numero de Bits \\
+    	  & Decimal & & Binary & Bit-length \\
     \hline      
-    Carry & 	& & \textit{0111} & \\
+    Carry over& 	& & \textit{0111} & \\
     $n_2$ & 10	& & 1010 & 4 \\
@@
 
-Para a subtração, o resultado no pior caso, irá demandar o mesmo número de bits
-do maior valor envolvido na operação, considerando o formato C2 ou não. Como
-exemplo, considere $n_3=1$, logo o resutlado da operação $n_2 - n_1$ irá
-demandar 3 bits, como pode ser verificado na Tabela \ref{tab:05}.
+In conclusion, in the worst case subtraction will require the same number of
+bits of the greatest operand, regardless of the use of C2 representation. As an
+example, consider the subtraction $7-1$ (table \ref{tab:05}).
 
@@
 	\centering
-    \caption{Comparacao da soma em base decimal e binaria.}
+    \caption{Binary subtraction}
     \begin{tabular}{crrrc}
     \hline
-    	  & Decimal & & Binario & Numero de Bits \\
+    	  & Decimal & & Binary & Bit-length \\
     \hline      
-    Carry & 	& & \textit{000} & \\
+    Carry over & 	& & \textit{000} & \\
     $n_1$ & 7	& & 111 & 3 \\
@@
 
-A multiplicação pode ser realizada da mesma forma que a soma e a subtração. Ela
-nada mais é que o resultado de múltiplas somas. Por exemplo, a operação $2
-\times 3 = 3 + 3 = 2 + 2 + 2$, ou seja 3 somado 2 vezes ou 2 somados 3 vezes.
-Vamos calcular a multiplicação $n_1 \times n_2$, a qual a operação está
-representada na Tabela \ref{tab:06}. Observe que são realizadas duas a duas, as
-somas, da mesma forma que o computador realiza a operação.
+\paragraph{Multiplication}
+If multiplication is thought of as a series of sums, we can apply much the same
+techniques. For example: $2 \times 3 = 3 + 3 = 2 + 2 + 2$. Table \ref{tab:06}
+describes a simple multiplication.
 
@@
 	\centering
-    \caption{Comparacao da soma em base decimal e binaria.}
+    \caption{Binary Multiplication}
     \begin{tabular}{crrrc}
     \hline
-    	  & Decimal & & Binario & Numero de Bits \\
+    	  & Decimal & & Binary & Bit-length \\
     \hline      
-    $n_2$ & 10	& & 1010 & 3 \\
+    $n_2$ & 10	& & 1010 & 3 [VERIFICAR]\\
     $n_1$ & \underline{$\times 7$} & & \underline{111} & 1 \\
@@
 
-Na multiplicação o comportamento do número de bits é diferente da soma e da
-subtração. No pior caso, o resultado é representado com o número de bits
-resultante da soma do número de bits de cada número envolvido na operação. Na
-operação $n_2 \times n_2$, ou seja $7 \times 7$, o resultado da operação, 49,
-demanda 6 bits para ser representado, pois 49 em binário é 110001.
+In multiplication the behavior of the bit-length is different from the sum and
+subtraction. In the worst case the product has a bit-length which is the sum of
+the bit-lengths of the factors. For example, in the product $7\times7$, the
+product is 6 bits long: $49$ which in binary is $110001$.
 
-Caso na multiplicação esteja envolvido algum número com sinal, torna-se
-necessário utilizar o complemento C2. Para ilustrar essa operação, vamos
-realizar a operação $n_2 \times -n_1$. Lembrando que 10 e -7, no formato C2, são
- iguais à 01010 e 11001, respectivamente. Na Tabela \ref{tab:07} é possível
-verificar a operação. A diferença básica é que os bits de sinal (em negrito),
-posicionados mais à esquerda, não são utilizados para executar as somas
-sucessivas. Apenas no final, esses bits são utilizados. Como 0 representa o
-positivo e 1, o negativo, o resultado terá sinal negativo. O resto das operações
-são iguais as operações já comentadas anteriormente. O número de bits
-necessários para representar o resultado, no pior caso, é igual a soma do número
-de bits envolvidos na multiplicação, sem contar o sinal, mais um bit, em
-decorrência do sinal do resultado.
+If a signed integer is involved in the multiplication, we need to use C2
+representation. To illustrate this case let's calculate $10\times-7$. in C2,
+this opration becomes $01010\times11001$. Table \ref{tab:07} contains the
+details. The basic difference is that the sign bits (in bold), positioned to
+the left, are not involved in the operation. Only in the end they are used to
+determine the sign of the product: $0$ meaning positive and $1$ negative. In
+the worst case the bit-length of the product is the sum of the bit-lengths of
+the factors plus 1 due to the sign.
 
@@
 	\centering
-    \caption{Comparacao da soma em base decimal e binaria.}
+    \caption{Multiplication involving a signed integer.}
     \begin{tabular}{crrrc}
     \hline
-    	  & Decimal & & Binario & Numero de Bits \\
+    	  & Decimal & & Binary & Bit-length \\
     \hline      
@@
 
-A divisão, ao contrário da multiplicação, demanda que seja executada sucessivas
-subtrações até encontrar um determinado resto. Quando exato, esse resto vale
-zero. Para ilustar o comportamento dos bits, considere $n_4=50$ na operação $n_4
-/ n_2$. O resultado está descrito na Tabela \ref{tab:08}. Observe que ocorre
-sucessivas subtrações em binário. O número de bits necessário para representar o
-resultado da divisão, no pior caso, é igual o número de bits do dividento,
-quando o dividendo é maior que o divisor. A cada subtração executada, soma-se 1
-ao quociente, que inicialmente vale 0. O processo de subtração acaba quando o
-valor do resto é zero.
+\paragraph{Division}
+Division, in contrast to multiplication, requires successive subtractions until
+a remainder is reached which may or may not be zero. To illustrate the
+procedure in binary, table \ref{tab:08} describes the division of $50$ by $10$.
+Note the successive binary subtractions. The bit-length of the quotient is, in
+the worst case, equal to the bit-length of the dividend when this is longer
+than that of the divisor. With each subtraction, we add $1$ to the quotient
+which is initially $0$. The process ends when the remainder is $0$ or less than
+the divisor.
+
+When the division involves a signed operand, we do the same as with the
+multiplication, the operations is executed on the unsigned operands and the
+sign is applied at the end.
 
@@
 	\centering
-    \caption{Comparacao da soma em base decimal e binaria.}
+    \caption{Binary division as a series of subtractions.}
     \begin{tabular}{crrrrr}
     \hline
-    	  & Decimal & & Binario & Quociente Decimal & Quociente Binario \\
+    	  & Decimal & & Binary & Decimal quotient & Binary quotient\\
     \hline  
@@
 \underline{+001} \\
-    Resto & 40      			& & 101000  			&1& 001
+    Remainder & 40      			& & 101000
+&1& 001
 \\
@@
 \underline{+001} \\
-    Resto & 30      			& & 011110  			&2& 010
+    Remainder & 30      			& & 011110
+&2& 010
 \\
@@
 \underline{+001} \\
-    Resto & 20      			& & 010100  			&3& 011
+    Remainder & 20      			& & 010100
+&3& 011
 \\
@@
 \underline{+001} \\
-    Resto & 10      			& & 001010  			&4& 100
+    Remainder & 10      			& & 001010
+&4& 100
 \\
@@
 \underline{+001} \\
-    Resto & 0      				& & 000000
+    Remainder & 0      				& & 000000
 &5& 101 \\
@@
 
-O mesmo ocorre quando o dividendo ou dividor possuem sinal. Para efetuar a
-divisão com números que possuam sinal, torna-se necessário utilizar o C2. Vamos
-ilustrar o processo, considerando a divisão $-n_4 / n_2$. Quando o dividendo vai
-ser subtraído o valor do divisor, quando no formato C2, o divisor deve ser
-convertido para C2. E ao invés de usar a subtração, realiza-se a soma. E como
-feito no caso da divisão sem sinal, a operação de soma termina, quando o resto é
-igual a zero. O processo de divisão de número com sinais pode ser visualizado no
-Tabela \ref{tab:08}. O quociente é 5, mas devido a diferença de sinais, esse
-deve ser convertido para -5.
-
-\begin{table}[ht]
-	\centering
-    \caption{Comparacao da soma em base decimal e binaria.}
-    \begin{tabular}{crrrrr}
-    \hline
-    	  & Decimal & & Binario & Quociente Decimal & Quociente Binario \\
-    \hline  
-    $n_4$ & -50 					& & 1001110
-&0& 000 \\
-    $n_2$ & \underline{+10} 	& & \underline{001010}  &\underline{+1}&
-\underline{+001} \\
-    Resto & -40      			& & 1011000  			&1& 001
-\\
-    $n_2$ & \underline{+10} 	& & \underline{001010}  &\underline{+1}&
-\underline{+001} \\
-    Resto & -30      			& & 1100010 			&2& 010
-\\
-    $n_2$ & \underline{+10} 	& & \underline{001010}  &\underline{+1}&
-\underline{+001} \\
-    Resto & -20      			& & 1101100  			&3& 011
-\\
-    $n_2$ & \underline{+10} 	& & \underline{001010}  &\underline{+1}&
-\underline{+001} \\
-    Resto & -10      			& & 1110110  			&4& 100
-\\
-    $n_2$ & \underline{+10} 	& & \underline{001010}  &\underline{+1}&
-\underline{+001} \\
-    Resto & 0      				& & 0000000
-&5& 101 \\
-    \hline
-	\end{tabular}
-    \label{tab:08}
-\end{table}
-
-Apresentamos até agora as quatro operações elementares, as quais verificamos o
-comportamento dos bits. Resumidamente podemos inferir sobre o comportamento do
-número de bits, sem sinal, no pior caso que reflete a utilização de bits, da
-seguinte forma:
+So far we have examined the 4 fundamental arithmetic operations. In summary the
+implication for memory allocation of the results are the folowing, in the worst
+case scenarios:
 
 \begin{itemize}
-\item Soma: demanda 1 bits a mais que o número de bits para representar o maior
-valor presente na operação;
-\item Subtração: demanda a mesma quantidade de bits para representar o do maior
-valor presente na operação;
-\item Multiplicação: demanda é igual a soma dos bits dos números envolvidos na
-operação;
-\item Divisão: demanda a mesma quantidade de bits do dividendo;
+\item Addition: requires 1 extra bit above the bit-length of the greatest
+operand;
+\item Subtraction: Requires the same number of bits as the greatest
+operand;
+\item Multiplication: Requires the sum of the bit-lenghts of the
+operands;
+\item Division: Requires the same bit-length as the dividend;
 \end{itemize}
 
-Quando os números apresentam sinal, eles demandam 1 bits a mais para o sinal.
-Nesse caso, na multiplicação, para o pior caso, o número de bits para
-representar o resultando será a soma do número de bits dos números envolvidos,
-menos 1.
+When operating with signed integers, an extra bit is used for the sign.
 
-Até o momento, as operações elementares foram executas apenas com números. Agora
-será demonstrado a sua aplicação sobre as tiras de bits. Essas tiras de bits são
-geradas pelos métodos SM ou VLB, os quais promovem a compactação dos dados e
-armazenam diversos números com economia de espaço em tiras de bits.
+The operations on single numbers (scalars) as exposed above are a simplification
+of the
+actual operations taking place on the bitstrings as we compute with compressed
+matrices. Additional details will be provided below when we discuss the
+operation with matrices. Note that we are restricting the example to operations
+which generate integer results. Handling of float operations and compression
+will be the subject of a subsequent paper.
 
-Antes de demonstrarmos o procedimento, algumas regras devem ser contempladas
-para que as operações possam ser realizadas. A primeira é que quando existir um
-número com sinal, todos os números deverão ser representados na forma de
-complemento de 2, como consequencia, existirá a demanda de 1 bit para cada
-número armazenado na tira. As operações só podem resultar números inteiros, pois
-a forma de armazenamento discutida atende apenas essa configuração.
+\begin{figure}[h]
+ \centering
+ \includegraphics[width=12cm]{./bitlength.png}
+ % simples.png: 4800x3600 pixel, 600dpi, 20.32x15.24 cm, bb=0 0 576 432
+ \caption{bitlength of the results for each of the four operations. The axes
+represent the values of the operands.}
+ \label{fig:bitlength}
+\end{figure}
 
-Para demonstrar a aplicação envolvendo tiras de bits, vamos considerar duas
-matrizes A e B, ambos as dimensões $3 \times 4$. Sendo definidas por:
+\subsection{Operation on Bitstrings}
+
+To illustrate operations with bitstrings, consider the matrices $A$ an $B$,
+both $3 \times 4$:
 
@@
 
-Após o processo de compressão, as matrizes A e B, antes com dimensão $3 \times
-4$, passam a ter dimensão $3 \times 1$. A matriz A e B após a compressão pelo
-método SM, podem ser visualizada em decinal e binário nas Equações \ref{eq:01} e
-\ref{eq:02}. As tiras possuem tamanhos de 4 bits e 5 bits, para as regiões de
-armazenamento, para a matriz A e B, respectivamente. Os "$\ldots$" representam
-os zeros omitidos em cada tira, pois os elementos são armazenados apenas em uma
-tira com 64 bits, na qual apenas alguns bits são utilizados, restando apenas
-zeros em desuso.
+After bitstring compression by the SM method, they become $3 \times 1$. They
+are show in decimal and binary from in \ref{eq:01} and
+\ref{eq:02}. The elements in $A$ and $B$ are 4 and 5 bits long, respectively
+(the ``supreme minimum''). the ellipses ("$\ldots$") in the binary matrix
+correspond the extra zeros to the left. Since the strings are written into 64
+bits integers, sometimes we end up with unused bits to the left.
 
@@
 
-Considere apenas a tira armazenada na primeira linha de cada matriz (AA e BB).
-Observe que para obter os valores de ambas as tiras, torna-se necessário
-conhecer o tamanho da região que os valores foram armazenados. Isso ocorre, pois
-elas foram geradas pelo método SM, o qual determina essa região. Para obter
-esses valores, de forma otimizada, torna-se necessário utilizar o que
-denominamos de máscara. A máscará irá permitir obter os valores armazenados
-dentro das tirar, de forma otimizada. Para esse exemplo, para otimizar as
-operações, a máscara irá coletar os valores da primeira e terceira região ao
-mesmo tempo, o que otimiza o processo. A máscara é armazenada em uma tira com o
-mesmo tamanho da tira que está armazenada nas matrizes. A máscara mencionada,
-para ser aplicada na matriz AA, está representada na Equação \ref{eq:06}. Para
-coletar as informações da segunda e quarta região da matriz AA, deve-se usar a
-máscara descrita na equação \ref{eq:07}. Já para a matriz BB deve-se usar as
-máscaras das equações \
-ref{eq:08} e \ref{eq:09}, para coletar as informações das primeira e terceira
-regiões e da segunda e quarta regiões, respectivamente.
+Consider now just the bitstring stored in the first line of each matrix ($AA$
+and $BB$). To recover the original elements we need to now the bit-lengthof
+the blocks of memory containing them, for the SM method they are all the same
+length. To obtain the values in an efficient way, we will use a binary mask.
+The mask will recover the first and third blocks at the same time to save time.
+The mask is stored as a bitstring the same length as those storing the matrix
+elements. The mask for the first and third elements of matrix $AA$ is depicted
+in \ref{eq:06}. To recover the second and fouth elements, we apply the mask
+shown on \ref{eq:07}. For Matrix $BB$, see \ref{eq:08} and \ref{eq:09} for the
+respective masks.
 
 \begin{equation}\label{eq:06}
-\text{máscara da matriz AA para posições 1 e 3} = \ldots0000111100001111
+\text{Matrix AA's mask for positions 1 and 3} = \ldots0000111100001111
 \end{equation}
@@
 \begin{equation}\label{eq:07}
-\text{máscara da matriz AA para posições 2 e 4} = \ldots1111000011110000
+\text{Matrix AA's mask for positions 2 and 4} = \ldots1111000011110000
 \end{equation}
@@
 \begin{equation}\label{eq:08}
-\text{máscara da matriz BB para posições 1 e 3} = \ldots00000111110000011111
+\text{Matrix BB's mask for positions 1 and 3} = \ldots00000111110000011111
 \end{equation}
@@
 \begin{equation}\label{eq:09}
-\text{máscara da matriz BB para posições 2 e 4} = \ldots11111000001111100000
+\text{Matrix BB's mask for positions 2 and 4} = \ldots11111000001111100000
 \end{equation}
 
-Uma vez definida a máscara, vamos aplicá-la na matriz A e B, para obter os
-valores armazenados na primeira e terceira regiões. Para usar a máscará, deve-se
-utilizar a função booleana AND. Vamos chamar de tempA e tempB, os valores
-coletados. O processo de coleta está ilustrado nas equações \ref{eq:10} e
-\ref{eq:11}. As região está destacadas. Verifique que apenas as regiões, na qual
-a máscara possui apenas 1 são preservadas, isso faz com que apenas esses valores
-sejam obtidos.
+Once defined the mask we can apply it to matrices $A$ and $B$. We use the mask
+by applying the boolean function AND. Let tempA and tempB the recovered
+elements. The recovering process is illustrated below. Note that only the
+positions where the mask is $1$ are retained.
 
@@
    A[1,1]			&=	\ldots0001|0011|0101|1000\\
-   \text{máscara A}	&=	\ldots0000|1111|0000|1111\\
+   \text{mask A}	&=	\ldots0000|1111|0000|1111\\
    tempA 			&=	\ldots0000|0011|0000|1000
@@
    B[1,1]			&= \ldots00001|00011|00111|01001 \\
-   \text{máscara B}	&= \ldots00000|11111|00000|11111 \\
+   \text{mask B}	&= \ldots00000|11111|00000|11111 \\
    tempB 			&= \ldots00000|00011|00000|01001
@@
 
-Suponha que desejamos calcular a soma das duas matrizes e armazená-las em uma
-matriz S. Para isso as regiões das tiras tempA e tempB devem ter o mesmo
-tamanho. Como para a soma, na pior das hipóteses, demanda-se 1 bit a mais que o
-maior valor armazenado, aqui esse valor é representado pela região na qual os
-valores são armazenados. Nesse caso, a região da tira resultante irá ter tamanho
-de 6 bits. As tiras, com as regiões corrigidas, são representadas nas equações
-\ref{eq:12} e \ref{eq:13}. Os bits inseridos para compensar o tamanho das
-regiões estão em negrito.
+\paragraph{Addition}
+
+Suppose we want to add both matrices and store it in a matrix $S$. for that the
+blocks of tempA and tempB must have the same length. As for the addition we
+need, in the worst case, an extra bit, the block length of the resulting
+bitstring will be 6. The string from each matrix, padded on the left towards
+this new length,  are shown in equations \ref{eq:12} and \ref{eq:13}.
 
@@
 
-Com as regiões igualadas, basta executar a operacão $S[1,1] = tempA + tempB$.
-Realizando o mesmo processo com tempA e tempB, mas na segunda e quarta região,
-podemos efetuar a operação soma de forma paralela, pois torna-se possível
-realizar 2 operações em paralelo, e o melhor, o resultado, nesse caso aa tira
-S[1,1], já está na forma compactada.
+Now that the blocks are matched in length we can perform the operation $S[1,1] =
+tempA + tempB$. The operation is repeated for the second and fourth regions.
+these operations can be done in parallel and the result, $S[1,1]$, is already
+in compressed form.
 
@@
 
-Na equação X, pode-se observar o resultado da soma para S[2,1] e S[3,1].
+We can see below the same operation performed for bitstrings $S[2,1]$ and
+$S[3,1]$.
 
@@
 
-Logo, como resultado da operação S= A + B, temos:
+Therefore, as the final result of the operation $S= A +
+B$ we have:
 
@@
 
-A operação de subtração pode ser realizada, mas para isso será utilizado a
-notação de complemento 2 - C2. Suponha que se deseja calcular $D = B - A$. Essa
-operação pode ser realizada da mesma forma, utilizando uma operação de soma $D =
-B + (-A)$, em que $-A$ será a matriz na forma C2. A matriz A na forma de
-complemento de 2, pode ser visualizada na Equação X. Ressalta-se que a os
-elementos serão calculados com 6 dígitos para serem armazenados nas tiras.
+\paragraph{Subtraction}
+To perform the subtraction, we need to use C2 notation. Supose  we need to
+calculate $D = B - A$. This operation can be done in the same way, by
+converting it to an addition: $D = B + (-A)$, where $-A$ will be converted to
+C2 (equation \ref{eq:ac2}). Note the elements will be calculated with six
+digits.
 
-\begin{equation}
+\begin{equation}\label{eq:ac2}
 	-A = \begin{bmatrix}
@@
 
-Para realizar a operação  $D = B + (-A)$, basta usar uma máscara para adequar as
-regiões que armazenam número na matriz B e realizar a operação como soma de C2.
+Now we apply the masks as before, but before we need to convert to C2 as well,
+in order to perform the addition.
 
-Como exemplo, vamos ilustrar a soma $B[1,1] + (-A[1,1])$, $B[2,1] + (-A[2,1])$ e
-$B[3,1] + (-A[3,1])$ . Utilizaremos o tempB, que correnponde a primeira linha da
-matriz B, ajustada para C2. Os números 1, destacados, são overflows e são
-removidos para não influenciarem no processamento. O resultado da operação $D =
-B + (-A)$ está representado abaixo:
+To exemplify, let's examine closely the operations $B[1,1] + (-A[1,1])$,
+$B[2,1] + (-A[2,1])$ and $B[3,1] + (-A[3,1])$. We use tempB to store each line
+of the matrix B converted to C2. The numbers $1$ marked in the bitstring, are
+overflows and must be removed. The result is shown below:
 
@@
 
-Removendo os overflows a matriz D, essa fica definida como:
+After the removal of the overflows, matrix $D$ becomes:
 
@@
 
-A multiplicação de matrizes é outra operação que é possível realizar com
-matrizes de bitstring. Agora vamos calcuar o produto matricial $P = A \times
-B^{t}$, sendo $B^t$ a matriz transposta da matriz B. Logo, temos que solucionar:
+\paragraph{Multiplication}
+Let's now examine matrix multiplication. Consider the product $P = A \times
+B^{t}$, where $B^t$ is the transpose of B. Thus the product becomes:
 
@@
 
-Utilizando as matrizes AA e BB, que correspondem as formas comprimidas das
-matrizes A e B, respectivamente, vamos cacular o produto. A multiplicação de
-matrizes bitstring ($\times\times$) fica definida como:
+Again, using $AA$ and $BB$ to denote the compressed versions of the matrices,
+and $\times\times$ to denote the multiplication of bitstring matrices, the
+compressed product becomes:
 
-\begin{equation}
+\begin{equation}\label{eq:prodbs}
 	P = AA \times BB = \begin{bmatrix}
@@
 \end{equation}
-
-Na equação X, as matrizes AA e BB estão no formato de bitstring.
 
@@
 
-Agora vamos demonstrar o cálculo do primeiro elemento da matriz P. A operação a
-ser realizada envolvendo as matrizes A e B, para calcular o primeiro elemento, é
-$1 \times 1 + 3 \times 3 + 5 \times 7 + 8 \times 9$. O mesmo ocorre em uma
-matriz de bitstring. Mas nesse caso os elementos devem ser obtidos das tiras,
-multiplicados e somandos para gerar o novo elemento. Para isso vamos ilustrar o
-processo. Novamente podemos utilizar a máscara A e máscara B para realizar a
-obtenção dos valores, como realizado em X e X. Uma vez que os valores foram
-obtidos, pode-se executar a multiplicação deles e posterior armazenagem. Suponha
-que tempA e tempB, armazenam os valores referentes as regiões 1 e 3, dos
-elementos A[1,1] e B[1,1]. Para que seja possível armazenar os valores, torna-se
-necessário calcular o tamanho da região que irá conter os valores. Como já
-demosntrado, para a multiplicação, no pior caso, a região equivale a soma das
-regiões das tiras de bitstring que estão sendo multiplicadas, nesse caso 4 e 5.
-Logo, a região teria que ter tamanho 9. Porém são realizadas 3 somas, e para
-cada uma delas, a região aumenta em 1 bit, logo a região para armazenar os
-produtos demanda 
-12 bits. Definido o tamano da região, vamos apenas realizar os cálculos e
-armazenálos na tira do produto. A multiplicação dos valores armazenados em tempA
-e tempB estão ilustrados nas Equações de  X até XX. Esses valores são
-armazenados nas variáveis temporárias tempPi, i = 1 até 4, que e que i
-corresponde a região dos números envolvidos no processo. Uma vez calculado os
-produtos, basta somá-los para obter o resultado.
+To calculate $P[1,1]$ which in uncompressed form is $1 \times 1 + 3 \times 3 + 5
+\times 7 + 8 \times 9$, we first must extract the numbers from the bitstrings
+and then proceed with the linear combination. The extraction will make use of
+masks as described before. Let tempA and tempB store the values of blocks 1 and
+3 of AA[1,1] and BB[1,1]. Now we need to determine the size of the blocks
+(bit-length) containing each element of the resulting matrix. As demonstrated
+before, in the worst case, the product will have a bit-length which is the sum
+of the bit-lengths of the operands. For this example this length is $4+5=9$,
+however, we also have three additions for each element,adding a total of 3
+extra bits, thus we end up with a bit-length of 12 for each element of the
+product. Having determined the bit-length of the product, we can now we can do
+the actual calculations and store the results.
+
+The linear combination of elements which generates $P[1,1]$ is detailed below.
+We use temporary variables $tempP_i$ to store the products before adding them,
+where $i$ is the product being calculated.
 
@@
  tempB&= \ldots\textbf{00001}|00011|00111|01001\\
- tempP1&= \ldots000000001
+ tempP_1&= \ldots000000001
 \end{align*}
@@
  tempB&= \ldots00001|\textbf{00011}|00111|01001\\
- tempP2&= \ldots000001001
+ tempP_2&= \ldots000001001
 \end{align*}
@@
  tempB&= \ldots00001|00011|\textbf{00111}|01001\\
- tempP3&= \ldots000100011
+ tempP_3&= \ldots000100011
 \end{align*}
@@
  tempB&= \ldots00001|00011|00111|\textbf{01001}\\
- tempP4&= \ldots001001000
+ tempP_4&= \ldots001001000
 \end{align*}
@@
 
-Portanto, o primeiro valor a ser armazenado na matriz P[1,1], na quarta região,
-é $000001010101$. A matriz P, após realizar todas as operações, fica definida
-como
+So, $P[1,1]$ is $000001010101$, and the entire matrix $P$ becomes:
 
@@
 
-Com os elementos armazenados na matriz P, pode-se verificar que  já estão no
-formato comprimido, o que torna a utilização da memória otimizada. A operação de
-divisão para matrizes é realizada usando uma inversa em um produto. Como até
-agora tratamos apenas de números inteiros, essa operação será estudada em outro
-trabalho.
+Again we can see that P is already in compressed form, which confirms that the
+entire operation was conducted without decompressing the data.
 
-Nos exemplo demonstrados, foi utilizado apenas as matrizes de bitstring criadas
-a partir do método SM, mas o mesmo procedimento pode ser aplicado ao método VLB.
-A diferença é que o resultado será armazenado no formato VLB.
+The division operation with matrices will be described in a subsequent paper
+after we describe how to represent and operate with floating-point numbers.
 
-Assim, foi demonstrado que para as operações elementares e para as operações de
-soma, subtração e multiplicação de matrizes o formato de compressão de bits
-string opde ser realizado.
+In the examples given above, we only used the SM method, but the same procedure
+can be easily adapted to the VLB method.
 
+Thus we complete our demonstration of how to perform basic arithmetic
+operations with bitstring compressed scalars and matrices.
+
+\bibliography{plos.bib}
 \end{document}

### 2013-03-25 bitmap.py

@@ -1,0 +1,77 @@
+import numpy as np
+import pylab as P
+from matplotlib import mpl, cm
+
+def check_sum(n):
+    op1, op2 = np.meshgrid(range(n),range(n))
+    @np.vectorize
+    def has_carry(o1,o2):
+        res = o1+o2
+        #return int(res.bit_length()>max([o1.bit_length(),o2.bit_length()]))
+        return res.bit_length()
+    return has_carry(op1,op2)
+
+def check_sub(n):
+    op1, op2 = np.meshgrid(range(n),range(n))
+    @np.vectorize
+    def has_carry(o1,o2):
+        res = o1-o2
+        #return int(res.bit_length()>max([o1.bit_length(),o2.bit_length()]))
+        return res.bit_length()
+    return has_carry(op1,op2)
+
+def check_mul(n):
+    op1, op2 = np.meshgrid(range(n),range(n))
+    @np.vectorize
+    def has_carry(o1,o2):
+        res = o1*o2
+        #return int(res.bit_length()>max([o1.bit_length(),o2.bit_length()]))
+        return res.bit_length()
+    return has_carry(op1,op2)
+
+def check_div(n):
+    op1, op2 = np.meshgrid(range(1,n),range(1,n))
+    @np.vectorize
+    def has_carry(o1,o2):
+        res = o1/o2
+        #return int(res.bit_length()>max([o1.bit_length(),o2.bit_length()]))
+        return res.bit_length()
+    return has_carry(op1,op2)
+
+def main():
+    #P.set_cmap('Paired')
+    tam = 256
+    #ncm = np.array([i.bit_length() for i in range(tam)])
+    #ncm *= 4
+    #lcm = mpl.colors.ListedColormap(np.squeeze(cm.Paired(ncm))[:,:3])
+    lcm = cm.Paired
+    summat = check_sum(tam)
+    submat = check_sub(tam)
+    mulmat = check_mul(tam)
+    divmat = check_div(tam)
+    fig = P.figure()
+    fig.add_subplot(221)
+    P.pcolor(summat, cmap=lcm)
+    P.axis([0,tam,0,tam])
+    P.colorbar()
+    P.title('Addition')
+    fig.add_subplot(222)
+    P.pcolor(submat, cmap=lcm)
+    P.axis([0,tam,0,tam])
+    P.colorbar()
+    P.title('Subtraction')
+    fig.add_subplot(223)
[…]

### 2013-09-13 manuscript.tex
atualizei

@@
 {\Large
-\textbf{Computable Compressed Matrices}
+\textbf{Computable Bitstring-compressed Matrices}
 }
@@
   \subfigure[$\alpha=1,\beta_1=1$]{
-  \includegraphics[scale=0.25,angle=-90,clip]{fig01}
+%   \includegraphics[scale=0.25,angle=-90,clip]{fig01}
   }
   \subfigure[$\alpha=1,\beta=32$]{
-  \includegraphics[scale=0.25,angle=-90,clip]{fig02}
+%   \includegraphics[scale=0.25,angle=-90,clip]{fig02}
   }
   \subfigure[$\alpha=32,\beta=1$]{
-  \includegraphics[scale=0.25,angle=-90,clip]{fig03}
+%   \includegraphics[scale=0.25,angle=-90,clip]{fig03}
   }
   \subfigure[$\alpha=64,\beta=64$]{
-  \includegraphics[scale=0.25,angle=-90]{fig04}
+%   \includegraphics[scale=0.25,angle=-90]{fig04}
   }
@@
    \subfigure[SM-VLB  Difference ($D = \eta_1 - \eta_2$)]{ 
-   \includegraphics[scale=0.18,clip]{fig07}
+%    \includegraphics[scale=0.18,clip]{fig07}
    }
    \subfigure[SM Efficiency  ($\eta_1$)]{
-   \includegraphics[scale=0.18,clip]{fig06}
+%    \includegraphics[scale=0.18,clip]{fig06}
    }
    \subfigure[VBL Efficiency ($\eta_2$)]{
-   \includegraphics[scale=0.18,clip]{fig05}
+%    \includegraphics[scale=0.18,clip]{fig05}
    }
    \subfigure[SM$\geq$VLB Efficiency ($\eta_1 \geq \eta_2$)]{
-   \includegraphics[scale=0.18,clip]{fig08}
+%    \includegraphics[scale=0.18,clip]{fig08}
    }
@@
   \subfigure[$\alpha_1=1,\beta_1=1,\alpha_2=1,\beta_2=1$]{
-  \includegraphics[scale=0.28,angle=-90,clip]{fig09}
+%   \includegraphics[scale=0.28,angle=-90,clip]{fig09}
   }
   \subfigure[$\alpha_1=1,\beta_1=32,\alpha_2=32,\beta_2=1$]{
-  \includegraphics[scale=0.28,angle=-90,clip]{fig10}
+%   \includegraphics[scale=0.28,angle=-90,clip]{fig10}
   }
   \subfigure[$\alpha_1=32,\beta_1=32,\alpha_2=32,\beta_2=32$]{
-  \includegraphics[scale=0.28,angle=-90,clip]{fig11}
+%   \includegraphics[scale=0.28,angle=-90,clip]{fig11}
   }
   \subfigure[$\alpha_1=64,\beta_1=32,\alpha_2=32,\beta_2=64$]{
-  \includegraphics[scale=0.28,angle=-90]{fig12}
+%   \includegraphics[scale=0.28,angle=-90]{fig12}
   }
@@
   \subfigure[$\alpha_1=64,\beta_1=48,\alpha_2=1,\beta_2=48$]{
-  \includegraphics[scale=0.28,angle=-90]{fig13}
+%   \includegraphics[scale=0.28,angle=-90]{fig13}
   }
   \subfigure[$\alpha_1=16,\beta_1=46,\alpha_2=49,\beta_2=64$]{
-  \includegraphics[scale=0.28,angle=-90]{fig14}
+%   \includegraphics[scale=0.28,angle=-90]{fig14}
   }
   \subfigure[$\alpha_1=16,\beta_1=16,\alpha_2=16,\beta_2=49$]{
-  \includegraphics[scale=0.28,angle=-90]{fig15}
+%   \includegraphics[scale=0.28,angle=-90]{fig15}
   }
   \subfigure[$\alpha_1=1,\beta_1=16,\alpha_2=1,\beta_2=49$]{
-  \includegraphics[scale=0.28,angle=-90]{fig16}
+%   \includegraphics[scale=0.28,angle=-90]{fig16}
   }
@@
   \subfigure[Efficiency histogram of the SM method ]{
-  \includegraphics[scale=0.25,angle=-90]{fig17}
+%   \includegraphics[scale=0.25,angle=-90]{fig17}
   \label{fig:17}
@@
   \subfigure[Efficiency histogram of the VLB method ]{
-  \includegraphics[scale=0.25,angle=-90]{fig18}
+%   \includegraphics[scale=0.25,angle=-90]{fig18}
   \label{fig:18}
@@
   \centering
-  \subfigure[Efficiencies of the SM and VLB methods, for constant bit-length matrices.]{\includegraphics[scale=0.3,clip]{fig19}\label{fig19}}
-  \subfigure[$\eta_1-\eta_2$ for matrices of constant bit-length.]{\includegraphics[scale=0.3,clip]{fig20}\label{fig20}}
+%   \subfigure[Efficiencies of the SM and VLB methods, for constant bit-length matrices.]{\includegraphics[scale=0.3,clip]{fig19}\label{fig19}}
+%   \subfigure[$\eta_1-\eta_2$ for matrices of constant bit-length.]{\includegraphics[scale=0.3,clip]{fig20}\label{fig20}}
   \caption{Compression efficiency of the SM and VLB methods for matrices of constant bit-length.}
@@
 \section{Conclusion}
-In this paper, we have focused in the compression of matrix data, since this is one of the most important application the authors foresee.
-However, the compression methodology presented can be applied to any numerical data structure, with gains to performance and memory footprint[citar tese Crysttian e possíveis artigos derivados].
+In this paper, we have focused in the compression of matrix data, since this is
+one of the most important application the authors foresee. However, the
+compression methodology presented can be applied to any numerical data
+structure, with gains to performance and memory footprint\cite{teseCrysttian}.
 
-Further discussions about doing computation with such compressed data-structures will be the subject of another manuscript (in preparation) in which we will present details about the implementation of the compression algorithm, and benchmarks on classical linear algebra tasks such as those in Linpack[ref].
+Further discussions about doing computation with such compressed data-structures
+will be the subject of another manuscript (in preparation) in which we will
+present details about the implementation of the compression algorithm, and
+benchmarks on classical linear algebra tasks such as those in
+Linpack\cite{dongarra1979linpack}.
 
@@
 \section*{Acknowledgments}
-We would like to thank Claudia Torres Code\c{c}o, Paulo Cezar Carvalho and Moacyr Silva for fruitful discussions and key ideas which helped improve the manuscript.
+We would like to thank Claudia Torres Code\c{c}o, Moacyr Silva and Paulo
+Cezar Carvalho for fruitful discussions and key ideas which helped improve the
+manuscript.
 

### 2013-09-13 suppl.tex
adicionei referência à figura.
edtei título
atualizei

@@
 \title{Computable Compressed Matrices\\Suplementary Material}
-\author{Crysttian A. Paixão}
+\author{Crysttian A. Paixão \and Flávio Codeço Coelho}
 
@@
 \maketitle
-
-\begin{abstract}
-Your abstract.
-\end{abstract}
 
@@
 Depending of the number involved in this operation, the result may be negative.
-This fact alone generates the need for an extra bit to represent the sign($+$ or
+This fact alone generates the need for an extra bit to represent the sign ($+$
+or
 $-$). The usual way to achieve this is to use the  two's complement
@@
 addition subtraction and multiplication operation the same as those for unsigned
-binary numbers. To exemplify let's subtract $7$ from $10$. To make use of the
-same algorithm of the sum we represent the operation as $10+(-7)$. First we
+binary numbers. This algorithm is what is actually implemented in CPUs. To
+exemplify let's subtract $7$ from $10$. To make use of the
+same algorithm of the sum , we represent the operation as $10+(-7)$. First we
 need to convert the operands to two's complement representation (C2):
@@
 This operation would be possible without the use of C2 representation as 10 is
-greater than 7 (see table \ref{tab:04} for this).
+greater than 7, and no signed integer is involved (see table \ref{tab:04} for
+this).
 
@@
     \hline      
-    $n_2$ & 10	& & 1010 & 3 [VERIFICAR]\\
-    $n_1$ & \underline{$\times 7$} & & \underline{111} & 1 \\
+    $n_2$ & 10	& & 1010 & 4\\
+    $n_1$ & \underline{$\times 7$} & & \underline{111} & 3 \\
     & & & 1010  & \\
@@
 representation. To illustrate this case let's calculate $10\times-7$. in C2,
-this opration becomes $01010\times11001$. Table \ref{tab:07} contains the
+this operation becomes $01010\times11001$. Table \ref{tab:07} contains the
 details. The basic difference is that the sign bits (in bold), positioned to
@@
     & & & 1010  & \\
-    & & & \underline{0000\ } & \\
-    & & & $01010$ & \\
-    & & & \underline{0000\ \ } & \\
+    & & & + \underline{0000\ } & \\
+    & & & $\,01010$ & \\
+    & & & + \underline{0000\ \ } & \\
     & & & 001010 & \\
-    & & & \underline{1010\ \ \ } & \\
+    & & & + \underline{1010\ \ \ } & \\
     Total& -70 & & \textbf{1}1011010 & 8\\
@@
 	\end{tabular}
-    \label{tab:07}
+    \label{tab:08}
 \end{table}
@@
 
+Figure \ref{fig:bitlength} shows the bit-lengths of results of the four
+operations with integer operands up to 8 bits in length. The conclusions listed
+above, are visually emphasized in the figures.
+
 When operating with signed integers, an extra bit is used for the sign.
@@
 actual operations taking place on the bitstrings as we compute with compressed
-matrices. Additional details will be provided below when we discuss the
+matrices. Additional details will be provided below, when we discuss the
 operation with matrices. Note that we are restricting the example to operations
@@
  \centering
- \includegraphics[width=12cm]{./bitlength.png}
+%  \includegraphics[width=12cm]{./bitlength.png}
  % simples.png: 4800x3600 pixel, 600dpi, 20.32x15.24 cm, bb=0 0 576 432
- \caption{bitlength of the results for each of the four operations. The axes
+ \caption{Bit-length of the results for each of the four operations. The axes
 represent the values of the operands.}
@@
 \ref{eq:02}. The elements in $A$ and $B$ are 4 and 5 bits long, respectively
-(the ``supreme minimum''). the ellipses ("$\ldots$") in the binary matrix
+(the ``supreme minimum''). The ellipses (``$\ldots$'') in the binary matrix
 correspond the extra zeros to the left. Since the strings are written into 64
@@
 
-\begin{equation}
-	S = \begin{bmatrix}
+\begin{align}
+	S &= \begin{bmatrix}
 			549649\\ 
@@
 		\end{bmatrix}
-        =
-        \begin{bmatrix}
-
+      \\ \nonumber
+      &= \begin{bmatrix}
 \ldots\underbrace{000010}_{2}\underbrace{000110}_{6}\underbrace{001100}_{12}
@@
 		\end{bmatrix}
-\end{equation}
+\end{align}
 
@@
 
-\begin{equation}\label{eq:ac2}
-	-A = \begin{bmatrix}
+\begin{align}\label{eq:ac2}
+	-A &= \begin{bmatrix}
 			4952\\ 
@@
 		\end{bmatrix}
-        =
-        \begin{bmatrix}
+        \\\nonumber
+        &= \begin{bmatrix}
  \ldots\underbrace{111111}_{-1}\underbrace{111101}_{-3}\underbrace{111011}_{-5}
@@
 
-\begin{equation}
-	D = \begin{bmatrix}
+\begin{align}
+	D &= \begin{bmatrix}
 			129\\ 
@@
 		\end{bmatrix}
-        =
-        \begin{bmatrix}
+      \\\nonumber
+        &=  \begin{bmatrix}
  \ldots\underbrace{000000}_{0}\underbrace{000000}_{0}\underbrace{000010}_{2}
@@
 \end{equation}
+\begin{small}
 \begin{equation}
         \times\times \\
-        \begin{bmatrix}
-
+        \begin{bmatrix}\mathsf{
 \ldots\underbrace{00001}_{1}\underbrace{00011}_{3}\underbrace{00111}_{7}
-\underbrace{01001}_{9}  &
-
+\underbrace{01001}_{9}}  &\mathsf{
 \ldots\underbrace{00001}_{1}\underbrace{00011}_{3}\underbrace{01110}_{14}
-\underbrace{01111}_{15} &
-
+\underbrace{01111}_{15}} & \mathsf{
 \ldots\underbrace{10100}_{20}\underbrace{11110}_{30}\underbrace{00010}_{2}
-\underbrace{00001}_{1}
-		\end{bmatrix}
+\underbrace{00001}_{1}}
+        \end{bmatrix}\nonumber
 \end{equation}
-
+\end{small}
 To calculate $P[1,1]$ which in uncompressed form is $1 \times 1 + 3 \times 3 + 5
@@
 
-\begin{equation}
-	P = \begin{bmatrix}
+\begin{align}
+	P &= \begin{bmatrix}
 			1426882688\\ 
@@
 		\end{bmatrix}
-\end{equation}
-\begin{equation}
-        =
-        \begin{bmatrix}
-
+      \\ \nonumber
+        &=        \begin{bmatrix}
 \ldots\underbrace{000001010101}_{117}\underbrace{000011001000}_{200}\underbrace{
 000010000000}_{128}\\
-
 \ldots\underbrace{000010110001}_{177}\underbrace{000100010001}_{273}\underbrace{
 001010101001}_{681}\\
-
 \ldots\underbrace{000011000001}_{193}\underbrace{000101001001}_{329}\underbrace{
@@
         \end{bmatrix}
-\end{equation}
+\end{align}
 

### 2013-09-25 plos.tex
recovered plos.tex from backup

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

\usepackage{psfrag}
\usepackage{subfigure}

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Matrix compression methods}
}
% Insert Author names, affiliations and corresponding author email.
\\
Author1$^{1}$, 
Author2$^{2}$, 
Author3$^{3,\ast}$
\\
\bf{1} Author1 Dept/Program/Center, Institution Name, City, State, Country
\\
\bf{2} Author2 Dept/Program/Center, Institution Name, City, State, Country
\\
\bf{3} Author3 Dept/Program/Center, Institution Name, City, State, Country
\\
$\ast$ E-mail: Corresponding author@institute.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}

\section*{Introduction}

The term compression can be defined as the converting process a data file in another file so that the compressed file size is less than the original \cite{salomon}. In many cases, data compression is required so that transmission process by internet can occur, even filing and information processing contained in the data file. Although it may be necessary to compress this file can be allocated on a device, for example, data allocation in a computer memory for processing.

Usually the techniques that involve data processing try to work around problems, arise in this case the the data file size, to perform processing. At present and due to the increased speed with which information is generated, it becomes necessary to develop methods which allow such management information in large scale, mostly so that no loss occurs. The information storage on a large scale, especially in digital form, can assist in advancing science \cite{lynch}.

There are different methods that perform this compression procedure, which can be classified based on their characteristics. We highlight the main information loss as characteristics during the compression process, the data manipulation procedure, methodology used, among others \cite{salomon2}.

If after compression, data suffer a information loss, this causes loss if the file is uncompressed. The compressed data can not be recovered in its original form. Therefore you can choose the method based on the ratio of the loss characteristics. The data that can be recovered in spite of presenting information loss do not invalidate its use. However, there are other processes that the data must be recovered in full.

According to the loss or not, the information manipulated, one can classify the data types on which compression can occur. When there is information loss, so that this does not undermine the use of the data, we have a method called nondestructive. An example of the data manipulated in this method are sound files (mp3), picture (JPEG) and movies (mpeg). As for the other data after being compressed, it can not have losses, they are called non-destructive. The types of manipulated file by this method are based on text and numbers \cite{salomon,salomon2}.

There is another method classification compression with respect to how it is done, its can be symmetrical or asymmetrical. In the case of symmetrical, time compression and decompression of data occurs synchronously, ie, it take the same time for being compressed and decompressed. In this classification, the fit into based dictionary-algorithm, such as LZW, which use dictionaries for performing the compression. Already another classification, the asymmetric demand different times in implementing the compression and decompression processes, as an example it can highlight the compressing audio method, mpeg \cite{salomon,salomon2}.

Another feature that is considered to distinguish compression types is related to the way in which the process runs. In this new approach, there are methods that are considered adaptive and non-adaptive. As the compression process is executed, the method can be adapted to optimize the processing. Considering this new feature, we can highlight the Golomb code among and other as non-adaptive. Already on the other hand, we have dictionary-based algorithms, LZ77 and LZ78, and adaptive as those based on Huffman coding, arithmetic coding, and others, can be both adaptive and non-adaptive \cite{salomon, salomon2}.

Besides the problem involved with compression, there is another related to the way in which data large volumes must be represented in memory. The procedures described up to now, in large part only function to compress the data file and in some cases access them. When these data should be accessed for manipulation, there is a need to be represented, in the computer case, in memory. From there arise a problem with the representation of this large volume of data in memory. Among the many techniques developed to solve this problem, there is the technique of memory-mapped files that seeks to use main memory to reference the data file in disk as the data are requested, they are loaded into main memory. This technique allows data large volumes can be handled, but still there is a limit, in which case the memory size. Another problem is related to access time data to disk, which is relatively higher compared to the memory access. Some programming languages ​​like Python, R and other, allow the use of this feature. 
In Python this feature is available in the standard library, already in the R this feature can be used by the package bigmemmory \cite{big}.

In this paper we present two methods that allow the compression of large data sets, such as the type of big data, for direct manipulation of files into memory. These methods allow direct access to data in compressed form, its handling and recovery occurs without loss. The proposed method also allows to use the techniques already presented to perform a subsequent compression process only if these data should be stored on physical media. We describe the process of compression and decompression, using as an example strips of bits to represent numerical arrays, the method sparse string \cite{salomon,salomon2}, to illustrate the methods. We also present a discussion of some results on the characteristics of each proposed method and elucidated the best options in which each can be applied. We have also emphasized that the methods allow information to be accessed after the compression process, which allows the handling and updating, if necessary. At the end of the paper we present their conclusions about using of 
the proposed methods.

% Results and Discussion can be combined.
\section*{Results}

Consider a matrix such that the elements are integers. The results will be presented that indicate the compression process propose two methods. These methods create a new matrix, so that the values ​​of the original matrix are stored in a new matrix, only using less computational resources, in this case memory to be represented.

To check the efficiency of these methods, we consider that this matrix is composed of numbers generated from random variables that follow a probability distributions mixture.

Now for simplicity and functionality will be used by a probability distributions mixture Beta. The probability density function (pdf) of the mixture of two Beta distributions is shown in equation \ref{eq:52}. The shape parameters $\alpha$ and $\beta$ will be correlated to the bit number to represent the values ​​in the data matrix to be compressed. Already the parameter $\lambda$ defines the proportion of each distribution make mixture. From this value range​​, one can by distributing generate other and different sets of values ​​to fill the matrix, then one can sweep all possibilities and estimating the efficiency of the compression process.

\begin{equation}\label{eq:52}
 f(x,\alpha_1,\beta_1,\alpha_2,\beta_2) = \lambda\frac{\Gamma(\alpha_1+\beta_1)}{\Gamma(\alpha_1)\Gamma(\beta_1)}x^{\alpha_1-1}(1-x)^{\beta_1-1} + (1-\lambda)\frac{\Gamma(\alpha_2+\beta_2)}{\Gamma(\alpha_2)\Gamma(\beta_2)}x^{\alpha_2-1}(1-x)^{\beta_2-1}
\end{equation}

Assume $ X_1 \sim Beta(\alpha_1,\beta_1)$ and $X_2 \sim Beta(\alpha_2,\beta_2)$, such that $X$ is mixture resulting of values $X_1$ and $X_2$, then $X$ has the probability density function described in Equation \ref{eq:52}. Therefore, their mean and variance for a variable $X$ that follows a Beta distribution mixture are shown in Equations \ref{eq:53} and \ref{eq:54}.

\begin{equation}\label{eq:53}
 E(X) = \lambda E(X_1) + (1-\lambda) E(X_2) = \lambda \frac{\alpha_1}{\alpha_1+\beta_1} + (1-\lambda)\frac{\alpha_2}{\alpha_2+\beta_2}
\end{equation}

\begin{equation}\label{eq:54}
 Var(X) = \lambda Var(X_1) + (1-\lambda) Var(X_2) + \lambda (1-\lambda) (E(X_1)^2 - E(X_2)^2)
 \end{equation}

From these values​​, which represent the expected value with respect to the parameters $\alpha$ and $\beta$, holds up a conversion for representing the bit number required to represent the values ​​in matrix to be compacted. With this, it becomes possible to describe the efficiency as a function of the parameters $\alpha$ and $\beta$ and the corresponding expectation and variance values ​​generated from this parameter. For a sample of size n, such that the sample fills a matrix with M with $r$ rows and $c$ columns ($n = r  \times c$), the expected value for the matrix M can be calculated by the equation \ref{eq:53}. From equation \ref{eq:55}, we can calculate the expected value for different values ​​of $\alpha$ and $\beta$ by performing a transformation to adjust the values ​​generated from the mixture of Betas distributions belong to the interval $[1,64]$. Recalling that the values ​​generated are truncated and only the integer values ​​will be used.

\begin{equation}\label{eq:55}
  \eta(\alpha,\beta) \approx 64 \times E(X) + 1
\end{equation}

Defined method for generation of samples to be analyzed, it now remains to estimate the expected efficiency for different values ​​of $\alpha$ and $\beta$. To do this, we use the equation \ref{eq:22} and \ref{eq:37}, regarding the efficiency of proposed methods 1 and 2. At first, we assume that $\lambda=0$, ie, the M values matrix  are only generated from a Beta distribution. In Figure \ref{fig:02} you can check the different distributions that can be generated from the Beta distribution, only changing the parameters $\alpha_1$ and $\beta_1$ . The histograms represented the distribution of the bit numbers used to represent a sample with 10,000 elements of an array.

\begin{figure}[ht]
  \centering
  \subfigure[$\alpha=1,\beta_1=1$]{
  \includegraphics[scale=0.25,angle=-90,clip]{fig02}
  }
  \subfigure[$\alpha=1,\beta=32$]{
  \includegraphics[scale=0.25,angle=-90,clip]{fig03}
  }
  \subfigure[$\alpha=32,\beta=1$]{
  \includegraphics[scale=0.25,angle=-90,clip]{fig04}
  }
  \subfigure[$\alpha=64,\beta=64$]{
  \includegraphics[scale=0.25,angle=-90]{fig05}
  }
  \caption{Histograms constructed from samples with 10,000 elements, generated from the Beta distribution. Below each histogram is possible to verify the parameters used.}
  \label{fig:02}
\end{figure}

As we intend to estimate the efficiency, the equations should be redefined in the sample function as can be seen in Equations \ref{eq:56} and \ref{eq:57}, methods 1 and 2, respectively, and $E(B)$ the expected value for the bit number (B) of number sample and $k$ largest expected value this sample.

\begin{equation}\label{eq:56}
 E(\zeta_1) = 1 - \frac{k}{64}
\end{equation}

\begin{equation}\label{eq:57}
 E(\zeta_2) = 1 - \frac{E(B)}{64} - \frac{k}{64}
\end{equation}

To define, based on a sample of the matrix M, which method should be used, we calculate the difference $R=E(\zeta_1)-E(\zeta_2)$. When $R>0$, the first method is more efficient than the method in the second data compression. At market $R<0$, the second method is more efficient. And otherwise, the two methods are the same.

In this first analysis, it is assumed that the values ​​of $\alpha$ and $\beta$ belong strictly to the interval $[1,64]$, the data are generated by modified Beta distribution $Beta(\alpha,\beta)$ and its representing the bit values required to represent elements of a matrix M. Generates three samples with size 10,000, it estimated efficiency values through the equations \ref{eq:56} and \ref{eq:57} and it calculated the difference $R$.

In Figure 3, you can verify the efficiency of the first (Figure (b)) and second (Figure (c)) methods, the relationship between them (Figure (a)) and the regions defined by the parameters alpha and beta, in which method is more efficient than the other (Figure (d)). In the Figures (b) and (c), it can observe that both methods are effective in the process of data compression, it reaching over 80 \% compression. Note that as the average (color scale) of the values ​​generated decreases, the compression rate increase, because the smaller the average value, the smaller bit number needed to represent these values. In the Figure (a) it is possible to verify the relationship between the methods defined by the difference $R$. Notice that most of the values ​​of the surface defined by R are negative, ie, the second method is more efficient than the first for most combinations. However, there is a small region, which can be seen in Figure (d), color white, indicating that for some values ​​generated from a modified 
beta distribution 
with parameters $\alpha$ and $\beta$, the first method is more efficient. This region in Figure (A) has red color and it indicates the average of data generated, then the bit number to represent a value in this region is larger and as a consequence first method makes more feasible.

\begin{figure}[h]
  \centering
%   \subfigure[$\alpha=1,\beta_1=1$]{
%   \includegraphics[scale=0.28,angle=-90,clip]{fig02}
%   }
%   \subfigure[$\alpha=1,\beta=32$]{
%   \includegraphics[scale=0.28,angle=-90,clip]{fig03}
%   }
%   \subfigure[$\alpha=32,\beta=1$]{
%   \includegraphics[scale=0.28,angle=-90,clip]{fig04}
%   }
%   \subfigure[$\alpha=64,\beta=64$]{
%   \includegraphics[scale=0.28,angle=-90]{fig05}
%   }
  \includegraphics[scale=0.7]{simples.eps}
  \caption{Comparisons between methods I and II in relation to efficiency in the compression process. In the Figure (a) can see the difference between the efficiencies of each method $R$. It is verified that practically the entire surface has $R<0$, that is, the second method is more feasible for compressing the information generated from a Beta distribution modified $Beta(\alpha,\beta)$. However there is a region, which can be seen in Figure (d), color white, its indicates the values ​​generated from values ​​determined by $\alpha$ and $\beta$, the first method is more efficient. This region can be identified in Figure (a) in red, indicating that the generated values ​​have larger average ​​and require more bits to be represented. This demand makes the first method more feasible for the values ​​generated in this configuration. In Figures (b) and (c), it can be seen that both methods are effective with respect to high data compression rates generated from a Beta distribution $Beta(\alpha,\beta)$. Note, in 
both figures, the value of the efficiency of the compression increases with reduction of the average number of generated. This is due to reducing the bit number required to represent those values.}
  \label{fig:03}
\end{figure}

Another analysis can be performed from the combination of two variables, $B_1$ and $B_2$, following a Beta distribution with parameters $Beta(\alpha_1$,$\beta_1)$ and $Beta(\alpha_2$,$\beta_2)$, respectively, it can calculate the bit number (B) to the mixture of these two distributions. To simplify operations, it is considered that $\lambda = 0.5$, ie, the mixture will have half the numbers generated from the distribution $Beta(\alpha_1,\beta_1)$ and the other part distribution $Beta(\alpha_1,\beta_1)$. The expected value for this mixture is shown in Equation \ref{eq:58}.

\begin{equation}\label{eq:58}
 E(B) = 0.5 E(B_1) + 0.5 E(B_2)
\end{equation}

Now, we must redefine the functions for calculating the efficiency. Some modifications should be considered, the first being related to Equation \ref{eq:56} which represents the expected value for efficiency using the first method. Thus, the function is defined only in terms of $B$, since for this method, the important is only the bit greater number than a demand matrix element M (Equation \ref{eq:59}). As for the second method​​, emphasizing that the constant $k$, is a function of $B$, again because they are the highest values ​​for represent an element of M (Equation \ref{eq:60}).

\begin{equation}\label{eq:59}
 E(\zeta_1) = 1 - \frac{k(E(B))}{64}
\end{equation}

\begin{equation}\label{eq:60}
 E(\zeta_2) = 1 - 0.5\frac{E(B_1)}{64} - 0.5\frac{E(B_2)}{64} - \frac{k(E(B))}{64}
\end{equation}

How the values ​​of the alphas and betas strictly belong to the interval $[1,64]$, ie, representing the values ​​of bits required to represent elements of a matrix M, generating a sample with size $n=10,000$ and estimated efficiency values through the equations \ref{eq:59} and \ref{eq:60}, and soon after, we calculate the efficiency difference $R$ of values ​​calculated.

Before you can calculate the difference of efficiencies, we will analyze some data histograms, constructed from samples with 10,000 elements, generated from beta mixtures distributions. For this, we consider that the values ​​of $\alpha_1$, $\beta_1$, $\alpha_2$ and $\beta_2$ described in Figures \ref{fig:02A} and \ref{fig:02B}. One can see that with different values ​​for the alphas and betas, it becomes possible to construct different histograms. Each histogram of this regards the bit distribution needed to fill a matrix. For each of the samples generated from alphas and betas, their efficiency is calculated methods I and II. The comparison between the methods can be seen in Table \ref{tab:06}. The results indicate that over 90 \% of the samples with method II was more efficient in the data compression process. This result indicates that for matrices, in which mixtures can be associated probability distributions method II is the most indicated. Analyzing the efficiencies histogram, Figure \ref{fig:1415} is 
verified that the method II (Figure \ref{fig:15)}, on average, presents greater efficiency values ​​in the data compression process of the method I (Figure \ref{fig:14}). Again confirming the results presented. Analyzing Figure \ref{fig:1617}, it can be seen that the regions but which the mean values ​​of mixed distributions, represented by $\mu_1$ and $\mu_2$, indicate that for high values​​, the method I is more appropriate (Figure \ref{fig:16}), provided that the variability of the data is small (Fig. \ref{fig:17}).

\begin{figure}[h]
  \centering
  \subfigure[$\alpha_1=1,\beta_1=1,\alpha_2=1,\beta_2=1$]{
  \includegraphics[scale=0.28,angle=-90,clip]{fig06}
  }
  \subfigure[$\alpha_1=1,\beta_1=32,\alpha_2=32,\beta_2=1$]{
  \includegraphics[scale=0.28,angle=-90,clip]{fig07}
  }
  \subfigure[$\alpha_1=32,\beta_1=32,\alpha_2=32,\beta_2=32$]{
  \includegraphics[scale=0.28,angle=-90,clip]{fig08}
  }
  \subfigure[$\alpha_1=64,\beta_1=32,\alpha_2=32,\beta_2=64$]{
  \includegraphics[scale=0.28,angle=-90]{fig09}
  }
  \caption{Histograms constructed from samples with 10,000 elements, generated from the combination of two distributions Betas. Below each histogram is possible to verify the parameters used..}
  \label{fig:02A}
\end{figure}

\begin{figure}[h]
  \centering
  \subfigure[$\alpha_1=64,\beta_1=48,\alpha_2=1,\beta_2=48$]{
  \includegraphics[scale=0.28,angle=-90]{fig10}
  }
  \subfigure[$\alpha_1=16,\beta_1=46,\alpha_2=49,\beta_2=64$]{
  \includegraphics[scale=0.28,angle=-90]{fig11}
  }
  \subfigure[$\alpha_1=16,\beta_1=16,\alpha_2=16,\beta_2=49$]{
  \includegraphics[scale=0.28,angle=-90]{fig12}
  }
  \subfigure[$\alpha_1=1,\beta_1=16,\alpha_2=1,\beta_2=49$]{
  \includegraphics[scale=0.28,angle=-90]{fig13}
  }
  \caption{Histograms constructed from samples with 10,000 elements, generated from the combination of two distributions Betas. Below each histogram is possible to verify the parameters used.}
  \label{fig:02B}
\end{figure}

\begin{table}[h]
 \centering
 \caption{Efficiency comparison in the compression data process methods I and II for sample data generated from a mixture of beta distributions parameterized alphas and betas. It was enumerated methods in which each method was more efficient.}
 \begin{tabular}{ccc}
  \hline 
  Methods  & Sample   & Percentage \\
  \hline
  I	   & 592	& 0.9034\% \\
  II	   & 64944	& 99.0966\% \\
  \hline
  Total    & 65536	& 100\% \\
  \hline
 \end{tabular}
 \label{tab:06}
\end{table}

\begin{figure}[h]
  \centering
  \subfigure[Efficiency histogram of the method I]{
  \includegraphics[scale=0.4,angle=-90]{fig14}
  \label{fig14}
  }
  \subfigure[Efficiency histogram of the method II]{
  \includegraphics[scale=0.4,angle=-90]{fig15}
  \label{fig15}
  }
  \caption{Efficiency histograms of the methods I and II. It is verified that the method II gives an average efficiency greater than method I.}
  \label{fig:1415}
\end{figure}

\begin{figure}[h]
  \centering
  \subfigure[Efficiency of the methods I and II as a function of the mean distributions mixed.]{
  \includegraphics[scale=0.4]{fig16}
  \label{fig16}
  }
  \subfigure[Efficiency of the methods I and II as a function of the variance distributions mixed.]{
  \includegraphics[scale=0.4]{fig17}
  \label{fig17}
  }
  \caption{Mean values ​​of the distributions mixed in which method I and II are more efficient. Note that only a small region with high values ​​for the mean distributions mixed that the method  I is more efficient, in other regions the method II is highlighted. With respect to variance of distributions mixed, one can verify that the method I is more efficient for low values ​​of the variance.}
  \label{fig:1617}
\end{figure}

% Até o momento, o método 2 é mais eficaz no processo de compressão dos dados. Isso ocorre, pois caso exista alguma variância no conjunto de dados da matriz M, o método 2 consegue reduzir a perda dos bits desnecessários para representar alguns números, promovendo a economia dos bits. Existem casos nos quais a variância entre os números é nula ou infíma. Nesses casos, o método 1 é mais eficaz, pois para um mesmo número de bits, pode-se representar diferentes números, por exemplo, com 3 bits é possível representar os valores de 0 até 7. Portanto, caso uma matriz seja constituídas por valores entre 0 e 7, o método 1 será mais eficaz, pois irá representar os números por meio de um número menor de bits. Para simular a situação descrita e as outras com as mesmas caracterísicas foi realizado uma simulação, gerando 10,000 valores para preenncher uma matriz M, de tal forma que esses valores demandem a mesma quantidade de bits para serem representados. Os resultados para essa simulação estão representados nas Figuras \
% ref{fig18} e \ref{fig19}. Na Figura \ref{fig18} pode-se verificar que o método 1 é mais eficiente no processo de compressão dos valores da matriz M para diferentes números de bits, pois não existem diferenças entre os números de bits necessários para representar cada um dos valores dos elementos da matriz M. Como possuem a mesma quantidade de bits, o método 1 torna-se mais eficaz nesse caso. Na Figura \ref{fig19}, pode-se observar o comportamento da diferença da eficiência entre os dois métodos em função do número de bits. Nessa figura é possível verificar que a medida que os valores são representados por mais bits, o método 1 é mais eficiente, mas quando os valores se aproximam de 64 bits, caem buscamente. Essa queda ocorre que quando o número de bits atinge o valor 64, nenhum dos dois métodos é eficaz para representar os valores, e portanto não existe compressão dos dados. Vale ressaltar que com esse experiment, o método II não foi mais eficaz em nenhuma situação do que o método I.

Until now, the method 2 is more effective in the data compression process. This occurs, because if there is some variance in the data set matrix M, the second method can reduce the unnecessary bits loss to represent some numbers, promoting the bit economy. There are cases where the variance between the numbers is null or insignificant. In such cases, the first method is more efficient, since for the same bit number can represent different numbers, for example, 3 bits can represent the values ​​0 through 7. Therefore, if one matrix is ​​constituted by values ​​between 0 and 7,  method I would be more efficient to represent the numbers by a smaller bit number. To simulate the situation described above and the other with the same characteristics a simulation was carried out, generating 10,000 values ​​for a matrix M, so that these values ​​demanding the same amount of bits to be represented. The results for this simulation are shown in Figures \ref{fig18} and \ref{fig19}. In Figure \ref{fig18} can be verified that the first method is more efficient in the compression process of the matrix M values ​​for different bit numbers since there are no differences between the bit numbers required to represent each element value of the matrix M. Since they have the same bit number, the first method becomes more effective in this case. In Figure \ref{fig19}, we can observe the behavior of the difference in efficiency between the two methods depending on the bit number. In this picture you can see that the measured values ​​are represented by more bits, method 1 is more efficient, but when the values ​​are close to 64-bit, decrease sharply. This decrease occurs when the bit number reaches the value 64, neither method is effective to represent the values​​, and so there is no compression of the data. It is noteworthy that in this experiment, the method II was no more effective in any situation than method I.

\begin{figure}[h]
  \centering
  \subfigure[Comparison of the efficiency of the methods I e II.]{\includegraphics[scale=0.3]{fig18}\label{fig18}}
  \subfigure[Difference in efficiency of the methods I e II.]{\includegraphics[scale=0.3]{fig19}\label{fig19}}
  \caption{Efficiency of the methods I and II the data compression process of a matrix of numbers which require the same amount of bits to be represented.}
  \label{fig:1819}
\end{figure}

% A simulação realizada anteriormente, ocorre em casos muitos específicos, mas mesmo assim, o método I consegue comprimir os valores gerados. Agora vamos analisar o comportamento do processo de compressão realizados sobre amostras geradas a partir da mistura de dois conjunto de dados. Os valores gerados para cada conjunto de dados, que juntos totalizam 10,000 valores, formam a matriz M. Todos os valores são representados por meio de duas quantidade de bits, sendo definidas como $b_1$ e $b_2$ para os métodos 1 e 2, respectivamente. A diferença agora é que existe um $\lambda$, que define a quantidade de elementos de cada grupo para formar a matriz M (Equação \ref{eq:61}).

The simulation performed previously, occurs in very special cases, but even so, the method I can compress values ​​generated. Now, let us consider the behavior of the compression process performed on samples generated from the mixing of two dataset. The values ​​generated for each set of data, which together total 10,000 values, forming the matrix M. All values ​​are represented by two bit quantity and is defined as $B_1$ and $B_2$ to methods I and II, respectively. The difference now is that there is a $\lambda$, which defines the element number in each group to form the matrix M (Equação \ref{eq:61}).

\begin{equation}\label{eq:61}
  n = \lambda b_1 + (1-\lambda) b_2
\end{equation}

% Os resultados podem ser verificados na Figura \ref{fig:2125}. Nas Figuras \ref{fig20}, \ref{fig22} e \ref{fig24}, pode-se verificar as superfícies que descrevem o compormento da eficiência dos processos 1 e 2 para os valores de $\lambda=0.1$, 0.5 e 0.9, respectivamente. Nessas figuras é possível verificar que existe um comportamento constante para o método 1,cujo comportamento da eficiência é crescente para valores próximos ou iguais de $b_1$ e $b_2$, já para os demais valores, nos quais ocorre uma diferença entre o número de bits, a eficiência cai. O método 2, diferente do método 1, apresenta maiores valores de eficiência nas regiões nas quais os número de bits $b_1$ e $b_2$ são discrepantes. A diferença das eficiências dos métodos podem ser verificadas nas Figuras \ref{fig21}, \ref{fig23} e \ref{fig25}. Observe que a curva assume valores positivos para a diferença, logo o método 1 é o melhor. Uma justificativa para esse comportamento é que os números de bits $b_1$ e $b_2$ são próximos. Já para os casos, 
% nos quais os valores de $b_1$ e $b_2$ são discrepantes, o método 2 é o mais eficiente.

The results can be verified in Figure \ref{fig: 2125}. In Figures \ref{fig20}, \ref{fig22} and \ref{fig24}, it can be seen surfaces that describe the behavior of the efficiency of method I and II for the values ​​of $\lambda = 0.1$, 0.5 and 0.9, respectively. In these figures, it can be seen that there is a constant behavior for method I, whose behavior is increasing efficiency to values ​​close or equal to $B_1$ and $B_2$, while for other values​​, where there is a difference between the bit number, the efficiency decrease. The method II, different of the method I, has higher efficiency values ​​in regions where the number of bits $B_1$ and $B_2 $ are discrepant. The difference in the efficiencies of the methods can be verified in Figures \ref{fig21}, \ref{fig23} and \ref{fig25}. Note that the curve takes positive values ​​for the difference, so the method is the best one. A justification for this behavior is that the bit numbers $B_1 $ and $ B_2 $ are close. But for the cases, where the values ​​of $B_1$ and $B_2 $ are discrepant, method II is the most efficient.

\begin{figure}[h]
  \centering
  \subfigure[Comparison of the efficiency of the methods I and II.($\lambda=0.1$)]{
  \includegraphics[scale=0.3,angle=-90]{fig20}
  \label{fig20}
  }
  \subfigure[Difference in efficiency of the methods I and II.]{
  \includegraphics[scale=0.3,angle=-90]{fig21}
  \label{fig21}
  }
  \subfigure[Comparison of the efficiency of the methods I and II.($\lambda=0.1$)]{
  \includegraphics[scale=0.3,angle=-90]{fig22}
  \label{fig22}
  }
  \subfigure[Difference in efficiency of the methods I and II.]{
  \includegraphics[scale=0.3,angle=-90]{fig23}
  \label{fig23}
  }
  \subfigure[Comparison of the efficiency of the methods I and II.($\lambda=0.1$)]{
  \includegraphics[scale=0.3,angle=-90]{fig24}
  \label{fig24}
  }\subfigure[Difference in efficiency of the methods I and II.]{
  \includegraphics[scale=0.3,angle=-90]{fig25}
  \label{fig25}
  }
  \caption{Efficiency of the methods I and II in the data compression process of a matrix of numbers which require the same amount of bits to be represented.}
  \label{fig:2125}
\end{figure}

The two methods presented allow the storage and information manipulation in the compressed form, which provides an data managing large amounts optimization. The next step in the work development is the implementation of the proposed methods, as well as tests to be performed on data large volumes.

\section*{Discussion}

\subsection*{Which method to use?}

To know the best methodology to be applied to compress the data matrix, it becomes necessary to evaluate the numbers present in the matrix and as a consequence verify the bits number behavior needed to represent them. The first methodology provides greater efficiency when the bit number to store the information in the matrix are close. The second approach is to try to minimize losses when the numbers have outliers. For example, when it is desired to store values ​​of different magnitudes, such as 0, 1 and 1023. In this values range​​, it can be seen that the values ​​0 and 1 only need 1 bit, while the number 1,023 demand 10 bits. Following the first methodology, we would have to divide the strip into regions of 10 bits, which for storing the values ​​0 and 1, causes loss occurs 9 bits (Equation \ref{eq:19}), not allowing for more efficient compression. For work around this problem the second methodology can be applied, promoting data compression. In this case, instead of using 10 bits would be used 
only 5, with 1 bit to store the value and the other 4 to store the size of the region that this amount will be allocated (Equation \ref{eq:20}). See that the second method requires only 24 bits to represent the three values​​, unlike the first, which requires 30 bits.

\begin{equation} \label{eq:19}
 0000000000000000000000000000000000\underbrace{1111111111}_{1023}\underbrace{\overbrace{000000000}^{\text{Lost}}1}_{1}\underbrace{\overbrace{000000000}^{\text{Lost}}0}_{0}
\end{equation}

\begin{equation} \label{eq:20}
 0000000000000000000000000000000000000000\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1}_{1}\underbrace{0001}_{1}\underbrace{0}_{0}\underbrace{0001}_{1}
\end{equation}

To verify the efficiency behavior ($\zeta$) compression process data, an analysis will be conducted on the bit number behavior used to store the information in a matrix with $r$ rows and $c$ columns. Consider the following equation \ref{eq:21}. This equation gives a percentage of the level of data compression.

\begin{equation} \label{eq:21}
 \zeta = \frac{\text{bits reserved } - \text{bits really needed}}{\text{bits reserved}}
\end{equation}

Consider a matrix $M_{r \times c}$, with $r$ rows and $c$ columns, we have in total $rc$ elements. To represent these elements in the worst case the, we have to reserve $64 \times rc$ bits. However, if these elements demanding only $b$ bits each represent, in total, would require $b \times rc$ bits. The function efficiency is presented in Equation \ref{eq:22}.

\begin{equation}\label{eq:22}
 \zeta = \frac{64 \times rc - b \times rc}{64 \times rc}
\end{equation}

Simplifying the Equation \ref{eq:22}, we obtain the equation \ref{eq:23}. The equation resulting from the simplification process does not depend on element number, but the bit number that is used.

\begin{equation} \label{eq:23}
 \zeta(b) = \frac{64  - b }{64}
\end{equation}

The equation \ref{eq:23} supplies in a direct way to efficiency in the worst and best case compression. In the worst case, $b = 64$, ie, at least one of the elements of the matrix M require 64 bits to be represented, resulting in $\zeta(64)=0\%$. As for the best case, $b=1$, the greater efficiency that can be achieved is $\zeta(1)=98.43\%$. For intermediate cases, it can be seen in Figure \ref{fig:01} efficiency behavior according to the bit number needed ($b$). This efficiency analysis  with respect to the first compression method presented in which all elements require the same bit number to be represented in memory.

\begin{figure}[h]
 \centering
 \psfrag{e}{$\xi$}
 \includegraphics[scale=0.5]{fig01}
 \caption{Compression efficiency behavior with respect to bit number.}
 \label{fig:01}
\end{figure}

For the second method, wherein the compression depends on the value to be stored, it should perform another analysis. This method considers that the elements demanding different bit number and this difference will determine the bit economy. We will analyze this configuration considering some numbers groups that can be represented by the same bit number as possible and when compared with the first method.

Again consider the matrix $M_{r \times c}$, only the $rc$ elements are divided into $g$ groups, where each group has $f_i$ numbers. For each element $b_i$ bits are reserved , corresponding to bits required to represent the desired number and $k$ bits, which represent the bit number to be reserved for storing the strip $b_i$ bits. The efficiency function has to be modified as shown in Equation \label{eq:24}.

\begin{equation}\label{eq:24}
 \zeta = \frac{64 \times rc - \sum_{i=1}^{g} ( b_i + k ) \times f_i }{64 \times rc} 
\end{equation}

Equations \ref{eq:25} to \ref{eq:27} are realized manipulations to adjust the equation. The equation that will provide the compression percentage  is apresented in Equation \ref{eq:30}.

\begin{equation} \label{eq:25}
 \zeta = \frac{64 \times rc - \sum_{i=1}^{g} ( b_i \times f_i + k \times f_i )}{64 \times rc} 
\end{equation}

\begin{equation} \label{eq:26}
 \zeta = \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  -\sum_{i=1}^{g}  k \times f_i }{64 \times rc} 
\end{equation}

\begin{equation} \label{eq:27}
 \zeta = \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  - k \times\sum_{i=1}^{g}  f_i }{64 \times rc} 
\end{equation}

\noindent as

\begin{equation} \label{eq:28}
 \sum_{i=1}^{g} f_i = rc
\end{equation}

\begin{equation} \label{eq:29}
 \zeta = \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  - k \times rc }{64 \times rc} 
\end{equation}

\begin{equation}\label{eq:30}
 \zeta = 1 - \frac{\sum_{i=1}^{g}  b_i \times f_i }{64 \times rc} - \frac{k}{64} 
\end{equation}

\noindent whereupon

\begin{equation} \label{eq:31}
 k = \pi(\text{ greater value of }M_{r \times c})
\end{equation}

In a first example, we suppose again with the matrix M with $rc$ elements, such that the half of elements demanding demand 1 bits to represented and the other half 64 bits to be represented. So in this case i = 2, $b_1=1 $bits, $b_2=64$ bits and $f_i = \frac{rc}{2}$ for $i = 1 \text{ and } 2$. As the largest demand 64-bit value, then $k=7$, applying the Equation \ref{eq:31}. Efficiency in the bit economy can be calculated using Equation \ref{eq:30}, after performing the operations, equations \ref{eq:32} to \ref{eq:36}, we obtain the approximate value of $38.29\%$ for compression efficient.

\begin{equation} \label{eq:32}
 \zeta = 1 - \frac{\sum_{i=1}^{2}  b_i \times f_i }{64 \times rc} - \frac{7}{64} 
\end{equation}

\begin{equation} \label{eq:33}
 \zeta = 1 - \frac{  1 \times \frac{rc}{2} + 64 \times \frac{rc}{2} }{64 \times rc} - \frac{7}{64} 
\end{equation}

\begin{equation} \label{eq:34}
 \zeta = 1 - \frac{  32.5  \times rc }{64 \times rc} - \frac{7}{64} 
\end{equation}

\begin{equation} \label{eq:35}
 \zeta \approx 1 - 0.5078 - 0.1093
\end{equation}

\begin{equation} \label{eq:36}
 \zeta \approx 38.29\%
\end{equation}

See that the economy efficiency in using the second method was approximately 38\%. What is will influence the size of each number groups, which can be considered as the frequency with which these numbers appear. In the problem presented earlier, we considered two number groups, in which each group had half of elements to be compressed. Let's change the proportion of the groups, but now we use the classical definition of probability to assist in the analysis.

According to the definition the Laplace to the probability refers to the relationship between the case numbers and probable possible case numbers, for example, the relationship between the elements of a subset A,  probable case number, with the elements of set $\Omega$ , the possible case number, such that $A \in \Omega$. Thus, the probability of the subset A is defined as:

\begin{equation}
 P(A) = \frac{\text{Probable case numbers}}{\text{Possible case numbers}} = \frac{\text{Element number in A}}{\text{Element number in }\Omega}
\end{equation}

With this definition it is possible to associate the problem studied. The sample $\Omega$ space is formed by the elements of the matrix to be compacted. The subsets are formed by number groups in the array. Therefore, following this reasoning, we can rewrite the equation \ref{eq:30}, being defined in Equation \ref{eq:37}. In Equation \ref{eq:37}, the term $\frac{f_i}{rc}$ is replaced by $p_i$, and represents the probability of elements from group i forming the matrix M.

\begin{equation}\label{eq:37}
 \zeta = 1 - \frac{\sum_{i=1}^{g}  b_i \times p_i }{64} - \frac{k}{64} 
\end{equation}

\noindent with

\begin{equation}\label{eq:38}
 p_i = \frac{f_i}{rc}
\end{equation}

With the equation \ref{eq:37} can analyze the group proportion influence in the compression efficiency. Consider again the matrix M with $r$ rows and $c$ columns, totaling $rc$ elements. Now consider two groups of numbers whose ratio is defined by Equation \ref{eq:39}.

\begin{equation} \label{eq:39}
 p_1 + p_2= 1\text{, } p_1 \geq 0 \text{ and } p_2 \geq 0
\end{equation}

\noindent Thus, $p_1$ and $p_2$ define the number proportion of the matrix M requiring 1 bit and 64 bits to be represented, respectively. Thus, efficiency is defined in Equation \ref{eq:40}.

\begin{equation}\label{eq:40}
 \zeta = 1 - \frac{1 \times p_1  + 64 \times p_2}{64} - \frac{7}{64} 
\end{equation}

Now, we can verify what the proportions allowing a higher and lower compression from the equation \ref{eq:40}. If $\zeta=1$, then the efficiency is maximum and if $\zeta=0$, a efficiency is minimum. To calculate the values ​​of $p_1$ and $p_2$ so that efficiency is maximum, we must solve the system shown in Equation \ref{eq:41}. This system appears from the combination of equations \ref{eq:37}, with $\zeta=0$, with two number sets, one representing the elements represented by 1 bit and other  64-bit, with probability of $p_1$ and $p_2$, respectively with the equation \ref{eq:39}.

\begin{equation}\label{eq:41}
  \left
  \{\begin{matrix}
    p_1 + p_2 = 1\\ 
    p_1+64p_2 = 7
  \end{matrix}
  \right.
\end{equation}

Solving the system, we have that when $p_1=0.9047$ and $p_2=0.0953$ efficiency is maximum. Therefore, when approximately 90\% of the numbers require only one bit to be represented and approximately 10\% require 64 bits, the efficiency is maximum in this case equal to 87.5\%. Already in case the efficiency is minimal, we must solve the system shown in Equation \ref{eq:42}.

\begin{equation}\label{eq:42}
  \left
  \{\begin{matrix}
    p_1 + p_2 = 1\\ 
    p_1+64p_2 = 57
  \end{matrix}
  \right.
\end{equation}

Thus, when $p_1=0.1111$ and $p_2=0.8889$ is efficiency minimal according to the second method. For other combinations verify to Table \ref{tab:02}. Evaluating the table, you can verify two proportions negative when ($p_1$,$p_2$) assume the values ​​(0,1) and (0.1,0.9). It can be interpreted as the need to allocate more memory to use the method, since for the lower efficiency is reached, the value should be $p_1=0.9047$ e $p_2=0.0953$. In Table, we have an efficiency of 8\% ​​in the case of data compression, following the second proposed method, it should have in matrix 20\% of the elements represented with 1 bit and 80\% represented by 64 bits. If these proportions were maintained and the method was applied first, the efficiency would be minimal in all cases, because the greatest value is represented by 64 bits, which makes the division of the strip to store other values​​.

\begin{table}[h]
 \centering
 \caption{Combinations $p_1$ and $p_2$ to calculate the efficiency.}
 \begin{tabular}{ccc}
  \hline 
  $p_1$  & $p_2$ & Efficiency \\
  \hline
  0.0	&1.0    &-0.109 \\
  0.1	&0.9	&-0.010 \\
  0.2	&0.8	&0.087 \\
  0.3	&0.7	&0.185 \\
  0.4	&0.6	&0.284 \\
  0.5	&0.5	&0.382 \\
  0.6	&0.4	&0.481 \\
  0.7	&0.3	&0.579 \\ 
  0.8	&0.2	&0.678 \\
  0.9	&0.1	&0.776 \\ 
  1.0	&0.0	&0.875 \\
  \hline
 \end{tabular}
 \label{tab:02}
\end{table}

Until the moment, we used only two number groups, just a question being examined is how different combinations of these numbers may influence the effectiveness of the compaction process by using two proposed methods.

As an example, we calculated the efficiency for different number groups with the same element number, or rather the same frequency. Note below the efficiencies calculated:

\begin{itemize}
  \item 3 number groups that require numbers 1, 32 and 64 bits, efficiency $\zeta=0.3854$,
  \item 5 number groups that require numbers 1, 16, 32, 48 and 64 bits, efficiency $\zeta=0.3875$,
  \item 8 number groups that require numbers 1, 8, 16, 24, 32, 40, 48, 56 and 64 bits, efficiency $\zeta=0.3888$
\end{itemize}

When groups are proportionality, ie, have the same frequency, as consequence the same probability of occur, the efficiency is almost the same, using the second method apresented.

Leveraging the idea frequency, which may be generalized as the probability that a number sets can occur, we use probability distributions, in this case the discrete to determine the element numbers in each groupie the frequency. With this, it becomes possible to evaluate, even in an approximate way the proposed methods efficiency behavior. In the case presented earlier, it can be seen that the distribution was uniform discrete was used to determine the bits number, since the probability of occurrence of each group was the same.

To illustrate the calculation, we use three probability distributions: Discrete Uniform, Binomial and Poisson. It is considered that the bit numbers of the numbers to be allocated to matrix follow these three distributions. The respective parameters, thus, relate to the bit number in the uniform discrete case $U(a=1,b=64)$, which states that the number probability is represented by 1, 2, 3, $\ldots$, 64 is the same, ie, $\frac{1}{64}$. In the binomial case, it is considered that $B(n,0.5)$, such that the success numbers n represents the maximum bit number to represent the matrix numbers being equal to 1, 8, 16, 32 and 64, with a probability equal to 0.5 assuming a bit value 0 or 1. In the Poisson distribution, the parameter Poisson ($\lambda=p$) being the bit numbers expected to represent the element matrix. For all the simulations, it is considered that $k=7$, that is, it is considered that there is a possibility of occur a number that requires 64 bits.

Thus, the efficiency is calculated to be as small as possible, because if the value of k decreases, efficiency increases, as can be seen in Equation \ref{eq:37}.

To analyze about this new viewpoint, consider a sample of $rc$ elements, corresponding to the matrix element number $M_{r \times c}$. Therefore, a sample generating $rc$ elements can completely fill the matrix and verify the efficiency of the compression method proposed.

Consider that the numbers composing the matrix suggested requiring a bit numbers which follow a discrete uniform distribution $U(a=1,b=64)$. To verify the efficiency, generates a sample with 100 ($M_{10 \times 10}$), 10,000 ($M_{100 \times 100}$) and 1,000,000 ($M_{1,000 \times 1.000}$) numbers . Since this is a simulation, are generated 1,000 samples, for each calculated efficiency and the computed final average. The results can be seen in Table \ref{tab:03}. If the numbers follow the distribution, the efficiency of the proposed method is approximately 40\%.

\begin{table}[h]
  \centering
  \caption{Results of the simulation process efficiency on matrices whose elements are the bits number determined by a Discrete Uniform distribution $U(a=1,b=64)$.}
 \begin{tabular}{cc}
    \hline
    Sample size& Efficiency \\
    \hline
     100	& 0.3760938 \\
     1,000 	& 0.3795953 \\
     10,000 	& 0.3826869 \\
    \hline
 \end{tabular}
 \label{tab:03}
\end{table}

We analyzed the performance another simulation, only now the bit number follows a binomial distribution $B(n,0.5) $, such that the successes number n represents the maximum bit number to represent the matrix numbers being equal 1, 8, 16, 32 and 64, with a probability equal to 0.5 assuming a bit value 0 or 1. In this case, we analyze the efficiency behavior as a function of the parameter $n$, the bit number required, in efficiency of data compression. Results can be seen in Table \ref{tab:04}. Note that when smaller the bit number to represent array elements, the efficiency level is high using the second method. However, for the case in which the elements require 64 bits, the efficiency is negative. This result indicates that using the second method, when elements require 64 bit this becomes impractical, because it would be require more memory to allocate the compressed matrix.

\begin{table}[h]
  \centering
  \caption{Results of the simulation process efficiency on matrices whose elements are the bit number determined by a binomial distribution $B(n,0.5)$, such that the successes number n represents the maximum bit number to represent the matrix numbers being equal to 1, 8, 16, 32 and 64, with a probability equal to 0.5 assuming a bit value 0 or 1.}
 \begin{tabular}{cccc}
    \hline
			& 		&Efficiency         & \\
    \hline
			&  		&Sample size & \\
    Strip size		& 100		& 1,000		    & 1,000,000 \\
    \hline
     1 			&  0.9760937	& 0.9764797 	    & 0.9765566 \\
     8			&  0.8109375	& 0.8128453 	    & 0.8125079 \\
     16			&  0.626875	& 0.6251859 	    & 0.6249333 \\
     32			&  0.2476562	& 0.2504953 	    & 0.2499712 \\
     64			&  -0.5004688	& -0.4999828 	    & -0.5000366 \\
    \hline
 \end{tabular}
 \label{tab:04}
\end{table}

It is now considered that the bit number distribution is described by a Poisson ($\lambda$). Now the parameter $\lambda$ can be considered as the expected bit number  to represent numbers in matrix. Values ​​greater than 64 generated in this simulation were considered as being equal to 64. The results for this simulation can be seen in Table \ref{tab:05}. Note that increasing the bit number to represent numbers increases, there is a loss of efficiency in the compression process. Another interesting fact is the emergence of a negative efficiency. In this case, for values ​​generated with a $\lambda=64$, the second method requires more memory to allocate the matrix formed by elements generated from this distribution.

\begin{table}[h]
  \centering
  \caption{Results of the simulation process efficiency on matrices whose elements are the bit number determined by a Poisson distribution ($\lambda$), such that $\lambda$ represents the bit number to represent the expected matrix numbers.}
 \begin{tabular}{cccc}
    \hline
			& 	&Efficiency         & \\
    \hline
			&  	&Size sample& \\
    $\lambda$		& 100	& 1,000		    & 1,000,000 \\
    \hline
     1	& 0.8746875 & 0.8749859   & 0.8749834 \\
     8 	& 0.7657812 & 0.7656016   & 0.7656138 \\
     16	& 0.6300000 & 0.6400594   & 0.6405952 \\
     32	& 0.3851562 & 0.3903531   & 0.3903477 \\
     64	& -0.0528125& -0.05984375 & -0.05953145 \\
    \hline
 \end{tabular}
 \label{tab:05}
\end{table}

As the bits amount that form the numbers composing the matrix M are considered random variables, we can perform an analysis of the expected values ​​and the variability of the data to verify the compression efficiency level.

Consider that the bit number (B) to represent the M matrix elements are defined as a random variable, $B \sim U(a=1,b=64)$, then applying the expected value in the equation \ref{eq:43}, we have

\begin{equation}\label{eq:43}
 E(\zeta) = 1 - \frac{\sum_{i=1}^{g}  b_i \times p_i }{64} - \frac{k}{64} 
\end{equation}

\noindent as $E(b_i) = \sum_{i} b_i \times p(b_i) = \frac{a+b}{2}$, it result in

\begin{equation}\label{eq:44}
 E(\zeta) = 1 - \frac{E(B)}{64} - \frac{k}{64} 
\end{equation}

Recovered the example used, consider that there are two groups, $b_1=1$ and $b_{64}=64$, implying that $k=7$. Since $E(X)=\frac{a + b}{2}$, when $X \sim U(a,b)$, the efficiency is defined in Equation \ref{eq:45}.

\begin{equation}\label{eq:45}
 E(\zeta) = 1 - \frac{\frac{1+64}{2}}{64} - \frac{7}{64} \approx 38.29\% 
\end{equation}

This result corroborates with the findings anteriorly found in Table \ref{tab:03}. The same principle can be applied when the bit number (B) is a random variable with Binomial and Poisson distributions already displayed. In the case of Binomial when $B \sim B(n=64,p)$, has

\begin{equation}\label{eq:46}
 E(\zeta) = 1 - \frac{\sum_{i=1}^{g}  b_i \times p_i }{64} - \frac{k}{64} 
\end{equation}

\noindent in this case $E(b_i) = \sum_{i} b_i \times p(b_i) = n \times p$

\begin{equation}\label{eq:47}
 E(\zeta) = 1 - \frac{64 \times p }{64} - \frac{k}{64} 
\end{equation}

Again, it is considered that $n=64$ and $p=0.5$ as the success probability of occurring a number that requires 64 bits to be represented, then $k=7$. As a result we obtain

\begin{equation}\label{eq:48}
 E(\zeta) = 1 - \frac{64 \times 0.5 }{64} - \frac{7}{64} \approx 39.05\%
\end{equation}

The value found can be seen in Table \ref{tab:04}, obtained through simulations. As for the Poisson case, the efficiency value is defined as

\begin{equation}\label{eq:49}
 E(\zeta) = 1 - \frac{\sum_{i=1}^{g}  b_i \times p_i }{64} - \frac{k}{64} 
\end{equation}

\noindent in this case $E(b_i) = \lambda$,

\begin{equation}\label{eq:50}
 E(\zeta) = 1 - \frac{\lambda}{64} - \frac{k}{64} 
\end{equation}

We consider $\lambda = 1$,

\begin{equation}\label{eq:51}
 E(\zeta) = 1 - \frac{\lambda}{64} - \frac{k}{64}  = 1 - \frac{1}{64} - \frac{7}{64} \approx 87.5\%
\end{equation}

The efficiency level calculated $\zeta \approx 87.5\%$ can be seen in Table \ref{tab:05}. The results presented show simulations and verify that the compression process occurs in different cases.

In all tests so far conducted, we analyzed the impact of the average data compression efficiency. The final analysis to be performed, consider the mean and variance of the numbers to be compressed. Thus, it is expected function and efficiency in calculating the mean and variance of matrix numbers. (Remover?)

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}

\section{Matrix compression}

The matrix compression process consists in reducing the bytes number required to allocate a matrix in memory. This process allows a greater amount of information to be stored using the same hardware available memory allocation traditional method. One of the advantages in applying this method is the possibility to work with large information amounts about data directly from memory without need for disk access, reading and writing, which burdens the processing.

The steps and implementations required to use the prepossessing method are described below. Consider a matrix $M_{r \times c}$ in which $r$ is the rows number and $c$ the columns number. Each element of this matrix, called $m_{ij}$, only stores an integer number. For an integer number being stored in the memory of a 64-bit processor, using the largest number of available bits, demand is 64 bits. Therefore, the bit number ($\Delta$) required to allocate all the elements of a matrix M in the traditional manner may be defined by:

\begin{equation}\label{eq:01}
  \Delta = l \times c \times 64
\end{equation}

The $\Delta$ represents a greater bit number than that actually used. Assume that in a matrix, for example, there are only elements with values ​​equal to 0 or 1. These values ​​are represented in binary by only one bit. Therefore, when one is allocated in variable that requires 64 bits to be stored in memory, there is a loss of 63 bits, which could be used to store other numbers or information. This example deals with the case more expressive with respect to waste memory. The bit number calculation $\delta$ really necessary to implement this example is represented in the Equation \ref{eq:02}.

\begin{equation}\label{eq:02}
  \delta = l \times c \times 1
\end{equation}

The bits economy $\xi$ can be represented by the difference between $\Delta$ and $\delta$ (Equations \ref{eq:03} to \ref{eq:06}).

\begin{equation}\label{eq:03}
 \xi = \Delta - \delta
\end{equation}

\begin{equation}\label{eq:04}
 \xi =  r \times c \times 64 - r \times c \times 1
\end{equation}

\begin{equation}\label{eq:05}
 \xi =  r \times c \times (64 -  1)
\end{equation}

\begin{equation} \label{eq:06}
 \xi =  r \times c \times 63
\end{equation}

Based on the Equations \ref{eq:01} to \ref{eq:06}, it is can to observe that the economy $\xi$ is high, because $r \times c$ elements require only $r \times c \times  1$ bits instead of $r \times c \times 64$ bits. The same happens with other values ​​that do not need 64 bits to be represented in memory, because it not always all bits are used. It can be said that the economy grows linearly with the number of array elements.

There are two ways to perform the manipulations of these bits to minimize wasted space. Each of these forms follows a specific methodology to be applied to certain cases.

\subsection*{The first method: Analyzing the greatest value to allocate in a matrix}

The first methodology consists in determining the value of the largest element $\theta$, which should be stored in the matrix and as a result determines the  largest bit number ($\pi$) really need to store that value. The bit number ($\pi$) can be directly approximated by the equation \ref{eq:07}. This approach consists in calculating the bit value needed to represent the value, as there is usually decimal, the value 1 which is added ensures, that the resulting bit number to be correct to store such value, the integer part of the operation.

\begin{equation} \label{eq:07}
 \pi(\theta) \approx \begin{cases}
	1, &  \text{ if } \theta = 0 \\ 
	\left( \frac{log(\theta)}{log(2)}  \right) + 1,  & \text{ if } \theta > 0  
	\end{cases}
\end{equation}

Each matrix element of M, denoted by $m_{ij}$, is represented by a strip composed of 64 bits. In the example mentioned, for values ​​0 and 1, the value of $\theta$ equals 1, ie, in a strip of 64 bits may be allocated 64 values. Each bit of this strip will represent one of the values ​​to be allocated. This fact is responsible for saving bits, for 64 values​​, which were traditionally stored in $r \times c \times 64$  bits are stored in only $r \times c$ bits.

Suppose that the greatest value to be stored in a matrix $\Phi$ is $\theta=1023$. Therefore, the bit number required to represent that really value is $\pi=10$, because the number 1023 in binary base is 1111111111, therefore, demand 10 bit and the other 54 bits zeros disregarding the right, to be represented in memory. Consider a matrix $\Phi$ represented in equation \ref{eq:08}.

\begin{equation} \label{eq:08}
  \Phi = \begin{bmatrix}
  900 & 1023 & 721 & 256 & 1 & 10 & 700 & 20 & \hdots\\ 
  \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \ddots 
  \end{bmatrix}
\end{equation}

The values ​​of the matrix $\Phi$ in binary are represented in Table \ref{tab:01}. It is possible
checking the bit number necessary to represent the values​​, then it can be concluded
that with 10 bits, it is can represent all numbers.

\begin{table}[h]
 \centering
 \caption{Some matrix elements $\Phi$ represented in binary base. It is can to verify that using 10 bits is possible to represent the all values.}
 \begin{tabular}{cccc} 
  \hline 
  Element & Value  & Binary & Bit number necessary\\
  \hline
  $\Phi_{1,1}$ & 900  & 1110000100 & 10\\
  $\Phi_{1,2}$ & 1023 & 1111111111 & 10\\
  $\Phi_{1,3}$ & 721  & 1011010001 & 10\\
  $\Phi_{1,4}$ & 256  & 100000000  & 9\\
  $\Phi_{1,5}$ & 1    & 1          & 1\\
  $\Phi_{1,6}$ & 10   & 1010       & 4\\
  $\Phi_{1,7}$ & 700  & 1010111100 & 10\\
  $\Phi_{1,8}$ & 20   & 10100      & 5\\
  \hline
 \end{tabular}
 \label{tab:01}
\end{table}

Consider a matrix $\phi$, such that the matrix will store the the matrix $\Phi$ values, only using the bit number actually required so that there is not wastage. According to this methodology, the first element of the matrix $\phi$, denoted by $\phi_{1,1}$, which has a strip of 64 bits associated, it will be partitioned into regions. Each of these regions will store the values ​​of the matrix $\Phi$. As it takes 10 bits to represent the values ​​of the matrix $\Phi$, the bit element strip $\phi_{1,1}$ will contain 6 regions, such that each region will bundle 10 bits. So overall, this strip will have 6 regions, totaling 60 bits. The 4 bits free remaining it would be used later. The elements bit strip of $\phi_{1,1}$ can be verified in the equation \ref{eq:09}. It is possible to observe the regions, the values ​​stored (binary and decimal), and the bits that are left initially.

\begin{equation}\label{eq:09}
 \phi_{1,1} = 0000\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\underbrace{0100000000}_{256}\underbrace{1011010001}_{721}\underbrace{1111111111}_{1023}\underbrace{1110000100}_{900} 
\end{equation}

The process of allocation values ​​is performed as follows: the element $\Phi_{1,1}=900$ is stored in the first region of the element strip $\phi_{1,1} $, which corresponds to bits 0 to 9. Already for the element $\Phi_{1,2}=1,023$ is allocated in the second region, from 10 to 19 bits and the same with the other elements. Thus, the elements $\Phi_{1,i}$, with $i=1,\ldots,7$ are stored on the strip element $\phi_{1,1}$. Already the seventh element, $\Phi_{1,7}=700$, should be divided into two parts, so that a part are stored in the strip element $\phi_{1,1}$ and the other in the element strip $\phi_{1,2}$. The process of division shown in equation \ref{eq:10}, it can be seen the place where the split occurs strip element $\Phi_{1,7}=$ 700 in the parts $a$ and $b$. In the equation \ref{eq:11}, it can verify the bits grouped, demonstrating that it is possible to recover the information.

\begin{equation}\label{eq:10}
  \underbrace{\underbrace{0000010100}_{20}\overbrace{101011}^{a}}_{\phi_{1,2}}|\underbrace{\overbrace{1100}^{b}\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\hdots}_{\phi_{1,1}}
\end{equation}

\begin{equation}\label{eq:11}
  \underbrace{0000010100}_{20}\underbrace{\overbrace{101011}^a|\overbrace{1100}^{b}}_{700}\underbrace{0000001010}_{10}\ldots
\end{equation}

In a general way, what happens is the storage of the matrix elements $\Phi$ in a matrix $\phi$, such that the bits required to store them in the array $\phi$ is much smaller than the matrix $\Phi$, due to reorganization of the data. Even with the compression of the data, the information can be retrieved, thus justifying the proposed procedure.

\subsection*{The second method: Analyzing the values ​​that will be allocated in a matrix}

A second proposed method tries to make a new form of data compression. Consider again the matrix described in Equation \ref{eq:08}. Suppose the largest value stored in it is number 1023. As already demonstrated, to store this value in binary are necessary 10 bits. Now instead of reserving a region on the strip to store each value, reserves two regions. Such regions, one is used to store the bit number needed to represent the value and the second region will store the respective value. The first region will have a fixed size, in the example cited, 4 bits. The 4 bits are required to store the bit number necessary to store the largest value, in this case 10 bits. The size of the second region is determined by the value stored in the first. And so the other values ​​are stored.

To illustrate this methodology, we will be using the matrix represented in Equation \ref{eq:08}, called $\Phi$. The largest element of $\Phi$ is 1,023, as soon as the bit number required to represent this value is $\pi=10$ bits. The element $\Phi_{1,1}=900$ is stored in bits of the element strip of $\phi_ {1,1}$. To store the element $\Phi_{1,1}$, it should create the two first regions of the strip element $\phi_{1,1} $. Initially element strip $\phi_{1,1}=0$ is shown in equation \ref{eq:12}.

\begin{equation} \label{eq:12}
 \phi_{1,1} = 0000000000000000000000000000000000000000000000000000000000000000
\end{equation}

The first region that will store the value 10, because to represent the number 900 are need 10 bits. Then, it will have 4 bits, because represent the value 10, it demand 4 bits. The 4 bits are manipulated of according to the bit number to represent the largest value of $\Phi$, in which case, the value is 1,023. The strip $\phi_{1,1}$, after storing the information of $\Phi_{1,1}$ is shown in Equation \ref{eq:13}

\begin{equation} \label{eq:13}
 \phi_{1,1} = 00000000000000000000000000000000000000000000000000\underbrace{\underbrace{1110000100}_{\text{value} = 900}\underbrace{1010}_{\text{10 bits}}}_{900}
\end{equation}

Adding the elements $\Phi_{1,2}=1023$ and $\Phi_{1,3}=721$, the strip of $\phi_{1,1}$ updated is represented in Equation \ref{eq:14}. You can check the regions that store the information from the elements.

\begin{equation} \label{eq:14}
 \phi_{1,1} = 
 0000000000000000000000\underbrace{1011010001}_{721}\underbrace{1010}_{10}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1110000100}_{900}\underbrace{1010}_{10}
\end{equation}

Until this moment, the second method is using more bits to store a value that the first methodology. With the insertion of the fourth element that the bit economy occurs, because $\Phi_{1,4}=256$ demand only 9 bits. The first region with 4 bits defines how many bits the second region will contain the stored value. So instead of storing $\Phi_{1,4}=10$ in 10 bits, as in the previous method, it demands 9 bits, but it takes a whole 13 bits. The economy becomes evident when the element $\Phi_{1,5}=1$ is stored. Note that this element requires only 1 bit to be stored, so instead of 10 bits, only 1 bit is used, more 4 bits to indicate that the value is stored in only 1 bit. Therefore, instead of the 10 bits required to store that value pair following the first methodology, now only 5 bits are needed. The strip of $\phi_{1,1}$, after insertion of the latter two values ​​are updated in Equation \ref{eq:15}.

\begin{equation} \label{eq:15}
 \phi_{1,1} =
 0000\underbrace{1}_{1}\underbrace{0001}_{1}\underbrace{100000000}_{256}\underbrace{1001}_{9}\underbrace{1011010001}_{721}\underbrace{1010}_{10}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1110000100}_{900}\underbrace{1010}_{10}
\end{equation}

Now the element $\Phi_{1,6}=10$ will be inserted in the strip $\phi_{1,1}$. The process need that 4 bits are used element strip $\phi_{1,1}$ and another 4 bits of the element $\phi_{1,2}$. The strips updated element $\phi_{1,1}$ and $\phi_{1,2}$ can be verify in Equations \ref{eq:16} and \ref{eq:17}.

\begin{equation} \label{eq:16}
 \phi_{1,1} =
 \underbrace{0100}_{4}\underbrace{1}_{1}\underbrace{0001}_{1}\underbrace{100000000}_{256}\underbrace{1001}_{9}\underbrace{1011010001}_{721}\underbrace{1010}_{10}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1110000100}_{900}\underbrace{1010}_{10}
\end{equation}

\begin{equation} \label{eq:17}
 \phi_{1,2} = 000000000000000000000000000000000000000000000000000000000000\underbrace{1010}_{10}
\end{equation}

Adding the last two elements $\Phi_{1,7}=700$ and $\Phi_{1,8}=20$ in the element strip $\phi_{1,2}$, which after being updated, it can be viewed at equation \ref{eq:18}.

\begin{equation} \label{eq:18}
 \phi_{1,2} = 00000000000000000000000000000000000\underbrace{10100}_{20}\underbrace{0101}_{5}\underbrace{1010111100}_{700}\underbrace{1010}_{10}\underbrace{1010}_{10}
\end{equation}

For storing the elements of matrix $\Phi$ in the matrix $\phi$ were used 87 bits instead of 80 bits used in the first method. The choice of which method to use will be discussed in the section Discussion. The proposed methods have different levels of compression, depending directly from the values ​​that must be stored.

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}

%\section*{References}
% The bibtex filename
\bibliography{plos.bib}

\section*{Figure Legends}
%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

\section*{Tables}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

\end{document}

### 2013-09-25 manuscript.tex
Introdução e método SM analisados.

@@
 
+\usepackage{todonotes}
+
 \begin{document}
@@
 
-Data compression is traditionally used to reduce storage resources usage and/or transmission costs\cite{salomon}.
-Compression techniques can be classified into lossy and lossless.
-Examples of lossy data compression are MP3 (audio), JPEG (image) and MPEG (video).
-In this paper we discuss the use of lossless compression for numerical data structures such as numerical arrays to achieve compression without losing the mathematical properties of the original data.
+Data compression is traditionally used to reduce storage resources usage and/or 
+transmission costs\cite{salomon}. Compression techniques can be classified into 
+lossy and lossless. Examples of lossy data compression are MP3 (audio), JPEG 
+(image) and MPEG (video). In this paper, we discuss the use of lossless 
+compression for numerical data structures such as numerical arrays to achieve 
+compression without losing the mathematical properties of the original data.
 
-Lossless compression methods usually exploit redundancies present in the data in order to find a shorter form of describing the same information content.
-For example, a dictionary-based compression, only stores the positions in which a given word occurs in a document, thus saving the space required to store all its repetitions\cite{salomon2}.
+Lossless compression methods usually exploit redundancies present in the data in 
+order to find a shorter form of describing the same information content. For 
+example, a dictionary-based compression, only stores the positions in which a 
+given word occurs in a document, thus saving the space required to store all its 
+repetitions\cite{salomon2}.
 
-Any kind of compression incurs some computational cost.
-Such costs often have to be paid twice since the data needs to be decompressed to be used for its original purpose.
-Sometimes computational costs are irrelevant, but the need to decompress for usage, can signify that the space saved with compression must be available when data is decompressed for usage, thus partially negating the advantages of compression.
+Any kind of compression incurs some computational cost. Such costs often have to 
+be paid twice since the data needs to be decompressed to be used for its 
+original purpose. Sometimes computational costs are irrelevant, but the need to 
+decompress for usage, can signify that the space saved with compression must be 
+available when data is decompressed for usage, thus partially negating the 
+advantages of compression.
 
-Most if not all existing lossless compression methods were developed under the following usage paradigm: \textit{produce} $\rightarrow$ \textit{compress} $\rightarrow$ \textit{store} $\rightarrow$ \textit{uncompress} $\rightarrow$ \textit{use}.
-The focus of the present work is to allow a slightly different usage:  \textit{produce} $\rightarrow$ \textit{compress} $\rightarrow$ \textit{perform mathematical manipulations} and \textit{decompress} (only for human reading).
+Most if not all existing lossless compression methods were developed under the 
+following usage paradigm: \textit{produce} $\rightarrow$ \textit{compress} 
+$\rightarrow$ \textit{store} $\rightarrow$ \textit{uncompress} $\rightarrow$ 
+\textit{use}. The focus of the present work is to allow a slightly different 
+usage:  \textit{produce} $\rightarrow$ \textit{compress} $\rightarrow$ 
+\textit{perform mathematical manipulations} and \textit{decompress} (only for 
+human reading).
 
-With the growth of data volumes and analytical demands, creative solutions are needed to efficiently store as well as consume it on demand.
-This issue is present in many areas of application, ranging from business to science\cite{lynch}, and is being called the \textit{Big Data} phenomenon.
-In the world of Big Data, the need to analyze data immediately after its coming into existence became the norm.
-And this analysis must take place, efficiently, within the confines of (RAM) memory.
-This kind of analyses are what is now known as streaming data analysis\cite{gaber2005mining}.
-Given a sufficiently dense stream of data, compression an decompression costs may become prohibitive.
-So having a way to compress data and keeping it compressed for the entire course of the analytical pipeline, is very desirable.
+With the growth of data volumes and analytical demands, creative solutions are 
+needed to efficiently store as well as consume it on demand. This issue is 
+present in many areas of application, ranging from business to 
+science\cite{lynch}, and is being called the \textit{Big Data} phenomenon. In 
+the world of Big Data, the need to analyze data immediately after its coming 
+into existence became the norm. And this analysis must take place, efficiently, 
+within the confines of (RAM) memory. This kind of analyses are what is now known 
+as streaming data analysis\cite{gaber2005mining}. Given a sufficiently dense 
+stream of data, compression an decompression costs may become prohibitive. So 
+having a way to compress data and keeping it compressed for the entire course of 
+the analytical pipeline, is very desirable.
 
-This paper will focus solely on numerical data which for the purpose of the applications is organized as   matrices.
-This is a most common data structure found in computational data analysis environments.
-The matrices compressed according to the methodologies proposed here should be able to undergo the same mathematical operations as the original uncompressed matrices, e.g.
-linear algebra manipulations.
-This way, the cost of compression is reduced to a single event of compression and no need of decompression except when displaying the results for human reading.
-The idea of operating with compressed arrays is relatively new\cite{yemliha2007compiler}, and it has yet to find mainstream applications to the field of numerical computations.
-One application which employs a form of compression is the sparse matrix linear algebra algorithms\cite{dodson1991sparse}, in this case there is no alteration in the standard encoding of the data, but only the non-zero elements of the matrices are stored and operated upon.
+This paper will focus solely on numerical data which for the purpose of the 
+applications is organized as matrices. This is a most common data structure 
+found in computational data analysis environments. The matrices compressed 
+according to the methodologies proposed here should be able to undergo the same 
+mathematical operations as the original uncompressed matrices, e.g. linear 
+algebra manipulations. This way, the cost of compression is reduced to a single 
+event of compression and no need of decompression, except when displaying the 
+results for human reading. The idea of operating with compressed arrays is 
+relatively new\cite{yemliha2007compiler}, and it has yet to find mainstream 
+applications to the field of numerical computations. One application which 
+employs a form of compression is the sparse matrix linear algebra 
+algorithms\cite{dodson1991sparse}, in this case there is no alteration in the 
+standard encoding of the data, but only the non-zero elements of the matrices 
+are stored and operated upon.
 
-Larger than RAM data structures can render traditional analytical algorithms impracticable.
-Currently, the technique most commonly used when dealing with large matrices for numerical computations, is memory mapping\cite{van2011numpy,big}.
-In memory mapping the matrix is allocated in a virtual contiguous address space which extends from memory into disk.
-Thus, larger than memory data structures can be manipulated as if they were in memory.
-This technique has a big performance penalty due to lower access speeds of disk when compared to RAM.
+Larger than RAM data structures can render traditional analytical algorithms 
+impracticable. Currently, the technique most commonly used when dealing with 
+large matrices for numerical computations, is memory 
+mapping\cite{van2011numpy,big}. In memory, mapping the matrix is allocated in a 
+virtual contiguous address space which extends from memory into disk. Thus, 
+larger than memory data structures can be manipulated as if they were in memory. 
+This technique has a big performance penalty due to lower access speeds of disk 
+when compared to RAM.
 
-In this paper we present two methods for the lossless compression of (numerical) arrays.
-The methods involve the encoding of the numbers as strings of bits of variable length.
-The methods resemble the arithmetic coding\cite{bodden2007arithmetic} algorithm, but is cheaper to compute.
-We describe the process of compression and decompression, and study their efficiency under different applications.
-We also discuss the efficiency of the compression as a function of the distribution of the elements of the matrix.
+In this paper, we present two methods for the lossless compression of 
+(numerical) arrays. The methods involve the encoding of the numbers as strings 
+of bits of variable length. The methods resemble the arithmetic 
+coding\cite{bodden2007arithmetic} algorithm, but is cheaper to compute. We 
+describe the process of compression and decompression, and study their 
+efficiency under different applications. We also discuss the efficiency of the 
+compression as a function of the distribution of the elements of the matrix. 
+\todo{Inseri isso} Also presented are some results in relation to the 
+traditional method of matrix allocation. We also consider the 
+arithmetic operations to measure the efficiency of the proposed methodology.
 
@@
 
-To maintain mathematical equivalence with the original data for any arithmetic operations, we need to  maintain the structure of the matrix, i.e., the ability to acess any element given its row $i$ and column $j$ and also the numeric nature of its elements.
-In order to achieve compression we decided to exploit inefficiencies in the conventional way matrices are allocated in memory.
-The examples in this paper will be restricted to matrices with positive integer elements.
+To maintain mathematical equivalence with the original data for any arithmetic 
+operations, we need to maintain the structure of the matrix, ie, the ability 
+to access any element given its row $i$ and column $j$ and also the numeric 
+nature of its elements. In order to achieve compression, we decided to exploit 
+inefficiencies in the conventional way matrices are allocated in memory. 
+\todo{adicionei aqui} The analyzes presented in this article, in the first 
+moment, are related to arrays of positive integers, but can be applied to 
+integers and real numbers with some adaptations, which are presented in due 
+course. \todo{removi isso The examples in this paper will be restricted to 
+matrices with positive integer elements, but}
 
-The compression method is as follows.
-Let $M_{r \times c}$ be a matrix, in which $r$ is the number of rows and $c$ the number of columns.
-Each element of this matrix, called $m_{ij}$, is a positive integer.
-In digital computers, all information is stored as binary code (base 2 numbers).
-However the conventional way to store arrays of integers is on a memory block sequence of fixed size (power of 2 numbers of bit), one for each element.
-The maximum size of a block is equal to the word size of the processor, which for most current CPUs is 64 bits.
-Some special number such as complex number may be encoded as two blocks instead of one.
-The size of the chunk of memory allocated to each number will determine their maximum  size (for integers) or their precision (for floating-point numbes).
-So for matrix $M$, the total memory allocated, assuming chunks of 64 bits, is given by $\mathcal{B} = r \times c \times 64$.
+The compression method is as follows. Let $M_{r \times c}$ be a matrix, in which 
+$r$ is the number of rows and $c$ the number of columns. Each element of this 
+matrix, called $m_{ij}$, is a positive integer. In digital computers, all 
+information is stored as binary code (base 2 numbers). However, the conventional 
+way to store arrays of integers is on a memory block sequence of fixed size 
+(power of 2 numbers of bit), one for each element. The maximum size of a block 
+is equal to the word size of the processor, which for most current CPUs is 64 
+bits. Some special number such as complex number may be encoded as two blocks 
+instead of one. The size of the chunk of memory allocated to each number will 
+determine their maximum  size (for integers) or their precision (for 
+floating-point numbers). So for matrix $M$, the total memory allocated, 
+assuming chunks of 64 bits, is given by $\mathcal{B} = r \times c \times 64$.
 
-The number of bits allocated $\mathcal{B}$, is larger than the absolute minimum number of bits required to represent all the elements of $M$, since smaller integers, when converted to base 2, require less digits.
-From now on, when the numerical base will be explicitly notated when necessary to avoid confusion between binary and decimal integers.
+The number of bits allocated $\mathcal{B}$, is larger than the absolute minimum 
+number of bits required to represent all the elements of $M$, since smaller 
+integers, when converted to base 2, require less digits. From now on, when the 
+numerical base will be explicitly notated when necessary to avoid confusion 
+between binary and decimal integers.
 
-Let's consider an extreme example: a matrix composed exclusively of 0s and 1s (base 10).
-If the matrix type is set to 64-bit integers, 63 bits will be wasted per element of the matrix, since the minimum number of bits needed to store such a matrix is $\mathfrak{b} = r \times c \times 1$.
-The potential economy of bits $\xi$ can be represented by $\xi = \mathcal{B} - \mathfrak{b} =  r \times c \times 63$.
+Let's consider an extreme example: a matrix composed exclusively of 0s and 1s 
+(base 10). If the matrix type is set to 64-bit integers, 63 bits will be wasted 
+per element of the matrix, since the minimum number of bits needed to store such 
+a matrix is $\mathfrak{b} = r \times c \times 1$. The potential economy of bits 
+$\xi$ can be represented by $\xi = \mathcal{B} - \mathfrak{b} =  r \times c 
+\times 63$.
 
-So it is evident that for any matrix whose greatest element requires less than 64 bits (or the fixed type of the matrix) to be represented, potential memory savings will grow linearly with the size of the matrix.
+So it is evident that for any matrix whose greatest element requires less than 
+64 bits (or the fixed type of the matrix) to be represented, potential memory 
+savings will grow linearly with the size of the matrix.
 
@@
 
-The SM method consists in determining the value of the greatest element of matrix $M$, which coincides with its supremum, $max M == \sup M$  and determine the minimum number of bits, $b(\sup M$)(Equation \ref{eq:01}), required to store it.
-We will use capital roman letters to denote uncompressed matrices and the corresponding lower case letter for the compressed version.
+The SM method consists in determining the value of the greatest element of 
+matrix $M$, which coincides with its supremum, $max M == \sup M$  and determine 
+the minimum number of bits, $b(\sup M$)(Equation \ref{eq:01}), required to store 
+it. We will use capital roman letters to denote uncompressed matrices and the 
+corresponding lower case letter for the compressed version.
 
@@
 	\end{cases}
-\end{equation}
+\end{equation}\todo{Será que não devemos trocar $\approx$ por = aqui? Note que 
+o número de bits não é aproximado e sim o exato.}
 
-The allocation of memory still happens in the usual way, i.e., in fixed size 64-bit chunks, only that now, in the space required for a single 64 bit integer, we can store for example, an entire $8\times 8$ matrix of $0_{10}$ and $1_{10}$.
+The allocation of memory still happens in the usual way, ie, in fixed size 
+64-bit chunks, only that now, in the space required for a single 64 bit integer, 
+we can store for example, an entire $8\times 8$ matrix of $0_{10}$ and 
+$1_{10}$.
 
-Let's look at a concrete example: suppose that the greatest value to be stored in a matrix $M$ is $\max M=1023$.
-Therefore, the number of bits required to represent it is $10$ $(1111111111)$.
-Let the first 8 elements of $M$ be:
+Let's look at a concrete example: suppose that the greatest value to be stored 
+in a matrix $M$ is $\max M=1023$. Therefore, the number of bits required to 
+represent it is $10$ $(1111111111)$. Let the first 8 elements of $M$ be:
 
@@
 
-These elements of $M$, in binary, are shown in Table \ref{tab:01}.
-It is evident that the number of bits required to represent any other element must be lower or equal to $10$.
-From now on the minimum number of bits required to represent a base 10 integer will be refered to as its bit-length.
+These elements of $M$, in binary, are shown in Table \ref{tab:01}. It is 
+evident that the number of bits required to represent any other element must be 
+lower or equal to $10$. From now on the minimum number of bits required to 
+represent a base 10 integer will be refered to as its bit-length.
 
@@
  \centering
- \caption{Some elements of $M$ represented in binary base.}
+ \caption{The first 8 elements of $M$ represented in binary base.}
  \begin{tabular}{cccc} 
@@
 
-To store matrix $M$ it first has to be converted to base $2$ ($M_2$).
-Then it will be unraveled by column (column major, e.g.
-in Fortran) or by row (row major, e.g.
-in C) and its elements will be written as  fixed size adjacent chunks of memory.
-The size of each chunk is determined by the type associated with the matrix (typically 64 bits, but always a power of 2).
+To store matrix $M$ it first has to be converted to base $2$ ($M_2$). Then it 
+will be unraveled by column (column major, e.g. in Fortran) or by row (row 
+major, e.g. in C) and its elements will be written as fixed size adjacent chunks 
+of memory. The size of each chunk is determined by the type associated with the 
+matrix (typically 64 bits, but always a power of 2).
 
-According to the SM method, having determined that each element will require at most 10 bits, we can divide the memory block corresponding to a single 64 bit integer into six 10-bit chunks which can each hold a single element of $M$.
-These 64-bit blocks will be called a bitstring.
-The remaining 4 bits will be used later.
-The number of bitstrings needed will be $\lfloor \frac{\dim(M) *b(\sup M)} {64} \rfloor +1$, where $dim(M)$ is the dimension of the matrix or its number of elements.
+According to the SM method, having determined that each element will require at 
+most 10 bits, we can divide the memory block corresponding to a single 64 bit 
+integer into six 10-bit chunks, which can each hold a single element of $M$. 
+These 64-bit blocks will be called a bitstring. The remaining 4 bits will be 
+used later. The number of bitstrings needed will be $\lfloor \frac{\dim(M) 
+*b(\sup M)} {64} \rfloor +1$, where $dim(M)$ is the dimension of the matrix or 
+its number of elements.
 
@@
  \end{equation*}
- \item Element $M_{1,7}=700=1010111100$ does not fit on the remaining 4 bits of the first bitstring.
-So it will straddle two bitstrings, i.e., it is divided in two segments $a$ and $b$, $a$ is written on the first bitstring and $b$ on the second.
+ \item Element $M_{1,7}=700=1010111100$ does not fit on the remaining 4 bits of 
+the first bitstring. So it will straddle two bitstrings, ie, it is divided in 
+two segments $a$ and $b$, $a$ is written on the first bitstring and $b$ on the 
+second.
  \begin{equation*}\label{eq:07}
@@
 
- Please notice that bitstrings are written from right to left.
+ Please, notice that bitstrings are written from right to left.
 
@@
 
-Thus the compressed matrix $m = M_2$ requires less memory than the conventional storage of $M$ as a 64-bit integer array.
+Thus the compressed matrix $m = M_2$ requires less memory than the conventional 
+storage of $M$ as a 64-bit integer array. An illustration of the compression 
+process and the allocation of the elements of matrix M in the matrix m is shown 
+in Figure \ref{fig:01}. Note that the matrix m has fewer columns that the 
+matrix M, then it demands less memory to be represented. We also emphasize that 
+the bits which represent the value 700 is divided in two parts and stored in 
+different strips in the matrix m.
 
@@
 
-In the SM method, there is still waste of space since for elements smaller than the supremum, a number of bits remain unused.
+In the SM method, there is still waste of space since for elements smaller than 
+the supremum, a number of bits remain unused.
 
-In the VBL method, the absolute minimal number of bits are used to store each value.
-However, if we are going to divide the biststrings into variable length chunks, we also need to reserve some extra bits to represent the size of each chunk, otherwise the elements cannot be recovered once they are stored.
+In the VBL method, the absolute minimal number of bits are used to store each 
+value. However, if we are going to divide the biststrings into variable length 
+chunks, we also need to reserve some extra bits to represent the size of each 
+chunk, otherwise the elements cannot be recovered once they are stored.
 
-Lets use again the matrix described in Equation \ref{eq:02}, where  the largest element is number 1023.
-Now instead of assigning one chunk of the bitstring to each element of $m$, we will assign two chunks: the first will store the number of bits required to store the element and the second  will store the actual element.
-The first chunk will have a fixed size, in this case, 4 bits.
-These 4 bits are the required space to store the bit-length of $\sup M $, in this case, 10.
+Lets use again the matrix described in Equation \ref{eq:02}, where  the largest 
+element is number 1023. Now instead of assigning one chunk of the bitstring to 
+each element of $m$, we will assign two chunks: the first will store the number 
+of bits required to store the element and the second  will store the actual 
+element. The first chunk will have a fixed size, in this case, 4 bits. These 4 
+bits are the required space to store the bit-length of $\sup M $, in this case, 
+10.
 
-Lets go through VLB compression step-by-step.
-The largest element of $M$ is 1023.
-Its bit-length is 10 which in turn is 4 bits long in base 2 (1010).
-Thus the fixed size chunk is 4 bits long for every element.
+Lets go through VLB compression step-by-step. The largest element of $M$ is 
+1023. Its bit-length is 10 which in turn is 4 bits long in base 2 (1010). Thus 
+the fixed size chunk is 4 bits long for every element.
 
@@
 
-%\section*{Figure Legends}
+\section*{Figure Legends}
+
+\begin{figure}[htp]
+  \includegraphics[scale=0.3,clip]{fig01.pdf}
+  \caption{Representation of the compression process and allocation of 
+elements of matrix M in the matrix m. The strip of value 700 is divided in two 
+parts and stores in different strips in the matrix m.}
+  \label{fig:01}
+\end{figure}
+
 %\begin{figure}[!ht]

### 2013-09-25 plos.tex
Introdução e método SM analisados.

@@
 \usepackage{color}
+
+\usepackage[colorinlistoftodos]{todonotes}
 
@@
 
-Usually the techniques that involve data processing try to work around problems, arise in this case the the data file size, to perform processing.
-At present and due to the increased speed with which information is generated, it becomes necessary to develop methods which allow such management information in large scale, mostly so that no loss occurs.
-The information storage on a large scale, especially in digital form, can assist in advancing science \cite{lynch}.
-
-There are different methods that perform this compression procedure, which can be classified based on their characteristics.
-We highlight the main information loss as characteristics during the compression process, the data manipulation procedure, methodology used, among others \cite{salomon2}.
-
-If after compression, data suffer a information loss, this causes loss if the file is uncompressed.
-The compressed data can not be recovered in its original form.
-Therefore you can choose the method based on the ratio of the loss characteristics.
-The data that can be recovered in spite of presenting information loss do not invalidate its use.
-However, there are other processes that the data must be recovered in full.
-
-According to the loss or not, the information manipulated, one can classify the data types on which compression can occur.
-When there is information loss, so that this does not undermine the use of the data, we have a method called nondestructive.
-An example of the data manipulated in this method are sound files (mp3), picture (JPEG) and movies (mpeg).
-As for the other data after being compressed, it can not have losses, they are called non-destructive.
-The types of manipulated file by this method are based on text and numbers \cite{salomon,salomon2}.
-
-There is another method classification compression with respect to how it is done, its can be symmetrical or asymmetrical.
-In the case of symmetrical, time compression and decompression of data occurs synchronously, ie, it take the same time for being compressed and decompressed.
-In this classification, the fit into based dictionary-algorithm, such as LZW, which use dictionaries for performing the compression.
-Already another classification, the asymmetric demand different times in implementing the compression and decompression processes, as an example it can highlight the compressing audio method, mpeg \cite{salomon,salomon2}.
-
-Another feature that is considered to distinguish compression types is related to the way in which the process runs.
-In this new approach, there are methods that are considered adaptive and non-adaptive.
-As the compression process is executed, the method can be adapted to optimize the processing.
-Considering this new feature, we can highlight the Golomb code among and other as non-adaptive.
-Already on the other hand, we have dictionary-based algorithms, LZ77 and LZ78, and adaptive as those based on Huffman coding, arithmetic coding, and others, can be both adaptive and non-adaptive \cite{salomon, salomon2}.
-
-Besides the problem involved with compression, there is another related to the way in which data large volumes must be represented in memory.
-The procedures described up to now, in large part only function to compress the data file and in some cases access them.
-When these data should be accessed for manipulation, there is a need to be represented, in the computer case, in memory.
-From there arise a problem with the representation of this large volume of data in memory.
-Among the many techniques developed to solve this problem, there is the technique of memory-mapped files that seeks to use main memory to reference the data file in disk as the data are requested, they are loaded into main memory.
-This technique allows data large volumes can be handled, but still there is a limit, in which case the memory size.
-Another problem is related to access time data to disk, which is relatively higher compared to the memory access.
-Some programming languages ​​like Python, R and other, allow the use of this feature.
-In Python this feature is available in the standard library, already in the R this feature can be used by the package bigmemmory \cite{big}.
-
-In this paper we present two methods that allow the compression of large data sets, such as the type of big data, for direct manipulation of files into memory.
-These methods allow direct access to data in compressed form, its handling and recovery occurs without loss.
-The proposed method also allows to use the techniques already presented to perform a subsequent compression process only if these data should be stored on physical media.
-We describe the process of compression and decompression, using as an example strips of bits to represent numerical arrays, the method sparse string \cite{salomon,salomon2}, to illustrate the methods.
-We also present a discussion of some results on the characteristics of each proposed method and elucidated the best options in which each can be applied.
-We have also emphasized that the methods allow information to be accessed after the compression process, which allows the handling and updating, if necessary.
-At the end of the paper we present their conclusions about using of
-the proposed methods.
+Usually the techniques that involve data processing try to work around problems, 
+arise in this case the the data file size, to perform processing. At present and 
+due to the increased speed with which information is generated, it becomes 
+necessary to develop methods which allow such management information in large 
+scale, mostly so that no loss occurs. The information storage on a large scale, 
+especially in digital form, can assist in advancing science \cite{lynch}.
+
+There are different methods that perform this compression procedure, which can 
+be classified based on their characteristics. We highlight the main information 
+loss as characteristics during the compression process, the data manipulation 
+procedure, methodology used, among others \cite{salomon2}.
+
+If after compression, data suffer a information loss, this causes loss if the 
+file is uncompressed. The compressed data can not be recovered in its original 
+form. Therefore you can choose the method based on the ratio of the loss 
+characteristics. The data that can be recovered in spite of presenting 
+information loss do not invalidate its use. However, there are other processes 
+that the data must be recovered in full.
+
+According to the loss or not, the information manipulated, one can classify the 
+data types on which compression can occur. When there is information loss, so 
+that this does not undermine the use of the data, we have a method called 
+nondestructive. An example of the data manipulated in this method are sound 
+files (mp3), picture (JPEG) and movies (mpeg). As for the other data after being 
+compressed, it can not have losses, they are called non-destructive. The types 
+of manipulated file by this method are based on text and numbers 
+\cite{salomon,salomon2}.
+
+There is another method classification compression with respect to how it is 
+done, its can be symmetrical or asymmetrical. In the case of symmetrical, time 
+compression and decompression of data occurs synchronously, ie, it take the same 
+time for being compressed and decompressed. In this classification, the fit into 
+based dictionary-algorithm, such as LZW, which use dictionaries for performing 
+the compression. Already another classification, the asymmetric demand different 
+times in implementing the compression and decompression processes, as an example 
+it can highlight the compressing audio method, mpeg \cite{salomon,salomon2}.
+
+Another feature that is considered to distinguish compression types is related 
+to the way in which the process runs. In this new approach, there are methods 
+that are considered adaptive and non-adaptive. As the compression process is 
+executed, the method can be adapted to optimize the processing. Considering this 
+new feature, we can highlight the Golomb code among and other as non-adaptive. 
+Already on the other hand, we have dictionary-based algorithms, LZ77 and LZ78, 
+and adaptive as those based on Huffman coding, arithmetic coding, and others, 
+can be both adaptive and non-adaptive \cite{salomon, salomon2}.
+
+Besides the problem involved with compression, there is another related to the 
+way in which data large volumes must be represented in memory. The procedures 
+described up to now, in large part only function to compress the data file and 
+in some cases access them. When these data should be accessed for manipulation, 
+there is a need to be represented, in the computer case, in memory. From there 
+arise a problem with the representation of this large volume of data in memory. 
+Among the many techniques developed to solve this problem, there is the 
+technique of memory-mapped files that seeks to use main memory to reference the 
+data file in disk as the data are requested, they are loaded into main memory. 
+This technique allows data large volumes can be handled, but still there is a 
+limit, in which case the memory size. Another problem is related to access time 
+data to disk, which is relatively higher compared to the memory access. Some 
+programming languages ​​like Python, R and other, allow the use of this feature. 
+In Python this feature is available in the standard library, already in the R 
+this feature can be used by the package bigmemmory \cite{big}.
+
+In this paper, we present two methods that allow the compression of large data 
+sets, such as the type of big data, for direct manipulation of files into 
+memory. These methods allow direct access to data in compressed form, its 
+handling and recovery occurs without loss. The proposed method also allows to 
+use the techniques already presented to perform a subsequent compression process 
+only if these data should be stored on physical media. We describe the process 
+of compression and decompression, using as an example strips of bits to 
+represent numerical arrays, the method sparse string \cite{salomon,salomon2}, to 
+illustrate the methods. We also present a discussion of some results on the 
+characteristics of each proposed method and elucidated the best options in which 
+each can be applied. We have also emphasized that the methods allow information 
+to be accessed after the compression process, which allows the handling and 
+updating, if necessary. At the end of the paper we present their conclusions 
+about using of the proposed methods.
 
@@
 
-Consider a matrix such that the elements are integers.
-The results will be presented that indicate the compression process propose two methods.
-These methods create a new matrix, so that the values ​​of the original matrix are stored in a new matrix, only using less computational resources, in this case memory to be represented.
-
-To check the efficiency of these methods, we consider that this matrix is composed of numbers generated from random variables that follow a probability distributions mixture.
-
-Now for simplicity and functionality will be used by a probability distributions mixture Beta.
-The probability density function (pdf) of the mixture of two Beta distributions is shown in equation \ref{eq:52}.
-The shape parameters $\alpha$ and $\beta$ will be correlated to the bit number to represent the values ​​in the data matrix to be compressed.
-Already the parameter $\lambda$ defines the proportion of each distribution make mixture.
-From this value range​​, one can by distributing generate other and different sets of values ​​to fill the matrix, then one can sweep all possibilities and estimating the efficiency of the compression process.
+Consider a matrix such that the elements are integers. The results will be 
+presented that indicate the compression process propose two methods. These 
+methods create a new matrix, so that the values ​​of the original matrix are 
+stored in a new matrix, only using less computational resources, in this case 
+memory to be represented.
+
+To check the efficiency of these methods, at first, we consider that this 
+matrix is composed of numbers generated from random variables that follow a 
+probability distributions mixture.
+
+Now for simplicity and functionality will be used by a probability distributions 
+mixture Beta. The probability density function (pdf) of the mixture of two Beta 
+distributions is shown in equation \ref{eq:52}. The shape parameters $\alpha$ 
+and $\beta$ will be correlated to the bit number to represent the values ​​in 
+the data matrix to be compressed. Already the parameter $\lambda$ defines the 
+proportion of each distribution make mixture. From this value range​​, one can 
+by distributing generate other and different sets of values ​​to fill the 
+matrix, then one can sweep all possibilities and estimating the efficiency of 
+the compression process.
 

### 2013-09-25 manuscript.tex
Método VLB revisado.

@@
 
-So far the VLB method is more wasteful than the SM, but when we add $M_{1,4} =256$  we start to save some space.
+So far the VLB method is more wasteful than the SM, but when we add $M_{1,4} 
+=256$, we start to save some space.
 
@@
 
- We used a total of 87 bits to store matrix $m$ with the VLB method instead of 80 bits using the SM method.
-However, as shall be seen later, the VLB method will be the most efficient for most matrices.
+ We used a total of 87 bits to store matrix $m$ with the VLB method instead of 
+80 bits using the SM method. However, as shall be seen later, the VLB method 
+will be the most efficient for most matrices. \todo{adicionei} With the method 
+VLB the resulting matrix is smaller than the original matrix.
 

### 2013-09-27 manuscript.tex
Tudo analisado, menos a conclusão

@@
 
+To store matrix $M$ it first has to be converted to base $2$ ($M_2$). Then it 
+will be unraveled by column (column major, e.g. in Fortran) or by row (row 
+major, e.g. in C) and its elements will be written as fixed size adjacent chunks 
+of memory. The size of each chunk is determined by the type associated with the 
+matrix (typically 64 bits, but always a power of 2).
+
+According to the SM method, having determined that each element will require at 
+most 10 bits, we can divide the memory block corresponding to a single 64 bit 
+integer into six 10-bit chunks, which can each hold a single element of $M$. 
+These 64-bit blocks will be called a bitstring. The remaining 4 bits will be 
+used later. The number of bitstrings needed will be $\lfloor \frac{\dim(M) 
+*b(\sup M)} {64} \rfloor +1$, where $dim(M)$ is the dimension of the matrix or 
+its number of elements.
+
+The final layout of the first 6 elements of $m$ in the first bitstring can be 
+seen in \ref{eq:03}.
+
+\begin{equation}\label{eq:03}
+ bitstring_1 = 0000\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\underbrace{0100000000}_{256}\underbrace{1011010001}_{721}\underbrace{1111111111}_{1023}\underbrace{1110000100}_{900} 
+\end{equation}
+
+Here is a step-by-step description of the application of the SM method to matrix $M$:
+
+\begin{enumerate}
+ \item Element $M_{1,1}=900 = m_{1,1} = 1110000100$ is stored in the first 
+10-bit chunk of the element strip $bitstring[1]$ , which corresponds to bits 0 
+to 9 (read from right to left). 
+\begin{equation*}\label{eq:04}
+bitstring_1 = 000000000000000000000000000000000000000000000000000000\underbrace{1110000100}_{900} 
+\end{equation*}
+ \item Element $M_{1,2}=1023$ is allocated in the second chunk, from bit 10 to 
+bit 19. 
+\begin{equation*}\label{eq:05}
+ bitstring_1 = 00000000000000000000000000000000000000000000\underbrace{1111111111}_{1023}\underbrace{1110000100}_{900} 
+\end{equation*}
+ \item Repeat for elements $M_{1,i}$ with $i=1,\ldots,6$ which are stored on the 
+remaining chunks. 
+ \begin{equation*}\label{eq:06}
+bitstring_1 = 0000\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\underbrace{0100000000}_{256}\underbrace{1011010001}_{721}\underbrace{1111111111}_{1023}\underbrace{1110000100}_{900} 
+ \end{equation*}
+ \item Element $M_{1,7}=700=1010111100$ does not fit on the remaining 4 bits of 
+the first bitstring. So it will straddle two bitstrings, ie, it is divided in 
+two segments $a$ and $b$, $a$ is written on the first bitstring and $b$ on the 
+second.
+ \begin{equation*}\label{eq:07}
+  \underbrace{\underbrace{0000010100}_{20}\overbrace{101011}^{b}}_{bitstring_2}|\underbrace{\overbrace{1100}^{a}\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\hdots}_{bitstring_1}
+\end{equation*}
+\end{enumerate}
+
+ Please, notice that bitstrings are written from right to left.
+
+\begin{equation*}\label{eq:8}
+  \underbrace{0000010100}_{20}\underbrace{\overbrace{101011}^a|\overbrace{1100}^{b}}_{700}\underbrace{0000001010}_{10}\ldots
+\end{equation*}
+
+Thus the compressed matrix $m = M_2$ requires less memory than the conventional 
+storage of $M$ as a 64-bit integer array. An illustration of the compression 
+process and the allocation of the elements of matrix M in the matrix m is shown 
+in Figure \ref{fig:01}. Note that the matrix m has fewer columns that the 
+matrix M, then it demands less memory to be represented. We also emphasize that 
+the bits which represent the value 700 is divided in two parts and stored in 
+different strips in the matrix m.
+
+\subsubsection*{Method 2: Variable Length Blocks (VLB)}
+
+In the SM method, there is still waste of space since for elements smaller than 
+the supremum, a number of bits remain unused.
+
+In the VBL method, the absolute minimal number of bits are used to store each 
+value. However, if we are going to divide the biststrings into variable length 
+chunks, we also need to reserve some extra bits to represent the size of each 
+chunk, otherwise the elements cannot be recovered once they are stored.
+
+Lets use again the matrix described in Equation \ref{eq:02}, where  the largest 
+element is number 1023. Now instead of assigning one chunk of the bitstring to 
+each element of $m$, we will assign two chunks: the first will store the number 
+of bits required to store the element and the second  will store the actual 
+element. The first chunk will have a fixed size, in this case, 4 bits. These 4 
+bits are the required space to store the bit-length of $\sup M $, in this case, 
+10.
+
+Lets go through VLB compression step-by-step. The largest element of $M$ is 
+1023. Its bit-length is 10 which in turn is 4 bits long in base 2 (1010). Thus 
+the fixed size chunk is 4 bits long for every element.
+
+\begin{enumerate}
+ \item The first element $M_{1,1}=900$ requires 10 bits to store, so we write 
+$10$ in the first chunk and $900$ in the second. 
+ \begin{equation*} \label{eq:09}
+ bitstring_1 = 00000000000000000000000000000000000000000000000000\underbrace{\underbrace{1110000100}_{\text{element} = 900}\underbrace{1010}_{\text{bit-length=10}}}_{M_{1,1}}
+\end{equation*}
+ \item Do the same for the next element, $M_{1,2}=1023$.
+  \begin{equation*} \label{eq:10}
+ bitstring_1 = \ldots00000000000000000000000000\underbrace{\underbrace{1111111111}_{\text{element} = 1023}\underbrace{1010}_{\text{bit-length=10}}}_{M_{1,2}}\underbrace{\underbrace{1110000100}_{\text{element} = 900}\underbrace{1010}_{\text{bit-length=10}}}_{M_{1,1}}
+\end{equation*}
+ \item Element $M_{1,3}=721$ is also added taking the bitstring to the state.
+ \begin{equation*} \label{eq:11}
+ bitstring_1 = 
+ 0000000000000000000000\underbrace{1011010001}_{721}\underbrace{1010}_{10}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1110000100}_{900}\underbrace{1010}_{10}
+\end{equation*}
+\end{enumerate}
+
+So far the VLB method is more wasteful than the SM, but when we add $M_{1,4} 
+=256$, we start to save some space.
+
+\begin{enumerate}
+ \item[4.] Element $M_{1,4} =256$ is added.  
+ \item [5.] Elements $M_{1,5} =1$ and $M_{1,6} =10$ are added requiring a total 
+of 13 bits instead of 20 with the SM method. With the addition of these elements 
+we require a second bitstring. 
+ \begin{align*} \label{eq:12}
+ bitstring_1 &=
+ \underbrace{0100}_{4}\underbrace{1}_{1}\underbrace{0001}_{1}\underbrace{100000000}_{256}\underbrace{1001}_{9}\underbrace{1011010001}_{721}\underbrace{1010}_{10}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1110000100}_{900}\underbrace{1010}_{10} \\
+ bitstring_2 &= 000000000000000000000000000000000000000000000000000000000000\underbrace{1010}_{10}
+\end{align*}
+ \item [6.] The remaining two elements are added $M_{1,7} =700$ and $M_{1,8} 
+=20$ in the second bit strip. 
+ \begin{equation*} \label{eq:13}
+ bitstring_2 = 00000000000000000000000000000000000\underbrace{10100}_{20}\underbrace{0101}_{5}\underbrace{1010111100}_{700}\underbrace{1010}_{10}\underbrace{1010}_{10}
+\end{equation*}
+\end{enumerate}
+
+ We used a total of 87 bits to store matrix $m$ with the VLB method instead of 
+80 bits using the SM method. However, as shall be seen later, the VLB method 
+will be the most efficient for most matrices. \todo{adicionei} With the method 
+VLB the resulting matrix is smaller than the original matrix.
+
+\subsection*{Compression Efficiency}
+ Compression efficiency depends of the data being compressed. Below, a formula 
+for calculating compression efficiency is derived for both methods. They will be 
+based on the following ratio:
+
+\begin{equation}\label{eq:14}
+  \eta=\frac{bits\, alocated-bits\, used}{bits\,alocated}
+\end{equation}
+
+ Where \textit{bits alocated} above mean total bits required for standard 
+storage of the matrix, without compression, while bits used mean total bits 
+requires to store the matrix after compression. From now on the efficiencies are 
+denoted by $\eta_1$ for the SM method and by $\eta_2$ for the VLB method.
+
+\subsubsection*{SM Method}
+
+Let $M_{r \times c}$ be the matrix we wish to compress. In comparison with a 
+conventional allocation (64-bit integers), we can apply Equation \ref{eq:14} to 
+calculate the efficiency of the SM method:
+
+\begin{align}\label{eq:15}
+ \eta_1 &= \frac{64 \times rc - b(max M) \times rc}{64 \times rc}\nonumber\\
+ &= \frac{64  - b(max M) }{64}
+\end{align}
+
+As we see in \ref{eq:15}, $\eta_1$ does not depend on size of the matrix, only 
+on the bit-length of $\max M$. If $b(\max M)=64$, $\eta_1$ is 0, ie, no 
+compression is possible. On the other extreme, if the matrix is composed 
+exclusively of 0s and 1s, maximal compression is achievable, 
+$\eta_1=0.9843$.\todo{Alterei a proporção para 0.98, pois é o maior valor que 
+pode ser obtido.}
+
+\subsubsection*{VLB Method}
+
+For the VLB method, compression depends on the value of each element of the 
+matrix. In this method bit-lentgh variability affects the compression ratio, so 
+the formula will have to include this information.
+
+Let the $rc$ elements of the matrix $M_{r \times c}$ be divided into $g$ groups, 
+each with $f_i$ numbers of bit-length $b_i = b(m_i)$. Thus $f_i$ is the 
+frequency of each bit-length present in $M$. Let $k = b(b(\max M))$, ie, the 
+bit length of the bit-length of $\max M$. The efficiency $\eta_2$ is shown 
+below.
+
+\begin{equation}\label{eq:16}
+ \eta_2 = \frac{64 \times rc - \sum_{i=1}^{g} ( b_i + k ) \times f_i }{64 \times rc} 
+\end{equation}
+
+We can further simplify Equation \ref{eq:16} to get at shorter expression for 
+the compression ratio.
+
+\begin{align}
+ \eta_2 &= \frac{64 \times rc - \sum_{i=1}^{g} ( b_i \times f_i + k \times f_i )}{64 \times rc} \nonumber \\
+  &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  -\sum_{i=1}^{g}  k \times f_i }{64 \times rc}\nonumber \\
+  &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  - k \times\sum_{i=1}^{g}  f_i }{64 \times rc}\nonumber 
+\end{align}
+
+ Knowing that $$\sum_{i=1}^{g} f_i = rc, $$ we can simplify the equation above, 
+obtaining (\ref{eq:17}).
+
+\begin{align}\label{eq:17}
+ \eta_2 &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  - k \times rc }{64 \times rc}\nonumber \\
+  &= 1 - \frac{\sum_{i=1}^{g}  b_i \times f_i }{64 \times rc} - \frac{k}{64}
+\end{align}
+
+For this method, the highest value obtained for the data compression is 0.9687, 
+when b = 1 and k = 1. However, the lower value is defined by a range of values, 
+​​such that the value of $k = \frac{\sum_{i=1}^{g}  b_i \times f_i}{rc} - 64$. 
+\todo{Adicionei esse trecho para ficar igual ao SM}
+
+% Results and Discussion can be combined.
+\section*{Results}
+
+\paragraph{Random Matrix Generation.}
+In both methods, compression efficiency depends on the distribution of the 
+bit-lengths $b(m_{i,j})$. Thus, in this section, a method to generate a variety 
+of random bit-length distributions is proposed.
+
+For simplicity we will model the distribution of as a mixture $X$ of two Beta 
+distributions, $B_1 \sim Beta(\alpha_1,\beta_1)$ and $B_2 \sim 
+Beta(\alpha_2,\beta_2)$, whose probability function is shown in Equation 
+\ref{eq:18}. Since the Beta distribution is defined only in the interval $[0,1] 
+\subset \mathbb{R}$ , we applied a simple transformation ($\lfloor 64 \times x 
+\rfloor + 1$) to the mixture in order to map it to the interval of $[1,64] 
+\subset \mathbb{Z}$.
+
+\begin{equation}\label{eq:18}
+ f(x) =  w \, Beta(\alpha_1,\beta_1) + (1-w)\,Beta(\alpha_2,\beta_2)
+\end{equation}
+
+The intention of using this mixture was to find a simple way to represent a 
+large variety of bit length distributions. The first two central moments of this 
+mixture are given in \ref{eq:19} and it will be used later to summarize our 
+numerical results.
+
+\begin{align}\label{eq:19}
+ E(X) &= w E(B_1) + (1-w) E(B_2) \nonumber \\
+ &= w \frac{\alpha_1}{\alpha_1+\beta_1} + (1-w)\frac{\alpha_2}{\alpha_2+\beta_2}\nonumber \\
+ Var(X) &= w Var(B_1) + (1-w) Var(B_2) + w (1-w) (E(B_1)^2 - E(B_2)^2)
+\end{align}
+
+In order to explore the compression efficiency of both methods, we generated 
+samples from the mixture defined above, varying its parameters. From now on, 
+when we mention Beta distribution we will mean the transformed version defined 
+above.
+
+From now on, we will apply Equations \ref{eq:15} and \ref{eq:17}, to determine 
+the compression efficiency of SM and VLB methods for random matrices generated 
+as describe above.
+
+With $w=0$, a single Beta distribution is used. In Figure \ref{fig:02030405}, 
+we show some distributions of bit-lengths for some combinations of $\alpha_1$ 
+and $\beta_1$. From the figure it can be seen that a large variety of unimodal 
+distributions can be generated in the interval $[1,64]$.
+
+As we are sampling from a large set distributions of bit-length, represented by 
+the mixture of betas presented above, in order to make our results more general, 
+we will base our analysis on the expected bit-length of a sample, since the 
+efficiency of both methods depends on it. So, from Equations \ref{eq:15} and 
+\ref{eq:17}, the expected efficiencies become:
+
+\begin{equation}\label{eq:20}
+ E(\eta_1) = 1 - \frac{k}{64}
+\end{equation}
+
+\begin{equation}\label{eq:21}
+ E(\eta_2) = 1 - \frac{E(b)}{64} - \frac{k}{64}
+\end{equation}
+
+\noindent where k, in (\ref{eq:21}), is set to 7 (the bit-length required to 
+represent the largest possible bit-length: 64). In (\ref{eq:20}), $k$ is the 
+bit-length of the greatest element, or in the worst case, 64.
+
+We will use the difference $D=E(\eta_1)-E(\eta_2)$ to compare the efficiency of 
+the two methods. Thus a positive $D$ will favor SM method while a negative $D$ 
+favors VLB method.
+
+The expected compression efficiency in the following numeric experiments, will 
+be calculated from $3$ matrices of dimension $10000$, generated as described, 
+and presented in tables and figures below.
+
+In Figure \ref{fig:06070809}, we can see the distribution of efficiencies and 
+their difference for a sample generated from a single Beta distribution of 
+bit-lengths. Note that both methods can achieve efficiencies greater than 80\% 
+for matrices with very small numbers. Also note that the VLB method is more 
+efficient in the majority of cases.
+
+Now let $w = 0.5$, ie, matrices will have elements with bit-lengths coming 
+from a mixture of beta distributions, $B_1\sim Beta(\alpha_1,\beta_1)$ and 
+$B_2\sim Beta(\alpha_2,\beta_2)$. The expected value for this mixture is shown 
+in Equation \ref{eq:22}.
+
+\begin{equation}\label{eq:22}
+ E(B) = 0.5 E(B_1) + 0.5 E(B_2)
+\end{equation}
+
+For bit-lengths coming from a mixture ($w>0$), let the expected efficiencies for 
+the SM and VLB methods be as given by Equations \ref{eq:23} and \ref{eq:24}. So 
+now, instead of having the efficiency be a function of greatest bit-length in 
+the sample (denoted as  $k$ in \ref{eq:15} and \ref{eq:17}), it will be a 
+function of $max\{E(B_1),E(B_2)\}$.
+
+\begin{equation}\label{eq:23}
+ E(\eta_1) = 1 - \frac{max\{E(B_1),E(B_2)\}}{64}
+\end{equation}
+
+\begin{equation}\label{eq:24}
+ E(\eta_2) = 1 - 0.5\frac{E(B_1)}{64} - 0.5\frac{E(B_2)}{64} - \frac{max\{E(B_1),E(B_2)\}}{64}
+\end{equation}
+
+As before, we generate 3 matrices of dimension $10,000$ for each 
+parameterization, calculate the average efficiencies (Equations \ref{eq:23} and 
+\ref{eq:24}) and their difference $D$.
+
+Before moving on to efficiency results and analyses, let's first inspect samples 
+from the mixture of transformed Beta distributions. Figures \ref{fig:10111213} 
+and \ref{fig:14151617}, show a few parameterizations and their resulting sample 
+distributions. It is important to note that from the mixture, we can now 
+generate bimodal distributions as well as the unimodal types tested before. 
+Since we are making statements about efficiency as a function of the expected 
+bit-length, it is important to verify if these statements hold for bimodal  
+distributions as well.
+
+After sampling uniformly ($[1,5,9,\ldots,64]$, $n=65,536$) the bit-length space 
+and comparing efficiencies, we summarized the results on Table \ref{tab:02}. In 
+it we see how many parameterizations (from our sample) favor each method. We can 
+also look at the distribution of efficiencies on our samples for each method 
+(Figure \ref{fig:1819}), which clearly demonstrate the greater expected 
+efficiency of method VLB (Figure \ref{fig:19}).
+
+As we have shown, the VLB method is more effective compressing most integer 
+datasets up to 64 bits in size. This is due to its ability to exploit the 
+variance in the data set and reduce the waste of bits in the representation of 
+some numbers. In specific cases, where the variance in the data null or too 
+small, method I will be more efficient. As a matter of fact, for matrices where 
+all elements have the same bit-length, SM method  will always be better, 
+regardless of bit-length (Figures \ref{fig20} and \ref{fig21}). There are only 
+exception, if for bit-length 64, where neither method is able to compress the 
+data. \todo{Alterei a última frase}
+
+\section*{Discussion}
+
+\subsection*{Calculating Efficiencies}
+
+To determine the best compression method to apply, it's necessary to inspect the 
+distribution of bit-lengths of matrix elements. When matrix elements are small 
+or have nearly-constant bit-length, the SM Method is better\todo{alterei de 
+best para better}, otherwise, the VLB method should be chosen.
+
+As an example, let $M_{r\times c}$ be a integer matrix, such that the half of 
+its elements have bit-length $1$ and the other half $64$. Recalling Equation 
+\ref{eq:17}, now we have two groups of elements (by bit-length), $b_1=1 $, 
+$b_2=64$ and $f_i = \frac{rc}{2}$ for $i = 1 \text{ and } 2$. As the greatest 
+bit-length is $64$, then $k=7$. Compression efficiency $\eta_2$ can be 
+calculated using Equation \ref{eq:17}. After plugging in our numbers, we obtain 
+a compression of $38.29\%$.
+
+\begin{equation*}\label{eq:25}
+ \eta_2 = 1 - \frac{\sum_{i=1}^{2}  b_i \times f_i }{64 \times rc} - \frac{7}{64} 
+\end{equation*}
+
+\begin{equation*}\label{eq:26}
+ \eta_2 = 1 - \frac{  1 \times \frac{rc}{2} + 64 \times \frac{rc}{2} }{64 \times rc} - \frac{7}{64} 
+\end{equation*}
+
+\begin{equation*}\label{eq:26}
+ \eta_2 = 1 - \frac{  32.5  \times rc }{64 \times rc} - \frac{7}{64} 
+\end{equation*}
+
+\begin{equation*}\label{eq:27}
+ \eta_2 \approx 1 - 0.5078 - 0.1093
+\end{equation*}
+
+\begin{equation*}\label{eq:28}
+ \eta_2 \approx 38.29\%
+\end{equation*}
+
+The efficiency of the VLB method is influenced by the relative size of the 
+bit-length groups. In this first example, we considered only two groups, each 
+comprised of half the matrix elements. Let's now vary the relative frequency of 
+the groups, $\frac{f_i}{rc}$, while sticking to two groups. Let's 
+also\todo{Acho que tinha que tirar esse Let's also, talvez colocar Now, we are 
+going to} assume that $\frac{f_i}{rc}$ is a good approximation to the 
+probability of a given bit-length in a matrix, which we will denote by $p_i$.
+
+With this definition, we can rewrite the Equation \ref{eq:17}, which becomes 
+\ref{eq:29}. In Equation \ref{eq:29}, the $\frac{f_i}{rc}$ is replaced by $p_i$, 
+representing the probability of elements from group $i$ in matrix M.
+
+\begin{equation}\label{eq:29}
+ \eta_2 = 1 - \frac{\sum_{i=1}^{g}  b_i \times p_i }{64} - \frac{k}{64} 
+\end{equation}
+
+\noindent with
+
+\begin{equation}\label{eq:30}
+ p_i = \frac{f_i}{rc}
+\end{equation}
+
+With the Equation \ref{eq:29} can analyze the influence of bit-length 
+probability in compression efficiency. In this example, $p_1$ and $p_2$ 
+represent the probability of elements of bit-lengths 1 and 64, respectively. 
+Thus, efficiency is defined in Equation \ref{eq:31}.
+
+\begin{equation}\label{eq:31}
+ \eta_2 = 1 - \frac{1 \times p_1  + 64 \times p_2}{64} - \frac{7}{64} 
+\end{equation}
+
+Now, we can determine which probabilities give us the best and worst compression 
+levels. When $\eta_2=1$, then the efficiency is maximal and if $\eta_2=0$, a 
+efficiency is minimal. To calculate the values ​​of $p_1$ and $p_2$ for both 
+extreme values of $\eta_2$,  we must solve the linear systems shown in Equations 
+\ref{eq:32} and \ref{eq:33}. The first equation on both systems come from the 
+law of total probability. The second comes from \ref{eq:31} after setting 
+$\eta_2$ to $1$ and $0$, respectively.
+
+\begin{equation}\label{eq:32}
+  \eta_2 = 1 :\left
+  \{\begin{matrix}
+    p_1 + p_2 = 1\\ 
+    p_1+64p_2 = 7
+  \end{matrix}
+  \right.
+\end{equation}
+
+Solving the system above, we find that  when $p_1=0.9047$ and $p_2=0.0953$, 
+efficiency is maximal, and in this particular case is equal to 87.5\%.
+
+\begin{equation}\label{eq:33}
+  \eta_2 = 0 :\left
+  \{\begin{matrix}
+    p_1 + p_2 = 1\\ 
+    p_1+64p_2 = 57
+  \end{matrix}
+  \right.
+\end{equation}
+
+Thus, when $p_1=0.1111$ and $p_2=0.8889$ the efficiency is minimal for the VLB 
+method. For other combinations see Table \ref{tab:03}. Looking at this table, 
+one can see two negative efficiencies, when ($p_1$,$p_2$) assume the values 
+​​(0,1) and (0.1,0.9). This correspond th cases, when the method increases, the 
+memory requirements instead of decreasing it.
+
+So far, we have examined only two groups (hence two probabilities) of bit-length 
+for the sake of simplicity. Before, we generalize to probability distributions 
+let's take a quick look at the efficiencies for more groups, with uniform 
+probability:
+
+\begin{itemize}
+  \item 3 groups with bit-lengths 1, 32 and 64 bits, efficiency $\eta_2=0.3854$,
+  \item 5 groups with bit-lengths 1, 16, 32, 48 and 64 bits, efficiency $\eta_2=0.3875$,
+  \item 8 groups with bit-lengths 1, 8, 16, 24, 32, 40, 48, 56 and 64 bits, efficiency $\eta_2=0.3888$
+\end{itemize}
+
+When the distribution of the group probabilities is uniform, ie, the groups 
+have approximately the same size, efficiency is basically the same, regardless 
+of the number of groups.
+
+Now, we can leverage the notion of bit-length probabilities, and study 
+efficiency when bit-lengths follow some commonly used discrete probability 
+distributions: Discrete Uniform, Binomial and Poisson. For all the experiments, 
+we assume $k=7$, that is, the maximum possible bit-length is 64 bits. Thus, 
+efficiency obtained will not be the best possible, since for that we would need 
+assume small values of $k$ (Equation \ref{eq:29}).
+
+\subsection*{Discrete Uniform}
+Let the bit-lengths of the matrices be distribute according to the Uniform 
+distribution $U(a=1,b=64)$, which means bit-lengths may take values in the set 
+$\{1, 2, 3, $\ldots$, 64\}$ with equal probability, ie, $\frac{1}{64}$.
+
+\paragraph{Theoretical Efficiency:}
+Let the random variable $B \sim U(a=1,b=64)$ represent the  bit-length of the 
+elements of matrix $M$. Then $E(b_i) = \sum_{i} b_i \times p(b_i) = 
+\frac{a+b}{2}$.  Applying this result to the expected compression efficiency of 
+VLB method (Equation \ref{eq:29}), we have
+
+\begin{equation}\label{eq:44}
+ E(\eta_2) = 1 - \frac{E(B)}{64} - \frac{k}{64} 
+\end{equation}
+
+\noindent assuming all bit-lengths are possible, i.e., $a=1$ and $b=64$, and 
+hence $k=7$, we can calculate $\eta_2$:
+
+\begin{equation}\label{eq:45}
+ E(\eta_2) = 1 - \frac{\frac{1+64}{2}}{64} - \frac{7}{64} \approx 38.28\% 
+\end{equation}
+
+This result agrees with the numerical estimates presented in Table \ref{tab:04}.
+
+\paragraph{Numerical Estimates:}
+
+To calculate the VLB efficiency, we generated a matrices with 100 ($M_{10 \times 
+10}$), 10,000 ($M_{100 \times 100}$) and 1,000,000 ($M_{1,000 \times 1,000}$) 
+elements with 1, 8, 16, 32 and 64 number of bits.  The average efficiency (Table 
+\ref{tab:03}) is calculated from a 1,000 replicates of each matrix size. As 
+expected the compression efficiency gets better with lower expected bit-length.
+
+\subsection*{Binomial Distribution}
+
+For the binomial distribution, we will use $Bin(n, p)$, with the number of 
+trials $n$ representing the greatest possible bit-length in the matrix, and $np$ 
+giving us the expected bit-length.
+
+\paragraph{Theoretical Efficiency:}
+Let bit-length $(B)$ be a random variable with Binomial distribution, $B \sim 
+B(n=64,p=0.5)$, $E(b_i) = \sum_{i} b_i \times p(b_i) = n \times p$ and the 
+efficiency becomes (with $k=7$):
+
+\begin{align}\label{eq:47}
+ E(\eta_2) &= 1 - \frac{64 \times p }{64} - \frac{k}{64}  \\
+  &= 1 - \frac{64 \times 0.5 }{64} - \frac{7}{64} \approx 39.05\% \nonumber
+\end{align}
+
+Which again agrees with estimates in Table \ref{tab:05}.
+
+\paragraph{Numerical Estimates:}
+
+For these experiments, the parameter n represents the maximum bit-length of 
+matrix elements and takes values in $\{1, 8, 16, 32, 64\}$. In this case, we 
+evaluate the efficiency as a function of the parameter $n$, and matrix size. 
+Even though efficiency does not depend on matrix size, we tried different sizes 
+to test the stability of the compression algorithm. Results are shown in Table 
+\ref{tab:05}. As expected, smaller bit-lengths lead to higher compression 
+efficiencies.
+
+\subsection*{Poisson Distribution}
+With bit-length derived from a Poisson($\lambda$), the parameter $\lambda$ 
+corresponds to the expected bit-length. For the purpose of this analysis this 
+Poisson distribution is truncated at 64.
+
+\paragraph{Theoretical Efficiency:}
+Let bit-length $B \sim Poisson(\lambda=32)$. In this case, $E(b_i) = \lambda$, 
+with $k=7$,  the efficiency becomes:
+
+\begin{align}
+ E(\eta_2) &= 1 - \frac{\lambda}{64} - \frac{k}{64} \label{eq:49} \\
+ &= 1 - \frac{32}{64} - \frac{7}{64} \approx 39.06\% \label{eq:50}
+\end{align}
+
+This result is in accordance to Table \ref{tab:06}.
+
+\paragraph{Numerical Estimates:}
+The results for this simulation can be seen in \ref{tab:06}. Note that 
+increasing the number of bits to represent numbers increases, there is a loss of 
+efficiency in the compression process. In this case, we did not simulate for 
+$\lambda=64$ since a large portion of the samples would fall above the maximum 
+bit-length, but \todo{coloquei ", but", aqui} we are considering for this 
+analysis.
+
+These results show that a good compression is guaranteed when bit-lengths are 
+distributed according to the tested distributions regardless of sample size.
+
+\todo{Parei aqui}
+
+\section{Conclusion}
+In this paper, we have focused in the compression of matrix data, since this is
+one of the most important application the authors foresee. However, the
+compression methodology presented can be applied to any numerical data
+structure, with gains to performance and memory footprint\cite{teseCrysttian}.
+
+Further discussions about doing computation with such compressed data-structures
+will be the subject of another manuscript (in preparation) in which we will
+present details about the implementation of the compression algorithm, and
+benchmarks on classical linear algebra tasks such as those in
+Linpack\cite{dongarra1979linpack}.
+
+For the compression calculations presented in this paper we limited bit-lenght of integers to 64 bits.
+However the compression would work in the same way as discussed for computer architectures with larger word sizes.
+
+Representation of floating point numbers is also possible within the proposed compression framework, but at the expense of precision in their representation.
+Although this may sound like a limitation, when we take into consideration that most experimental data have fewer ``significant'' digits than the maximal precision available in modern computers, fairly good compression may still be achievable for floats.
+
+% You may title this section "Methods" or "Models". 
+% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
+% authors may use "Analysis"
+
+% Do NOT remove this, even if you are not including acknowledgments
+\section*{Acknowledgments}
+We would like to thank Claudia Torres Code\c{c}o, Moacyr Silva and Paulo
+Cezar Carvalho for fruitful discussions and key ideas which helped improve the
+manuscript.
+
+%\section*{References}
+% The bibtex filename
+\bibliography{plos.bib}
+
+\section*{Figure Legends}
+
+\begin{figure}[h]
+  \includegraphics[scale=0.3,clip]{fig01}
+  \caption{Representation of the compression process and allocation of 
+elements of matrix M in the matrix m. The strip of value 700 is divided in two 
+parts and stores in different strips in the matrix m.}
+  \label{fig:01}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$\alpha=1,\beta_1=1$]{
+   \includegraphics[scale=0.25,angle=-90,clip]{fig02}
+  }
+  \subfigure[$\alpha=1,\beta=32$]{
+   \includegraphics[scale=0.25,angle=-90,clip]{fig03}
+  }
+  \subfigure[$\alpha=32,\beta=1$]{
+   \includegraphics[scale=0.25,angle=-90,clip]{fig04}
+  }
+  \subfigure[$\alpha=64,\beta=64$]{
+   \includegraphics[scale=0.25,angle=-90]{fig05}
+  }
+  \caption{Histograms constructed from samples with 10,000 elements, generated 
+from a Beta distribution. Below each histogram is possible to verify the 
+parameters used.With these two parameters it is possible to perform different 
+combinations of numbers to fill the arrays to be compressed.}
+  \label{fig:02030405}
+\end{figure}\todo{Adicionei a última frase}
+
+\begin{figure}[h]
+  \centering
+   \subfigure[SM-VLB  Difference ($D = \eta_1 - \eta_2$)]{ 
+    \includegraphics[scale=0.19,clip]{fig08}
+   }
+   \subfigure[SM Efficiency  ($\eta_1$)]{
+    \includegraphics[scale=0.19,clip]{fig07}
+   }
+   \subfigure[VBL Efficiency ($\eta_2$)]{
+    \includegraphics[scale=0.19,clip]{fig06}
+   }
+   \subfigure[SM$\geq$VLB Efficiency ($\eta_1 \geq \eta_2$)]{
+    \includegraphics[scale=0.19,clip]{fig09}
+   }
+  \caption{Comparing compression efficiency of methods 1 and 2. Color scale in 
+represent average bit-length. In (a) we can see the difference $D = \eta_1 - 
+\eta_2$. It can be seen that for most combination of $\alpha$ and $\beta$, 
+$D<0$, meaning the second method is more efficient to compress a sample of 
+numbers with bit-lengths coming from a $Beta(\alpha,\beta)$ distribution. 
+However, there is a small region in parameter space, which is shown in white on 
+(d), where the SM method is more efficient. This region corresponds to the dots 
+in red in (a), where the average bit-length is higher. In panels (b) and (c), we 
+can see the efficiencies of SM and VLB methods, respectively.}
+  \label{fig:06070809}
+\end{figure}
+\todo{A escala de cor é a mesma em todas as figuras, no texto estava dizendo 
+que era apenas na a, b e c. Eu alterei.}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$\alpha_1=1,\beta_1=1,\alpha_2=1,\beta_2=1$]{
+   \includegraphics[scale=0.28,angle=-90,clip]{fig10}
+  }
+  \subfigure[$\alpha_1=1,\beta_1=32,\alpha_2=32,\beta_2=1$]{
+   \includegraphics[scale=0.28,angle=-90,clip]{fig11}
+  }
+  \subfigure[$\alpha_1=32,\beta_1=32,\alpha_2=32,\beta_2=32$]{
+   \includegraphics[scale=0.28,angle=-90,clip]{fig12}
+  }
+  \subfigure[$\alpha_1=64,\beta_1=32,\alpha_2=32,\beta_2=64$]{
+   \includegraphics[scale=0.28,angle=-90]{fig13}
+  }
+  \caption{Histograms constructed from samples with 10,000 elements, generated 
+from the mixture of two Beta distributions with $w=0.5$. Below each histogram 
+are the parameters of the mixture.}
+  \label{fig:10111213}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$\alpha_1=64,\beta_1=48,\alpha_2=1,\beta_2=48$]{
+   \includegraphics[scale=0.28,angle=-90]{fig14}
+  }
+  \subfigure[$\alpha_1=16,\beta_1=46,\alpha_2=49,\beta_2=64$]{
+   \includegraphics[scale=0.28,angle=-90]{fig15}
+  }
+  \subfigure[$\alpha_1=16,\beta_1=16,\alpha_2=16,\beta_2=49$]{
+   \includegraphics[scale=0.28,angle=-90]{fig16}
+  }
+  \subfigure[$\alpha_1=1,\beta_1=16,\alpha_2=1,\beta_2=49$]{
+   \includegraphics[scale=0.28,angle=-90]{fig17}
+  }
+  \caption{Histograms constructed from samples with 10,000 elements, generated 
+from the mixture of two Beta distributions with $w=0.5$. Below each histogram 
+are the parameters of the mixture.}
+  \label{fig:14151617}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[Efficiency histogram of the SM method ]{
+   \includegraphics[scale=0.25,angle=-90]{fig18}
+  \label{fig:18}
+  }
+  \subfigure[Efficiency histogram of the VLB method ]{
+   \includegraphics[scale=0.25,angle=-90]{fig19}
+  \label{fig:19}
+  }
+  \caption{Efficiency histograms of the SM and VLB methods. Note that the VLB 
+method has a greater average efficiency than SM method.}
+  \label{fig:1819}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+   \subfigure[Efficiencies of the SM and VLB methods, for constant bit-length 
+matrices.]{\includegraphics[scale=0.3,clip]{fig20}\label{fig20}}
+   \subfigure[$\eta_1-\eta_2$ for matrices of constant 
+bit-length.]{\includegraphics[scale=0.3,clip]{fig21}\label{fig21}}
+  \caption{Compression efficiency of the SM and VLB methods for matrices of 
+constant bit-length.}
+  \label{fig:2021}
+\end{figure}
+
+%\begin{figure}[!ht]
+%\begin{center}
+%%\includegraphics[width=4in]{figure_name.2.eps}
+%\end{center}
+%\caption{
+%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
+%should be left justified, as specified by the options to the caption 
+%package.
+%}
+%\label{Figure_label}
+%\end{figure}
+
+\section*{Tables}
+%\begin{table}[!ht]
+%\caption{
+%\bf{Table title}}
+%\begin{tabular}{|c|c|c|}
+%table information
+%\end{tabular}
+%\begin{flushleft}Table caption
+%\end{flushleft}
+%\label{tab:label}
+% \end{table}
+
 \begin{table}[h]
@@
 
-To store matrix $M$ it first has to be converted to base $2$ ($M_2$). Then it 
-will be unraveled by column (column major, e.g. in Fortran) or by row (row 
-major, e.g. in C) and its elements will be written as fixed size adjacent chunks 
-of memory. The size of each chunk is determined by the type associated with the 
-matrix (typically 64 bits, but always a power of 2).
-
-According to the SM method, having determined that each element will require at 
-most 10 bits, we can divide the memory block corresponding to a single 64 bit 
-integer into six 10-bit chunks, which can each hold a single element of $M$. 
-These 64-bit blocks will be called a bitstring. The remaining 4 bits will be 
-used later. The number of bitstrings needed will be $\lfloor \frac{\dim(M) 
-*b(\sup M)} {64} \rfloor +1$, where $dim(M)$ is the dimension of the matrix or 
-its number of elements.
-
-The final layout of the first 6 elements of $m$ in the first bitstring can be seen in \ref{eq:03}.
-
-\begin{equation}\label{eq:03}
- bitstring_1 = 0000\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\underbrace{0100000000}_{256}\underbrace{1011010001}_{721}\underbrace{1111111111}_{1023}\underbrace{1110000100}_{900} 
-\end{equation}
-
-Here is a step-by-step description of the application of the SM method to matrix $M$:
-
-\begin{enumerate}
- \item Element $M_{1,1}=900 = m_{1,1} = 1110000100$ is stored in the first 10-bit chunk of the element strip $bitstring[1]$ , which corresponds to bits 0 to 9 (read from right to left).
-\begin{equation*}\label{eq:04}
-bitstring_1 = 000000000000000000000000000000000000000000000000000000\underbrace{1110000100}_{900} 
-\end{equation*}
- \item Element $M_{1,2}=1023$ is allocated in the second chunk, from bit 10 to bit 19.
-\begin{equation*}\label{eq:05}
- bitstring_1 = 00000000000000000000000000000000000000000000\underbrace{1111111111}_{1023}\underbrace{1110000100}_{900} 
-\end{equation*}
- \item Repeat for elements $M_{1,i}$ with $i=1,\ldots,6$ which are stored on the remaining chunks.
- \begin{equation*}\label{eq:06}
-bitstring_1 = 0000\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\underbrace{0100000000}_{256}\underbrace{1011010001}_{721}\underbrace{1111111111}_{1023}\underbrace{1110000100}_{900} 
- \end{equation*}
- \item Element $M_{1,7}=700=1010111100$ does not fit on the remaining 4 bits of 
-the first bitstring. So it will straddle two bitstrings, ie, it is divided in 
-two segments $a$ and $b$, $a$ is written on the first bitstring and $b$ on the 
-second.
- \begin{equation*}\label{eq:07}
-  \underbrace{\underbrace{0000010100}_{20}\overbrace{101011}^{b}}_{bitstring_2}|\underbrace{\overbrace{1100}^{a}\underbrace{0000001010}_{10}\underbrace{0000000001}_{1}\hdots}_{bitstring_1}
-\end{equation*}
-\end{enumerate}
-
- Please, notice that bitstrings are written from right to left.
-
-\begin{equation*}\label{eq:8}
-  \underbrace{0000010100}_{20}\underbrace{\overbrace{101011}^a|\overbrace{1100}^{b}}_{700}\underbrace{0000001010}_{10}\ldots
-\end{equation*}
-
-Thus the compressed matrix $m = M_2$ requires less memory than the conventional 
-storage of $M$ as a 64-bit integer array. An illustration of the compression 
-process and the allocation of the elements of matrix M in the matrix m is shown 
-in Figure \ref{fig:01}. Note that the matrix m has fewer columns that the 
-matrix M, then it demands less memory to be represented. We also emphasize that 
-the bits which represent the value 700 is divided in two parts and stored in 
-different strips in the matrix m.
-
-\subsubsection*{Method 2: Variable Length Blocks (VLB)}
-
-In the SM method, there is still waste of space since for elements smaller than 
-the supremum, a number of bits remain unused.
-
-In the VBL method, the absolute minimal number of bits are used to store each 
-value. However, if we are going to divide the biststrings into variable length 
-chunks, we also need to reserve some extra bits to represent the size of each 
-chunk, otherwise the elements cannot be recovered once they are stored.
-
-Lets use again the matrix described in Equation \ref{eq:02}, where  the largest 
-element is number 1023. Now instead of assigning one chunk of the bitstring to 
-each element of $m$, we will assign two chunks: the first will store the number 
-of bits required to store the element and the second  will store the actual 
-element. The first chunk will have a fixed size, in this case, 4 bits. These 4 
-bits are the required space to store the bit-length of $\sup M $, in this case, 
-10.
-
-Lets go through VLB compression step-by-step. The largest element of $M$ is 
-1023. Its bit-length is 10 which in turn is 4 bits long in base 2 (1010). Thus 
-the fixed size chunk is 4 bits long for every element.
-
-\begin{enumerate}
- \item The first element $M_{1,1}=900$ requires 10 bits to store, so we write $10$ in the first chunk and $900$ in the second.
- \begin{equation*} \label{eq:09}
- bitstring_1 = 00000000000000000000000000000000000000000000000000\underbrace{\underbrace{1110000100}_{\text{element} = 900}\underbrace{1010}_{\text{bit-length=10}}}_{M_{1,1}}
-\end{equation*}
- \item Do the same for the next element, $M_{1,2}=1023$.
-  \begin{equation*} \label{eq:10}
- bitstring_1 = \ldots00000000000000000000000000\underbrace{\underbrace{1111111111}_{\text{element} = 1023}\underbrace{1010}_{\text{bit-length=10}}}_{M_{1,2}}\underbrace{\underbrace{1110000100}_{\text{element} = 900}\underbrace{1010}_{\text{bit-length=10}}}_{M_{1,1}}
-\end{equation*}
- \item Element $M_{1,3}=721$ is also added taking the bitstring to the state.
- \begin{equation*} \label{eq:11}
- bitstring_1 = 
- 0000000000000000000000\underbrace{1011010001}_{721}\underbrace{1010}_{10}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1110000100}_{900}\underbrace{1010}_{10}
-\end{equation*}
-\end{enumerate}
-
-So far the VLB method is more wasteful than the SM, but when we add $M_{1,4} 
-=256$, we start to save some space.
-
-\begin{enumerate}
- \item[4.] Element $M_{1,4} =256$ is added.  
- \item [5.] Elements $M_{1,5} =1$ and $M_{1,6} =10$ are added requiring a total of 13 bits instead of 20 with the SM method. With the addition of these elements we require a second bitstring.
- \begin{align*} \label{eq:12}
- bitstring_1 &=
- \underbrace{0100}_{4}\underbrace{1}_{1}\underbrace{0001}_{1}\underbrace{100000000}_{256}\underbrace{1001}_{9}\underbrace{1011010001}_{721}\underbrace{1010}_{10}\underbrace{1111111111}_{1023}\underbrace{1010}_{10}\underbrace{1110000100}_{900}\underbrace{1010}_{10} \\
- bitstring_2 &= 000000000000000000000000000000000000000000000000000000000000\underbrace{1010}_{10}
-\end{align*}
- \item [6.] The remaining two elements are added $M_{1,7} =700$ and $M_{1,8} =20$ in the second bit strip.
- \begin{equation*} \label{eq:13}
- bitstring_2 = 00000000000000000000000000000000000\underbrace{10100}_{20}\underbrace{0101}_{5}\underbrace{1010111100}_{700}\underbrace{1010}_{10}\underbrace{1010}_{10}
-\end{equation*}
-\end{enumerate}
-
- We used a total of 87 bits to store matrix $m$ with the VLB method instead of 
-80 bits using the SM method. However, as shall be seen later, the VLB method 
-will be the most efficient for most matrices. \todo{adicionei} With the method 
-VLB the resulting matrix is smaller than the original matrix.
-
- \subsection*{Compression Efficiency}
- Compression efficiency depends of the data being compressed. Below, a formula for calculating compression efficiency is derived for both methods. They will be based on the following ratio:
-
- \begin{equation}\label{eq:14}
-  \eta=\frac{bits\, alocated-bits\, used}{bits\,alocated}
- \end{equation}
-
- Where \textit{bits alocated} above mean total bits required for standard storage of the matrix, without compression, while bits used mean total bits requires to store the matrix after compression.
-From now on the efficiencies are denoted by $\eta_1$ for the SM method and by $\eta_2$ for the VLB method.
- \subsubsection*{SM Method}
- Let $M_{r \times c}$ be the matrix we wish to compress.
-In comparison with a conventional allocation (64-bit integers), we can apply Equation \ref{eq:14} to caluculate the efficiency of the SM method:
-
-\begin{align}\label{eq:15}
- \eta_1 &= \frac{64 \times rc - b(max M) \times rc}{64 \times rc}\nonumber\\
- &= \frac{64  - b(max M) }{64}
-\end{align}
-
-As we see in \ref{eq:15}, $\eta_1$ does not depend on size of the matrix, only on the bit-length of $\max M$.
-If $b(\max M)=64$, $\eta_1$ is 0, i.e., no compression is possible.
-On the other extreme, if the matrix is composed exclusively of 0s and 1s, maximal compression is achievable, $\eta_1=1$.
-
- \subsubsection*{VLB Method}
-
- For the VLB method, compression depends on the value of each element of the matrix.
-In this method bit-lentgh variability affects the compression ratio, so the formula will have to include this information.
-
-Let the $rc$ elements of the matrix $M_{r \times c}$ be divided into $g$ groups, each with $f_i$ numbers of bit-length $b_i = b(m_i)$.
-Thus $f_i$ is the frequency of each bit-length present in $M$.
-Let $k = b(b(\max M))$, i.e., the bit length of the bit-length of $\max M$.
-The efficiency $\eta_2$ is shown below.
-
-\begin{equation}\label{eq:16}
- \eta_2 = \frac{64 \times rc - \sum_{i=1}^{g} ( b_i + k ) \times f_i }{64 \times rc} 
-\end{equation}
-
-We can further simplify Equation \ref{eq:16} to get at shorter expression for the compression ratio.
-
-\begin{align}
- \eta_2 &= \frac{64 \times rc - \sum_{i=1}^{g} ( b_i \times f_i + k \times f_i )}{64 \times rc} \nonumber \\
-  &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  -\sum_{i=1}^{g}  k \times f_i }{64 \times rc}\nonumber \\
-  &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  - k \times\sum_{i=1}^{g}  f_i }{64 \times rc}\nonumber 
-\end{align}
-
- Knowing that $$\sum_{i=1}^{g} f_i = rc, $$ we can simplify the equation above, obtaining (\ref{eq:17}).
-
-\begin{align}\label{eq:17}
- \eta_2 &= \frac{64 \times rc - \sum_{i=1}^{g}  b_i \times f_i  - k \times rc }{64 \times rc}\nonumber \\
-  &= 1 - \frac{\sum_{i=1}^{g}  b_i \times f_i }{64 \times rc} - \frac{k}{64}
-\end{align}
-
-% Results and Discussion can be combined.
-\section*{Results}
-
-\paragraph{Random Matrix Generation.}
-In both methods, compression efficiency depends on the distribution of the bit-lengths $b(m_{i,j})$.
-Thus, in this section, a method to generate a variety of random bit-length distributions is proposed.
-
-For simplicity we will model the distribution of  as a mixture $X$ of two Beta distributions, $B_1 \sim Beta(\alpha_1,\beta_1)$ and $B_2 \sim Beta(\alpha_2,\beta_2)$, whose probability function is shown in Equation \ref{eq:18}.
-Since the Beta distribution is defined only in the interval $[0,1] \subset \mathbb{R}$ , we applied a simple transformation ($\lfloor 64 \times x \rfloor + 1$) to the mixture in order to map it to the interval of $[1,64] \subset \mathbb{Z}$.
-
-\begin{equation}\label{eq:18}
- F(x) =  w \, Beta(\alpha_1,\beta_1) + (1-w)\,Beta(\alpha_2,\beta_2)
-\end{equation}
-
-The intention of using this mixture was to find a simple way to represent a large variety of bit length distributions.
-The first two central moments of this mixture are given in \ref{eq:19} and will be used later to summarize our numerical results.
-
-\begin{align}\label{eq:19}
- E(X) &= w E(B_1) + (1-w) E(B_2) \nonumber \\
- &= w \frac{\alpha_1}{\alpha_1+\beta_1} + (1-w)\frac{\alpha_2}{\alpha_2+\beta_2}\nonumber \\
- Var(X) &= w Var(B_1) + (1-w) Var(B_2) + w (1-w) (E(B_1)^2 - E(B_2)^2)
-\end{align}
-
-In order to explore the compression efficiency of both methods, we generated samples from the mixture defined above, varying its parameters.
-From now on, when we mention Beta distribution we will mean the transformed version defined above.
-
-From now on we will apply Equations \ref{eq:15} and \ref{eq:17}, to determine the compression efficiency of SM and VLB methods for random matrices generated as describe above.
-
-With $w=0$, a single Beta distribution is used.
-In Figure \ref{fig:01020304}, we show some distributions of bit-lengths for some combinations of $\alpha_1$ and $\beta_1$.
-From the figure it can be seen that a large variety of unimodal distributions can be generated in the interval $[1,64]$.
-
-\begin{figure}[ht]
-  \centering
-  \subfigure[$\alpha=1,\beta_1=1$]{
-%   \includegraphics[scale=0.25,angle=-90,clip]{fig01}
-  }
-  \subfigure[$\alpha=1,\beta=32$]{
-%   \includegraphics[scale=0.25,angle=-90,clip]{fig02}
-  }
-  \subfigure[$\alpha=32,\beta=1$]{
-%   \includegraphics[scale=0.25,angle=-90,clip]{fig03}
-  }
-  \subfigure[$\alpha=64,\beta=64$]{
-%   \includegraphics[scale=0.25,angle=-90]{fig04}
-  }
-  \caption{Histograms constructed from samples with 10,000 elements, generated from a Beta distribution. Below each histogram is possible to verify the parameters used.}
-  \label{fig:01020304}
-\end{figure}
-
-As we are sampling from a large set distributions of bit-length, represented by the mixture of betas presented above, in order to make our results more general, we will base our analysis on the expected bit-length of a sample, since the efficiency of both methods depends on it.
-So, from Equations \ref{eq:15} and \ref{eq:17}, the expected efficiencies become:
-
-\begin{equation}\label{eq:20}
- E(\eta_1) = 1 - \frac{k}{64}
-\end{equation}
-
-\begin{equation}\label{eq:21}
- E(\eta_2) = 1 - \frac{E(b)}{64} - \frac{k}{64}
-\end{equation}
-
-\noindent where k, in (\ref{eq:21}), is set to 7 (the bit-length required to represent the largest possible bit-length: 64).
-In (\ref{eq:20}), $k$ is the bit-length of the greatest element, or in the worst case, 64.
-
-We will use the difference $D=E(\eta_1)-E(\eta_2)$ to compare the efficiency of the two methods. Thus a positive $D$ will favor SM method while a negative $D$ favors VLB method.
-
-The expected compression efficiency in the following numeric experiments, will be calculated from $3$ matrices of dimension $10000$, generated as described, and presented in tables and figures below.
-
-In Figure \ref{fig:06070809}, we can see the distribution of efficiencies and their difference for a sample generated from a single Beta distribution of bit-lengths.
-Note that both methods can achieve efficiencies greater than 80\% for matrices with very small numbers.
-Also note that the VLB method is more efficient in the majority of cases.
-
-\begin{figure}[h]
-  \centering
-   \subfigure[SM-VLB  Difference ($D = \eta_1 - \eta_2$)]{ 
-%    \includegraphics[scale=0.18,clip]{fig07}
-   }
-   \subfigure[SM Efficiency  ($\eta_1$)]{
-%    \includegraphics[scale=0.18,clip]{fig06}
-   }
-   \subfigure[VBL Efficiency ($\eta_2$)]{
-%    \includegraphics[scale=0.18,clip]{fig05}
-   }
-   \subfigure[SM$\geq$VLB Efficiency ($\eta_1 \geq \eta_2$)]{
-%    \includegraphics[scale=0.18,clip]{fig08}
-   }
-  \caption{Comparing compression efficiency of methods 1 and 2.
-Color scale in (a), (b) and (c) represent average bit-length.
-In (a) we can see the difference $D = \eta_1 - \eta_2$.
-It can be seen that for most combination of $\alpha$ and $\beta$, $D<0$, meaning the second method is more efficent to compress a sample of numbers with bit-lengths coming from a $Beta(\alpha,\beta)$ distribution.
-However, there is a small region in parameter space, which is shown in white on (d), where the SM method is more efficient.
-This region corresponds to the dots in red in (a), where the average bit-length is higher.
-In panels (b) and (c), we can see the efficiencies of SM and VLB methods, respectively.}
-  \label{fig:06070809}
-\end{figure}
-
-Now let $w = 0.5$, i.e., matrices will have elements with bit-lengths comming from a mixture of beta distributions, $B_1\sim Beta(\alpha_1,\beta_1)$ and $B_2\sim Beta(\alpha_2,\beta_2)$.
-The expected value for this mixture is shown in Equation \ref{eq:22}.
-
-\begin{equation}\label{eq:22}
- E(B) = 0.5 E(B_1) + 0.5 E(B_2)
-\end{equation}
-
-For bit-lengths coming from a mixture ($w>0$), let the expected efficiencies for the SM and VLB methods be as given by Equations \ref{eq:23} and \ref{eq:24}.
-So now, instead of having the efficiency be a function of greatest bit-length in the sample (denoted as  $k$ in \ref{eq:15} and \ref{eq:17}), it will be a function of $max\{E(B_1),E(B_2)\}$.
-
-\begin{equation}\label{eq:23}
- E(\eta_1) = 1 - \frac{max\{E(B_1),E(B_2)\}}{64}
-\end{equation}
-
-\begin{equation}\label{eq:24}
- E(\eta_2) = 1 - 0.5\frac{E(B_1)}{64} - 0.5\frac{E(B_2)}{64} - \frac{max\{E(B_1),E(B_2)\}}{64}
-\end{equation}
-
-As before, we generate 3 matrices of dimension $10,000$ for each parameterization, calculate the average efficiencies (Equations \ref{eq:23} and \ref{eq:24}) and their diference $D$.
-
-Before moving on to efficiency results and analyses, let's first inspect samples from the mixture of transformed Beta distributions.
-Figures \ref{fig:09101112} and \ref{fig:13141516}, show a few parameterizations and their resulting sample distributions.
-It is important to note that from the mixture we can now generate bimodal distributions as well as the unimodal types tested before.
-Since we are making statements about efficiency as a function of the expected bit-length, it is important to verify if these statements hold for bimodal  distributions as well.
-
-After sampling uniformly ($[1,5,9,\ldots,64]$, $n=65,536$) the bit-length space and comparing efficiencies, we summarized the results on Table \ref{tab:02}.
-In it we see how many parameterizations (from our sample) favor each method.
-We can also look at the distribution of efficiencies on our samples for each method (Figure \ref{fig:1718}), which clearly demonstrate the greater expected efficiency of method VLB (Figure \ref{fig:18}).
-
-\begin{figure}[h]
-  \centering
-  \subfigure[$\alpha_1=1,\beta_1=1,\alpha_2=1,\beta_2=1$]{
-%   \includegraphics[scale=0.28,angle=-90,clip]{fig09}
-  }
-  \subfigure[$\alpha_1=1,\beta_1=32,\alpha_2=32,\beta_2=1$]{
-%   \includegraphics[scale=0.28,angle=-90,clip]{fig10}
-  }
-  \subfigure[$\alpha_1=32,\beta_1=32,\alpha_2=32,\beta_2=32$]{
-%   \includegraphics[scale=0.28,angle=-90,clip]{fig11}
-  }
-  \subfigure[$\alpha_1=64,\beta_1=32,\alpha_2=32,\beta_2=64$]{
-%   \includegraphics[scale=0.28,angle=-90]{fig12}
-  }
-  \caption{Histograms constructed from samples with 10,000 elements, generated from the mixture of two Beta distributions with $w=0.5$. Below each histogram are the parameters of the mixture.}
-  \label{fig:09101112}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \subfigure[$\alpha_1=64,\beta_1=48,\alpha_2=1,\beta_2=48$]{
-%   \includegraphics[scale=0.28,angle=-90]{fig13}
-  }
-  \subfigure[$\alpha_1=16,\beta_1=46,\alpha_2=49,\beta_2=64$]{
-%   \includegraphics[scale=0.28,angle=-90]{fig14}
-  }
-  \subfigure[$\alpha_1=16,\beta_1=16,\alpha_2=16,\beta_2=49$]{
-%   \includegraphics[scale=0.28,angle=-90]{fig15}
-  }
-  \subfigure[$\alpha_1=1,\beta_1=16,\alpha_2=1,\beta_2=49$]{
-%   \includegraphics[scale=0.28,angle=-90]{fig16}
-  }
-  \caption{Histograms constructed from samples with 10,000 elements, generated from the mixture of two Beta distributions with $w=0.5$. Below each histogram are the parameters of the mixture.}
-  \label{fig:13141516}
-\end{figure}
-
 \begin{table}[h]
  \centering
- \caption{Efficiency comparison of SM and VLB methods for parameters covering uniformly the support of $B$.
-Column $n$ shows the number of parameter combinations with which each method has superior compression.}
- \begin{tabular}{ccc}
+ \caption{Efficiency comparison of SM and VLB methods for parameters covering 
+uniformly the support of $B$. Column $n$ shows the number of parameter 
+combinations with which each method has superior compression.}
+ \begin{tabular}{crr}
   \hline 
@@
 \end{table}
-
-\begin{figure}[h]
-  \centering
-  \subfigure[Efficiency histogram of the SM method ]{
-%   \includegraphics[scale=0.25,angle=-90]{fig17}
-  \label{fig:17}
-  }
-  \subfigure[Efficiency histogram of the VLB method ]{
-%   \includegraphics[scale=0.25,angle=-90]{fig18}
-  \label{fig:18}
-  }
-  \caption{Efficiency histograms of the SM and VLB methods. Note that the VLB method has a greater average efficiency than SM method .}
-  \label{fig:1718}
-\end{figure}
-
-As we have shown, the VLB method is more effective compressing most integer datasets up to 64 bits in size.
-This is due to its ability to exploit the variance in the data set and reduce the waste of bits in the representation of some numbers.
-In specific cases where the variance in the data null or too small, method I will be more efficient.
-As a matter of fact, for matrices where all elements have the same bit-length, SM method  will always be better, regardless of bit-length (Figures \ref{fig19} and \ref{fig20}), The only exception if for bit-length 64 where neither method is able to compress the data.
-
-\begin{figure}[h]
-  \centering
-%   \subfigure[Efficiencies of the SM and VLB methods, for constant bit-length matrices.]{\includegraphics[scale=0.3,clip]{fig19}\label{fig19}}
-%   \subfigure[$\eta_1-\eta_2$ for matrices of constant bit-length.]{\includegraphics[scale=0.3,clip]{fig20}\label{fig20}}
-  \caption{Compression efficiency of the SM and VLB methods for matrices of constant bit-length.}
-  \label{fig:1920}
-\end{figure}
-
-\section*{Discussion}
-
-\subsection*{Calculating Efficiencies}
-
-To determine the best compression method to apply, it's necessary to inspect the distribution of bit-lengths of matrix elements.
-When matrix elements are small or have nearly-constant bit-length, the SM Method is best, otherwise, the VLB method should be chosen.
-
-As an example, let $M_{r\times c}$ be a integer matrix such that the half of its elements have bit-length $1$ and the other half $64$.
-Recalling Equation \ref{eq:17}, now we have two groups of elements (by bit-length), $b_1=1 $, $b_2=64$ and $f_i = \frac{rc}{2}$ for $i = 1 \text{ and } 2$.
-As the greatest bit-length is $64$, then $k=7$.
-Compression efficiency $\eta_2$ can be calculated using Equation \ref{eq:17}.
-After plugging in our numbers, we obtain a compression of $38.29\%$.
-
-\begin{equation*}\label{eq:25}
- \eta_2 = 1 - \frac{\sum_{i=1}^{2}  b_i \times f_i }{64 \times rc} - \frac{7}{64} 
-\end{equation*}
-
-\begin{equation*}\label{eq:26}
- \eta_2 = 1 - \frac{  1 \times \frac{rc}{2} + 64 \times \frac{rc}{2} }{64 \times rc} - \frac{7}{64} 
-\end{equation*}
-
-\begin{equation*}\label{eq:26}
- \eta_2 = 1 - \frac{  32.5  \times rc }{64 \times rc} - \frac{7}{64} 
-\end{equation*}
-
-\begin{equation*}\label{eq:27}
- \eta_2 \approx 1 - 0.5078 - 0.1093
-\end{equation*}
-
-\begin{equation*}\label{eq:28}
- \eta_2 \approx 38.29\%
-\end{equation*}
-
-The efficiency of the VLB method is influenced by the relative size of the bit-length groups.
-In this first example we considered only two groups, each comprised of half the matrix elements.
-Let's now vary the relative frequency of the groups, $\frac{f_i}{rc}$, while sticking to two groups.
-Let's also assume that $\frac{f_i}{rc}$ is a good approximation to the probability of a given bit-length in a matrix, which we will denote by $p_i$.
-
-With this definition we can rewrite the Equation \ref{eq:17}, which becomes \ref{eq:29}.
-In Equation \ref{eq:29}, the $\frac{f_i}{rc}$ is replaced by $p_i$, representing the probability of elements from group $i$ in matrix M.
-
-\begin{equation}\label{eq:29}
- \eta_2 = 1 - \frac{\sum_{i=1}^{g}  b_i \times p_i }{64} - \frac{k}{64} 
-\end{equation}
-
-\noindent with
-
-\begin{equation}\label{eq:30}
- p_i = \frac{f_i}{rc}
-\end{equation}
-
-With the Equation \ref{eq:29} can analyze the influence of bit-length probability in compression efficiency.
-In this example, $p_1$ and $p_2$ represent the probability of elements of bit-lengths 1 and 64, respectively.
-Thus, efficiency is defined in Equation \ref{eq:31}.
-
-\begin{equation}\label{eq:31}
- \eta_2 = 1 - \frac{1 \times p_1  + 64 \times p_2}{64} - \frac{7}{64} 
-\end{equation}
-
-Now, we can determine which probabilities give us the best and worst compression levels.
-When $\eta_2=1$, then the efficiency is maximal and if $\eta_2=0$, a efficiency is minimal.
-To calculate the values ​​of $p_1$ and $p_2$ for both extreme values of $\eta_2$,  we must solve the linear systems shown in Equations \ref{eq:32} and \ref{eq:33}.
-The first equation on both systems come from the law of total probability.
-The second comes from \ref{eq:31} after setting $\eta_2$ to $1$ and $0$, respectively.
-
-\begin{equation}\label{eq:32}
-  \eta_2 = 1 :\left
-  \{\begin{matrix}
-    p_1 + p_2 = 1\\ 
-    p_1+64p_2 = 7
-  \end{matrix}
-  \right.
-\end{equation}
-
-Solving the system above, we find that  when $p_1=0.9047$ and $p_2=0.0953$, efficiency is maximal, and in this particular case is equal to 87.5\%.
-
-\begin{equation}\label{eq:33}
-  \eta_2 = 0 :\left
-  \{\begin{matrix}
-    p_1 + p_2 = 1\\ 
-    p_1+64p_2 = 57
-  \end{matrix}
-  \right.
-\end{equation}
-
-Thus, when $p_1=0.1111$ and $p_2=0.8889$ the efficiency is minimal for the VLB method.
-For other combinations see Table \ref{tab:02}.
-Looking at this table, one can see two negative efficiencies, when ($p_1$,$p_2$) assume the values ​​(0,1) and (0.1,0.9).
-This correspond th cases when the method increases the memory requirements instead of decreasing it.
 
@@
  \end{tabular}
- \label{tab:02}
+ \label{tab:03}
 \end{table}
-
-So far, we have examined only two groups (hence two probabilities) of bit-length for the sake of simplicity.
-Before we generalize to probability distributions let's take a quick look at the efficiencies for more groups, with uniform probability:
-
-\begin{itemize}
-  \item 3 groups with bit-lengths 1, 32 and 64 bits, efficiency $\eta_2=0.3854$,
-  \item 5 groups with bit-lengths 1, 16, 32, 48 and 64 bits, efficiency $\eta_2=0.3875$,
-  \item 8 groups with bit-lengths 1, 8, 16, 24, 32, 40, 48, 56 and 64 bits, efficiency $\eta_2=0.3888$
-\end{itemize}
-
-When the distribution of the group probabilities is uniform, i.e., the groups have approximately the same size, efficiency is basically the same, regardless of the number of groups.
-
-Now we can leverage the notion of bit-length probabilities, and study efficiency when bit-lengths follow some  commonly used discrete probability distributions: Discrete Uniform, Binomial and Poisson.
-For all the experiments, we assume $k=7$, that is, the maximum possible bit-length is 64 bits.
-Thus, efficiency obtained will not be the best possible, since for that we would need assume small values of $k$ (Equation \ref{eq:29}).
-
-\subsection*{Discrete Uniform}
-Let the bit-lengths of the matrices be distribute according to the Uniform distribution $U(a=1,b=64)$, which means bit-lengths may take values in the set $\{1, 2, 3, $\ldots$, 64\}$ with equal probability, i.e., $\frac{1}{64}$.
-
-\paragraph{Theoretical Efficiency:}
-Let the random variable $B \sim U(a=1,b=64)$ represent the  bit-length of the elements of matrix $M$.
-Then $E(b_i) = \sum_{i} b_i \times p(b_i) = \frac{a+b}{2}$.
-Applying this result to the expected compression efficiency of VLB method (Equation \ref{eq:29}), we have
-
-\begin{equation}\label{eq:44}
- E(\eta_2) = 1 - \frac{E(B)}{64} - \frac{k}{64} 
-\end{equation}
-
-assuming all bit-lengths are possible, i.e., $a=1$ and $b=64$, and hence $k=7$, we can calculate $\eta_2$:
-
-\begin{equation}\label{eq:45}
- E(\eta_2) = 1 - \frac{\frac{1+64}{2}}{64} - \frac{7}{64} \approx 38.28\% 
-\end{equation}
-
-This result agrees with the numerical estimates presented in Table \ref{tab:03}.
-
-\paragraph{Numerical Estimates:}
-To calculate the VLB efficiency, we generated a matrices with 100 ($M_{10 \times 10}$), 10,000 ($M_{100 \times 100}$) and 1,000,000 ($M_{1,000 \times 1,000}$) elements with 1, 8, 16, 32 and 64 number of bitss.
-The average efficiency (Table \ref{tab:03}) is calculated from a 1,000 replicates of each matrix size.
-As expected the compression effiency gets better with lower expected bit-length.
 
@@
   \centering
-  \caption{Compression efficiency of VLB method of samples with bit-lengths coming from a Discrete Uniform distribution $U(a=1,b=64)$.
-Average efficiency ($\overline{\eta_2}\pm\textrm{SD}$) were calculated over a 1,000 replicates.}
+  \caption{Compression efficiency of VLB method of samples with bit-lengths 
+coming from a Discrete Uniform distribution $U(a=1,b=64)$. Average efficiency 
+($\overline{\eta_2}\pm\textrm{SD}$) were calculated over a 1,000 replicates.}
  \begin{tabular}{cccc}
@@
  \end{tabular}
- \label{tab:03}
+ \label{tab:04}
 \end{table}
-
-\subsection*{Binomial Distribution}
-
-For the binomial distribution, we will use $Bin(n, p)$, with the number of trials $n$ representing the greatest possible bit-length in the matrix, and $np$ giving us the expected bit-length.
-
-\paragraph{Theoretical Efficiency:}
-Let bit-length $(B)$ be a random variable with Binomial distribution, $B \sim B(n=64,p=0.5)$, $E(b_i) = \sum_{i} b_i \times p(b_i) = n \times p$ and the eficiency becomes (with $k=7$):
-
-\begin{align}\label{eq:47}
- E(\eta_2) &= 1 - \frac{64 \times p }{64} - \frac{k}{64}  \\
-  &= 1 - \frac{64 \times 0.5 }{64} - \frac{7}{64} \approx 39.05\% \nonumber
-\end{align}
-
-Which again agrees with estimates in Table \ref{tab:04}.
-
-\paragraph{Numerical Estimates:}
-For these experiments, the parameter n represents the maximum bit-length of matrix elements and takes values in $\{1, 8, 16, 32, 64\}$.
-In this case, we evaluate the efficiency as a function of the parameter $n$, and matrix size.
-Even though efficiency does not depend on matrix size, we tried different sizes to test the stability of the compression algorithm.
-Results are shown in Table \ref{tab:04}.
-As expected, smaller bit-lengths lead to higher compression efficiencies.
 
@@
   \centering
-  \caption{Compression efficiency with bit-lengths distributed according to a binomial distribution $B(n,0.5)$.
-Parameter $n \in \{1, 8, 16, 32, 64 \} $ represents the maximum bit-length.
-Since $p=0.5$ the expected bit-length is $n/2$ (first column)}
+  \caption{Compression efficiency with bit-lengths distributed according to a 
+binomial distribution $B(n,0.5)$. Parameter $n \in \{1, 8, 16, 32, 64 \} $ 
+represents the maximum bit-length. Since $p=0.5$ the expected bit-length is 
+$n/2$ (first column)}
    \begin{tabular}{cccc}
@@
   \end{tabular}
-  \label{tab:04}
+  \label{tab:05}
 \end{table}
-
-\subsection*{Poisson Distribution}
-With bit-length derived from a Poisson($\lambda$), the parameter $\lambda$ corresponds to the expected bit-length. For the purpose of this analysis this Poisson distribution is truncated at 64. 
-\paragraph{Theoretical Efficiency:}
-Let bit-length $B \sim Poisson(\lambda=32)$. In this case, $E(b_i) = \lambda$, with $k=7$,  the efficiency becomes:
-
-\begin{align}
- E(\eta_2) &= 1 - \frac{\lambda}{64} - \frac{k}{64} \label{eq:49} \\
- &= 1 - \frac{32}{64} - \frac{7}{64} \approx 39.06\% \label{eq:50}
-\end{align}
-
-This result is in accordance to Table \ref{tab:05}.
-\paragraph{Numerical Estimates:}
-The results for this simulation can be seen in \ref{tab:05}.
-Note that increasing the number of bits to represent numbers increases, there is a loss of efficiency in the compression process.
-In this case we did not simulate for $\lambda=64$ since a large portion of the samples would fall above the maximum bit-length we are considering for this analysis.
 
@@
   \centering
-   \caption{Compression efficiency with bit-lengths distributed according to a Poisson distribution ($\lambda$), where $\lambda$ represents the expected bit-length.}
+  \caption{Compression efficiency with bit-lengths distributed according to a 
+Poisson distribution ($\lambda$), where $\lambda$ represents the expected 
+bit-length.}
   \begin{tabular}{cccc}
@@
   \end{tabular}
-  \label{tab:05}
+  \label{tab:06}
 \end{table}
 
-These results show that a good compression is guaranteed when bit-lengths are distributed according to the tested distributions regardless of sample size.
-
-\section{Conclusion}
-In this paper, we have focused in the compression of matrix data, since this is
-one of the most important application the authors foresee. However, the
-compression methodology presented can be applied to any numerical data
-structure, with gains to performance and memory footprint\cite{teseCrysttian}.
-
-Further discussions about doing computation with such compressed data-structures
-will be the subject of another manuscript (in preparation) in which we will
-present details about the implementation of the compression algorithm, and
-benchmarks on classical linear algebra tasks such as those in
-Linpack\cite{dongarra1979linpack}.
-
-For the compression calculations presented in this paper we limited bit-lenght of integers to 64 bits.
-However the compression would work in the same way as discussed for computer architectures with larger word sizes.
-
-Representation of floating point numbers is also possible within the proposed compression framework, but at the expense of precision in their representation.
-Although this may sound like a limitation, when we take into consideration that most experimental data have fewer ``significant'' digits than the maximal precision available in modern computers, fairly good compression may still be achievable for floats.
-
-% You may title this section "Methods" or "Models". 
-% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
-% authors may use "Analysis"
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgments}
-We would like to thank Claudia Torres Code\c{c}o, Moacyr Silva and Paulo
-Cezar Carvalho for fruitful discussions and key ideas which helped improve the
-manuscript.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{plos.bib}
-
-\section*{Figure Legends}
-
-\begin{figure}[htp]
-  \includegraphics[scale=0.3,clip]{fig01.pdf}
-  \caption{Representation of the compression process and allocation of 
-elements of matrix M in the matrix m. The strip of value 700 is divided in two 
-parts and stores in different strips in the matrix m.}
-  \label{fig:01}
-\end{figure}
-
-%\begin{figure}[!ht]
-%\begin{center}
-%%\includegraphics[width=4in]{figure_name.2.eps}
-%\end{center}
-%\caption{
-%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
-%should be left justified, as specified by the options to the caption 
-%package.
-%}
-%\label{Figure_label}
-%\end{figure}
-
-%\section*{Tables}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
 \end{document}

### 2013-10-02 suppl.tex
Suplemento revisado...

@@
 \usepackage{cite}
+\usepackage{todonotes}
 
@@
 bitstring compressed arrays. We will consider integer matrices compressed via
-both the SM and the VLB methods.
+both the SM and the VLB methods, \todo{inseri esse texto} but the proposed 
+methods can be applied to integers and real numbers with sign.
 
@@
 more space to store and the operation in itself gets more complex, as in place
-operations are not possible. In the following examples we will explore these
+operations are not possible. In the following examples, we will explore these
 kinds of operations.
@@
 where $n_1=7$ and $n_2=10$. This addition is illustrated in table \ref{tab:01}
-in both decimal and binary. Note that when we add $1+1$, a carry over is
-generated to be added to bit on the left. In total we generate three bits in
+in both decimal and binary. Note that, when we add $1+1$, a carry over is
+generated to be added to bit on the left. In total, we generate three bits in
 carry over.
 
-In Table \ref{tab:02} we show the addition of $7 + 7$ which generates a 4 bits
+In Table \ref{tab:02}, we show the addition of $7 + 7$ which generates a 4 bits
 long number. In integer additions, the results may, in the very least, be as
@@
 \end{table}
-
+\todo{Traduzi o título que estava em português.}
 \begin{table}[ht]
     \centering
-    \caption{Comparacao da soma em base decimal e binaria.}
+    \caption{Comparison of the sum in binary and decimal base.}
     \begin{tabular}{crrrc}
@@
 \paragraph{Subtraction}
+
 Depending of the number involved in this operation, the result may be negative.
 This fact alone generates the need for an extra bit to represent the sign ($+$
-or
-$-$). The usual way to achieve this is to use the  two's complement
+or $-$). The usual way to achieve this is to use the  two's complement
 \cite{flores1963logic} representation which has the advantage of making the
 addition subtraction and multiplication operation the same as those for unsigned
-binary numbers. This algorithm is what is actually implemented in CPUs. To
-exemplify let's subtract $7$ from $10$. To make use of the
+binary numbers. This algorithm \todo{removi is what} is actually implemented in 
+CPUs. To exemplify let's subtract $7$ from $10$. To make use of the
 same algorithm of the sum , we represent the operation as $10+(-7)$. First we
@@
 In multiplication the behavior of the bit-length is different from the sum and
-subtraction. In the worst case the product has a bit-length which is the sum of
+subtraction. In the worst case, the product has a bit-length which is the sum of
 the bit-lengths of the factors. For example, in the product $7\times7$, the
@@
 So far we have examined the 4 fundamental arithmetic operations. In summary the
-implication for memory allocation of the results are the folowing, in the worst
+implication for memory allocation of the results are the following, in the worst
 case scenarios:
@@
 operand;
-\item Multiplication: Requires the sum of the bit-lenghts of the
+\item Multiplication: Requires the sum of the bit-lengths of the
 operands;
@@
 
-Figure \ref{fig:bitlength} shows the bit-lengths of results of the four
+\todo{adicionei esse trecho - As can be seen in} As can be seen in figure 
+\ref{fig:bitlength}, the bit-lengths of results of the four
 operations with integer operands up to 8 bits in length. The conclusions listed
@@
 The operations on single numbers (scalars) as exposed above are a simplification
-of the
-actual operations taking place on the bitstrings as we compute with compressed
-matrices. Additional details will be provided below, when we discuss the
-operation with matrices. Note that we are restricting the example to operations
-which generate integer results. Handling of float operations and compression
-will be the subject of a subsequent paper.
+of the actual operations taking place on the bitstrings as we compute with 
+compressed matrices. Additional details will be provided below, when we discuss 
+the operation with matrices. Note that we are restricting the example to 
+operations which generate integer results. Handling of float operations and 
+compression will be the subject of a subsequent paper.
 
@@
  \centering
-%  \includegraphics[width=12cm]{./bitlength.png}
- % simples.png: 4800x3600 pixel, 600dpi, 20.32x15.24 cm, bb=0 0 576 432
+ \includegraphics[width=12cm]{./bitlength.png}
  \caption{Bit-length of the results for each of the four operations. The axes
@@
 
-e
+and
 
@@
 Consider now just the bitstring stored in the first line of each matrix ($AA$
-and $BB$). To recover the original elements we need to now the bit-lengthof
+and $BB$). To recover the original elements we need to now the bit-length of
 the blocks of memory containing them, for the SM method they are all the same
@@
 elements. The mask for the first and third elements of matrix $AA$ is depicted
-in \ref{eq:06}. To recover the second and fouth elements, we apply the mask
+in \ref{eq:06}. To recover the second and fourth elements, we apply the mask
 shown on \ref{eq:07}. For Matrix $BB$, see \ref{eq:08} and \ref{eq:09} for the
@@
 Once defined the mask we can apply it to matrices $A$ and $B$. We use the mask
-by applying the boolean function AND. Let tempA and tempB the recovered
+by applying the boolean function AND. Let $tempA$ and $tempB$ the recovered
 elements. The recovering process is illustrated below. Note that only the
@@
 
-%\begin{equation}\label{eq:10}
   \begin{align*}
@@
   \end{align*}
-%\end{equation}
 
-%\begin{equation}\label{eq:11}
   \begin{align*}
@@
   \end{align*}
-%\end{equation}
 
@@
 Suppose we want to add both matrices and store it in a matrix $S$. for that the
-blocks of tempA and tempB must have the same length. As for the addition we
+blocks of $tempA$ and $tempB$ must have the same length. As for the addition we
 need, in the worst case, an extra bit, the block length of the resulting
@@
 
-Therefore, as the final result of the operation $S= A +
-B$ we have:
+Therefore, as the final result of the operation $S= A + B$, we have:
 
 \begin{align}
-	S &= \begin{bmatrix}
+	S &=\begin{bmatrix}
 	      549649\\ 
@@
       \\ \nonumber      
-      &= \begin{bmatrix}
+      &=\begin{bmatrix}
   \ldots\underbrace{000010}_{2}\underbrace{000110}_{6}\underbrace{001100}_{12}
@@
 \paragraph{Subtraction}
-To perform the subtraction, we need to use C2 notation. Supose  we need to
+To perform the subtraction, we need to use C2 notation. Suppose  we need to
 calculate $D = B - A$. This operation can be done in the same way, by
@@
         \\\nonumber
-        &= \begin{bmatrix}
+        &=\begin{bmatrix}
 \ldots\underbrace{111111}_{-1}\underbrace{111101}_{-3}\underbrace{111011}_{-5}
@@
 To exemplify, let's examine closely the operations $B[1,1] + (-A[1,1])$,
-$B[2,1] + (-A[2,1])$ and $B[3,1] + (-A[3,1])$. We use tempB to store each line
+$B[2,1] + (-A[2,1])$ and $B[3,1] + (-A[3,1])$. We use $tempB$ to store each line
 of the matrix B converted to C2. The numbers $1$ marked in the bitstring, are
@@
     tempB	&= \ldots010100|011110|000010|000001 \\
-+(-A[3,1])&= \ldots111101|111001|110110|110101\\ 
+    +(-A[3,1])	&= \ldots111101|111001|110110|110101\\ 
     D[3,1]	&= \ldots\sout{1}010001|\sout{1}110111|\sout{1}111000|110110
@@
 \begin{align}
-	D &= \begin{bmatrix}
+	D &=\begin{bmatrix}
 		129\\ 
@@
 	\\\nonumber
-        &=  \begin{bmatrix}
+	&=\begin{bmatrix}
 \ldots\underbrace{000000}_{0}\underbrace{000000}_{0}\underbrace{000010}_{2}
@@
 \paragraph{Multiplication}
+
 Let's now examine matrix multiplication. Consider the product $P = A \times
@@
 \begin{equation}\label{eq:prodbs}
-	P = AA \times BB = \begin{bmatrix}
+	P = AA \times BB =\begin{bmatrix}
 			    4952\\ 
@@
     	\begin{bmatrix}
-
 \ldots\underbrace{0001}_{1}\underbrace{0011}_{3}\underbrace{0101}_{5}\underbrace
 {1000}_{8}\\
-
 \ldots\underbrace{1100}_{12}\underbrace{1110}_{14}\underbrace{0110}_{6}
 \underbrace{1001}_{9}\\
-
 \ldots\underbrace{0011}_{3}\underbrace{0111}_{7}\underbrace{1010}_{10}
@@
 \end{equation}
+
 \begin{small}
@@
 \end{small}
+
 To calculate $P[1,1]$ which in uncompressed form is $1 \times 1 + 3 \times 3 + 5
@@
 and then proceed with the linear combination. The extraction will make use of
-masks as described before. Let tempA and tempB store the values of blocks 1 and
-3 of AA[1,1] and BB[1,1]. Now we need to determine the size of the blocks
+masks as described before. Let $tempA$ and $tempB$ store the values of blocks 1 
+and 3 of AA[1,1] and BB[1,1]. Now we need to determine the size of the blocks
 (bit-length) containing each element of the resulting matrix. As demonstrated
@@
 extra bits, thus we end up with a bit-length of 12 for each element of the
-product. Having determined the bit-length of the product, we can now we can do
-the actual calculations and store the results.
+product. Having determined the bit-length of the product, we can now \todo{we 
+can - removi} do the actual calculations and store the results.
 
@@
 
-Again we can see that P is already in compressed form, which confirms that the
+Again, we can see that P is already in compressed form, which confirms that the
 entire operation was conducted without decompressing the data.

### 2013-11-24 manuscript.tex
Atualizacao do artigo

@@
 
-\todo{Parei aqui}
-
-\section{Conclusion}
+\section*{Library Implementation }
+
+The methodology employed in this this paper includes the SM method, which was implemented in a library for Fortran. The 
+overloaded operations were assignment, addition, subtraction, multiplication, transpose, and maximum of a matrix. 
+Although only few significant operations of matrices presented ovearload, any other operation can be performed, since 
+the methods for inserting and collecting an element in a matrix of bitstring were implemented. With these two methods, 
+as long as they are properly adapted into algorithms, it becomes possible to implement any other desired operation.
+
+The method was also extended to allow for the allocation of real and integer numbers, besides positive integers, inside 
+the matrix bistring. Some adjustments were made so that the other groups of numbers could be represented. In order to 
+represent the integers with signs, it is necessary to use two's complement, as it is applied to thein accordance to 
+traditional methods.
+
+The SM method uses a range of bits to store the values in such a way that
+the size or number of bits used is important to represent the highest value in the bitstring matrix. The adaptation 
+consists in determining the largest value to be stored considering the absolute value of the number. That done,
+when the number is negative, it is necessary to rewrite it since, due to the sign, this will be written in the form of 
+two's complement. The purpose of this conversion is to optimize library processing. Take the following as an example: 
+the number 5, when represented in a 32-bit integer, is defined as:
+
+\begin{equation}\nonumber
+ 5 = 00000000 00000000 00000000 00000101
+\end{equation}
+
+However, if the number is negative, this is represented in the form of two's complement. Therefore, the number -5 is 
+represented as:
+
+\begin{equation}\nonumber
+ -5 = 11111111 11111111 11111111 11111011
+\end{equation}
+
+Note that the bit positioned more to the left should be used to inform that the number is -5. For the purpose of 
+optimization of the library, it was chosen to conversion of positive numbers to negative numbers and use of 1 
+bit to represent the sign. (Dúvidas aqui)
+
+Observe que os bits posicionados mais à esquerda devem ser utilizados para 
+informar que o número é o -5. Para efeito de otimização da biblioteca, foi 
+escolhido a conversão dos números negativos para positivos e a utilização de 1 
+bit para representar o sinal.
+
+In turn, in order to represent real numbers, a conversion procedure of real numbers to integer numbers was used. The 
+real number is multiplied by a power of ten and is then immediately truncated. The real numbers in a binary system of 64 
+bits are represented using a bit to represent the sign of the number, 11 bits to the exponent, and 52 bits to the 
+mantissa. Note that 1 bit is still used implicitly, according to the IEEE-754 standard. In summary, the mantissa is 
+represented in the base $2^{-1}$, $2^{-2}$, $2^{-3}$, $\hdots$, $2^{-52}$. Depending on the value to be represented, 
+loss of information can take place, which is related to the accuracy of the type used. In this case, considering a real 
+number with double precision, the demand of 64 bits to represent each one occurs, and in some cases, the number to be 
+represented is an approximation. As an example, consider that the number to be stored is 1.109. Following the 
+traditional system, the number 1,109 is represented by
+
+\begin{equation}\nonumber
+  0011111111110001101111100111011011001000101101000011100101011000
+\end{equation}
+
+We consider that the power used is $10^3$, that is, the number which needs to be stored would be 1109. In binary, that 
+number is represented by 10001010101. Note that the number of bits required is smaller, but depending on numerical 
+accuracy, the compression methods suggested are ineffective. When the power of 10 is used, it must be applied across the 
+matrix. This conversion allows the proposed methods to be applied to a set of real numbers. It should be taken into 
+account that the greater the power of 10 used, which is associated with the conversion of the real number, the larger 
+the set of bits will be required to represent the converted number. This affects the efficiency of data compression. The 
+use of the library to represent real numbers is conditioned to a study of the necessary level of precision to solve the 
+problem.
+
+\section*{Library Application}
+
+In order to create an application, some elementary matrix operations. Different operations have been performed by 
+measuring the execution time of each. We basically, considered the attribution of a constant to all the elements of a 
+matrix; the addition, subtraction and multiplication of the elements of a matrix by a constant, the addition, 
+subtraction and multiplication of matrices. In addition to that, the calculation of  the transposed matrix and the 
+maximum an array were also considered. This measure was used to evaluate the efficiency of the method implemented when 
+compared to traditional methods.
+
+One of the characteristics of bitstring matrices is that the operations per column, in the case of implementation in 
+Fortran, can be parallelized. This reduces the execution time, compared to traditional methods. Therefore, in order to 
+achieve this optimization, threads implementations were used for the implemented operations.
+
+As a test methodology, all the described operations were also compared to traditional methods. The matrix dimensions n x 
+n used were $n = 1,000$, $6,000$, $11,000$ to $20,000$. We emphasize that the first matrix,  created with a 1,000 
+dimension has $10^6$ elements, while the one with a 20,000 dimension has $4 \times 10^8$ elements. With the increase of 
+the size of the matrices, a larger amount of memory must be reserved to allocate the numbers. The compression effect of 
+the matrices can be verified in the results since by the proposed method the matrix is ​​directly loaded into the main 
+memory, avoiding the need for access to the secondary memory.
+
+Another library application was the implementation of the algorithm in order to deal with the problem of arrays of 
+collaborative filtering \cite{cf}. These arrays have large dimensions due to the number of users and products. Each 
+element of this matrix stores the rating of a user with respect to a specific product. With this information, it is 
+possible to specify a product for a determined customer.
+
+The manipulation of these matrices, depending on the amount of memory to be allocated, are one of the possible 
+applications of the methods proposed in this work, since it promotes the compression and data allocation in the main 
+memory. The allocation of the matrix within the main memory makes the access time to the elements much smaller than when 
+the matrix is stored in secondary memory. This latter type of storage is used in some already mentioned techniques. As 
+an example, let us examine the manipulation of a CF matrix, evaluating the access time to the elements and especially 
+the size of the memory required to run such operation. The algorithms applied to this matrix were the average per-use 
+and the bias from mean \cite{cf}, this last one of the simplest algorithms for predicting the rate of users.
+
+An example of the application of the methodology presented is shown in Figure \ref{cf}. In this figure, the matrix with 
+normal and bitstring formats are displayed. In both matrices, the information is represented and can be accessed. 
+Therefore, it becomes possible to apply the selected algorithms.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.4,trim=5cm 0cm 5cm 0cm,clip]{cf}
+  \caption{Illustration of the process of the bistring model application to matrices of collaborative filttering of the 
+algorithms per use average (PUA) and of the bias from mean (BFM) in a matrix represented in the traditional and 
+bitstring formats.}
+  \label{cf}
+\end{figure}
+
+Falar dos dados do GraphLab, os usados foram o do netflix, e medir o tempo 
+médio da executacao por umas 10 vezes.
+
+Os dados utilizados nos testes realizados estão disponíveis em XXX. Esses dados 
+correspondem a ....
+
+\section*{Conclusion}
+
 In this paper, we have focused in the compression of matrix data, since this is
@@
 
-For the compression calculations presented in this paper we limited bit-lenght of integers to 64 bits.
-However the compression would work in the same way as discussed for computer architectures with larger word sizes.
-
-Representation of floating point numbers is also possible within the proposed compression framework, but at the expense of precision in their representation.
-Although this may sound like a limitation, when we take into consideration that most experimental data have fewer ``significant'' digits than the maximal precision available in modern computers, fairly good compression may still be achievable for floats.
+For the compression calculations presented in this paper we limited bit-lenght of integers to 64 bits. However the 
+compression would work in the same way as discussed for computer architectures with larger word sizes.
+
+Representation of floating point numbers is also possible within the proposed compression framework, but at the expense 
+of precision in their representation. Although this may sound like a limitation, when we take into consideration that 
+most experimental data have fewer ``significant'' digits than the maximal precision available in modern computers, 
+fairly good compression may still be achievable for floats.
 

### 2013-11-24 plos.tex
Atualizacao do artigo

--- 2013-09-25 plos.tex
@@
 \\
-Author1$^{1}$, 
-Author2$^{2}$, 
-Author3$^{3,\ast}$
+Crysttian Arantes Paix\~{a}o$^{1\ast}$, 
+Fl\'{a}vio Code\c{c}o Coelho$^{2}$, 
 \\
-\bf{1} Author1 Dept/Program/Center, Institution Name, City, State, Country
+\bf{1} Crysttian Arantes Paix\~{a}o Applied Mathematics School, Getulio Vargas Foundation, Rio de Janeiro, RJ, Brazil
 \\
-\bf{2} Author2 Dept/Program/Center, Institution Name, City, State, Country
+\bf{2} Fl\'{a}vio Code\c{c}o Coelho Applied Mathematics School, Getulio Vargas Foundation, Rio de Janeiro, RJ, Brazil
 \\
-\bf{3} Author3 Dept/Program/Center, Institution Name, City, State, Country
-\\
-$\ast$ E-mail: Corresponding author@institute.edu
+$\ast$ E-mail: Corresponding crysttian.paixao@fgv.br
 \end{flushleft}

### 2013-11-29 manuscript.tex
atualizacao

@@
 with the traditional representation of matrices.
+
 % Please keep the Author Summary between 150 and 200 words
@@
 As a test methodology, all the described operations were also compared to traditional methods. The matrix dimensions n x 
-n used were $n = 1,000$, $6,000$, $11,000$ to $20,000$. We emphasize that the first matrix,  created with a 1,000 
-dimension has $10^6$ elements, while the one with a 20,000 dimension has $4 \times 10^8$ elements. With the increase of 
-the size of the matrices, a larger amount of memory must be reserved to allocate the numbers. The compression effect of 
-the matrices can be verified in the results since by the proposed method the matrix is ​​directly loaded into the main 
-memory, avoiding the need for access to the secondary memory.
+n used were $n = 10$, $100$, $1,000$, $10,000$, $20,000$ to $100,000$, by $10,000$. We emphasize that the matrix,  
+created with a 1,000 dimension has $10^6$ elements, while the one with a 100,000 dimension has $1 \times 10^{10}$ 
+elements. With the increase of the size of the matrices, a larger amount of memory must be reserved to allocate the 
+numbers. The compression effect of the matrices can be verified in the results since by the proposed method the matrix 
+is ​​directly loaded into the main memory, avoiding the need for access to the secondary memory.
+
+Inserir os resultados
 
@@
 
-Falar dos dados do GraphLab, os usados foram o do netflix, e medir o tempo 
-médio da executacao por umas 10 vezes.
-
-Os dados utilizados nos testes realizados estão disponíveis em XXX. Esses dados 
-correspondem a ....
+Inserir o segundo conjunto de resultados aqui.
 

### 2013-12-03 manuscript.tex
Adicionando as últimas figuras

@@
 the groups, $\frac{f_i}{rc}$, while sticking to two groups. Let's 
-also\todo{Acho que tinha que tirar esse Let's also, talvez colocar Now, we are 
+also \todo{Acho que tinha que tirar esse Let's also, talvez colocar Now, we are 
 going to} assume that $\frac{f_i}{rc}$ is a good approximation to the 
@@
 bit to represent the sign. (Dúvidas aqui)
-
-Observe que os bits posicionados mais à esquerda devem ser utilizados para 
-informar que o número é o -5. Para efeito de otimização da biblioteca, foi 
-escolhido a conversão dos números negativos para positivos e a utilização de 1 
-bit para representar o sinal.
 
@@
 
-Inserir os resultados
+XXXXXXXXX
 
@@
 
-An example of the application of the methodology presented is shown in Figure \ref{cf}. In this figure, the matrix with 
-normal and bitstring formats are displayed. In both matrices, the information is represented and can be accessed. 
+An example of the application of the methodology presented is shown in Figure \ref{fig22}. In this figure, the matrix 
+with normal and bitstring formats are displayed. In both matrices, the information is represented and can be accessed. 
 Therefore, it becomes possible to apply the selected algorithms.
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.4,trim=5cm 0cm 5cm 0cm,clip]{cf}
-  \caption{Illustration of the process of the bistring model application to matrices of collaborative filttering of the 
-algorithms per use average (PUA) and of the bias from mean (BFM) in a matrix represented in the traditional and 
-bitstring formats.}
-  \label{cf}
-\end{figure}
-
-Inserir o segundo conjunto de resultados aqui.
+yyyyyyyyyyyyyy
 
@@
   \label{fig:2021}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.4,trim=5cm 0cm 5cm 0cm,clip]{fig22}
+  \caption{Illustration of the process of the bistring model application to matrices of collaborative filttering of the 
+algorithms per use average (PUA) and of the bias from mean (BFM) in a matrix represented in the traditional and 
+bitstring formats.}
+  \label{fig22}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig23}
+  \caption{Tempo de processamento do método Per User Average em função do número de linhas de uma matriz com 3,000 
+colunas. A linha verde na vetical indica que o momento aproximado que o método tradicional demanda acesso a disco para 
+armazenar a matriz a ser analisada, onerando o processo. O método de bitstring consegue representar a matriz 
+integralmente na memória, sem a necessiada de acesso a disco.}
+  \label{fig23}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig24}
+  \caption{Tempo de processamento do método Per User Average em função do número de linhas de uma matriz com 3,000 
+colunas. A linha verde na vetical indica que o momento aproximado que o método tradicional demanda acesso a disco para 
+armazenar a matriz a ser analisada, onerando o processo. Para dimensões inferiores a aproximadamente 400 liinhas, o 
+método tracional é mais rápido que o de bitstring, ressaltando que o custo computacional dos dois métodos é linear.}
+  \label{fig24}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig25}
+  \caption{Relação entre o tempo de processamento pelo método de bisttring e o método tracional em função do número de 
+linhas de uma matriz com 3,000 colunas. Verifica-se que para a matriz com um número inferior a aproximadamente 400 
+linhas, o método tradicional é mais rápido, porém quando o número de linhas é superior, o método de bitstring é mais 
+rápido, por causa da representação compactada em  memória da matriz.}
+  \label{fig25}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig26}
+  \caption{xxx.}
+  \label{fig23}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig27}
+  \caption{xxx.}
+  \label{fig24}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig28}
+  \caption{xxx.}
+  \label{fig25}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$A=5$]{
+  \includegraphics[scale=0.4,clip]{fig29}
+  }
+  \subfigure[$A=A+5$]{
+  \includegraphics[scale=0.4,clip]{fig30}
+  }
+  \subfigure[$A=A+A$]{
+  \includegraphics[scale=0.4,clip]{fig31}
+  }
+  \subfigure[$A=A-8$]{
+  \includegraphics[scale=0.4,clip]{fig32}
+  }
+  \caption{Z}
+  \label{fig:29303132}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$A=5$]{
+  \includegraphics[scale=0.4,clip]{fig33}
+  }
+  \subfigure[$A=A+5$]{
+  \includegraphics[scale=0.4,clip]{fig34}
+  }
+  \subfigure[$A=A+A$]{
+  \includegraphics[scale=0.4,clip]{fig35}
+  }
+  \subfigure[$A=A-8$]{
+  \includegraphics[scale=0.4,clip]{fig36}
+  }
+  \caption{Z}
+  \label{fig:33343536}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$A=5$]{
+   \includegraphics[scale=0.4,clip]{fig37}
+  }
+  \subfigure[$A=A+5$]{
+   \includegraphics[scale=0.4,clip]{fig38}
+  }
+  \subfigure[$A=A+A$]{
+   \includegraphics[scale=0.4,clip]{fig39}
+  }
+  \subfigure[$A=A-8$]{
+   \includegraphics[scale=0.4,clip]{fig40}
+  }
+  \caption{R}
+  \label{fig:37383940}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$A=5$]{
+   \includegraphics[scale=0.4,clip]{fig41}
+  }
+  \subfigure[$A=A+5$]{
+   \includegraphics[scale=0.4,clip]{fig42}
+  }
+  \subfigure[$A=A+A$]{
+   \includegraphics[scale=0.4,clip]{fig43}
+  }
+  \subfigure[$A=A-8$]{
+   \includegraphics[scale=0.4,clip]{fig44}
+  }
+  \caption{R}
+  \label{fig:41424344}
 \end{figure}
@@
 
+\newpage
+
 \section*{Tables}

### 2014-01-27 manuscript.tex
Figuras ordenadas corretamente.
Atualizado o texto...

--- 2013-11-29 manuscript.tex
@@
 
-Inserir os resultados
+In Figures \ref{fig:23242526} and \ref{fig:27282930} the results from the processing time (in milliseconds) 
+for the operations that involved only integer ($Z$) and positive integer numbers ($Z^+$) are represented. 
+The traditional method (Normal) and the Bitstring method (Bitstring $Z$, integer numbers and Bitstring Zp, 
+positive integer numbers) were used. The operations were attribution ($A=5$, A is a matrix with dimension n by n); 
+sum of the elements of a matrix with a constant ($A=A+5$); sum of matrices ($A=A+A$); andsubtraction of the elements 
+of a matrix with a constant ($A=A-8$). These operations are represented in Figure \ref{fig:23242526}. 
+Figure \ref{fig:27282930}, in turn, shows the results for the multiplicationoperations of matrices ($A = A \times A$); 
+for the multiplication of the the elements of a matrix by a constant ($A=A \times 2$); 
+and for the calculation of transpose and maximum of a matrix.
+
+The results show that the bitstring method is faster than the traditional ones only for the attribution.
+This occurs because the element of the matrix is used more than once to store the numbers, differently 
+from the traditional method. Due to the number of manipulations  the bitstring method does, the other 
+operations  demand more time to be processed, but this can be optimized. When applied to positive integers, 
+the bitstring model works faster than when applied to integers because it does not need to do operations 
+related to the signals.
+
+The results for the real numbers are represented in Figures \ref{fig:31323334} and \ref{fig:35363738}. 
+By repeating the operations tested in the integer numbers, it was possible to see that the bitstring method 
+cannot overcome the traditional method. This happensbecause its application to  the real numbers, as alternative 
+to maintain the compression, asks for a conversion of real to integer numbers in order for them to be stored and 
+when accessed, another conversion is made. This is a problem from the implementation process, but it can be optimized.
 
@@
 
-An example of the application of the methodology presented is shown in Figure \ref{cf}. In this figure, the matrix with 
-normal and bitstring formats are displayed. In both matrices, the information is represented and can be accessed. 
+An example of the application of the methodology presented is shown in Figure \ref{fig22}. In this figure, the matrix 
+with normal and bitstring formats are displayed. In both matrices, the information is represented and can be accessed. 
 Therefore, it becomes possible to apply the selected algorithms.
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.4,trim=5cm 0cm 5cm 0cm,clip]{cf}
-  \caption{Illustration of the process of the bistring model application to matrices of collaborative filttering of the 
-algorithms per use average (PUA) and of the bias from mean (BFM) in a matrix represented in the traditional and 
-bitstring formats.}
-  \label{cf}
-\end{figure}
-
-Inserir o segundo conjunto de resultados aqui.
+In order to test a possible model application, we considered a matrix with the following dimensions: 95,000 users 
+(lines) by 3,000 movies (columns), with the ratings for Netflix’s movies
+\footnote{http://graphlab.org/downloads/datasets/}. In order to evaluate a larger dataset, we modified the 
+NetFlix data so we could increase the dimension of the matrix to be analyzed. The 95,000 users correspond to 
+thelines of the matrix. In orderto gradually increase the number of lines, we performed consecutive re-samplings 
+and added them to the original matrix . As a results, we obtained a  matrix with 600,000 lines and 3,000 columns. 
+The operations used with the matrix were the Per User Average (PUA) and the rating estimation of a movie i using 
+the Bias From Mean (BFM) \cite{cf}. The analyses were  made, considering the created matrix with different dimensions, 
+with n from 40,000 to 600,000 by 20,000. It is important to emphasize that the number of columns is constant and 
+equal to 3,000. The implementations of this problem were made in Fortran and both the developedlibrary based in 
+the method SM and in the traditional form were used. Codes were inserted in the library in order to optimize the 
+multicore processing using threads. No additional configuration was necessary to use the multicore processing. 
+since the library detects the characteristics of the processor uses them in the simulations. In turn, for the 
+traditional implementation, weused the optimization process and parallelization, which are made available by 
+the processor,as the vectorization and multicore parallelization.
+
+The results are represented in Figures \ref{fig39}, \ref{fig40} and \ref{fig41} for the PUA Method and in Figures 26, 27 and 28 for the 
+BFM Methods.  A computer with 8 GB of RAM Memory, 16 GB of swap, processor core i7 model 870\todo{confirmar modelo}
+and operational 
+system Linux Debian 7.0 was used to perform this simulation. In Figure \ref{fig39}, it is possible to verify the approximate 
+moment  when the  the storage in disk (swaap) begins to be used. This is indicated by the vertical green line. 
+At this moment, the bitstring method becomes faster than the traditional method because it does not require disk 
+storage due tocompressed memory representation. Note that, in the Figure \ref{fig40}, the traditional method is faster 
+than the bitstring one before the beginning of the disk access. However, after it becomes necessary use the disk 
+so that the matrix can be accessed and stored, , the traditional method  becames slower (Figura \ref{fig41}). 
+These results 
+indicate that the bistring method can promote the optimization of the memory use because the access to the main 
+memories (RAM and cache) are faster than the access to secondary memories (disk). In Figures \ref{fig42}, 
+\ref{fig43} and \ref{fig44} the 
+results of the processing time for  the BFM method are represented. Once again, a vertical green line indicates 
+the approximate moment when the disk access begins. There is a difference in relation to the PUA method since the 
+BMF method uses the PUA values to perform its calculation. Therefore, the BMF method demands more memory to be 
+processed, which reduces the matrix dimensions involved in the process, represented in the main memory. As 
+indicated in Figure \ref{fig41}, for the PUA method, and Figure \ref{fig44}, for the BMF method, it is possible 
+to verify that 
+before the access to the disk, the traditional method is faster than the bitstring one, but once the access 
+is initiated,  the bitstring method becames faster.
+
+In spite of the simplicity of the application, it is relevant since  it allows for large-sized matrices to be 
+allocated in memory and manipulated.
 
@@
   \label{fig:2021}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.4,trim=5cm 0cm 5cm 0cm,clip]{fig22}
+  \caption{Illustration of the process of the bistring model application to matrices of collaborative filttering of the 
+algorithms per use average (PUA) and of the bias from mean (BFM) in a matrix represented in the traditional and 
+bitstring formats.}
+  \label{fig22}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$A=5$]{
+  \includegraphics[scale=0.4,clip]{fig23}
+  }
+  \subfigure[$A=A+5$]{
+  \includegraphics[scale=0.4,clip]{fig24}
+  }
+  \subfigure[$A=A+A$]{
+  \includegraphics[scale=0.4,clip]{fig25}
+  }
+  \subfigure[$A=A-8$]{
+  \includegraphics[scale=0.4,clip]{fig26}
+  }
+  \caption{Processing time of the following operations: attribution ($A=5$, A is the matrix with an n by n dimension); 
+  element sum of a matrix with a constant ($A=A+5$);sum of matrices ($A=A+A$); and subtraction of the elements of a 
+  matrix by a constant ($A=A-8$) involving integer numbers ($Z$) and positive integer numbers ($Z^+$). The bitstring 
+  method presented the best overall performance only for the attribution operations, because the storage of 
+  different numbers occurs in one element of a matrix, which reduces the accessing time. For the other operations, 
+  the bitstring method is slower due to the number ofoperations which are necessary to be  executed in order to 
+  access the elements. This can still be optimized in the implemented library.}
+  \label{fig:23242526}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$A = A \times A$]{
+  \includegraphics[scale=0.4,clip]{fig27}
+  }
+  \subfigure[$A=A \times 2$]{
+  \includegraphics[scale=0.4,clip]{fig28}
+  }
+  \subfigure[$transpose(A)$]{
+  \includegraphics[scale=0.4,clip]{fig29}
+  }
+  \subfigure[$maximum(A)$]{
+  \includegraphics[scale=0.4,clip]{fig30}
+  }
+  \caption{Processing time of matrix multiplication ($A = A \times A$, A is the matrix with a n by n dimension), 
+  element multiplication of matrix by constant ($A=A \times 2$), the calculation of transpose and maximum of a 
+  matrix involve integer numbers ($Z$) and positive integer numbers ($Z^+$). In this cases, the traditional 
+  method was faster than the bitstring model.}
+  \label{fig:27282930}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$A=5.0000$]{
+   \includegraphics[scale=0.4,clip]{fig31}
+  }
+  \subfigure[$A=A+5.0000$]{
+   \includegraphics[scale=0.4,clip]{fig32}
+  }
+  \subfigure[$A=A+A$]{
+   \includegraphics[scale=0.4,clip]{fig33}
+  }
+  \subfigure[$A=A-8.0000$]{
+   \includegraphics[scale=0.4,clip]{fig34}
+  }
+  \caption{Processing time of the following operations: attribution ($A=5.0000$, A is the matrix with an n by n dimension); 
+  element sum of a matrix with a constant ($A=A+5.000$);sum of matrices ($A=A+A$); and subtraction of the elements of a 
+  matrix by a constant ($A=A-8.0000$) involving real numbers ($R$) and positive integer numbers ($R^+$). The bitstring 
+  method presented the best overall performance only for the attribution operations, because the storage of 
+  different numbers occurs in one element of a matrix, which reduces the accessing time. For the other operations, 
+  the bitstring method is slower due to the number ofoperations which are necessary to be  executed in order to 
+  access the elements. This can still be optimized in the implemented library.}
+  \label{fig:31323334}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$A = A \times A$]{
+   \includegraphics[scale=0.4,clip]{fig35}
+  }
+  \subfigure[$A=A \times 2$]{
+   \includegraphics[scale=0.4,clip]{fig36}
+  }
+  \subfigure[$transpose(A)$]{
+   \includegraphics[scale=0.4,clip]{fig37}
+  }
+  \subfigure[$maximum(A)$]{
+   \includegraphics[scale=0.4,clip]{fig38}
+  }
+  \caption{Processing time of matrix multiplication ($A = A \times A$, A is the matrix with a n by n dimension), 
+  element multiplication of matrix by constant ($A=A \times 2.0000$), the calculation of transpose and maximum of a 
+  matrix involve integer numbers ($R$) and positive integer numbers ($R^+$). In this cases, the traditional 
+  method was faster than the bitstring model.}
+  \label{fig:35363738}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig39}
+  \caption{Processing time for the Per User Average method (PUA) in function of the numbers of linesof a matrix 
+  with 3,000 columns. The vertical green line indicates the approximate moment when the traditional method  
+  demands  disk access in order to store the matrix to be analyzed, which  overloads the process. The bitstring 
+  method fully represents a matrix  in memory, without the necessity to access the  disk.}
+  \label{fig39}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig40}
+  \caption{Processing time for the Per User Average method (PUA) in function of the  numbers of lines of a matrix 
+  with 3,000 columns.  The vertical green line indicates the approximate moment when the traditional method  
+  demands disk access in order to store the matrixto be analyzed, which  overloads the process. For dimensions 
+  inferior to approximately 400 lines, the traditional method is faster than the bitstring one. It is important 
+  to emphasize that the computational cost of both methods is linear.}
+  \label{fig40}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig41}
+  \caption{Relation between the processing time using the bitstring method and the traditional method in function 
+  of the line numbers of a matrix with 3,000 columns. It is possible to verify that for a matrix with less than 400 
+  lines, the traditional method is faster. However, when the number of lines is greater  than 400 lines, the best 
+  method is the bitstring because of the compressed representation in the matrix memory.}
+  \label{fig41}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig42}
+  \caption{Processing time for the Bias From Mean method (BMF) in function of the numbers of linesof a matrix 
+  with 3,000 columns. The vertical green line indicates the approximate moment when the traditional method  
+  demands  disk access in order to store the matrix to be analyzed, which  overloads the process. The bitstring 
+  method fully represents a matrix  in memory, without the necessity to access the  disk.}
+  \label{fig42}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig43}
+  \caption{Processing time for the Bias From Mean method (BMF) in function of the  numbers of lines of a matrix 
+  with 3,000 columns.  The vertical green line indicates the approximate moment when the traditional method  
+  demands disk access in order to store the matrixto be analyzed, which  overloads the process. For dimensions 
+  inferior to approximately 400 lines, the traditional method is faster than the bitstring one. It is important 
+  to emphasize that the computational cost of both methods is linear.}
+  \label{fig43}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig44}
+  \caption{Relation between the processing time using the bitstring method and the traditional method in function 
+  of the line numbers of a matrix with 3,000 columns. It is possible to verify that for a matrix with less than 400 
+  lines, the traditional method is faster. However, when the number of lines is greater  than 400 lines, the best 
+  method is the bitstring because of the compressed representation in the matrix memory.}
+  \label{fig44}
 \end{figure}
@@
 
+\newpage
+
+\clearpage
+
 \section*{Tables}

### 2014-01-27 manuscript.tex
atualizacao referencias figuras

@@
 
-The results are represented in Figures \ref{fig39}, \ref{fig40} and \ref{fig41} for the PUA Method and in Figures 26, 27 and 28 for the 
-BFM Methods.  A computer with 8 GB of RAM Memory, 16 GB of swap, processor core i7 model 870\todo{confirmar modelo}
+The results are represented in Figures \ref{fig39}, \ref{fig40} and \ref{fig41} for the PUA Method and in Figures 
+\ref{fig42}, \ref{fig43} and \ref{fig44} for the 
+BFM Methods.  A computer with 8 GB of RAM Memory, 16 GB of swap, processor core i7 model 870  \todo{confirmar modelo}
 and operational 
@@
 Cezar Carvalho for fruitful discussions and key ideas which helped improve the
-manuscript.
+manuscript.\todo{Adicionar a Carol}
 

### 2014-02-07 manuscript.tex
fiz a revisão do texto removendo todos os "\todo"s e acrescentando outros

@@
 efficiency under different applications. We also discuss the efficiency of the 
-compression as a function of the distribution of the elements of the matrix. 
-\todo{Inseri isso} Also presented are some results in relation to the 
-traditional method of matrix allocation. We also consider the 
-arithmetic operations to measure the efficiency of the proposed methodology.
+compression as a function of the distribution of the elements of the 
+matrix, some results comparing our method with the traditional method of matrix 
+allocation, and arithmetic operations as a benchmark measure the efficiency of 
+operating with bitstring-compressed matrices.
 
@@
 inefficiencies in the conventional way matrices are allocated in memory. 
-\todo{adicionei aqui} The analyzes presented in this article, in the first 
-moment, are related to arrays of positive integers, but can be applied to 
-integers and real numbers with some adaptations, which are presented in due 
-course. \todo{removi isso The examples in this paper will be restricted to 
-matrices with positive integer elements, but}
+ The analyses presented in this article concern arrays of positive integers, 
+but can be applied to signed integers and real numbers with few adaptations.
 
@@
 way to store arrays of integers is on a memory block sequence of fixed size 
-(power of 2 numbers of bit), one for each element. The maximum size of a block 
+(power of 2 number of bits), one for each element. The maximum size of a block 
 is equal to the word size of the processor, which for most current CPUs is 64 
@@
 \begin{equation} \label{eq:01}
- b(\sup M) \approx \begin{cases}
+ b(\sup M) = \begin{cases}
 	1, &  \text{ if } \sup M \in \{0,1\} \\ 
@@
 80 bits using the SM method. However, as shall be seen later, the VLB method 
-will be the most efficient for most matrices. \todo{adicionei} With the method 
-VLB the resulting matrix is smaller than the original matrix.
+will be the most efficient for most matrices. With the VLB 
+method the resulting matrix is smaller than the original matrix.
 
@@
 exclusively of 0s and 1s, maximal compression is achievable, 
-$\eta_1=0.9843$.\todo{Alterei a proporção para 0.98, pois é o maior valor que 
-pode ser obtido.}
+$\eta_1=0.9843$.
 
@@
 ​​such that the value of $k = \frac{\sum_{i=1}^{g}  b_i \times f_i}{rc} - 64$.
-\todo{Adicionei esse trecho para ficar igual ao SM}
 
@@
 regardless of bit-length (Figures \ref{fig20} and \ref{fig21}). There are only 
-exception, if for bit-length 64, where neither method is able to compress the 
-data. \todo{Alterei a última frase}
+one exception: for bit-length 64, where neither method is able to compress 
+the data.
 
@@
 distribution of bit-lengths of matrix elements. When matrix elements are small 
-or have nearly-constant bit-length, the SM Method is better\todo{alterei de 
-best para better}, otherwise, the VLB method should be chosen.
+or have nearly-constant bit-length, the SM method is better, otherwise, the VLB 
+method should be chosen.
 
@@
 comprised of half the matrix elements. Let's now vary the relative frequency of 
-the groups, $\frac{f_i}{rc}$, while sticking to two groups. Let's 
-also \todo{Acho que tinha que tirar esse Let's also, talvez colocar Now, we are 
-going to} assume that $\frac{f_i}{rc}$ is a good approximation to the 
+the groups, $\frac{f_i}{rc}$, while sticking to two groups. Here, we assume 
+that $\frac{f_i}{rc}$ is a good approximation to the 
 probability of a given bit-length in a matrix, which we will denote by $p_i$.
@@
 \paragraph{Numerical Estimates:}
-The results for this simulation can be seen in \ref{tab:06}. Note that 
+The results for this simulation can be seen in table \ref{tab:06}. Note that 
 increasing the number of bits to represent numbers increases, there is a loss of 
@@
 $\lambda=64$ since a large portion of the samples would fall above the maximum 
-bit-length, but \todo{coloquei ", but", aqui} we are considering for this 
-analysis.
+bit-length.
 
@@
 
-\section*{Library Implementation }
-
-The methodology employed in this this paper includes the SM method, which was implemented in a library for Fortran. The 
-overloaded operations were assignment, addition, subtraction, multiplication, transpose, and maximum of a matrix. 
-Although only few significant operations of matrices presented ovearload, any other operation can be performed, since 
-the methods for inserting and collecting an element in a matrix of bitstring were implemented. With these two methods, 
-as long as they are properly adapted into algorithms, it becomes possible to implement any other desired operation.
-
-The method was also extended to allow for the allocation of real and integer numbers, besides positive integers, inside 
-the matrix bistring. Some adjustments were made so that the other groups of numbers could be represented. In order to 
-represent the integers with signs, it is necessary to use two's complement, as it is applied to thein accordance to 
-traditional methods.
-
-The SM method uses a range of bits to store the values in such a way that
+\section*{Implementation }\todo{acho que esta seção toda deve ir para o 
+material suplementar.}
+
+This paper proposes the bitstring-compression methodologies. The SM 
+method was implemented as a Fortran library which is available under an 
+open-source license\todo{colocar o link para o repositório aqui}. The 
+implementation extends the standard Matrix type of Fortran, overloading 
+operators such as assignment, addition, subtraction, multiplication, transpose, 
+and maximum of a matrix. 
+Although only few significant matrix operations have been ovearloaded in this 
+example implementation, any other operation can be performed, since 
+the methods for inserting and collecting an element in a matrix of bitstring 
+were also implemented. With these two methods, as long as they are properly 
+adapted into algorithms, it is possible to implement any other desired 
+operation.
+
+The SM method was also extended to allow for the allocation of real and 
+signed integer matrices, besides positive integer ones. Some adjustments were 
+made so that the other types of numbers could be represented. In order to 
+represent the signed integers, it was necessary to use two's complement, as it 
+is applied to thein accordance to traditional methods\todo{acho que esta 
+explicação ficou muito solta. Não precisamos explica a implementação no 
+artigo, podemos colocar na documentação do código}.
+
+The SM method uses a fixed number of bits to store the values in such a way that
 the size or number of bits used is important to represent the highest value in the bitstring matrix. The adaptation 
@@
 
-\section*{Library Application}
+\section*{Library Application}\todo{Também deve ir para o material suplementar}
 

### 2014-02-12 manuscript.tex
Alterações realizadas

@@
 	\end{cases}
-\end{equation}\todo{Será que não devemos trocar $\approx$ por = aqui? Note que 
-o número de bits não é aproximado e sim o exato.}
+\end{equation}
 
@@
 
-\section*{Implementation }\todo{acho que esta seção toda deve ir para o 
-material suplementar.}
-
-This paper proposes the bitstring-compression methodologies. The SM 
-method was implemented as a Fortran library which is available under an 
-open-source license\todo{colocar o link para o repositório aqui}. The 
-implementation extends the standard Matrix type of Fortran, overloading 
-operators such as assignment, addition, subtraction, multiplication, transpose, 
-and maximum of a matrix. 
-Although only few significant matrix operations have been ovearloaded in this 
-example implementation, any other operation can be performed, since 
-the methods for inserting and collecting an element in a matrix of bitstring 
-were also implemented. With these two methods, as long as they are properly 
-adapted into algorithms, it is possible to implement any other desired 
-operation.
-
-The SM method was also extended to allow for the allocation of real and 
-signed integer matrices, besides positive integer ones. Some adjustments were 
-made so that the other types of numbers could be represented. In order to 
-represent the signed integers, it was necessary to use two's complement, as it 
-is applied to thein accordance to traditional methods\todo{acho que esta 
-explicação ficou muito solta. Não precisamos explica a implementação no 
-artigo, podemos colocar na documentação do código}.
-
-The SM method uses a fixed number of bits to store the values in such a way that
-the size or number of bits used is important to represent the highest value in the bitstring matrix. The adaptation 
-consists in determining the largest value to be stored considering the absolute value of the number. That done,
-when the number is negative, it is necessary to rewrite it since, due to the sign, this will be written in the form of 
-two's complement. The purpose of this conversion is to optimize library processing. Take the following as an example: 
-the number 5, when represented in a 32-bit integer, is defined as:
-
-\begin{equation}\nonumber
- 5 = 00000000 00000000 00000000 00000101
-\end{equation}
-
-However, if the number is negative, this is represented in the form of two's complement. Therefore, the number -5 is 
-represented as:
-
-\begin{equation}\nonumber
- -5 = 11111111 11111111 11111111 11111011
-\end{equation}
-
-Note that the bit positioned more to the left should be used to inform that the number is -5. For the purpose of 
-optimization of the library, it was chosen to conversion of positive numbers to negative numbers and use of 1 
-bit to represent the sign. (Dúvidas aqui)
-
-In turn, in order to represent real numbers, a conversion procedure of real numbers to integer numbers was used. The 
-real number is multiplied by a power of ten and is then immediately truncated. The real numbers in a binary system of 64 
-bits are represented using a bit to represent the sign of the number, 11 bits to the exponent, and 52 bits to the 
-mantissa. Note that 1 bit is still used implicitly, according to the IEEE-754 standard. In summary, the mantissa is 
-represented in the base $2^{-1}$, $2^{-2}$, $2^{-3}$, $\hdots$, $2^{-52}$. Depending on the value to be represented, 
-loss of information can take place, which is related to the accuracy of the type used. In this case, considering a real 
-number with double precision, the demand of 64 bits to represent each one occurs, and in some cases, the number to be 
-represented is an approximation. As an example, consider that the number to be stored is 1.109. Following the 
-traditional system, the number 1,109 is represented by
-
-\begin{equation}\nonumber
-  0011111111110001101111100111011011001000101101000011100101011000
-\end{equation}
-
-We consider that the power used is $10^3$, that is, the number which needs to be stored would be 1109. In binary, that 
-number is represented by 10001010101. Note that the number of bits required is smaller, but depending on numerical 
-accuracy, the compression methods suggested are ineffective. When the power of 10 is used, it must be applied across the 
-matrix. This conversion allows the proposed methods to be applied to a set of real numbers. It should be taken into 
-account that the greater the power of 10 used, which is associated with the conversion of the real number, the larger 
-the set of bits will be required to represent the converted number. This affects the efficiency of data compression. The 
-use of the library to represent real numbers is conditioned to a study of the necessary level of precision to solve the 
-problem.
-
-\section*{Library Application}\todo{Também deve ir para o material suplementar}
-
-In order to create an application, some elementary matrix operations. Different operations have been performed by 
-measuring the execution time of each. We basically, considered the attribution of a constant to all the elements of a 
-matrix; the addition, subtraction and multiplication of the elements of a matrix by a constant, the addition, 
-subtraction and multiplication of matrices. In addition to that, the calculation of  the transposed matrix and the 
-maximum an array were also considered. This measure was used to evaluate the efficiency of the method implemented when 
-compared to traditional methods.
-
-One of the characteristics of bitstring matrices is that the operations per column, in the case of implementation in 
-Fortran, can be parallelized. This reduces the execution time, compared to traditional methods. Therefore, in order to 
-achieve this optimization, threads implementations were used for the implemented operations.
-
-As a test methodology, all the described operations were also compared to traditional methods. The matrix dimensions n x 
-n used were $n = 10$, $100$, $1,000$, $10,000$, $20,000$ to $100,000$, by $10,000$. We emphasize that the matrix,  
-created with a 1,000 dimension has $10^6$ elements, while the one with a 100,000 dimension has $1 \times 10^{10}$ 
-elements. With the increase of the size of the matrices, a larger amount of memory must be reserved to allocate the 
-numbers. The compression effect of the matrices can be verified in the results since by the proposed method the matrix 
-is ​​directly loaded into the main memory, avoiding the need for access to the secondary memory.
-
-In Figures \ref{fig:23242526} and \ref{fig:27282930} the results from the processing time (in milliseconds) 
-for the operations that involved only integer ($Z$) and positive integer numbers ($Z^+$) are represented. 
-The traditional method (Normal) and the Bitstring method (Bitstring $Z$, integer numbers and Bitstring Zp, 
-positive integer numbers) were used. The operations were attribution ($A=5$, A is a matrix with dimension n by n); 
-sum of the elements of a matrix with a constant ($A=A+5$); sum of matrices ($A=A+A$); andsubtraction of the elements 
-of a matrix with a constant ($A=A-8$). These operations are represented in Figure \ref{fig:23242526}. 
-Figure \ref{fig:27282930}, in turn, shows the results for the multiplicationoperations of matrices ($A = A \times A$); 
-for the multiplication of the the elements of a matrix by a constant ($A=A \times 2$); 
-and for the calculation of transpose and maximum of a matrix.
-
-The results show that the bitstring method is faster than the traditional ones only for the attribution.
-This occurs because the element of the matrix is used more than once to store the numbers, differently 
-from the traditional method. Due to the number of manipulations  the bitstring method does, the other 
-operations  demand more time to be processed, but this can be optimized. When applied to positive integers, 
-the bitstring model works faster than when applied to integers because it does not need to do operations 
-related to the signals.
-
-The results for the real numbers are represented in Figures \ref{fig:31323334} and \ref{fig:35363738}. 
-By repeating the operations tested in the integer numbers, it was possible to see that the bitstring method 
-cannot overcome the traditional method. This happensbecause its application to  the real numbers, as alternative 
-to maintain the compression, asks for a conversion of real to integer numbers in order for them to be stored and 
-when accessed, another conversion is made. This is a problem from the implementation process, but it can be optimized.
-
-Another library application was the implementation of the algorithm in order to deal with the problem of arrays of 
-collaborative filtering \cite{cf}. These arrays have large dimensions due to the number of users and products. Each 
-element of this matrix stores the rating of a user with respect to a specific product. With this information, it is 
-possible to specify a product for a determined customer.
-
-The manipulation of these matrices, depending on the amount of memory to be allocated, are one of the possible 
-applications of the methods proposed in this work, since it promotes the compression and data allocation in the main 
-memory. The allocation of the matrix within the main memory makes the access time to the elements much smaller than when 
-the matrix is stored in secondary memory. This latter type of storage is used in some already mentioned techniques. As 
-an example, let us examine the manipulation of a CF matrix, evaluating the access time to the elements and especially 
-the size of the memory required to run such operation. The algorithms applied to this matrix were the average per-use 
-and the bias from mean \cite{cf}, this last one of the simplest algorithms for predicting the rate of users.
-
-An example of the application of the methodology presented is shown in Figure \ref{fig22}. In this figure, the matrix 
-with normal and bitstring formats are displayed. In both matrices, the information is represented and can be accessed. 
-Therefore, it becomes possible to apply the selected algorithms.
-
-In order to test a possible model application, we considered a matrix with the following dimensions: 95,000 users 
-(lines) by 3,000 movies (columns), with the ratings for Netflix’s movies
-\footnote{http://graphlab.org/downloads/datasets/}. In order to evaluate a larger dataset, we modified the 
-NetFlix data so we could increase the dimension of the matrix to be analyzed. The 95,000 users correspond to 
-thelines of the matrix. In orderto gradually increase the number of lines, we performed consecutive re-samplings 
-and added them to the original matrix . As a results, we obtained a  matrix with 600,000 lines and 3,000 columns. 
-The operations used with the matrix were the Per User Average (PUA) and the rating estimation of a movie i using 
-the Bias From Mean (BFM) \cite{cf}. The analyses were  made, considering the created matrix with different dimensions, 
-with n from 40,000 to 600,000 by 20,000. It is important to emphasize that the number of columns is constant and 
-equal to 3,000. The implementations of this problem were made in Fortran and both the developedlibrary based in 
-the method SM and in the traditional form were used. Codes were inserted in the library in order to optimize the 
-multicore processing using threads. No additional configuration was necessary to use the multicore processing. 
-since the library detects the characteristics of the processor uses them in the simulations. In turn, for the 
-traditional implementation, weused the optimization process and parallelization, which are made available by 
-the processor,as the vectorization and multicore parallelization.
-
-The results are represented in Figures \ref{fig39}, \ref{fig40} and \ref{fig41} for the PUA Method and in Figures 
-\ref{fig42}, \ref{fig43} and \ref{fig44} for the 
-BFM Methods.  A computer with 8 GB of RAM Memory, 16 GB of swap, processor core i7 model 870  \todo{confirmar modelo}
-and operational 
-system Linux Debian 7.0 was used to perform this simulation. In Figure \ref{fig39}, it is possible to verify the approximate 
-moment  when the  the storage in disk (swaap) begins to be used. This is indicated by the vertical green line. 
-At this moment, the bitstring method becomes faster than the traditional method because it does not require disk 
-storage due tocompressed memory representation. Note that, in the Figure \ref{fig40}, the traditional method is faster 
-than the bitstring one before the beginning of the disk access. However, after it becomes necessary use the disk 
-so that the matrix can be accessed and stored, , the traditional method  becames slower (Figura \ref{fig41}). 
-These results 
-indicate that the bistring method can promote the optimization of the memory use because the access to the main 
-memories (RAM and cache) are faster than the access to secondary memories (disk). In Figures \ref{fig42}, 
-\ref{fig43} and \ref{fig44} the 
-results of the processing time for  the BFM method are represented. Once again, a vertical green line indicates 
-the approximate moment when the disk access begins. There is a difference in relation to the PUA method since the 
-BMF method uses the PUA values to perform its calculation. Therefore, the BMF method demands more memory to be 
-processed, which reduces the matrix dimensions involved in the process, represented in the main memory. As 
-indicated in Figure \ref{fig41}, for the PUA method, and Figure \ref{fig44}, for the BMF method, it is possible 
-to verify that 
-before the access to the disk, the traditional method is faster than the bitstring one, but once the access 
-is initiated,  the bitstring method becames faster.
-
-In spite of the simplicity of the application, it is relevant since  it allows for large-sized matrices to be 
-allocated in memory and manipulated.
-
 \section*{Conclusion}
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.4,trim=5cm 0cm 5cm 0cm,clip]{fig22}
-  \caption{Illustration of the process of the bistring model application to matrices of collaborative filttering of the 
-algorithms per use average (PUA) and of the bias from mean (BFM) in a matrix represented in the traditional and 
-bitstring formats.}
-  \label{fig22}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \subfigure[$A=5$]{
-  \includegraphics[scale=0.4,clip]{fig23}
-  }
-  \subfigure[$A=A+5$]{
-  \includegraphics[scale=0.4,clip]{fig24}
-  }
-  \subfigure[$A=A+A$]{
-  \includegraphics[scale=0.4,clip]{fig25}
-  }
-  \subfigure[$A=A-8$]{
-  \includegraphics[scale=0.4,clip]{fig26}
-  }
-  \caption{Processing time of the following operations: attribution ($A=5$, A is the matrix with an n by n dimension); 
-  element sum of a matrix with a constant ($A=A+5$);sum of matrices ($A=A+A$); and subtraction of the elements of a 
-  matrix by a constant ($A=A-8$) involving integer numbers ($Z$) and positive integer numbers ($Z^+$). The bitstring 
-  method presented the best overall performance only for the attribution operations, because the storage of 
-  different numbers occurs in one element of a matrix, which reduces the accessing time. For the other operations, 
-  the bitstring method is slower due to the number ofoperations which are necessary to be  executed in order to 
-  access the elements. This can still be optimized in the implemented library.}
-  \label{fig:23242526}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \subfigure[$A = A \times A$]{
-  \includegraphics[scale=0.4,clip]{fig27}
-  }
-  \subfigure[$A=A \times 2$]{
-  \includegraphics[scale=0.4,clip]{fig28}
-  }
-  \subfigure[$transpose(A)$]{
-  \includegraphics[scale=0.4,clip]{fig29}
-  }
-  \subfigure[$maximum(A)$]{
-  \includegraphics[scale=0.4,clip]{fig30}
-  }
-  \caption{Processing time of matrix multiplication ($A = A \times A$, A is the matrix with a n by n dimension), 
-  element multiplication of matrix by constant ($A=A \times 2$), the calculation of transpose and maximum of a 
-  matrix involve integer numbers ($Z$) and positive integer numbers ($Z^+$). In this cases, the traditional 
-  method was faster than the bitstring model.}
-  \label{fig:27282930}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \subfigure[$A=5.0000$]{
-   \includegraphics[scale=0.4,clip]{fig31}
-  }
-  \subfigure[$A=A+5.0000$]{
-   \includegraphics[scale=0.4,clip]{fig32}
-  }
-  \subfigure[$A=A+A$]{
-   \includegraphics[scale=0.4,clip]{fig33}
-  }
-  \subfigure[$A=A-8.0000$]{
-   \includegraphics[scale=0.4,clip]{fig34}
-  }
-  \caption{Processing time of the following operations: attribution ($A=5.0000$, A is the matrix with an n by n dimension); 
-  element sum of a matrix with a constant ($A=A+5.000$);sum of matrices ($A=A+A$); and subtraction of the elements of a 
-  matrix by a constant ($A=A-8.0000$) involving real numbers ($R$) and positive integer numbers ($R^+$). The bitstring 
-  method presented the best overall performance only for the attribution operations, because the storage of 
-  different numbers occurs in one element of a matrix, which reduces the accessing time. For the other operations, 
-  the bitstring method is slower due to the number ofoperations which are necessary to be  executed in order to 
-  access the elements. This can still be optimized in the implemented library.}
-  \label{fig:31323334}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \subfigure[$A = A \times A$]{
-   \includegraphics[scale=0.4,clip]{fig35}
-  }
-  \subfigure[$A=A \times 2$]{
-   \includegraphics[scale=0.4,clip]{fig36}
-  }
-  \subfigure[$transpose(A)$]{
-   \includegraphics[scale=0.4,clip]{fig37}
-  }
-  \subfigure[$maximum(A)$]{
-   \includegraphics[scale=0.4,clip]{fig38}
-  }
-  \caption{Processing time of matrix multiplication ($A = A \times A$, A is the matrix with a n by n dimension), 
-  element multiplication of matrix by constant ($A=A \times 2.0000$), the calculation of transpose and maximum of a 
-  matrix involve integer numbers ($R$) and positive integer numbers ($R^+$). In this cases, the traditional 
-  method was faster than the bitstring model.}
-  \label{fig:35363738}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.6,clip]{fig39}
-  \caption{Processing time for the Per User Average method (PUA) in function of the numbers of linesof a matrix 
-  with 3,000 columns. The vertical green line indicates the approximate moment when the traditional method  
-  demands  disk access in order to store the matrix to be analyzed, which  overloads the process. The bitstring 
-  method fully represents a matrix  in memory, without the necessity to access the  disk.}
-  \label{fig39}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.6,clip]{fig40}
-  \caption{Processing time for the Per User Average method (PUA) in function of the  numbers of lines of a matrix 
-  with 3,000 columns.  The vertical green line indicates the approximate moment when the traditional method  
-  demands disk access in order to store the matrixto be analyzed, which  overloads the process. For dimensions 
-  inferior to approximately 400 lines, the traditional method is faster than the bitstring one. It is important 
-  to emphasize that the computational cost of both methods is linear.}
-  \label{fig40}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.6,clip]{fig41}
-  \caption{Relation between the processing time using the bitstring method and the traditional method in function 
-  of the line numbers of a matrix with 3,000 columns. It is possible to verify that for a matrix with less than 400 
-  lines, the traditional method is faster. However, when the number of lines is greater  than 400 lines, the best 
-  method is the bitstring because of the compressed representation in the matrix memory.}
-  \label{fig41}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.6,clip]{fig42}
-  \caption{Processing time for the Bias From Mean method (BMF) in function of the numbers of linesof a matrix 
-  with 3,000 columns. The vertical green line indicates the approximate moment when the traditional method  
-  demands  disk access in order to store the matrix to be analyzed, which  overloads the process. The bitstring 
-  method fully represents a matrix  in memory, without the necessity to access the  disk.}
-  \label{fig42}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.6,clip]{fig43}
-  \caption{Processing time for the Bias From Mean method (BMF) in function of the  numbers of lines of a matrix 
-  with 3,000 columns.  The vertical green line indicates the approximate moment when the traditional method  
-  demands disk access in order to store the matrixto be analyzed, which  overloads the process. For dimensions 
-  inferior to approximately 400 lines, the traditional method is faster than the bitstring one. It is important 
-  to emphasize that the computational cost of both methods is linear.}
-  \label{fig43}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.6,clip]{fig44}
-  \caption{Relation between the processing time using the bitstring method and the traditional method in function 
-  of the line numbers of a matrix with 3,000 columns. It is possible to verify that for a matrix with less than 400 
-  lines, the traditional method is faster. However, when the number of lines is greater  than 400 lines, the best 
-  method is the bitstring because of the compressed representation in the matrix memory.}
-  \label{fig44}
-\end{figure}
-
-%\begin{figure}[!ht]
-%\begin{center}
-%%\includegraphics[width=4in]{figure_name.2.eps}
-%\end{center}
-%\caption{
-%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
-%should be left justified, as specified by the options to the caption 
-%package.
-%}
-%\label{Figure_label}
-%\end{figure}
-
-\newpage
-
-\clearpage
-
 \section*{Tables}

### 2014-02-12 suppl.tex
Alterações realizadas

@@
 %
-\documentclass[12pt]{article}
+\documentclass[10pt]{article}
 
@@
 
+\usepackage{psfrag}
+\usepackage{subfigure}
+
 \usepackage[normalem]{ulem}
+
+% Text layout
+\topmargin 0.0cm
+\oddsidemargin 0.5cm
+\evensidemargin 0.5cm
+\textwidth 16cm 
+\textheight 21cm
 
@@
 kinds of operations.
+
+It also presents an application of the developed library. The results of 
+operations implemented and implementation are presented in a matrix problem of 
+collaborative filtering.\todo{Inseri para colocar a parte da implementação.}
 
@@
 
+\section{Implementation}
+
+This paper proposes the bitstring-compression methodologies. The SM 
+method was implemented as a Fortran library which is available under an 
+open-source license\todo{colocar o link para o repositório aqui}\todo{Acho 
+melhor mandar para o artigo e anexar a biblioteca, para depois disponibilizar 
+o código.}. The 
+implementation extends the standard Matrix type of Fortran, overloading 
+operators such as assignment, addition, subtraction, multiplication, transpose, 
+and maximum of a matrix. 
+Although only few significant matrix operations have been ovearloaded in this 
+example implementation, any other operation can be performed, since 
+the methods for inserting and collecting an element in a matrix of bitstring 
+were also implemented. With these two methods, as long as they are properly 
+adapted into algorithms, it is possible to implement any other desired 
+operation.
+
+The SM method was also extended to allow for the allocation of real and 
+signed integer matrices, besides positive integer ones. Some adjustments were 
+made so that the other types of numbers could be represented. In order to 
+represent the signed integers, it was necessary to use two's complement, as it 
+is applied to thein accordance to traditional methods\todo{acho que esta 
+explicação ficou muito solta. Não precisamos explica a implementação no 
+artigo, podemos colocar na documentação do código}. \todo{Eu deixaria, pois é 
+apenas um apontamento com relação as operações, pois no arquivo que 
+compartilhei com você eu coloco os códigos fontes com comentários.}
+
+The SM method uses a fixed number of bits to store the values in such a way that
+the size or number of bits used is important to represent the highest value in 
+the bitstring matrix. The adaptation 
+consists in determining the largest value to be stored considering the absolute 
+value of the number. That done,
+when the number is negative, it is necessary to rewrite it since, due to the 
+sign, this will be written in the form of 
+two's complement. The purpose of this conversion is to optimize library 
+processing. Take the following as an example: 
+the number 5, when represented in a 32-bit integer, is defined as:
+
+\begin{equation}\nonumber
+ 5 = 00000000 00000000 00000000 00000101
+\end{equation}
+
+However, if the number is negative, this is represented in the form of two's 
+complement. Therefore, the number -5 is 
+represented as:
+
+\begin{equation}\nonumber
+ -5 = 11111111 11111111 11111111 11111011
+\end{equation}
+
+Note that the bit positioned more to the left should be used to inform that the 
+number is -5. For the purpose of 
+optimization of the library, it was chosen to conversion of positive numbers to 
+negative numbers and use of 1 
+bit to represent the sign. (Dúvidas aqui)
+
+In turn, in order to represent real numbers, a conversion procedure of real 
+numbers to integer numbers was used. The 
+real number is multiplied by a power of ten and is then immediately truncated. 
+The real numbers in a binary system of 64 
+bits are represented using a bit to represent the sign of the number, 11 bits to 
+the exponent, and 52 bits to the 
+mantissa. Note that 1 bit is still used implicitly, according to the IEEE-754 
+standard. In summary, the mantissa is 
+represented in the base $2^{-1}$, $2^{-2}$, $2^{-3}$, $\hdots$, $2^{-52}$. 
+Depending on the value to be represented, 
+loss of information can take place, which is related to the accuracy of the type 
+used. In this case, considering a real 
+number with double precision, the demand of 64 bits to represent each one 
+occurs, and in some cases, the number to be 
+represented is an approximation. As an example, consider that the number to be 
+stored is 1.109. Following the 
+traditional system, the number 1,109 is represented by
+
+\begin{equation}\nonumber
+  0011111111110001101111100111011011001000101101000011100101011000
+\end{equation}
+
+We consider that the power used is $10^3$, that is, the number which needs to be 
+stored would be 1109. In binary, that 
+number is represented by 10001010101. Note that the number of bits required is 
+smaller, but depending on numerical 
+accuracy, the compression methods suggested are ineffective. When the power of 
+10 is used, it must be applied across the 
+matrix. This conversion allows the proposed methods to be applied to a set of 
+real numbers. It should be taken into 
+account that the greater the power of 10 used, which is associated with the 
+conversion of the real number, the larger 
+the set of bits will be required to represent the converted number. This affects 
+the efficiency of data compression. The 
+use of the library to represent real numbers is conditioned to a study of the 
+necessary level of precision to solve the 
+problem.
+
+\section*{Library Application}
+
+In order to create an application, some elementary matrix operations. Different 
+operations have been performed by 
+measuring the execution time of each. We basically, considered the attribution 
+of a constant to all the elements of a 
+matrix; the addition, subtraction and multiplication of the elements of a matrix 
+by a constant, the addition, 
+subtraction and multiplication of matrices. In addition to that, the calculation 
+of  the transposed matrix and the 
+maximum an array were also considered. This measure was used to evaluate the 
+efficiency of the method implemented when 
+compared to traditional methods.
+
+One of the characteristics of bitstring matrices is that the operations per 
+column, in the case of implementation in 
+Fortran, can be parallelized. This reduces the execution time, compared to 
+traditional methods. Therefore, in order to 
+achieve this optimization, threads implementations were used for the implemented 
+operations.
+
+As a test methodology, all the described operations were also compared to 
+traditional methods. The matrix dimensions n x 
+n used were $n = 10$, $100$, $1,000$, $10,000$, $20,000$ to $100,000$, by 
+$10,000$. We emphasize that the matrix,  
+created with a 1,000 dimension has $10^6$ elements, while the one with a 100,000 
+dimension has $1 \times 10^{10}$ 
+elements. With the increase of the size of the matrices, a larger amount of 
+memory must be reserved to allocate the 
+numbers. The compression effect of the matrices can be verified in the results 
+since by the proposed method the matrix 
+is ​​directly loaded into the main memory, avoiding the need for access to the 
+secondary memory.
+
+In Figures \ref{fig:23242526} and \ref{fig:27282930} the results from the 
+processing time (in milliseconds) 
+for the operations that involved only integer ($Z$) and positive integer numbers 
+($Z^+$) are represented. 
+The traditional method (Normal) and the Bitstring method (Bitstring $Z$, integer 
+numbers and Bitstring Zp, 
+positive integer numbers) were used. The operations were attribution ($A=5$, A 
+is a matrix with dimension n by n); 
+sum of the elements of a matrix with a constant ($A=A+5$); sum of matrices 
+($A=A+A$); andsubtraction of the elements 
+of a matrix with a constant ($A=A-8$). These operations are represented in 
+Figure \ref{fig:23242526}. 
+Figure \ref{fig:27282930}, in turn, shows the results for the 
+multiplicationoperations of matrices ($A = A \times A$); 
+for the multiplication of the the elements of a matrix by a constant ($A=A 
+\times 2$); 
+and for the calculation of transpose and maximum of a matrix.
+
+The results show that the bitstring method is faster than the traditional ones 
+only for the attribution.
+This occurs because the element of the matrix is used more than once to store 
+the numbers, differently 
+from the traditional method. Due to the number of manipulations  the bitstring 
+method does, the other 
+operations  demand more time to be processed, but this can be optimized. When 
+applied to positive integers, 
+the bitstring model works faster than when applied to integers because it does 
+not need to do operations 
+related to the signals.
+
+The results for the real numbers are represented in Figures \ref{fig:31323334} 
+and \ref{fig:35363738}. 
+By repeating the operations tested in the integer numbers, it was possible to 
+see that the bitstring method 
+cannot overcome the traditional method. This happensbecause its application to  
+the real numbers, as alternative 
+to maintain the compression, asks for a conversion of real to integer numbers in 
+order for them to be stored and 
+when accessed, another conversion is made. This is a problem from the 
+implementation process, but it can be optimized.
+
+Another library application was the implementation of the algorithm in order to 
+deal with the problem of arrays of 
+collaborative filtering \cite{cf}. These arrays have large dimensions due to the 
+number of users and products. Each 
+element of this matrix stores the rating of a user with respect to a specific 
+product. With this information, it is 
+possible to specify a product for a determined customer.
+
+The manipulation of these matrices, depending on the amount of memory to be 
+allocated, are one of the possible 
+applications of the methods proposed in this work, since it promotes the 
+compression and data allocation in the main 
+memory. The allocation of the matrix within the main memory makes the access 
+time to the elements much smaller than when 
+the matrix is stored in secondary memory. This latter type of storage is used in 
+some already mentioned techniques. As 
+an example, let us examine the manipulation of a CF matrix, evaluating the 
+access time to the elements and especially 
+the size of the memory required to run such operation. The algorithms applied to 
+this matrix were the average per-use 
+and the bias from mean \cite{cf}, this last one of the simplest algorithms for 
+predicting the rate of users.
+
+An example of the application of the methodology presented is shown in Figure 
+\ref{fig22}. In this figure, the matrix 
+with normal and bitstring formats are displayed. In both matrices, the 
+information is represented and can be accessed. 
+Therefore, it becomes possible to apply the selected algorithms.
+
+In order to test a possible model application, we considered a matrix with the 
+following dimensions: 95,000 users 
+(lines) by 3,000 movies (columns), with the ratings for Netflix’s movies
+\footnote{http://graphlab.org/downloads/datasets/}. In order to evaluate a 
+larger dataset, we modified the 
+NetFlix data so we could increase the dimension of the matrix to be analyzed. 
+The 95,000 users correspond to 
+thelines of the matrix. In orderto gradually increase the number of lines, we 
+performed consecutive re-samplings 
+and added them to the original matrix . As a results, we obtained a  matrix with 
+600,000 lines and 3,000 columns. 
+The operations used with the matrix were the Per User Average (PUA) and the 
+rating estimation of a movie i using 
+the Bias From Mean (BFM) \cite{cf}. The analyses were  made, considering the 
+created matrix with different dimensions, 
+with n from 40,000 to 600,000 by 20,000. It is important to emphasize that the 
+number of columns is constant and 
+equal to 3,000. The implementations of this problem were made in Fortran and 
+both the developedlibrary based in 
+the method SM and in the traditional form were used. Codes were inserted in the 
+library in order to optimize the 
+multicore processing using threads. No additional configuration was necessary to 
+use the multicore processing. 
+since the library detects the characteristics of the processor uses them in the 
+simulations. In turn, for the 
+traditional implementation, weused the optimization process and parallelization, 
+which are made available by 
+the processor,as the vectorization and multicore parallelization.
+
+The results are represented in Figures \ref{fig39}, \ref{fig40} and \ref{fig41} 
+for the PUA Method and in Figures 
+\ref{fig42}, \ref{fig43} and \ref{fig44} for the 
+BFM Methods.  A computer with 8 GB of RAM Memory, 16 GB of swap, processor core 
+i7 model 870  \todo{confirmar modelo}
+and operational 
+system Linux Debian 7.0 was used to perform this simulation. In Figure 
+\ref{fig39}, it is possible to verify the approximate 
+moment  when the  the storage in disk (swaap) begins to be used. This is 
+indicated by the vertical green line. 
+At this moment, the bitstring method becomes faster than the traditional method 
+because it does not require disk 
+storage due tocompressed memory representation. Note that, in the Figure 
+\ref{fig40}, the traditional method is faster 
+than the bitstring one before the beginning of the disk access. However, after 
+it becomes necessary use the disk 
+so that the matrix can be accessed and stored, , the traditional method  becames 
+slower (Figura \ref{fig41}). 
+These results 
+indicate that the bistring method can promote the optimization of the memory use 
+because the access to the main 
+memories (RAM and cache) are faster than the access to secondary memories 
+(disk). In Figures \ref{fig42}, 
+\ref{fig43} and \ref{fig44} the 
+results of the processing time for  the BFM method are represented. Once again, 
+a vertical green line indicates 
+the approximate moment when the disk access begins. There is a difference in 
+relation to the PUA method since the 
+BMF method uses the PUA values to perform its calculation. Therefore, the BMF 
+method demands more memory to be 
+processed, which reduces the matrix dimensions involved in the process, 
+represented in the main memory. As 
+indicated in Figure \ref{fig41}, for the PUA method, and Figure \ref{fig44}, for 
+the BMF method, it is possible 
+to verify that 
+before the access to the disk, the traditional method is faster than the 
+bitstring one, but once the access 
+is initiated,  the bitstring method becames faster.
+
+In spite of the simplicity of the application, it is relevant since  it allows 
+for large-sized matrices to be 
+allocated in memory and manipulated.
+
+\newpage
+
+\section*{Figure Legends}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.4,trim=5cm 0cm 5cm 0cm,clip]{fig22}
+  \caption{Illustration of the process of the bistring model application to 
+matrices of collaborative filttering of the 
+algorithms per use average (PUA) and of the bias from mean (BFM) in a matrix 
+represented in the traditional and 
+bitstring formats.}
+  \label{fig22}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$A=5$]{
+  \includegraphics[scale=0.4,clip]{fig23}
+  }
+  \subfigure[$A=A+5$]{
+  \includegraphics[scale=0.4,clip]{fig24}
+  }
+  \subfigure[$A=A+A$]{
+  \includegraphics[scale=0.4,clip]{fig25}
+  }
+  \subfigure[$A=A-8$]{
+  \includegraphics[scale=0.4,clip]{fig26}
+  }
+  \caption{Processing time of the following operations: attribution ($A=5$, A is 
+the matrix with an n by n dimension); 
+  element sum of a matrix with a constant ($A=A+5$);sum of matrices ($A=A+A$); 
+and subtraction of the elements of a 
+  matrix by a constant ($A=A-8$) involving integer numbers ($Z$) and positive 
+integer numbers ($Z^+$). The bitstring 
+  method presented the best overall performance only for the attribution 
+operations, because the storage of 
+  different numbers occurs in one element of a matrix, which reduces the 
+accessing time. For the other operations, 
+  the bitstring method is slower due to the number ofoperations which are 
+necessary to be  executed in order to 
+  access the elements. This can still be optimized in the implemented library.}
+  \label{fig:23242526}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$A = A \times A$]{
+  \includegraphics[scale=0.4,clip]{fig27}
+  }
+  \subfigure[$A=A \times 2$]{
+  \includegraphics[scale=0.4,clip]{fig28}
+  }
+  \subfigure[$transpose(A)$]{
+  \includegraphics[scale=0.4,clip]{fig29}
+  }
+  \subfigure[$maximum(A)$]{
+  \includegraphics[scale=0.4,clip]{fig30}
+  }
+  \caption{Processing time of matrix multiplication ($A = A \times A$, A is the 
+matrix with a n by n dimension), 
+  element multiplication of matrix by constant ($A=A \times 2$), the calculation 
+of transpose and maximum of a 
+  matrix involve integer numbers ($Z$) and positive integer numbers ($Z^+$). In 
+this cases, the traditional 
+  method was faster than the bitstring model.}
+  \label{fig:27282930}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$A=5.0000$]{
+   \includegraphics[scale=0.4,clip]{fig31}
+  }
+  \subfigure[$A=A+5.0000$]{
+   \includegraphics[scale=0.4,clip]{fig32}
+  }
+  \subfigure[$A=A+A$]{
+   \includegraphics[scale=0.4,clip]{fig33}
+  }
+  \subfigure[$A=A-8.0000$]{
+   \includegraphics[scale=0.4,clip]{fig34}
+  }
+  \caption{Processing time of the following operations: attribution ($A=5.0000$, 
+A is the matrix with an n by n dimension); 
+  element sum of a matrix with a constant ($A=A+5.000$);sum of matrices 
+($A=A+A$); and subtraction of the elements of a 
+  matrix by a constant ($A=A-8.0000$) involving real numbers ($R$) and positive 
+integer numbers ($R^+$). The bitstring 
+  method presented the best overall performance only for the attribution 
+operations, because the storage of 
+  different numbers occurs in one element of a matrix, which reduces the 
+accessing time. For the other operations, 
+  the bitstring method is slower due to the number ofoperations which are 
+necessary to be  executed in order to 
+  access the elements. This can still be optimized in the implemented library.}
+  \label{fig:31323334}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[$A = A \times A$]{
+   \includegraphics[scale=0.4,clip]{fig35}
+  }
+  \subfigure[$A=A \times 2$]{
+   \includegraphics[scale=0.4,clip]{fig36}
+  }
+  \subfigure[$transpose(A)$]{
+   \includegraphics[scale=0.4,clip]{fig37}
+  }
+  \subfigure[$maximum(A)$]{
+   \includegraphics[scale=0.4,clip]{fig38}
+  }
+  \caption{Processing time of matrix multiplication ($A = A \times A$, A is the 
+matrix with a n by n dimension), 
+  element multiplication of matrix by constant ($A=A \times 2.0000$), the 
+calculation of transpose and maximum of a 
+  matrix involve integer numbers ($R$) and positive integer numbers ($R^+$). In 
+this cases, the traditional 
+  method was faster than the bitstring model.}
+  \label{fig:35363738}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig39}
+  \caption{Processing time for the Per User Average method (PUA) in function of 
+the numbers of linesof a matrix 
+  with 3,000 columns. The vertical green line indicates the approximate moment 
+when the traditional method  
+  demands  disk access in order to store the matrix to be analyzed, which  
+overloads the process. The bitstring 
+  method fully represents a matrix  in memory, without the necessity to access 
+the  disk.}
+  \label{fig39}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig40}
+  \caption{Processing time for the Per User Average method (PUA) in function of 
+the  numbers of lines of a matrix 
+  with 3,000 columns.  The vertical green line indicates the approximate moment 
+when the traditional method  
+  demands disk access in order to store the matrixto be analyzed, which  
+overloads the process. For dimensions 
+  inferior to approximately 400 lines, the traditional method is faster than the 
+bitstring one. It is important 
+  to emphasize that the computational cost of both methods is linear.}
+  \label{fig40}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig41}
+  \caption{Relation between the processing time using the bitstring method and 
+the traditional method in function 
+  of the line numbers of a matrix with 3,000 columns. It is possible to verify 
+that for a matrix with less than 400 
+  lines, the traditional method is faster. However, when the number of lines is 
+greater  than 400 lines, the best 
+  method is the bitstring because of the compressed representation in the matrix 
+memory.}
+  \label{fig41}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig42}
+  \caption{Processing time for the Bias From Mean method (BMF) in function of 
+the numbers of linesof a matrix 
+  with 3,000 columns. The vertical green line indicates the approximate moment 
+when the traditional method  
+  demands  disk access in order to store the matrix to be analyzed, which  
+overloads the process. The bitstring 
+  method fully represents a matrix  in memory, without the necessity to access 
+the  disk.}
+  \label{fig42}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig43}
+  \caption{Processing time for the Bias From Mean method (BMF) in function of 
+the  numbers of lines of a matrix 
+  with 3,000 columns.  The vertical green line indicates the approximate moment 
+when the traditional method  
+  demands disk access in order to store the matrixto be analyzed, which  
+overloads the process. For dimensions 
+  inferior to approximately 400 lines, the traditional method is faster than the 
+bitstring one. It is important 
+  to emphasize that the computational cost of both methods is linear.}
+  \label{fig43}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6,clip]{fig44}
+  \caption{Relation between the processing time using the bitstring method and 
+the traditional method in function 
+  of the line numbers of a matrix with 3,000 columns. It is possible to verify 
+that for a matrix with less than 400 
+  lines, the traditional method is faster. However, when the number of lines is 
+greater  than 400 lines, the best 
+  method is the bitstring because of the compressed representation in the matrix 
+memory.}
+  \label{fig44}
+\end{figure}
+
+%\begin{figure}[!ht]
+%\begin{center}
+%%\includegraphics[width=4in]{figure_name.2.eps}
+%\end{center}
+%\caption{
+%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
+%should be left justified, as specified by the options to the caption 
+%package.
+%}
+%\label{Figure_label}
+%\end{figure}
+
+\newpage
+
+\clearpage
+
 \bibliography{plos.bib}

### 2014-04-04 suppl.tex
revisei o suplemento e o Manuscrito. Deixei alguns todos no Suplemento para informação. por favor remova @crysttian

@@
 bitstring compressed arrays. We will consider integer matrices compressed via
-both the SM and the VLB methods, \todo{inseri esse texto} but the proposed 
+both the SM and the VLB methods, but the proposed 
 methods can be applied to integers and real numbers with sign.
@@
 
-It also presents an application of the developed library. The results of 
-operations implemented and implementation are presented in a matrix problem of 
-collaborative filtering.\todo{Inseri para colocar a parte da implementação.}
+It also presents an application of the developed library. We apply our 
+implementation to classical machine learning problem: collaborative filtering.
 
@@
     \centering
-    \caption{Comparison of the sum in binary and decimal base.}
+    \caption{Comparing binary and decimal sum of 7+7.}
     \begin{tabular}{crrrc}
     \hline
-    	  & Decimal & & Binario & Bit-length \\
+    	  & Decimal & & Binary & Bit-length \\
     \hline      
@@
 addition subtraction and multiplication operation the same as those for unsigned
-binary numbers. This algorithm \todo{removi is what} is actually implemented in 
+binary numbers. This algorithm is actually implemented in 
 CPUs. To exemplify let's subtract $7$ from $10$. To make use of the
@@
 
-\todo{adicionei esse trecho - As can be seen in} As can be seen in figure 
+As can be seen in figure 
 \ref{fig:bitlength}, the bit-lengths of results of the four
@@
 extra bits, thus we end up with a bit-length of 12 for each element of the
-product. Having determined the bit-length of the product, we can now \todo{we 
-can - removi} do the actual calculations and store the results.
+product. Having determined the bit-length of the product, we can now do the 
+actual calculations and store the results.
 
@@
 method was implemented as a Fortran library which is available under an 
-open-source license\todo{colocar o link para o repositório aqui}\todo{Acho 
-melhor mandar para o artigo e anexar a biblioteca, para depois disponibilizar 
-o código.}. The 
+open-source license. The 
 implementation extends the standard Matrix type of Fortran, overloading 
@@
 signed integer matrices, besides positive integer ones. Some adjustments were 
-made so that the other types of numbers could be represented. In order to 
-represent the signed integers, it was necessary to use two's complement, as it 
-is applied to thein accordance to traditional methods\todo{acho que esta 
-explicação ficou muito solta. Não precisamos explica a implementação no 
-artigo, podemos colocar na documentação do código}. \todo{Eu deixaria, pois é 
-apenas um apontamento com relação as operações, pois no arquivo que 
-compartilhei com você eu coloco os códigos fontes com comentários.}
+made so that the other types of numbers could be represented. \todo{removi 
+``in order to represent...''}
 
@@
 BFM Methods.  A computer with 8 GB of RAM Memory, 16 GB of swap, processor core 
-i7 model 870  \todo{confirmar modelo}
+i7 model 870
 and operational 
